### 2020-10-19 jss-paper/jss-paper.Rmd
Initial draft

---
documentclass: jss
author:
  - name: Mladen Jovanović
    affiliation: |
      | University of Belgrade
    address: |
      | Faculty of Sport and Physical Education
      | Belgrade, Serbia
    email: \email{coach.mladen.jovanovic@gmail.com}
  - name: Jason D. Vescovi
    affiliation: |
      | University of Toronto
    address: |
      | Faculty of Kinesiology and Physical Education
      | Graduate School of Exercise Science
      | Toronto, ON Canada
    email: \email{vescovij@gmail.com}
title:
  formatted: "\\pkg{shorts}: An R Package for Modeling Short Sprints"
abstract: >
  The abstract of the article.
keywords:
  formatted: [sprints, sports, modeling]
preamble: >
  \usepackage{amsmath}
output:
  bookdown::pdf_book:
    base_format: rticles::jss_article
citation_package: natbib
bibliography: [references.bib, packages.bib]
biblio-style: apalike
link-citations: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
require(knitr)
require(shorts)
require(tidyverse)

my_random_seed <- 1667
set.seed(my_random_seed)

options("width"=96)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
  dpi = 600,
  out.width = "90%",
  fig.align = 'center',
  fig.width = 6,
  fig.height = 6 * 0.618,  # 1 / phi
  fig.show = "hold",
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  width = 96
)

# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'shorts'
), 'packages.bib')
```

# Introduction

Short sprint speed is one of the most distinguishable and admired physical traits in sports, and one that have puzzled coaches and sports scientists (if they were available back then) from the time when two humanoids raced against each other (and other beasts). Short sprints are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end. This is approximately less than 6 seconds in duration (why? ATP/CP? --- REF), but the distance can vary based on the level of the athlete (i.e. lower level athletes can reach this point earlier, around 30-40m, while elite sprinters much later, around 50-60m --- REF).

Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time) originally proposed by @doi:10.1098/rspb.1927.0035, and more recently popularized by @greenePredictingSprintDynamics1986, @clarkNFLCombine40Yard2017, @chellyLegPowerHopping2001, @morinSimpleMethodComputing2019, @morinInterpretingPowerForceVelocityProfiles2016, @haugenPowerForceVelocityProfilingSprinting2020, @haugenSprintMechanicalVariables2019, @haugenSprintMechanicalProperties2020, @doi:10.1098/rspb.1927.0035, @haugenDifferenceStartImpact2012, and @samozinoSimpleMethodMeasuring2016:

\begin{equation}
  v(t) = MSS \times (1 - e^{-\frac{t}{TAU}}) (\#eq:velocity-time)
\end{equation}

The parameters of the equation \@ref(eq:velocity-time) are \textit{maximum sprinting speed} (MSS; expressed in $ms^-1$) and \textit{relative acceleration} (TAU). Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in $ms^-2$) \@ref(eq:maximal-acceleration).

\begin{equation}
  MAC = \frac{MSS}{TAU}(\#eq:maximal-acceleration)
\end{equation}

Although TAU is used in the equations, and later estimated, it is preferred to use MAC instead since it is easier to grasp, particularly for less math inclined coaches.

By derivating equation \@ref(eq:velocity-time), we can get equation for acceleration \@ref(eq:acceleration-time).

\begin{equation}
  a(t) = \frac{MSS}{TAU} \times e^{-\frac{t}{TAU}}  (\#eq:acceleration-time)
\end{equation}

By integrating equation \@ref(eq:velocity-time), we can get equation for distance \@ref(eq:distance-time).

\begin{equation}
  d(t) = MSS \times (t + TAU \times e^{-\frac{t}{TAU}}) - MSS \times TAU  (\#eq:distance-time)
\end{equation}

## Four athletes

Let's consider four athletes with different levels of MSS (high versus low maximal sprinting speed) and MAC (high versus low maximal acceleration; as mentioned previously, using MAC is preferred over using TAU) (Table \@ref(tab:four-athletes-table)).

```{r four-athletes-table, echo=FALSE}
athletes <- tribble(
  ~Athlete, ~MSS, ~MAC,
  "Athlete A", 12, 10,
  "Athlete B", 12, 6,
  "Athlete C", 8, 10,
  "Athlete D", 8, 6
)

athletes <- athletes %>%
  mutate(TAU = MSS / MAC)

knitr::kable(
  athletes,
  caption = "Four athletes with different MSS and MAC parameters.",
  digits = 2)
```

The Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6s).

```{r four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds."}
kinematics <- expand_grid(
  athletes,
  time = seq(0, 6, length.out = 1000)
) %>%
  mutate(
    distance = predict_distance_at_time(time, MSS, TAU),
    velocity = predict_velocity_at_time(time, MSS, TAU),
    acceleration = predict_acceleration_at_time(time, MSS, TAU)
  )

kinematics_long <- kinematics %>%
  pivot_longer(
    cols = c("distance", "velocity", "acceleration"),
    names_to = "variable" ) %>%
  mutate(
    variable = factor(
      variable,
      levels = c("distance", "velocity", "acceleration"),
      labels = c("Distance (m)", "Velocity (m/s)", "Acceleration (m/s/s)"))
  )

gg <- ggplot(
  kinematics_long,
  aes(x = time, y = value, color = Athlete)
) +
  theme_bw() +
  geom_line() +
  facet_wrap(~variable, scales = "free_y") +
  ylab(NULL) +
  xlab("Time (s)")

plot(gg)
```

If we plot acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is, according to the mathematical model, linear.

```{r four-athletes-profile, echo=FALSE, fig.cap="Acceleration-Velocity profile of four athletes with different MSS and MAC parameters."}
gg <- ggplot(
  kinematics,
  aes(x = velocity, y = acceleration, color = Athlete)
) +
  theme_bw() +
  geom_line() +
  ylab("Acceleration (m/s/s)") +
  xlab("Velocity (m/s)")

plot(gg)
```

# Estimation using \pkg{shorts} package

Short sprints profiling usually comes in two types: (1) split times using timing gates, and (2) velocity using radar gun.

### 2020-10-19 jss-paper/paper.Rmd
Initial draft

---
citation_package: natbib
bibliography:
- references.bib
- packages.bib
biblio-style: apalike
link-citations: yes
editor_options:
  chunk_output_type: console
output:
  bookdown::pdf_document2: default
output_dir: "docs"
---

```{r setup, include=FALSE}
require(knitr)
require(shorts)
require(tidyverse)
require(bookdown)

my_random_seed <- 1667
set.seed(my_random_seed)

options("width" = 80)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
  dpi = 600,
  out.width = "90%",
  fig.align = "center",
  fig.width = 6,
  fig.height = 6 * 0.618, # 1 / phi
  fig.show = "hold",
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  width = 80
)

# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
), "packages.bib")
```

# Introduction

Short sprint speed is one of the most distinguishable and admired physical traits in sports, and one that have puzzled coaches and sports scientists from the time when two humanoids raced against each other (if they were available back then). Short sprints are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end. This is approximately less than 6 seconds in duration (why? ATP/CP? --- REF), but the distance can vary based on the level of the athlete (i.e. lower level athletes can reach this point earlier, around 30-40m, while elite sprinters much later, around 50-60m --- REF).

Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time) originally proposed by @doi:10.1098/rspb.1927.0035, and more recently popularized by @greenePredictingSprintDynamics1986, @clarkNFLCombine40Yard2017, @chellyLegPowerHopping2001, @morinSimpleMethodComputing2019, @morinInterpretingPowerForceVelocityProfiles2016, @haugenPowerForceVelocityProfilingSprinting2020, @haugenSprintMechanicalVariables2019, @haugenSprintMechanicalProperties2020, @doi:10.1098/rspb.1927.0035, @haugenDifferenceStartImpact2012, and @samozinoSimpleMethodMeasuring2016:

\begin{equation}
  v(t) = MSS \times (1 - e^{-\frac{t}{TAU}}) (\#eq:velocity-time)
\end{equation}

The parameters of the equation \@ref(eq:velocity-time) are \textit{maximum sprinting speed} (MSS; expressed in $ms^-1$) and \textit{relative acceleration} (TAU). Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in $ms^-2$) \@ref(eq:maximal-acceleration).

\begin{equation}
  MAC = \frac{MSS}{TAU}(\#eq:maximal-acceleration)
\end{equation}

Although TAU is used in the equations, and later estimated, it is preferred to use MAC instead since it is easier to grasp, particularly for less math inclined coaches.

By derivating equation \@ref(eq:velocity-time), we can get equation for acceleration \@ref(eq:acceleration-time).

\begin{equation}
  a(t) = \frac{MSS}{TAU} \times e^{-\frac{t}{TAU}}  (\#eq:acceleration-time)
\end{equation}

By integrating equation \@ref(eq:velocity-time), we can get equation for distance \@ref(eq:distance-time).

\begin{equation}
  d(t) = MSS \times (t + TAU \times e^{-\frac{t}{TAU}}) - MSS \times TAU  (\#eq:distance-time)
\end{equation}

Let's consider four athletes with different levels of MSS (high versus low maximal sprinting speed) and MAC (high versus low maximal acceleration; as mentioned previously, using MAC is preferred over using TAU) (Table \@ref(tab:four-athletes-table)).

```{r four-athletes-table, echo=FALSE}
athletes <- tribble(
  ~Athlete, ~MSS, ~MAC,
  "Athlete A", 12, 10,
  "Athlete B", 12, 6,
  "Athlete C", 8, 10,
  "Athlete D", 8, 6
)

athletes <- athletes %>%
  mutate(TAU = MSS / MAC)

knitr::kable(
  athletes,
  caption = "Four athletes with different MSS and MAC parameters.",
  digits = 2
)
```

The Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6s).

```{r four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds."}
kinematics <- expand_grid(
  athletes,
  time = seq(0, 6, length.out = 1000)
) %>%
  mutate(
    distance = predict_distance_at_time(time, MSS, TAU),
    velocity = predict_velocity_at_time(time, MSS, TAU),
    acceleration = predict_acceleration_at_time(time, MSS, TAU)
  )

kinematics_long <- kinematics %>%
  pivot_longer(
    cols = c("distance", "velocity", "acceleration"),
    names_to = "variable"
  ) %>%
  mutate(
    variable = factor(
      variable,
      levels = c("distance", "velocity", "acceleration"),
      labels = c("Distance (m)", "Velocity (m/s)", "Acceleration (m/s/s)")
    )
  )

gg <- ggplot(
  kinematics_long,
  aes(x = time, y = value, color = Athlete)
) +
  theme_bw() +
  geom_line(alpha = 0.7) +
  facet_wrap(~variable, scales = "free_y") +
  ylab(NULL) +
  xlab("Time (s)") +
  theme(legend.position = "top")

plot(gg)
```

If we plot acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is, according to the mathematical model, linear.

```{r four-athletes-profile, echo=FALSE, fig.cap="Acceleration-Velocity profile of four athletes with different MSS and MAC parameters."}
gg <- ggplot(
  kinematics,
  aes(x = velocity, y = acceleration, color = Athlete)
) +
  theme_bw(10) +
  geom_line(alpha = 0.7) +
  ylab("Acceleration (m/s/s)") +
  xlab("Velocity (m/s)") +
  theme(legend.position = "top")

plot(gg)
```

# Estimation using `shorts` package

Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at 0m, 10m, 20m, 30m, 40m), and (2) getting a velocity trace using radar gun (usually at around 100Hz). Estimation of the MSS and TAU parameters from the equation \@ref(eq:velocity-time) is performed in `shorts` package using non-linear least squares regression implemented in the `nls` function in the base R [@R-base] and `nlme` function in the `nlme` package [@R-nlme] for the mixed-effect models.

## Estimating short sprint parameters using split times

Let's consider an example of an athlete with MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$m/s^2$ performing 40m sprint with timing gates positioned at each 10m split. When it comes to split times, distance is a predictor, and time is outcome variable, thus the equation \@ref(eq:velocity-time) takes another form:

\begin{equation}
  t(d) = TAU \times W(-e^{\frac{-d}{MSS \times TAU}} - 1) + \frac{d}{MSS} + TAU (\#eq:time-distance)
\end{equation}

$W$ in the equation \@ref(eq:velocity-time) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `shorts::model_using_splits` function:

```{r}
require(shorts)

split_distance <- c(10, 20, 30, 40)

# Generate times by using:
# round(predict_time_at_distance(distance = split_distance, MSS = 9, TAU = 1.3), 2)

split_time <- c(2.17, 3.43, 4.60, 5.73)

m1 <- shorts::model_using_splits(
  distance = split_distance,
  time = split_time
)

m1
```

Maximal relative power (PMAX) from the output is estimated using $\frac{MSS \times MAC}{4}$, which disregards the air resistance (which we will deal with soon). `time_correction` and `distance_corection` parameters will be covered later in the article.

Besides providing *residual standard error* (RSE), `shorts` functions provide additional model fit estimators. Additional information can be gained by exploring the returned object, particularly object returned from the `nls` functions:

```{r}
summary(m1$model)
```

Once we have estimated MSS and TAU, we can use `shorts::predict_` family of functions to predict various relationships (i.e. time at distance, acceleration at distance, velocity at time):

```{r}
# Predict time at distance
shorts::predict_time_at_distance(
  distance = split_distance,
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU
)

# Predict acceleration at time
shorts::predict_acceleration_at_time(
  time = c(0, 1, 2, 3, 4, 5, 6),
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU
)
```

Here is a plot of observed and model predicted split times, with the help of the `ggplot2` package [@R-ggplot2]:

```{r}
require(ggplot2)

df <- data.frame(
  distance = split_distance,
  time = split_time,
  pred_time = shorts::predict_time_at_distance(
    distance = split_distance,
    MSS = m1$parameters$MSS,
    TAU = m1$parameters$TAU
  )
)

ggplot(df, aes(x = distance)) +
  theme_bw(10) +
  geom_line(aes(y = time)) +
  geom_point(aes(y = pred_time), color = "red") +
  xlab("Distance (m)") +
  ylab("Time (sec)")
```

### Air resistance and the calculation of force and mechanical power

To estimate force production at distance or time (using `shorts::predict_force_at_distance` and `shorts::predict_force_at_time` functions), and later power production (using `shorts::predict_power_at_distance` and `shorts::predict_power_at_time` functions), one needs to take into account the air resistance. Air resistance (in Newtons) is estimated using `shorts::get_air_resistance` function, which takes velocity, body mass (in kg), body height (in meters), barometric pressure (in Torrs), air temperature (in Celzius), and wind velocity (in $ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):

```{r}
shorts::get_air_resistance(
  velocity = 5,
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)
```

When estimating force and power, one can set the air resistance parameters using `...`:

```{r}
# To calculate horizontal force produced
shorts::predict_force_at_distance(
  distance = split_distance,
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to shorts::get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

# To calculate power produced
shorts::predict_power_at_distance(
  distance = split_distance,
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to shorts::get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)
```

The easiest way to get all kinematics for 0-6sec short sprints is to use `shorts::predict_kinematics` functions:

```{r}
df <- shorts::predict_kinematics(
  m1,
  max_time = 6,
  frequency = 100,
  # Additional parameters forwarded to shorts::get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

head(df)
```

Plotting the model predictions can be done once we convert data from wide to long with the help of `dplyr`[@R-dplyr], `tidyr` [@R-tidyr], and `tidyverse` [@R-tidyverse] packages:

```{r}
require(tidyverse)

df <- pivot_longer(data = df, cols = -2)

ggplot(df, aes(x = distance, y = value)) +
  theme_bw(10) +
  facet_wrap(~name, scales = "free_y") +
  geom_line(alpha = 0.7) +
  ylab(NULL) +
  xlab("Distance (m)")
```

### Utility functions

Sports scientists and coaches might be interested in finding distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range. To help finding these threshold, `shorts` package comes with `shorts::find_` family of functions:

```{r}
# Finds distance where 90% of maximum sprinting speed is reached
shorts::find_velocity_critical_distance(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  percent = 0.9
)

# Finds maximal power and distance (this time using air resistance)
shorts::find_max_power_distance(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to shorts::get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

# Finds distance over 90% power range
shorts::find_power_critical_distance(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to shorts::get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)
```

### Mixed-effects model

Short sprints are often times performed with a group of athlete (e.g., soccer club) representing a single strata of interest. Sports scientists can estimate individual profiles (i.e., for each individual), or utilize mixed-effects models. To perform mixed-effects models in `shorts` for split times, one can use `shorts::mixed_model_using_splits` function. To demonstrate this functionality, we will load the `split_times` dataset provided in the `shorts` package:

```{r}
data(split_times)

head(split_times)

# Mixed model
m2 <- shorts::mixed_model_using_splits(
  data = split_times,
  distance = "distance",
  time = "time",
  athlete = "athlete",

  # Select random effects
  # Default is MSS and TAU
  random = MSS + TAU ~ 1
)

m2
```

Additional information about mixed-effects model performed using the `nlme` package [@R-nlme] can be obtained using `summary`:

```{r}
summary(m2)
```

To following plot contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:

```{r}
df <- predict_kinematics(m2, max_time = 10)

df <- pivot_longer(df, cols = c(-1, -3))

ggplot(
  filter(df, distance < 40),
  aes(x = distance, y = value, group = athlete, color = athlete)
) +
  theme_bw(10) +
  facet_wrap(~name, scales = "free_y") +
  geom_line(alpha = 0.7) +
  ylab(NULL) +
  xlab("Distance (m)") +
  theme(legend.position = "top")
```

## Estimating short sprint parameters using radar gun

Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as an target or outcome variable. Thus the equation \@ref(eq:velocity-time) is used to estimate MSS and TAU.

Let's consider the same example of an athlete with MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$m/s^2$ performing 40m sprint with velocity estimated using radar run (in this case with 1Hz sampling rate).

```{r}
sprint_time <- seq(0, 6, 1)

# Generate velocity by using:
# round(predict_velocity_at_time(time = sprint_time, MSS = 9, TAU = 1.3), 2)

sprint_velocity <- c(0.00, 4.83, 7.07, 8.10, 8.59, 8.81, 8.91)

m3 <- shorts::model_using_radar(
  velocity = sprint_velocity,
  time = sprint_time
)

m3
```

Both split and radar gun models allow the use of *weighted* non-linear regression. For example, we can give more weight to shorter distance or faster velocities. Weighted non-linear regression is performed by setting `weights` parameter:

```{r}
m3_weighted <- shorts::model_using_radar(
  velocity = sprint_velocity,
  time = sprint_time,
  weights = 1 / (sprint_velocity + 1)
)

m3_weighted
```

### Mixed-effects model

Mixed-effects model using radar data is done using `shorts::mixed_model_using_radar` function. To perform mixed model, let's load data that comes with `shorts` package.

```{r}
data("radar_gun_data")

head(radar_gun_data)

m4 <- shorts::mixed_model_using_radar(
  radar_gun_data,
  time = "time",
  velocity = "velocity",
  athlete = "athlete"
)

m4
```

# Problems with estimation

With the previous examples, estimation can look simple. Unfortunately, in real life things are a bit more *messy*, and this messiness can bias the estimated MSS and TAU parameters. In the next section, we will demonstrate how real life issues can introduce biases, but we will also present potential solutions within the `shorts` package.

## Problems with time sync with radar gun

One source of messiness in the estimation using radar gun, beside measurement error, is the time synchronization, where zero velocity happens at $t=0$. Let's use our athlete and add and deduct 0.5second to simulate bad synchronization and its effect on estimated MSS and TAU.

```{r}
df <- tibble(
  `true time` = sprint_time,
  velocity = sprint_velocity,
  `0.5s added` = `true time` + 0.5,
  `0.5s deducted` = `true time` - 0.5
)

head(df)
```

```{r}
plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")

ggplot(plot_df, aes(x = value, y = velocity, color = `Sync issue`)) +
  theme_bw(10) +
  geom_line(alpha = 0.7) +
  xlab("Time (s)") +
  ylab("Velocity (m/s)")
```

The following three models estimate MSS and TAU from the three datasets:

```{r}
# Without synchronization issues
m5 <- shorts::model_using_radar(
  velocity = df$velocity,
  time = df$`true time`
)

# With time added
m6 <- shorts::model_using_radar(
  velocity = df$velocity,
  time = df$`0.5s added`
)

# With time deducted
m7 <- shorts::model_using_radar(
  velocity = df$velocity,
  time = df$`0.5s deducted`
)

rbind(
  data.frame(
    model = "True time",
    t(coef(m5))
  ),
  data.frame(
    model = "Added 0.5s time",
    t(coef(m6))
  ),
  data.frame(
    model = "Deducted 0.5s time",
    t(coef(m7))
  )
)
```

As can be seen from the example, it's the TAU estimate that is mostly affected by a bad synchronization of time with velocity. Solution implemented in `shorts` package involves estimation of the *time correction* parameters using the following equation:

\begin{equation}
  v(t) = MSS \times (1 - e^{-\frac{t + time \; correction}{TAU}}) (\#eq:velocity-time-correction)
\end{equation}

This model is utilizes using the `shorts::model_using_radar_with_time_correction` function:

```{r}
# With time added
m8 <- shorts::model_using_radar_with_time_correction(
  velocity = df$velocity,
  time = df$`0.5s added`
)
coef(m8)

# With time deducted
m9 <- shorts::model_using_radar_with_time_correction(
  velocity = df$velocity,
  time = df$`0.5s deducted`
)
coef(m9)
```

When using `shorts::predict_` family of functions, one can provide estimated time correction to get predictions at original time scale.

```{r}
# Using the true time
round(
  predict_velocity_at_time(
    time = df$`true time`,
    MSS = m5$parameters$MSS,
    TAU = m5$parameters$TAU
  ),
  2
)

# Using time with sync issues
round(
  predict_velocity_at_time(
    time = df$`0.5s added`,
    MSS = m8$parameters$MSS,
    TAU = m8$parameters$TAU,
    time_correction = m8$parameters$time_correction
  ),
  2
)
```

### Mixed-model approach

When it comes to mixed-model approach, time correction can be modeled as fixed effect or random effect using the `shorts::mixed_model_using_radar_with_time_correction` function.

```{r}
# Adding 0.5s to radar_gun_data
radar_gun_data$time <- radar_gun_data$time + 0.5

# Mixed model with time correction being fixed effect
m10 <- shorts::mixed_model_using_radar_with_time_correction(
  radar_gun_data,
  time = "time",
  velocity = "velocity",
  athlete = "athlete",
  random = MSS + TAU ~ 1
)

m10

# Mixed model with time correction being random effect
m11 <- shorts::mixed_model_using_radar_with_time_correction(
  radar_gun_data,
  time = "time",
  velocity = "velocity",
  athlete = "athlete",
  random = MSS + TAU + time_correction ~ 1
)

m11
```

## Problems at the start when using split times

Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$m/s^2$. Let's call them John and Jack. The are both performing 40m sprint using timing gates set at 5, 10, 20, 30, and 40m. The initial timing gate at the start serves the purpose of activating the timing system (i.e., when they cross the beam).

John is our *theoretical model*, while Jack is *street smart guy*, most likely playing soccer, and decides to move slightly back (i.e. 0.5m) and use body rocking to initiate the start. Let's see how their sprints differ.

```{r}
MSS <- 9
TAU <- 1.3
MAC <- MSS / TAU

split_times <- tibble(
  distance = c(5, 10, 20, 30, 40),
  john_time = shorts::predict_time_at_distance(distance, MSS, TAU),

  # Jack's performance
  jack_distance = distance + 0.5,
  jack_true_time = shorts::predict_time_at_distance(jack_distance, MSS, TAU),
  time_05m = shorts::predict_time_at_distance(0.5, MSS, TAU),
  jack_time = jack_true_time - time_05m
)

split_times
```

As expected, Jack performs better than John, even if they have same short sprint capabilities. Let's see how this affects the MSS and TAU estimates:

```{r}
# Since this is a perfect simulation and stats::nls will complain
# we need to add very small noise, or measurement error to the times
set.seed(1667)
rand_noise <- rnorm(nrow(split_times), 0, 10^-5)
split_times$john_time <- split_times$john_time + rand_noise
split_times$jack_time <- split_times$jack_time + rand_noise

john_profile <- shorts::model_using_splits(
  distance = split_times$distance,
  time = split_times$john_time
)

jack_profile <- shorts::model_using_splits(
  distance = split_times$distance,
  time = split_times$jack_time
)

sprint_parameters <- rbind(
  unlist(john_profile$parameters),
  unlist(jack_profile$parameters)
)

rownames(sprint_parameters) <- c("John", "Jack")

round(sprint_parameters, 2)
```

As can be seen from the results, *cheating* (or not having consistent start) at the begin of the sprint yields biased estimates, particularly for the TAU, MAC and PMAX. Let's create a small simulation to estimate effects of cheating distance over combinations of MSS and MAC on their estimates as well as residuals.

Simulated splits are 5, 10, 20, 30, 40, and 50m, with MSS and MAC varying from 6 to 9, and cheating distance varying from 0 to 1m.

```{r}
sim_df <- expand.grid(
  MSS = c(6, 7, 8, 9),
  MAC = c(6, 7, 8, 9),
  cheat_distance = c(
    seq(0, 0.001, length.out = 20),
    seq(0.001, 0.01, length.out = 20),
    seq(0.01, 0.1, length.out = 20),
    seq(0.1, 1, length.out = 20)
  ),
  distance = c(5, 10, 20, 30, 40, 50)
)

sim_df <- sim_df %>%
  mutate(
    TAU = MSS / MAC,
    PMAX = MSS * MAC / 4,
    true_distance = distance + cheat_distance,
    true_time = shorts::predict_time_at_distance(true_distance, MSS, TAU),
    stolen_time = shorts::predict_time_at_distance(cheat_distance, MSS, TAU),
    time = true_time - stolen_time
  )

# Add small noise to allow model fit
set.seed(1667)
rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
sim_df$time <- sim_df$time + rand_noise

head(round(sim_df, 2))
```

Now when we have a simulation dataset, we can check the model estimates and predictions, given the cheating distance:

```{r}
# Prediction wrapper
pred_wrapper <- function(data) {
  model <- shorts::model_using_splits(
    distance = data$distance,
    time = data$time
  )

  params <- data.frame(t(unlist(model$parameters)))

  predicted_time <- shorts::predict_time_at_distance(
    distance = data$distance,
    MSS = model$parameters$MSS,
    TAU = model$parameters$TAU
  )

  colnames(params) <- c(
    "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
    "est_time_correction", "est_distance_correction"
  )

  cbind(data, params, data.frame(predicted_time = as.numeric(predicted_time)))
}

# estimated parameters and predited time
model_df <- sim_df %>%
  group_by(MSS, TAU, cheat_distance) %>%
  do(pred_wrapper(.)) %>%
  ungroup()

# Prediction residuals
model_df$residuals <- model_df$predicted_time - model_df$time
```

The following image demonstrates the effect of cheating distance on estimated MSS:

```{r}
# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, cheat_distance) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(df, aes(x = cheat_distance, y = est_MSS, color = MAC_string)) +
  theme_bw() +
  geom_line(alpha = 0.7) +
  facet_wrap(~MSS_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated MSS (m/s)") +
  theme(legend.title = element_blank())
```

As can be seen from the image, MSS is underestimated as cheating distance increases. The following image demonstrates the effect of cheating distance on estimated MAC:

```{r}
# MAC
ggplot(df, aes(x = cheat_distance, y = est_MAC, color = MSS_string)) +
  theme_bw() +
  geom_line(alpha = 0.7) +
  facet_wrap(~MAC_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated MAC (m/s/s)") +
  theme(legend.title = element_blank())
```

MAC (and also TAU) are highly affected by the cheating distance, and from the figure we can notice that MAC is overestimated as cheat distance increases.

And finally, the following image demonstrates the effect of cheating distance on estimated PMAX:

```{r}
# PMAX
ggplot(df, aes(x = cheat_distance, y = est_PMAX, color = MSS_string)) +
  theme_bw() +
  geom_line(alpha = 0.7) +
  facet_wrap(~MAC_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated PMAX (W/kg)") +
  theme(legend.title = element_blank())
```

Estimated PMAX is also overestimated as cheat distance increases.

Model residuals are also affected by cheating distance. The shape of residuals depends on splits utilized, but here we can see the effect of the cheating distance on the model residuals per split distance:

```{r}
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, cheat_distance)
  )

ggplot(model_df, aes(y = residuals, x = distance, color = cheat_distance, group = group)) +
  theme_bw() +
  geom_line(alpha = 0.3) +
  facet_grid(MSS_string ~ MAC_string) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (sec)")
```

If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler images conveying issues with residuals when there is cheating at the start:

```{r}
ggplot(model_df, aes(y = residuals, x = distance, color = cheat_distance, group = group)) +
  theme_bw() +
  geom_line(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (sec)")
```

Clearly, cheating and sprint start timing issues and inconsistencies can result in biased parameters and predictions. This is particularly important in testing and comparing test results implementing different starting techniques and timing initiations (**REFERENCE**). Since speed trait is one of the hardest to improve, the effects of start inconsistencies can mask effects of the training intervention. It is thus of crucial importance to take these into account, either by standardizing the start (which is always a must), or by implementing the following techniques.

### How to deal with the cheating or sprint start timing issues?

Besides standardizing sprint starts, particularly for longitudinal monitoring of the training effects, one can implement additional strategies to make model estimates more robust.

The simplest strategy is to used validity study results (i.e. comparing timing technique of interest to a gold standard measurement) and to use correction factors (**REFERENCE**). One way to implement correction is to use *time correction* in a same manner used with the radar gun synchronization issues. Thus, equation \@ref(time-distance) gets another form:

\begin{equation}
  t(d) = TAU \times W(-e^{\frac{-d}{MSS \times TAU}} - 1) + \frac{d}{MSS} + TAU - time \; correction (\#eq:time-correction)
\end{equation}

If using validity studies, this time correction can be provided (i.e., 0.3-0.5seconds; *REFERENCE*), which is oftentimes the case, or it can me estimated as MSS and TAU is estimated. To estimate time correction, we use `shorts::model_using_splits_with_time_correction` function. Here is how we can estimate Jack parameters using these two methods:

```{r}
jack_profile_fixed_time <- shorts::model_using_splits(
  distance = split_times$distance,
  time = split_times$jack_time,
  time_correction = 0.3
)

jack_profile_time_estimated <- shorts::model_using_splits_with_time_correction(
  distance = split_times$distance,
  time = split_times$jack_time
)

jack_parameters <- rbind(
  unlist(john_profile$parameters),
  unlist(jack_profile$parameters),
  unlist(jack_profile_fixed_time$parameters),
  unlist(jack_profile_time_estimated$parameters)
)

rownames(jack_parameters) <- c(
  "John",
  "Jack - No corrections",
  "Jack - Fixed time correction",
  "Jack - Estimated time correction"
)

jack_parameters
```

In Jack's case, both fixed time correction and time correction estimation yield parameters closer to John's (i.e. true parameters).

Another model definition, which is novel approach implemented in the `shorts` packages, is to ulitize *distance correction*, besides time correction. Thus, equation \@ref(time-distance) gets another form:

\begin{equation}
  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction (\#eq:distance-correction)
\end{equation}

This model is implemented in `shorts::model_using_splits_with_corrections` function. Let's check this model estimates:

```{r}
jack_profile_distance_correction <- shorts::model_using_splits_with_corrections(
  distance = split_times$distance,
  time = split_times$jack_time
)

jack_parameters <- rbind(
  unlist(john_profile$parameters),
  unlist(jack_profile$parameters),
  unlist(jack_profile_fixed_time$parameters),
  unlist(jack_profile_time_estimated$parameters),
  unlist(jack_profile_distance_correction$parameters)
)

rownames(jack_parameters) <- c(
  "John",
  "Jack - No corrections",
  "Jack - Fixed time correction",
  "Jack - Estimated time correction",
  "Jack - Estimated distance correction"
)

jack_parameters
```

As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters. There are few issues with this model definition. Besides being novel and still not validated with practical data, distance correction model has four parameters to estimated, which implies that one needs at least five sprint splits. This imposes practical limitations, since acquiring six timing gate (one for the start and five to splits) might be practically troublesome.

We will get back to these issues later, but let's see how these models perform using simulated data with varying cheating distance. The following code contains the wrapper that performs all four models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction):

```{r}
pred_wrapper <- function(data) {
  no_correction <- shorts::model_using_splits(
    distance = data$distance,
    time = data$time
  )

  fixed_correction <- shorts::model_using_splits(
    distance = data$distance,
    time = data$time,
    time_correction = 0.3
  )

  time_correction <- shorts::model_using_splits_with_time_correction(
    distance = data$distance,
    time = data$time,
    control = nls.control(tol = 1)
  )

  time_dist_correction <- shorts::model_using_splits_with_corrections(
    distance = data$distance,
    time = data$time,
    control = nls.control(tol = 1)
  )

  params <- rbind(
    data.frame(
      model = "No correction",
      t(unlist(no_correction$parameters))
    ),
    data.frame(
      model = "Fixed correction",
      t(unlist(fixed_correction$parameters))
    ),
    data.frame(
      model = "Time correction",
      t(unlist(time_correction$parameters))
    ),
    data.frame(
      model = "Time and distance correction",
      t(unlist(time_dist_correction$parameters))
    )
  )

  colnames(params) <- c(
    "model", "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
    "est_time_correction", "est_distance_correction"
  )

  df <- expand_grid(
    data,
    params
  )

  df$predicted_time <- shorts::predict_time_at_distance(
    distance = df$distance,
    MSS = df$est_MSS,
    TAU = df$est_TAU,
    time_correction = df$est_time_correction,
    distance_correction = df$est_distance_correction
  )

  df$residuals <- df$predicted_time - df$time
  return(df)
}

# estimated parameters and predited time
model_df <- sim_df %>%
  group_by(MSS, TAU, cheat_distance) %>%
  do(pred_wrapper(.)) %>%
  ungroup()
```

As can be seen from the next figure, estimated time correction model estimates MSS almost perfectly, while estimated time and distance correction model estimates MSS perfectly.

```{r}
model_df$model <- factor(
  model_df$model,
  levels = c(
    "No correction",
    "Fixed correction",
    "Time correction",
    "Time and distance correction"
  )
)
# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, cheat_distance, model) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(df, aes(x = cheat_distance, y = est_MSS, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated MSS (m/s)") +
  theme(legend.title = element_blank())
```

Same conclusions can be said for the MAC parameter. Time and distance corrections model performs perfectly, while time correction model performs almost as good (although worse than for MSS parameter).

```{r}
# MAC
ggplot(df, aes(x = cheat_distance, y = est_MAC, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated MAC (m/s/s)") +
  theme(legend.title = element_blank())
```

When it comes to PMAX, same conclusion can be reached as for MAC.

```{r}
# PMAX
ggplot(df, aes(x = cheat_distance, y = est_PMAX, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated PMAX (W/kg)") +
  theme(legend.title = element_blank())
```

The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the stolen time; indicated by the dashed line on the figure).

```{r}
# time_correction
ggplot(df, aes(x = cheat_distance, y = est_time_correction, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_line(aes(y = stolen_time), color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated time correction (sec)") +
  theme(legend.title = element_blank())
```

The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., cheat distance; indicated by the dashed line on the figure, which represents *identity line* since cheat distance is already on the x-axis).

```{r}
# distance_correction
ggplot(df, aes(x = cheat_distance, y = est_distance_correction, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_abline(slope = 1, color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated distance correction (m)") +
  theme(legend.title = element_blank())
```

The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models, particularly without issues at specific distance zones.

```{r}
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, cheat_distance)
  )

ggplot(model_df, aes(y = residuals, x = distance, color = cheat_distance, group = group)) +
  theme_bw() +
  geom_line(alpha = 0.3) +
  facet_wrap(~model) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (sec)")
```

Given the simulation data, time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model in the face of start cheating. Since time correction model is simpler and demand three parameters to be estimated, it might be practically more useful than time and distance correction model, which demand four parameters estimation and thus more than five timing gates and sprint splits.

Time correction and time and distance corrections are also implemented in the mixed-models using `shorts::mixed_model_using_splits_with_time_correction` and `shorts::mixed_model_using_splits_with_corrections`. We will showcase their use at the end of this article.

### Simulation of additional starting issues

Starting behind the initial timing gate represent only one issue (i.e. cheating). In this section, we will simulate few other issues to check how sensitive are the presented models to perturbations that can happen in the practice.

#### Triggering the gate before the start

One issue that might happen in certain timing gates setups is an individual triggering timing system before the sprint is initiated. This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is *reaction time* (RT) involved. Both of these scenarios represent *time lag* that is added to the split times. Let's simulate the effect of this time lag on model estimates and predictions.

```{r}
sim_df <- expand.grid(
  MSS = c(6, 7, 8, 9),
  MAC = c(6, 7, 8, 9),
  time_lag = seq(0, 1, length.out = 50),
  distance = c(5, 10, 20, 30, 40, 50)
)

sim_df <- sim_df %>%
  mutate(
    TAU = MSS / MAC,
    PMAX = MSS * MAC / 4,
    true_time = shorts::predict_time_at_distance(distance, MSS, TAU),
    time = true_time + time_lag
  )

# Add small noise to allow model fit
set.seed(1667)
rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
sim_df$time <- sim_df$time + rand_noise

head(round(sim_df, 2))
```

```{r}
# estimated parameters and predicted time
model_df <- sim_df %>%
  group_by(MSS, TAU, time_lag) %>%
  do(pred_wrapper(.)) %>%
  ungroup()
```

From the figure below it can be seen that time lag affects estimated MSS for the the model without correction and fixed correction model. Time correction and time and distance corrections model correctly estimated MSS.

```{r}
model_df$model <- factor(
  model_df$model,
  levels = c(
    "No correction",
    "Fixed correction",
    "Time correction",
    "Time and distance correction"
  )
)
# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, time_lag, model) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(df, aes(x = time_lag, y = est_MSS, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
  xlab("Time lag (sec)") +
  ylab("estimated MSS (m/s)") +
  theme(legend.title = element_blank())
```

From the figure below it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction model. Time correction and time and distance corrections model correctly estimated MAC.

```{r}
# MAC
ggplot(df, aes(x = time_lag, y = est_MAC, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Time lag (sec)") +
  ylab("estimated MAC (m/s/s)") +
  theme(legend.title = element_blank())
```

Two figure below depicts correctly identified time lag (i.e. using time correction parameter) using time correction and time and distance corrections models.

```{r}
# time_correction
ggplot(df, aes(x = time_lag, y = est_time_correction, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_abline(slope = -1, color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Time lag (sec)") +
  ylab("estimated time correction (sec)") +
  theme(legend.title = element_blank())
```

The next figure depicts estimated distance correction for the time and distance correction model. The estimated distance correction parameters looks jumpy due random noise that we have to inject to allow model fit.

```{r}
# distance_correction
ggplot(df, aes(x = time_lag, y = est_distance_correction, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Cheat distance (m)") +
  ylab("estimated distance correction (m)") +
  theme(legend.title = element_blank())
```

The following figure depicts residuals (i.e., predicted time minus observed time).

```{r}
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, time_lag)
  )

ggplot(model_df, aes(y = residuals, x = distance, color = time_lag, group = group)) +
  theme_bw() +
  geom_line(alpha = 0.3) +
  facet_wrap(~model) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (sec)")
```

#### Bad position of the timing gates

Another common issue with timing gates in the practical field settings is the bad measurement of the distance and thus bad positions of the timing gates. In this example, we will simulate the scenario where the initial timing gate is not positioned at $d=0$, but rather with some error (i.e., -0.5 to 0.5m).

```{r}
sim_df <- expand.grid(
  MSS = c(6, 7, 8, 9),
  MAC = c(6, 7, 8, 9),
  displacement = seq(-0.5, 0.5, length.out = 51),
  distance = c(5, 10, 20, 30, 40, 50)
)

sim_df <- sim_df %>%
  mutate(
    TAU = MSS / MAC,
    PMAX = MSS * MAC / 4,
    true_distance = distance + displacement,
    time = shorts::predict_time_at_distance(true_distance, MSS, TAU)
  )

# Add small noise to allow model fit
set.seed(1667)
rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
sim_df$time <- sim_df$time + rand_noise

head(round(sim_df, 2))
```

```{r}
# estimated parameters and predicted time
model_df <- sim_df %>%
  group_by(MSS, TAU, displacement) %>%
  do(pred_wrapper(.)) %>%
  ungroup()
```

Figure below depicts effect of timing gate displacement on the MSS parameter. Time correction and time and distance correction models are almost identical in their correct estimation of the MSS.

```{r}
model_df$model <- factor(
  model_df$model,
  levels = c(
    "No correction",
    "Fixed correction",
    "Time correction",
    "Time and distance correction"
  )
)

# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, displacement, model) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(df, aes(x = displacement, y = est_MSS, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
  xlab("Displacement (m)") +
  ylab("estimated MSS (m/s)") +
  theme(legend.title = element_blank())
```

Figure below depicts effect of timing gate displacement on the MSS parameter. Time and distance correction model is the only model that estimated MAC correctly.

```{r}
# MAC
ggplot(df, aes(x = displacement, y = est_MAC, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Displacement (m)") +
  ylab("estimated MAC (m/s/s)") +
  theme(legend.title = element_blank())
```

Figures below depicts estimated time and distance corrections parameters.

```{r}
# time_correction
ggplot(df, aes(x = displacement, y = est_time_correction, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Displacement (m)") +
  ylab("estimated time correction (sec)") +
  theme(legend.title = element_blank())
```

```{r}
# distance_correction
ggplot(df, aes(x = displacement, y = est_distance_correction, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_abline(slope = 1, color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Displacement (m)") +
  ylab("estimated distance correction (m)") +
  theme(legend.title = element_blank())
```

The following figure depicts residuals (i.e., predicted time minus observed time).

```{r}
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, displacement)
  )

ggplot(model_df, aes(y = residuals, x = distance, color = displacement, group = group)) +
  theme_bw() +
  geom_line(alpha = 0.3) +
  facet_wrap(~model) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (sec)")
```

Additional sources of issues with the timing gate involve combination of the previous three. For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence. More elaborate simulation is beyond the scope of the current paper.

# Leave-one-out Cross-Validation

To estimate parameter stability, model over-fitting, and performance on the unseen data, `shorts` model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018]. LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation. This process is repeated for each observations in the model dataset. LOOCV allows one to check estimated parameters stability, and model performance on the unseen data.

Let's perform LOOCV using Jack's data and time correction model:

```{r}
jack_LOOCV <- shorts::model_using_splits_with_time_correction(
  distance = split_times$distance,
  time = split_times$jack_time,
  LOOCV = TRUE
)

jack_LOOCV
```

The model print output provides training dataset estimates and model performance, as well as LOOCV estimates and model performance.

Let's plot estimated parameters across LOOCV folds:

```{r}
df <- jack_LOOCV$LOOCV$parameters

df <- pivot_longer(df, cols = 1:6, names_to = "parameter")

df$parameter <- factor(
  df$parameter,
  levels = c(
    "MSS",
    "TAU",
    "MAC",
    "PMAX",
    "time_correction",
    "distance_correction"
  )
)

ggplot(df, aes(x = value)) +
  theme_bw() +
  geom_boxplot() +
  facet_wrap(~parameter, scales = "free_x") +
  xlab(NULL) +
  ylab(NULL) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

Here is the plot of the training and LOOCV residuals:

```{r}
df <- data.frame(
  distance = jack_LOOCV$data$distance,
  time = jack_LOOCV$data$time,
  pred_time = jack_LOOCV$data$pred_time,
  LOOCV_time = jack_LOOCV$LOOCV$data$pred_time
)

df <- df %>%
  pivot_longer(cols = c("pred_time", "LOOCV_time"))

df$resid <- df$value - df$time

ggplot(df, aes(x = distance, y = resid, color = name)) +
  theme_bw() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point() +
  theme(legend.title = element_blank()) +
  xlab("Distance (m)") +
  ylab("Predicted - observed time (sec)")
```

As expected, model has more issues predicting unseen short or long sprints. Please note, that since LOOCV removes one observation, if model estimates three parameters, one needs to have at least five observations, since we need to make sure that model can be estimated once a single observation is removed. LOOCV can also be implemented with the mixed-effects models in the `shorts` package.

# Example analysis

Let's utilize demonstrated functionalities of the `shorts` package using real-world data. The first dataset comes from Usain Bolt run from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data set for 52 female soccer athletes which comes with the `shorts` package.

## Usain Bolt's run from London 2017

The following dataset represents Usain Bolt's race in the finals at the IAAF World Championship held in London, 2017. Since reaction time enters the splits, we want to see how will that affect the model estimates, and particularly, will estimated time correction model pick reaction time up.

For the sake of this analysis, only 10m splits over 60m are used.

```{r}
# Taken from London 2017
# https://barrunning.com/resources/SiteUploads/20180715/London%202017%20REPORT%20FOR%20THE%20100%20m%20Men.pdf
bolt_reaction_time <- 0.183

bolt_distance <- c(10, 20, 30, 40, 50, 60)
bolt_time <- c(1.963, 2.983, 3.883, 4.763, 5.643, 6.493)

# No corrections model
bolt_m1 <- shorts::model_using_splits(
  distance = bolt_distance,
  time = bolt_time
)

# Model with reaction time as fixed time correction
bolt_m2 <- shorts::model_using_splits(
  distance = bolt_distance,
  time = bolt_time,
  time_correction = -bolt_reaction_time
)

# Model with estimated time correction
bolt_m3 <- shorts::model_using_splits_with_time_correction(
  distance = bolt_distance,
  time = bolt_time
)

# Model with estimated time correction, but deducted reaction time
bolt_m4 <- shorts::model_using_splits_with_time_correction(
  distance = bolt_distance,
  time = bolt_time - bolt_reaction_time
)

# Model with estimated time and distance corrections
bolt_m5 <- shorts::model_using_splits_with_corrections(
  distance = bolt_distance,
  time = bolt_time
)

# Model with estimated time and distance corrections and deducted reaction time
bolt_m6 <- shorts::model_using_splits_with_corrections(
  distance = bolt_distance,
  time = bolt_time - bolt_reaction_time
)

bolt_model <- rbind(
  data.frame(
    model = "No correction",
    t(coef(bolt_m1))
  ),
  data.frame(
    model = "No correction - RT",
    t(coef(bolt_m2))
  ),
  data.frame(
    model = "Time correction",
    t(coef(bolt_m3))
  ),
  data.frame(
    model = "Time correction - RT",
    t(coef(bolt_m4))
  ),
  data.frame(
    model = "Distance correction",
    t(coef(bolt_m5))
  ),
  data.frame(
    model = "Distance correction - RT",
    t(coef(bolt_m6))
  )
)

rownames(bolt_model) <- bolt_model$model

round(bolt_model[-1], 3)
```

## Vescovi data

Vescovi's data set represents a  sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players). Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset.  Soccer players were older (24.6±3.6 vs. 18.9±2.7 yr, p < 0.000), however there were no differences for height (167.3±5.9 vs. 167.0±5.7 cm, p = 0.886), body mass (62.5±5.9 vs. 64.0±9.4 kg, p = 0.500) or any sprint interval time (p > 0.650).

The protocol for assessing linear sprint speed has been described previously [@vescoviImpactMaximumSpeed2014; @vescoviLocomotorHeartRateMetabolic2016; @vescoviSprintSpeedCharacteristics2012] and was identical for each cohort.  Briefly, all athletes performed a standardized warm-up that included general exercises such as jogging, shuffling, multi-directional movements, and dynamic stretching exercises. Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, and 35 meters at a height of approximately 1.0 meter. Participants stood with their lead foot positioned approximately 5 cm behind the initial infrared beam (i.e., start line). Only forward movement was permitted (no leaning or rocking backwards) and timing started when the laser of the starting gate was triggered. The best 35 m time, and all associated split times were kept for analysis.  The assessment of linear sprints using infrared timing gates does not require familiarization [@moirInfluenceFamiliarizationReliability2004].

To analyze this dataset, we will utilize mixed-effects models, although individual models can be utilized as well.

```{r}
data("vescovi")

# Convert data to long
df <- vescovi %>%
  select(1:13) %>%
  # slice(1:10) %>%
  pivot_longer(cols = 9:13, names_to = "distance", values_to = "time") %>%
  mutate(
    distance = as.numeric(str_extract(distance, "^[0-9]+"))
  )

head(df)
```

Here we will utilize the following models: no corrections model, fixed time correction model (using 0.3s heuristic rule of thumb), estimated time correction as a fixed effect model, estimated time correction as a random effect model, estimated distance correction as fixed effect model (and time correction as random effect), and estimated distance correction as random effect model.

```{r}
no_corrections <- shorts::mixed_model_using_splits(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete"
)

fixed_correction <- shorts::mixed_model_using_splits(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  time_correction = 0.3
)

time_correction_fixed <- shorts::mixed_model_using_splits_with_time_correction(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU ~ 1
)

time_correction_random <- shorts::mixed_model_using_splits_with_time_correction(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU + time_correction ~ 1
)

time_distance_correction_fixed <- shorts::mixed_model_using_splits_with_corrections(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU + time_correction ~ 1
)

time_distance_correction_random <- shorts::mixed_model_using_splits_with_corrections(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU + time_correction + distance_correction ~ 1
)
```

The following image represents model fit estimator RSE for each model. As can be seen, RSE drops the more flexible the model.

```{r}
model_fit <- rbind(
  data.frame(
    model = "No corrections",
    t(unlist(no_corrections$model_fit))
  ),
  data.frame(
    model = "Fixed correction",
    t(unlist(fixed_correction$model_fit))
  ),
  data.frame(
    model = "Time correction fixed",
    t(unlist(time_correction_fixed$model_fit))
  ),
  data.frame(
    model = "Time correction random",
    t(unlist(time_correction_random$model_fit))
  ),
  data.frame(
    model = "Time and distance correction fixed",
    t(unlist(time_distance_correction_fixed$model_fit))
  ),
  data.frame(
    model = "Time and distance correction random",
    t(unlist(time_distance_correction_random$model_fit))
  )
)

model_fit$model <- factor(
  model_fit$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

ggplot(model_fit, aes(x = RSE, y = model)) +
  theme_bw() +
  geom_point() +
  xlab("RSE (s)") +
  ylab(NULL)
```

The following image depicts estimated parameters for each model:

```{r}
est_params <- rbind(
  data.frame(
    model = "No corrections",
    no_corrections$parameters$random
  ),
  data.frame(
    model = "Fixed correction",
    fixed_correction$parameters$random
  ),
  data.frame(
    model = "Time correction fixed",
    time_correction_fixed$parameters$random
  ),
  data.frame(
    model = "Time correction random",
    time_correction_random$parameters$random
  ),
  data.frame(
    model = "Time and distance correction fixed",
    time_distance_correction_fixed$parameters$random
  ),
  data.frame(
    model = "Time and distance correction random",
    time_distance_correction_random$parameters$random
  )
)

est_params$model <- factor(
  est_params$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

est_params <- est_params %>%
  pivot_longer(cols = -(1:2), names_to = "parameter")

est_params$parameter <- factor(
  est_params$parameter,
  levels = c(
    "MSS",
    "TAU",
    "MAC",
    "PMAX",
    "time_correction",
    "distance_correction"
  )
)

ggplot(est_params, aes(y = model, x = value)) +
  theme_bw(8) +
  geom_boxplot() +
  facet_wrap(~parameter, scales = "free_x") +
  xlab(NULL) +
  ylab(NULL)
```

The following image depicts model residuals across distance splits. To provide practical magnitude of the residuals, we have used between subject observed time SD multiplied with 0.2 and -0.2. This provides practical anchor for the residual magnitude, often referred to as *smallest worthwhile change* (SWC) or *smallest effect size of interest* (SESOI) [@jovanovicBmbstatsBootstrapMagnitudebased2020]. If the residuals are within this magnitude band, they have no practical significance, andthus the model is useful in making practically useful predictions.

Error bars represent residual bias and $\pm$ 1SD.

```{r}
model_resid <- rbind(
  data.frame(
    model = "No corrections",
    no_corrections$data
  ),
  data.frame(
    model = "Fixed correction",
    fixed_correction$data
  ),
  data.frame(
    model = "Time correction fixed",
    time_correction_fixed$data
  ),
  data.frame(
    model = "Time correction random",
    time_correction_random$data
  ),
  data.frame(
    model = "Time and distance correction fixed",
    time_distance_correction_fixed$data
  ),
  data.frame(
    model = "Time and distance correction random",
    time_distance_correction_random$data
  )
)

model_resid$model <- factor(
  model_resid$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

model_resid$resid <- model_resid$pred_time - model_resid$time

# Create SWC / SESOI band
model_SESOI <- model_resid %>%
  group_by(model, distance) %>%
  summarise(
    bias = mean(resid),
    variance = sd(resid),
    upper = bias + variance,
    lower = bias - variance,
    MAD = mean(abs(resid)),
    SESOI_upper = sd(time) * 0.2,
    SESOI_lower = -sd(time) * 0.2
  )

# Plot
ggplot(model_resid, aes(y = model)) +
  theme_bw(8) +
  geom_vline(
    data = model_SESOI,
    aes(xintercept = SESOI_lower), color = "blue", alpha = 0.5, linetype = "dashed"
  ) +
  geom_vline(
    data = model_SESOI,
    aes(xintercept = SESOI_upper), color = "blue", alpha = 0.5, linetype = "dashed"
  ) +
  geom_vline(xintercept = 0, color = "blue", alpha = 0.5) +
  geom_jitter(aes(x = resid), alpha = 0.2, height = 0.25) +
  geom_errorbarh(data = model_SESOI, aes(xmin = lower, xmax = upper), height = 0.1) +
  geom_point(data = model_SESOI, aes(x = bias)) +
  facet_wrap(~distance, scales = "free_x") +
  xlab("Predicted time - observed time (sec)") +
  ylab(NULL)
```

The following figure depicts model residuals estimators (bias, or mean residual; variance, or SD of the residuals, and MAD, or mean absolute difference).

```{r}
df <- model_SESOI %>%
  pivot_longer(cols = -(1:2), names_to = "estimator") %>%
  filter(estimator %in% c("bias", "variance", "MAD"))

df$model <- factor(
  df$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

df$estimator <- factor(
  df$estimator,
  levels = c("bias", "variance", "MAD")
)

ggplot(df, aes(x = value, y = model)) +
  theme_bw() +
  geom_point() +
  facet_grid(distance ~ estimator, scales = "free_x") +
  xlab(NULL) +
  ylab(NULL)
```

Which model should should be used? Although providing a better fit (using RSE as an estimator of model fit), time and distance correction models often estimate these parameters that are hard to interpret. Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for inconsistencies and/or cheating at the sprint starts.

We are hoping that the `shorts` package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising a measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation inconsistencies.

# References

### 2020-10-20 paper.Rmd
folder re-arrange

--- 2020-10-19 jss-paper/paper.Rmd
@@
 ---
+documentclass: jss
+author:
+  - name: Mladen Jovanović
+    affiliation: |
+      | University of Belgrade
+    address: |
+      | Faculty of Sport and Physical Education
+      | Belgrade, Serbia
+    email: \email{coach.mladen.jovanovic@gmail.com}
+  - name: Jason D. Vescovi
+    affiliation: |
+      | University of Toronto
+    address: |
+      | Faculty of Kinesiology and Physical Education
+      | Graduate School of Exercise Science
+      | Toronto, ON Canada
+    email: \email{vescovij@gmail.com}
+title:
+  formatted: "\\pkg{shorts}: An R Package for Modeling Short Sprints"
+abstract: >
+  The abstract of the article.
+keywords:
+  formatted: [sprints, sports, modeling]
+preamble: >
+  \usepackage{amsmath}
+output:
+  bookdown::pdf_book:
+    base_format: rticles::jss_article
 citation_package: natbib
-bibliography:
-- references.bib
-- packages.bib
+bibliography: [references.bib, packages.bib]
 biblio-style: apalike
@@
   chunk_output_type: console
-output:
-  bookdown::pdf_document2: default
-output_dir: "docs"
 ---
@@
 ), "packages.bib")
+
+# Set rounding
+op <- options()
+options(digits = 2)
 ```
@@
 
-Short sprint speed is one of the most distinguishable and admired physical traits in sports, and one that have puzzled coaches and sports scientists from the time when two humanoids raced against each other (if they were available back then).
-Short sprints are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-This is approximately less than 6 seconds in duration (why?
-ATP/CP?
---- REF), but the distance can vary based on the level of the athlete (i.e.
-lower level athletes can reach this point earlier, around 30-40m, while elite sprinters much later, around 50-60m --- REF).
-
-Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time) originally proposed by @doi:10.1098/rspb.1927.0035, and more recently popularized by @greenePredictingSprintDynamics1986, @clarkNFLCombine40Yard2017, @chellyLegPowerHopping2001, @morinSimpleMethodComputing2019, @morinInterpretingPowerForceVelocityProfiles2016, @haugenPowerForceVelocityProfilingSprinting2020, @haugenSprintMechanicalVariables2019, @haugenSprintMechanicalProperties2020, @doi:10.1098/rspb.1927.0035, @haugenDifferenceStartImpact2012, and @samozinoSimpleMethodMeasuring2016:
+Short sprint performance is one of the most distinguishable and admired physical trait in sports.
+Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
+Peak anaerobic power is achieved within the first few seconds (<5 sec) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)[@brownAssessmentLinearSprinting2004].
+
+Describing individual differences in short sprint performance, quantifying training intervention effects, but more importantly, understanding short sprint *bottle necks* (i.e.
+limiting factors) has been one of the prominent tasks of sports science profession since its establishment.
+The \pkg{shorts} package [@R-shorts], written in the R language [@R-base], represents an open-source tool that aims to help sports scientist in achieving these goals through mathematical modeling.
+
+In the current paper, we will provide explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
+
+# Mathematical model
+
+Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time) originally proposed by @doi:10.1098/rspb.1927.0035, and more recently popularized by @clarkNFLCombine40Yard2017, and @samozinoSimpleMethodMeasuring2016.
+Equation \@ref(eq:velocity-time) represents function for instantaneous horizontal velocity $v$ given the time $t$ and two model parameters:
 
@@
 
-The parameters of the equation \@ref(eq:velocity-time) are \textit{maximum sprinting speed} (MSS; expressed in $ms^-1$) and \textit{relative acceleration} (TAU).
-Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in $ms^-2$) \@ref(eq:maximal-acceleration).
+The parameters of the equation \@ref(eq:velocity-time) are \textit{maximum sprinting speed} (MSS; expressed in $ms^{-1}$) and \textit{relative acceleration} (TAU).
+Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in $ms^{-2}$) \@ref(eq:maximal-acceleration).
 
@@
 
-By derivating equation \@ref(eq:velocity-time), we can get equation for acceleration \@ref(eq:acceleration-time).
+By derivating equation \@ref(eq:velocity-time), we can get equation for horizontal acceleration \@ref(eq:acceleration-time).
 
@@
 
-By integrating equation \@ref(eq:velocity-time), we can get equation for distance \@ref(eq:distance-time).
+By integrating equation \@ref(eq:velocity-time), we can get equation for distance covered \@ref(eq:distance-time).
 
@@
       levels = c("distance", "velocity", "acceleration"),
-      labels = c("Distance (m)", "Velocity (m/s)", "Acceleration (m/s/s)")
+      labels = c(
+        expression("Distance (m"*")"),
+        expression("Velocity (ms"^-1*")"),
+        expression("Acceleration (ms"^-2*")"))
     )
@@
   geom_line(alpha = 0.7) +
-  facet_wrap(~variable, scales = "free_y") +
+  facet_wrap(
+    ~variable,
+    scales = "free_y",
+    labeller = label_parsed) +
   ylab(NULL) +
   xlab("Time (s)") +
-  theme(legend.position = "top")
+  theme(legend.position = "top", legend.title = element_blank())
 
@@
 
-If we plot acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is, according to the mathematical model, linear.
+If we plot acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
 
@@
   geom_line(alpha = 0.7) +
-  ylab("Acceleration (m/s/s)") +
-  xlab("Velocity (m/s)") +
-  theme(legend.position = "top")
+  ylab(expression("Acceleration (ms"^-2*")")) +
+  xlab(expression("Velocity (ms"^-1*")")) +
+  theme(legend.position = "top", legend.title = element_blank())
 
@@
 
-# Estimation using `shorts` package
-
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at 0m, 10m, 20m, 30m, 40m), and (2) getting a velocity trace using radar gun (usually at around 100Hz).
-Estimation of the MSS and TAU parameters from the equation \@ref(eq:velocity-time) is performed in `shorts` package using non-linear least squares regression implemented in the `nls` function in the base R [@R-base] and `nlme` function in the `nlme` package [@R-nlme] for the mixed-effect models.
+# Estimation using shorts package
+
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), and (2) getting a velocity trace using a radar gun.
+Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in \pkg{shorts} package using non-linear least squares regression implemented in the `nls` function in the \pkg{base} R [@R-base] and `nlme` function in the \pkg{nlme} package [@R-nlme] for the mixed-effect models.
 
@@
 
-Let's consider an example of an athlete with MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$m/s^2$ performing 40m sprint with timing gates positioned at each 10m split.
+Let's consider an example of an athlete with MSS equal to 9 $ms^-1$, TAU equal to 1.3, and MAC equal to 6.92 $ms^-2$ performing 40m sprint with timing gates positioned at each 10m split.
 When it comes to split times, distance is a predictor, and time is outcome variable, thus the equation \@ref(eq:velocity-time) takes another form:
@@
 
-$W$ in the equation \@ref(eq:velocity-time) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `shorts::model_using_splits` function:
+$W$ in equation \@ref(eq:velocity-time) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `shorts::model_using_splits` function:
 
@@
 split_distance <- c(10, 20, 30, 40)
-
-# Generate times by using:
-# round(predict_time_at_distance(distance = split_distance, MSS = 9, TAU = 1.3), 2)
 
@@
 
-Maximal relative power (PMAX) from the output is estimated using $\frac{MSS \times MAC}{4}$, which disregards the air resistance (which we will deal with soon).
+Maximal relative power (PMAX) from the output is estimated using $\frac{MSS \times MAC}{4}$, which disregards the air resistance.
 `time_correction` and `distance_corection` parameters will be covered later in the article.
 
-Besides providing *residual standard error* (RSE), `shorts` functions provide additional model fit estimators.
-Additional information can be gained by exploring the returned object, particularly object returned from the `nls` functions:
+Besides providing *residual standard error* (RSE), \pkg{shorts} functions provide additional model fit estimators.
+Additional information can be gained by exploring the returned object, particularly object returned from the `nls` function:
 
@@
 
-Here is a plot of observed and model predicted split times, with the help of the `ggplot2` package [@R-ggplot2]:
+Here is a plot of observed and model predicted split times, with the help of the \pkg{ggplot2} package [@R-ggplot2]:
 
@@
   xlab("Distance (m)") +
-  ylab("Time (sec)")
+  ylab("Time (s)")
 ```
@@
 
-To estimate force production at distance or time (using `shorts::predict_force_at_distance` and `shorts::predict_force_at_time` functions), and later power production (using `shorts::predict_power_at_distance` and `shorts::predict_power_at_time` functions), one needs to take into account the air resistance.
-Air resistance (in Newtons) is estimated using `shorts::get_air_resistance` function, which takes velocity, body mass (in kg), body height (in meters), barometric pressure (in Torrs), air temperature (in Celzius), and wind velocity (in $ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
+To estimate force production at distance or time (using `shorts::predict_force_at_distance` and `shorts::predict_force_at_time` functions), as well as power production (using `shorts::predict_power_at_distance` and `shorts::predict_power_at_time` functions), one needs to take into account the air resistance.
+Air resistance (N) is estimated using `shorts::get_air_resistance` function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
 
@@
 
-When estimating force and power, one can set the air resistance parameters using `...`:
+When estimating force and power, the air resistance parameters can be set using `"..."`, which are forwarded to the `shorts::get_air_resistance`:
 
@@
 
-The easiest way to get all kinematics for 0-6sec short sprints is to use `shorts::predict_kinematics` functions:
+The easiest way to get all kinematics for short sprints is to use `shorts::predict_kinematics` functions:
 
@@
 
-Plotting the model predictions can be done once we convert data from wide to long with the help of `dplyr`[@R-dplyr], `tidyr` [@R-tidyr], and `tidyverse` [@R-tidyverse] packages:
+Plotting the model predictions can be done once we convert data from wide to long with the help of \pkg{dplyr} [@R-dplyr], \pkg{tidyr} [@R-tidyr], and \pkg{tidyverse} [@R-tidyverse] packages:
 
@@
 
-Sports scientists and coaches might be interested in finding distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
-To help finding these threshold, `shorts` package comes with `shorts::find_` family of functions:
+Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
+To identify these values, \pkg{shorts} package comes with `shorts::find_` family of functions:
 
@@
 
-Short sprints are often times performed with a group of athlete (e.g., soccer club) representing a single strata of interest.
-Sports scientists can estimate individual profiles (i.e., for each individual), or utilize mixed-effects models.
-To perform mixed-effects models in `shorts` for split times, one can use `shorts::mixed_model_using_splits` function.
-To demonstrate this functionality, we will load the `split_times` dataset provided in the `shorts` package:
+Sprint performance is often evaluated with a group of athletes (e.g., soccer club) representing a single strata of interest.
+Sports scientists can estimate individual profiles, or utilize mixed-effects models.
+To perform mixed-effects models in \pkg{shorts} for split times, one can use `shorts::mixed_model_using_splits` function.
+To demonstrate this functionality, we load the `split_times` dataset provided in the \pkg{shorts} package:
 
@@
 
-To following plot contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:
+The following plot contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:
 
@@
   xlab("Distance (m)") +
-  theme(legend.position = "top")
+  theme(legend.position = "top", legend.title = element_blank())
 ```
@@
 Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as an target or outcome variable.
-Thus the equation \@ref(eq:velocity-time) is used to estimate MSS and TAU.
-
-Let's consider the same example of an athlete with MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$m/s^2$ performing 40m sprint with velocity estimated using radar run (in this case with 1Hz sampling rate).
+Thus equation \@ref(eq:velocity-time) is used to estimate MSS and TAU.
+
+Let's consider the same example of an athlete with MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$ms^-2$ performing 40m sprint with velocity estimated using radar run (in this case with 1Hz sampling rate).
 
@@
 
-Mixed-effects model using radar data is done using `shorts::mixed_model_using_radar` function. To perform mixed model, let's load data that comes with `shorts` package.
+Mixed-effects model using radar data is done using `shorts::mixed_model_using_radar` function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
 
@@
 
-With the previous examples, estimation can look simple.
-Unfortunately, in real life things are a bit more *messy*, and this messiness can bias the estimated MSS and TAU parameters.
-In the next section, we will demonstrate how real life issues can introduce biases, but we will also present potential solutions within the `shorts` package.
+There is a challenge when collecting sprint data that could have a substantial impact on modeled outcomes.
+To ensure accurate parameter outcomes, the initial force production must be synced with start time.
+Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \pkg{shorts} package.
 
@@
 
-One source of messiness in the estimation using radar gun, beside measurement error, is the time synchronization, where zero velocity happens at $t=0$.
-Let's use our athlete and add and deduct 0.5second to simulate bad synchronization and its effect on estimated MSS and TAU.
+One source of error in the modeled estimation using radar gun is the time synchronization.
+In theory, synchronization is ideal when sprint is initiated at $t=0$ (i.e., $v(t=0) = 0$).
+In practice, this is often not the case.
+Let's use our athlete and add and deduct 0.5second to simulate an error in synchronization and its effect on estimated MSS and TAU.
 
@@
   xlab("Time (s)") +
-  ylab("Velocity (m/s)")
+  ylab(expression("Velocity ("*ms^-1*")"))
 ```
@@
 
-As can be seen from the example, it's the TAU estimate that is mostly affected by a bad synchronization of time with velocity.
-Solution implemented in `shorts` package involves estimation of the *time correction* parameters using the following equation:
+As can be seen from the example, TAU is affected by an error in synchronization of time with velocity.
+The potential solution incorporated into the \pkg{shorts} package involves estimation of the *time correction* parameter using the following equation:
 
@@
 
-This model is utilizes using the `shorts::model_using_radar_with_time_correction` function:
+This model is incorporated in the `shorts::model_using_radar_with_time_correction` function:
 
@@
 
-When it comes to mixed-model approach, time correction can be modeled as fixed effect or random effect using the `shorts::mixed_model_using_radar_with_time_correction` function.
+When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the `shorts::mixed_model_using_radar_with_time_correction` function.
 
@@
 
-Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$m/s^2$.
+Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$ms^-2$.
 Let's call them John and Jack.
 The are both performing 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
-The initial timing gate at the start serves the purpose of activating the timing system (i.e., when they cross the beam).
-
-John is our *theoretical model*, while Jack is *street smart guy*, most likely playing soccer, and decides to move slightly back (i.e.
-0.5m) and use body rocking to initiate the start.
+The initial timing gate at the start (i.e., $d=0m$) serves to activate the timing system (i.e., when they cross the beam).
+
+John represents the *theoretical model*, in which we assume that the initial force production and the timing initiation are perfectly synchronized.
+Jack, on the other hand, represents a *practical model*, and decides to move slightly behind the initial timing gate (i.e.
+for 0.5m) and use body rocking to initiate the sprint start.
+In other words, Jack is using a *flying start*, a common scenario when testing field sports athletes.
 Let's see how their sprints differ.
@@
 
-As expected, Jack performs better than John, even if they have same short sprint capabilities. Let's see how this affects the MSS and TAU estimates:
+And here is a graphical representation of the sprint splits:
+
+```{r}
+plot_df <- split_times %>%
+  select(distance, john_time, jack_time) %>%
+  rename(John = john_time, Jack = jack_time) %>%
+  pivot_longer(cols = -1, names_to = "athlete", values_to = "time") %>%
+  mutate(distance = factor(distance))
+
+ggplot(plot_df, aes(x = distance, y = time, color = athlete, group = athlete)) +
+  theme_bw() +
+  geom_point() +
+  geom_line() +
+  xlab("Distance (m)") +
+  ylab("Time (s)")
+```
+
+Using the following code, we can see how different are the estimated MSS and TAU parameters:
 
@@
 
-As can be seen from the results, *cheating* (or not having consistent start) at the begin of the sprint yields biased estimates, particularly for the TAU, MAC and PMAX.
-Let's create a small simulation to estimate effects of cheating distance over combinations of MSS and MAC on their estimates as well as residuals.
-
-Simulated splits are 5, 10, 20, 30, 40, and 50m, with MSS and MAC varying from 6 to 9, and cheating distance varying from 0 to 1m.
+As can be seen from the results, flying starts yields biased estimates, particularly for the TAU, MAC and PMAX.
+
+Below is a simulation sprint with 5, 10, 20, 30, 40, and 50m splits, with MSS and MAC varying from 6 to 9 ($ms^-1$ and $ms^-2$ respectively), and flying start distance varying from 0 to 1m.
 
@@
   MAC = c(6, 7, 8, 9),
-  cheat_distance = c(
+  flying_start_distance = c(
     seq(0, 0.001, length.out = 20),
@@
     PMAX = MSS * MAC / 4,
-    true_distance = distance + cheat_distance,
+    true_distance = distance + flying_start_distance,
     true_time = shorts::predict_time_at_distance(true_distance, MSS, TAU),
-    stolen_time = shorts::predict_time_at_distance(cheat_distance, MSS, TAU),
+    stolen_time = shorts::predict_time_at_distance(flying_start_distance, MSS, TAU),
     time = true_time - stolen_time
@@
 
-Now when we have a simulation dataset, we can check the model estimates and predictions, given the cheating distance:
+Now when we have a simulation dataset, we can check the model estimates and predictions, given the flying start distance:
 
@@
 model_df <- sim_df %>%
-  group_by(MSS, TAU, cheat_distance) %>%
+  group_by(MSS, TAU, flying_start_distance) %>%
   do(pred_wrapper(.)) %>%
@@
 
-The following image demonstrates the effect of cheating distance on estimated MSS:
+The following image demonstrates the effect of flying start distance on estimated MSS:
 
@@
 df <- model_df %>%
-  group_by(MSS, TAU, cheat_distance) %>%
+  group_by(MSS, TAU, flying_start_distance) %>%
   slice(1) %>%
@@
 # MSS
-ggplot(df, aes(x = cheat_distance, y = est_MSS, color = MAC_string)) +
+ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = MAC_string)) +
   theme_bw() +
@@
   facet_wrap(~MSS_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated MSS (m/s)") +
@@
 
-As can be seen from the image, MSS is underestimated as cheating distance increases. The following image demonstrates the effect of cheating distance on estimated MAC:
+As can be seen from the image, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
 
@@
 # MAC
-ggplot(df, aes(x = cheat_distance, y = est_MAC, color = MSS_string)) +
+ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = MSS_string)) +
   theme_bw() +
@@
   facet_wrap(~MAC_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated MAC (m/s/s)") +
@@
 
-MAC (and also TAU) are highly affected by the cheating distance, and from the figure we can notice that MAC is overestimated as cheat distance increases.
-
-And finally, the following image demonstrates the effect of cheating distance on estimated PMAX:
+MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as cheat distance increases.
+
+And finally, the following image demonstrates the effect of flying start distance on estimated PMAX:
 
@@
 # PMAX
-ggplot(df, aes(x = cheat_distance, y = est_PMAX, color = MSS_string)) +
+ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)) +
   theme_bw() +
@@
   facet_wrap(~MAC_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated PMAX (W/kg)") +
@@
 
-Model residuals are also affected by cheating distance.
-The shape of residuals depends on splits utilized, but here we can see the effect of the cheating distance on the model residuals per split distance:
+Model residuals are also affected by flying start distance.
+The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40m versus 5, 15, 30m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simulation:
 
@@
     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-    group = paste(MSS, MAC, cheat_distance)
-  )
-
-ggplot(model_df, aes(y = residuals, x = distance, color = cheat_distance, group = group)) +
+    group = paste(MSS, MAC, flying_start_distance)
+  )
+
+ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
   theme_bw() +
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (sec)")
-```
-
-If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler images conveying issues with residuals when there is cheating at the start:
-
-```{r}
-ggplot(model_df, aes(y = residuals, x = distance, color = cheat_distance, group = group)) +
+  ylab("Predicted time - observed time (s)")
+```
+
+If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler images conveying issues with residuals when there is a flying start:
+
+```{r}
+ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
   theme_bw() +
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (sec)")
-```
-
-Clearly, cheating and sprint start timing issues and inconsistencies can result in biased parameters and predictions.
-This is particularly important in testing and comparing test results implementing different starting techniques and timing initiations (**REFERENCE**).
-Since speed trait is one of the hardest to improve, the effects of start inconsistencies can mask effects of the training intervention.
-It is thus of crucial importance to take these into account, either by standardizing the start (which is always a must), or by implementing the following techniques.
-
-### How to deal with the cheating or sprint start timing issues?
-
-Besides standardizing sprint starts, particularly for longitudinal monitoring of the training effects, one can implement additional strategies to make model estimates more robust.
-
-The simplest strategy is to used validity study results (i.e.
-comparing timing technique of interest to a gold standard measurement) and to use correction factors (**REFERENCE**).
-One way to implement correction is to use *time correction* in a same manner used with the radar gun synchronization issues.
-Thus, equation \@ref(time-distance) gets another form:
+  ylab("Predicted time - observed time (s)")
+```
+
+Clearly, any type of flying start where there is a difference between initial force production and start time can result in biased parameters and predictions.
+Since maximal sprint speed is difficult to improve, the effects of start inconsistencies can mask effects of the training intervention.
+It is thus crucial to standardize the start when testing and implementing the following techniques when using the \pkg{shorts} package.
+
+### How to overcome missing the initial force production when using timing gates?
+
+A potential solution is to use a correction factor - the recommendation in the literature is +0.5sec [@haugenSprintMechanicalProperties2020; @haugenSprintMechanicalVariables2019].
+Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 sec [@haugenDifferenceStartImpact2012].
+So, while a timing correction factor is warranted to avoid subsequent errors in estimates of kinetic variables (e.g., overestimate power), a correction factor that is too large will have the opposite effect (e.g., underestimate power).
+
+When implementing *time correction*, equation \@ref(eq:time-distance) gets another form:
 
@@
 
-If using validity studies, this time correction can be provided (i.e., 0.3-0.5seconds; *REFERENCE*), which is oftentimes the case, or it can me estimated as MSS and TAU is estimated.
+Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides estimation of this parameter from the data provide, together with MSS and TAU.
 To estimate time correction, we use `shorts::model_using_splits_with_time_correction` function.
-Here is how we can estimate Jack parameters using these two methods:
-
-```{r}
-jack_profile_fixed_time <- shorts::model_using_splits(
+Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5s) or estimated time correction:
+
+```{r}
+jack_profile_fixed_time_short <- shorts::model_using_splits(
   distance = split_times$distance,
@@
 
+jack_profile_fixed_time_long <- shorts::model_using_splits(
+  distance = split_times$distance,
+  time = split_times$jack_time,
+  time_correction = 0.5
+)
+
 jack_profile_time_estimated <- shorts::model_using_splits_with_time_correction(
@@
   unlist(jack_profile$parameters),
-  unlist(jack_profile_fixed_time$parameters),
+  unlist(jack_profile_fixed_time_short$parameters),
+  unlist(jack_profile_fixed_time_long$parameters),
   unlist(jack_profile_time_estimated$parameters)
@@
   "Jack - No corrections",
-  "Jack - Fixed time correction",
+  "Jack - Fixed time correction (+0.3s)",
+  "Jack - Fixed time correction (+0.5s)",
   "Jack - Estimated time correction"
@@
 
-In Jack's case, both fixed time correction and time correction estimation yield parameters closer to John's (i.e. true parameters).
-
-Another model definition, which is novel approach implemented in the `shorts` packages, is to ulitize *distance correction*, besides time correction.
-Thus, equation \@ref(time-distance) gets another form:
+In Jack's case, both +0.3s fixed time correction and time correction estimation yield parameters closer to John's (i.e. true parameters).
+
+Another model definition, which is novel approach implemented in the \pkg{shorts} packages, is to ulitize *distance correction*, besides time correction.
+Thus, equation \@ref(eq:time-distance) gets another form:
 
@@
   unlist(jack_profile$parameters),
-  unlist(jack_profile_fixed_time$parameters),
+  unlist(jack_profile_fixed_time_short$parameters),
+  unlist(jack_profile_fixed_time_long$parameters),
   unlist(jack_profile_time_estimated$parameters),
@@
   "Jack - No corrections",
-  "Jack - Fixed time correction",
+  "Jack - Fixed time correction (+0.3s)",
+  "Jack - Fixed time correction (+0.5s)",
   "Jack - Estimated time correction",
@@
 There are few issues with this model definition.
-Besides being novel and still not validated with practical data, distance correction model has four parameters to estimated, which implies that one needs at least five sprint splits.
-This imposes practical limitations, since acquiring six timing gate (one for the start and five to splits) might be practically troublesome.
-
-We will get back to these issues later, but let's see how these models perform using simulated data with varying cheating distance.
+Besides being novel and still not validated with practical data, distance correction model has four parameters to estimate, which implies that one needs at least five sprint splits.
+This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.
+
+We will get back to these issues later, but let's see how these models perform using simulated data with varying flying start distance.
 The following code contains the wrapper that performs all four models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction):
@@
 
-  fixed_correction <- shorts::model_using_splits(
+  fixed_correction_short <- shorts::model_using_splits(
     distance = data$distance,
@@
     time_correction = 0.3
+  )
+
+  fixed_correction_long <- shorts::model_using_splits(
+    distance = data$distance,
+    time = data$time,
+    time_correction = 0.5
   )
@@
     data.frame(
-      model = "Fixed correction",
-      t(unlist(fixed_correction$parameters))
+      model = "Fixed correction +0.3s",
+      t(unlist(fixed_correction_short$parameters))
+    ),
+    data.frame(
+      model = "Fixed correction +0.5s",
+      t(unlist(fixed_correction_long$parameters))
     ),
@@
     "No correction",
-    "Fixed correction",
+    "Fixed correction +0.3s",
+    "Fixed correction +0.5s",
     "Time correction",
@@
 df <- model_df %>%
-  group_by(MSS, TAU, cheat_distance, model) %>%
+  group_by(MSS, TAU, flying_start_distance, model) %>%
   slice(1) %>%
@@
 # MSS
-ggplot(df, aes(x = cheat_distance, y = est_MSS, color = model)) +
+ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = model)) +
   theme_bw(8) +
@@
   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated MSS (m/s)") +
@@
 
-Same conclusions can be said for the MAC parameter. Time and distance corrections model performs perfectly, while time correction model performs almost as good (although worse than for MSS parameter).
+Same conclusions can be said for the MAC parameter.
+Time and distance corrections model performs perfectly, while time correction model performs almost as good (although worse than for the MSS parameter).
 
@@
 # MAC
-ggplot(df, aes(x = cheat_distance, y = est_MAC, color = model)) +
+ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = model)) +
   theme_bw(8) +
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated MAC (m/s/s)") +
@@
 
-When it comes to PMAX, same conclusion can be reached as for MAC.
+When it comes to PMAX, same conclusion can be reached.
 
@@
 # PMAX
-ggplot(df, aes(x = cheat_distance, y = est_PMAX, color = model)) +
+ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = model)) +
   theme_bw(8) +
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated PMAX (W/kg)") +
@@
 
-The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the stolen time; indicated by the dashed line on the figure).
+The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the *stolen time*; indicated by the dashed line on the figure).
 
@@
 # time_correction
-ggplot(df, aes(x = cheat_distance, y = est_time_correction, color = model)) +
+ggplot(df, aes(x = flying_start_distance, y = est_time_correction, color = model)) +
   theme_bw(8) +
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
-  ylab("estimated time correction (sec)") +
+  xlab("Flying start distance (m)") +
+  ylab("estimated time correction (s)") +
   theme(legend.title = element_blank())
@@
 
-The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., cheat distance; indicated by the dashed line on the figure, which represents *identity line* since cheat distance is already on the x-axis).
+The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents *identity line* since flying start distance is already on the x-axis).
 
@@
 # distance_correction
-ggplot(df, aes(x = cheat_distance, y = est_distance_correction, color = model)) +
+ggplot(df, aes(x = flying_start_distance, y = est_distance_correction, color = model)) +
   theme_bw(8) +
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Flying start distance (m)") +
   ylab("estimated distance correction (m)") +
@@
 
-The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models, particularly without issues at specific distance zones.
+The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models:
 
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (sec)")
-```
-
-Given the simulation data, time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model in the face of start cheating.
+  ylab("Predicted time - observed time (s)")
+```
+
+Given the simulation data, time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model in the face of flying start issues.
 Since time correction model is simpler and demand three parameters to be estimated, it might be practically more useful than time and distance correction model, which demand four parameters estimation and thus more than five timing gates and sprint splits.
@@
 Starting behind the initial timing gate represent only one issue (i.e.
-cheating).
-In this section, we will simulate few other issues to check how sensitive are the presented models to perturbations that can happen in the practice.
+flying start).
+In this section, we simulate a few other issues to check how sensitive the presented models are to perturbations that can happen in practical settings during field testing.
 
@@
 
-One issue that might happen in certain timing gates setups is an individual triggering timing system before the sprint is initiated.
+One issue that might happen in certain timing gates setups is an individual triggering timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
 This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is *reaction time* (RT) involved.
@@
   MAC = c(6, 7, 8, 9),
-  time_lag = seq(0, 1, length.out = 50),
+  time_lag = seq(0, 0.5, length.out = 50),
   distance = c(5, 10, 20, 30, 40, 50)
@@
 set.seed(1667)
-rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
+rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
 sim_df$time <- sim_df$time + rand_noise
@@
   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-  xlab("Time lag (sec)") +
+  xlab("Time lag (s)") +
   ylab("estimated MSS (m/s)") +
@@
 
-From the figure below it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction model.
+From the figure below it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
 Time correction and time and distance corrections model correctly estimated MAC.
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Time lag (sec)") +
+  xlab("Time lag (s)") +
   ylab("estimated MAC (m/s/s)") +
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Time lag (sec)") +
-  ylab("estimated time correction (sec)") +
+  xlab("Time lag (s)") +
+  ylab("estimated time correction (s)") +
   theme(legend.title = element_blank())
@@
 The next figure depicts estimated distance correction for the time and distance correction model.
-The estimated distance correction parameters looks jumpy due random noise that we have to inject to allow model fit.
+The estimated distance correction parameters looks jumpy due to random noise that we have to added to allow model fit.
 
@@
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Cheat distance (m)") +
+  xlab("Time lag (s)") +
   ylab("estimated distance correction (m)") +
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (sec)")
-```
+  ylab("Predicted time - observed time (s)")
+```
+
+Additional scenario can create time lag: if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
+In this case, time lag is a negative number since it reduces the split times.
+This scenario will not be simulated in the current paper.
 
@@
 In this example, we will simulate the scenario where the initial timing gate is not positioned at $d=0$, but rather with some error (i.e., -0.5 to 0.5m).
+Although this scenario represents poor procedure implementation, we have included this simulation to depict effects on estimated parameters.
 
@@
   xlab("Displacement (m)") +
-  ylab("estimated time correction (sec)") +
+  ylab("estimated time correction (s)") +
   theme(legend.title = element_blank())
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (sec)")
-```
-
-Additional sources of issues with the timing gate involve combination of the previous three.
+  ylab("Predicted time - observed time (s)")
+```
+
+Additional sources of issues with the timing gates involve combination of the previous scenarios, as well as effect of number and distances of the timing gates [@haugenPowerForceVelocityProfilingSprinting2020; also see @haugenDifferenceStartImpact2012].
 For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence.
@@
 
-To estimate parameter stability, model over-fitting, and performance on the unseen data, `shorts` model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
+To estimate parameter stability, model over-fitting, and performance on the unseen data, \pkg{shorts} model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
 LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation.
@@
   xlab("Distance (m)") +
-  ylab("Predicted - observed time (sec)")
+  ylab("Predicted - observed time (s)")
 ```
@@
 Please note, that since LOOCV removes one observation, if model estimates three parameters, one needs to have at least five observations, since we need to make sure that model can be estimated once a single observation is removed.
-LOOCV can also be implemented with the mixed-effects models in the `shorts` package.
+LOOCV can also be implemented with the mixed-effects models in the \pkg{shorts} package.
 
@@
 
-Let's utilize demonstrated functionalities of the `shorts` package using real-world data.
-The first dataset comes from Usain Bolt run from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data set for 52 female soccer athletes which comes with the `shorts` package.
+Let's utilize demonstrated functionalities of the \pkg{shorts} package using real-world data.
+The first dataset comes from Usain Bolt run from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data set for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package.
 
@@
 ```{r}
-# Taken from London 2017
-# https://barrunning.com/resources/SiteUploads/20180715/London%202017%20REPORT%20FOR%20THE%20100%20m%20Men.pdf
 bolt_reaction_time <- 0.183
@@
 Soccer players were older (24.6±3.6 vs.
-18.9±2.7 yr, p < 0.000), however there were no differences for height (167.3±5.9 vs.
+18.9±2.7 yr, p < 0.0001), however there were no differences for height (167.3±5.9 vs.
 167.0±5.7 cm, p = 0.886), body mass (62.5±5.9 vs.
@@
 Briefly, all athletes performed a standardized warm-up that included general exercises such as jogging, shuffling, multi-directional movements, and dynamic stretching exercises.
-Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, and 35 meters at a height of approximately 1.0 meter.
+Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, 30, and 35 meters at a height of approximately 1.0 meter.
 Participants stood with their lead foot positioned approximately 5 cm behind the initial infrared beam (i.e., start line).
@@
 The best 35 m time, and all associated split times were kept for analysis.
-The assessment of linear sprints using infrared timing gates does not require familiarization [@moirInfluenceFamiliarizationReliability2004].
-
-To analyze this dataset, we will utilize mixed-effects models, although individual models can be utilized as well.
+
+Below is the mixed-effects models analysis of the dataset.
 
@@
 
-The following image represents model fit estimator RSE for each model. As can be seen, RSE drops the more flexible the model.
+The following image represents model fit estimator RSE for each model. As can be seen, RSE is reduced the more flexible the model.
 
@@
   data.frame(
-    model = "Fixed correction",
+    model = "Fixed correction +0.3s",
     t(unlist(fixed_correction$model_fit))
@@
     "No corrections",
-    "Fixed correction",
+    "Fixed correction +0.3s",
     "Time correction fixed",
@@
   data.frame(
-    model = "Fixed correction",
+    model = "Fixed correction +0.3s",
     fixed_correction$parameters$random
@@
 This provides practical anchor for the residual magnitude, often referred to as *smallest worthwhile change* (SWC) or *smallest effect size of interest* (SESOI) [@jovanovicBmbstatsBootstrapMagnitudebased2020].
-If the residuals are within this magnitude band, they have no practical significance, andthus the model is useful in making practically useful predictions.
+If the residuals are within this magnitude band, they have no practical significance, and thus the model is good in making practically useful predictions.
 
@@
   data.frame(
-    model = "Fixed correction",
+    model = "Fixed correction +0.3s",
     fixed_correction$data
@@
   geom_vline(xintercept = 0, color = "blue", alpha = 0.5) +
-  geom_jitter(aes(x = resid), alpha = 0.2, height = 0.25) +
-  geom_errorbarh(data = model_SESOI, aes(xmin = lower, xmax = upper), height = 0.1) +
-  geom_point(data = model_SESOI, aes(x = bias)) +
+  geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
+  geom_errorbarh(data = model_SESOI, aes(xmin = lower, xmax = upper), height = 0.1, color = "black") +
+  geom_point(data = model_SESOI, aes(x = bias), color = "black") +
   facet_wrap(~distance, scales = "free_x") +
-  xlab("Predicted time - observed time (sec)") +
+  xlab("Predicted time - observed time (s)") +
   ylab(NULL)
@@
 ggplot(df, aes(x = value, y = model)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_point() +
@@
 Which model should should be used?
-Although providing a better fit (using RSE as an estimator of model fit), time and distance correction models often estimate these parameters that are hard to interpret.
-Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for inconsistencies and/or cheating at the sprint starts.
-
-We are hoping that the `shorts` package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising a measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation inconsistencies.
+Although providing a better fit (using RSE as an estimator of model fit), time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction).
+Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for timing inconsistencies.
+
+We are hoping that the \pkg{shorts} package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
+
+```{r include=FALSE}
+# Return original options
+options(op) 
+```
 

### 2020-10-20 paper.Rmd
Reformat

@@
 
-options("width" = 80)
+options("width" = 60)
 
@@
   dpi = 600,
-  out.width = "90%",
+  out.width = "100%",
   fig.align = "center",
@@
   message = FALSE,
-  width = 80
+  width = 60
 )
@@
 op <- options()
-options(digits = 2)
+options(digits = 3)
 ```
@@
       labels = c(
-        expression("Distance (m"*")"),
-        expression("Velocity (ms"^-1*")"),
-        expression("Acceleration (ms"^-2*")"))
+        expression("Distance (m" * ")"),
+        expression("Velocity (ms"^-1 * ")"),
+        expression("Acceleration (ms"^-2 * ")")
+      )
     )
@@
 ) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.7) +
@@
     scales = "free_y",
-    labeller = label_parsed) +
+    labeller = label_parsed
+  ) +
   ylab(NULL) +
@@
 ) +
-  theme_bw(10) +
-  geom_line(alpha = 0.7) +
-  ylab(expression("Acceleration (ms"^-2*")")) +
-  xlab(expression("Velocity (ms"^-1*")")) +
+  theme_bw(8) +
+  geom_line(alpha = 0.7) +
+  ylab(expression("Acceleration (ms"^-2 * ")")) +
+  xlab(expression("Velocity (ms"^-1 * ")")) +
   theme(legend.position = "top", legend.title = element_blank())
@@
 ggplot(df, aes(x = distance)) +
-  theme_bw(10) +
+  theme_bw(8) +
   geom_line(aes(y = time)) +
@@
 ggplot(df, aes(x = distance, y = value)) +
-  theme_bw(10) +
+  theme_bw(8) +
   facet_wrap(~name, scales = "free_y") +
@@
 ) +
-  theme_bw(10) +
+  theme_bw(8) +
   facet_wrap(~name, scales = "free_y") +
@@
 sprint_time <- seq(0, 6, 1)
-
-# Generate velocity by using:
-# round(predict_velocity_at_time(time = sprint_time, MSS = 9, TAU = 1.3), 2)
 
@@
 ggplot(plot_df, aes(x = value, y = velocity, color = `Sync issue`)) +
-  theme_bw(10) +
+  theme_bw(8) +
   geom_line(alpha = 0.7) +
   xlab("Time (s)") +
-  ylab(expression("Velocity ("*ms^-1*")"))
+  ylab(expression("Velocity (" * ms^-1 * ")"))
 ```
@@
 # Using the true time
-round(
 predict_velocity_at_time(
@@
   TAU = m5$parameters$TAU
-  ),
-  2
 )
@@
 # Using time with sync issues
-round(
 predict_velocity_at_time(
@@
   time_correction = m8$parameters$time_correction
-  ),
-  2
 )
@@
 ggplot(plot_df, aes(x = distance, y = time, color = athlete, group = athlete)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_point() +
@@
 
-round(sprint_parameters, 2)
+sprint_parameters
 ```
@@
 
-head(round(sim_df, 2))
+head(sim_df)
 ```
@@
 ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = MAC_string)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.7) +
@@
 ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = MSS_string)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.7) +
@@
 ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.7) +
@@
 ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.3) +
@@
 ggplot(model_df, aes(y = residuals, x = distance, color = time_lag, group = group)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.3) +
@@
 
-head(round(sim_df, 2))
+head(sim_df, 2)
 ```
@@
 ggplot(model_df, aes(y = residuals, x = distance, color = displacement, group = group)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_line(alpha = 0.3) +
@@
 ggplot(df, aes(x = value)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_boxplot() +
@@
 ggplot(df, aes(x = distance, y = resid, color = name)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_hline(yintercept = 0, linetype = "dashed") +
@@
 
-rownames(bolt_model) <- bolt_model$model
-
-round(bolt_model[-1], 3)
+bolt_model
 ```
@@
 ggplot(model_fit, aes(x = RSE, y = model)) +
-  theme_bw() +
+  theme_bw(8) +
   geom_point() +

### 2020-10-20 paper.tex
Reformat

\documentclass[
]{jss}

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Mladen Jovanović\\University of Belgrade \And Jason D. Vescovi\\University of Toronto
}
\title{\pkg{shorts}: An R Package for Modeling Short Sprints}

\Plainauthor{Mladen Jovanović, Jason D. Vescovi}

\Abstract{
The abstract of the article.
}

\Keywords{sprints, sports, modeling}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Mladen Jovanović\\
    University of Belgrade\\
    Faculty of Sport and Physical Education\\
Belgrade, Serbia\\
  E-mail: \email{coach.mladen.jovanovic@gmail.com}\\

      Jason D. Vescovi\\
    University of Toronto\\
    Faculty of Kinesiology and Physical Education\\
Graduate School of Exercise Science\\
Toronto, ON Canada\\
  E-mail: \email{vescovij@gmail.com}\\

  }

% Pandoc header

\usepackage{amsmath}

\begin{document}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Short sprint performance is one of the most distinguishable and admired physical trait in sports. Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end. Peak anaerobic power is achieved within the first few seconds (\textless5 sec) of maximal efforts \citep{mangineSpeedForcePower2014}, whereas the ability to achieve maximal sprint speed varies based on the type of sport. For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)\citep{brownAssessmentLinearSprinting2004}.

Describing individual differences in short sprint performance, quantifying training intervention effects, but more importantly, understanding short sprint \emph{bottle necks} (i.e.~limiting factors) has been one of the prominent tasks of sports science profession since its establishment. The \pkg{shorts} package \citep{R-shorts}, written in the R language \citep{R-base}, represents an open-source tool that aims to help sports scientist in achieving these goals through mathematical modeling.

In the current paper, we will provide explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.

\hypertarget{mathematical-model}{%
\section{Mathematical model}\label{mathematical-model}}

Short sprints have been modeled using the mono-exponential equation \eqref{eq:velocity-time} originally proposed by \citet{doi:10.1098/rspb.1927.0035}, and more recently popularized by \citet{clarkNFLCombine40Yard2017}, and \citet{samozinoSimpleMethodMeasuring2016}. Equation \eqref{eq:velocity-time} represents function for instantaneous horizontal velocity \(v\) given the time \(t\) and two model parameters:

\begin{equation}
  v(t) = MSS \times (1 - e^{-\frac{t}{TAU}}) \label{eq:velocity-time}
\end{equation}

The parameters of the equation \eqref{eq:velocity-time} are \textit{maximum sprinting speed} (MSS; expressed in \(ms^{-1}\)) and \textit{relative acceleration} (TAU). Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in \(ms^{-2}\)) \eqref{eq:maximal-acceleration}.

\begin{equation}
  MAC = \frac{MSS}{TAU}\label{eq:maximal-acceleration}
\end{equation}

Although TAU is used in the equations, and later estimated, it is preferred to use MAC instead since it is easier to grasp, particularly for less math inclined coaches.

By derivating equation \eqref{eq:velocity-time}, we can get equation for horizontal acceleration \eqref{eq:acceleration-time}.

\begin{equation}
  a(t) = \frac{MSS}{TAU} \times e^{-\frac{t}{TAU}}  \label{eq:acceleration-time}
\end{equation}

By integrating equation \eqref{eq:velocity-time}, we can get equation for distance covered \eqref{eq:distance-time}.

\begin{equation}
  d(t) = MSS \times (t + TAU \times e^{-\frac{t}{TAU}}) - MSS \times TAU  \label{eq:distance-time}
\end{equation}

Let's consider four athletes with different levels of MSS (high versus low maximal sprinting speed) and MAC (high versus low maximal acceleration; as mentioned previously, using MAC is preferred over using TAU) (Table \ref{tab:four-athletes-table}).

\begin{CodeChunk}
\begin{table}

\caption{\label{tab:four-athletes-table}Four athletes with different MSS and MAC parameters.}
\centering
\begin{tabular}[t]{l|r|r|r}
\hline
Athlete & MSS & MAC & TAU\\
\hline
Athlete A & 12 & 10 & 1.20\\
\hline
Athlete B & 12 & 6 & 2.00\\
\hline
Athlete C & 8 & 10 & 0.80\\
\hline
Athlete D & 8 & 6 & 1.33\\
\hline
\end{tabular}
\end{table}

\end{CodeChunk}

The Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6s).

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/four-athletes-kinematics-1}

}

\caption[Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds]{Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds.}\label{fig:four-athletes-kinematics}
\end{figure}
\end{CodeChunk}

If we plot acceleration against velocity (Figure \ref{fig:four-athletes-profile}), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/four-athletes-profile-1}

}

\caption[Acceleration-Velocity profile of four athletes with different MSS and MAC parameters]{Acceleration-Velocity profile of four athletes with different MSS and MAC parameters.}\label{fig:four-athletes-profile}
\end{figure}
\end{CodeChunk}

\hypertarget{estimation-using-shorts-package}{%
\section{Estimation using shorts package}\label{estimation-using-shorts-package}}

Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), and (2) getting a velocity trace using a radar gun. Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \pkg{shorts} package using non-linear least squares regression implemented in the \texttt{nls} function in the \pkg{base} R \citep{R-base} and \texttt{nlme} function in the \pkg{nlme} package \citep{R-nlme} for the mixed-effect models.

\hypertarget{estimating-short-sprint-parameters-using-split-times}{%
\subsection{Estimating short sprint parameters using split times}\label{estimating-short-sprint-parameters-using-split-times}}

Let's consider an example of an athlete with MSS equal to 9 \(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^-2\) performing 40m sprint with timing gates positioned at each 10m split. When it comes to split times, distance is a predictor, and time is outcome variable, thus the equation \eqref{eq:velocity-time} takes another form:

\begin{equation}
  t(d) = TAU \times W(-e^{\frac{-d}{MSS \times TAU}} - 1) + \frac{d}{MSS} + TAU \label{eq:time-distance}
\end{equation}

\(W\) in equation \eqref{eq:velocity-time} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{shorts::model\_using\_splits} function:

\begin{CodeChunk}
\begin{CodeInput}
R> require(shorts)
R> 
R> split_distance <- c(10, 20, 30, 40)
R> 
R> split_time <- c(2.17, 3.43, 4.60, 5.73)
R> 
R> m1 <- shorts::model_using_splits(
+   distance = split_distance,
+   time = split_time
+ )
R> 
R> m1
\end{CodeInput}
\begin{CodeOutput}
#> Estimated model parameters
#> --------------------------
#>                 MSS                 TAU                 MAC 
#>                9.01                1.31                6.89 
#>                PMAX     time_correction distance_correction 
#>               15.52                0.00                0.00 
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>   0.00249   1.00000  -0.00178   0.00265   0.00265   0.00176 
#>       MAE      MAPE 
#>   0.00157   0.04742
\end{CodeOutput}
\end{CodeChunk}

Maximal relative power (PMAX) from the output is estimated using \(\frac{MSS \times MAC}{4}\), which disregards the air resistance. \texttt{time\_correction} and \texttt{distance\_corection} parameters will be covered later in the article.

Besides providing \emph{residual standard error} (RSE), \pkg{shorts} functions provide additional model fit estimators. Additional information can be gained by exploring the returned object, particularly object returned from the \texttt{nls} function:

\begin{CodeChunk}
\begin{CodeInput}
R> summary(m1$model)
\end{CodeInput}
\begin{CodeOutput}
#> 
#> Formula: corrected_time ~ TAU * I(LambertW::W(-exp(1)^(-distance/(MSS * 
#>     TAU) - 1))) + distance/MSS + TAU
#> 
#> Parameters:
#>     Estimate Std. Error t value Pr(>|t|)    
#> MSS   9.0121     0.0155     581  3.0e-06 ***
#> TAU   1.3083     0.0066     198  2.5e-05 ***
#> ---
#> Signif. codes:  
#> 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
#> 
#> Residual standard error: 0.00249 on 2 degrees of freedom
#> 
#> Number of iterations to convergence: 4 
#> Achieved convergence tolerance: 3.11e-06
\end{CodeOutput}
\end{CodeChunk}

Once we have estimated MSS and TAU, we can use \texttt{shorts::predict\_} family of functions to predict various relationships (i.e.~time at distance, acceleration at distance, velocity at time):

\begin{CodeChunk}
\begin{CodeInput}
R> # Predict time at distance
R> shorts::predict_time_at_distance(
+   distance = split_distance,
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 2.17 3.43 4.60 5.73
\end{CodeOutput}
\begin{CodeInput}
R> 
R> # Predict acceleration at time
R> shorts::predict_acceleration_at_time(
+   time = c(0, 1, 2, 3, 4, 5, 6),
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 6.8884 3.2075 1.4935 0.6954 0.3238 0.1508 0.0702
\end{CodeOutput}
\end{CodeChunk}

Here is a plot of observed and model predicted split times, with the help of the \pkg{ggplot2} package \citep{R-ggplot2}:

\begin{CodeChunk}
\begin{CodeInput}
R> require(ggplot2)
R> 
R> df <- data.frame(
+   distance = split_distance,
+   time = split_time,
+   pred_time = shorts::predict_time_at_distance(
+     distance = split_distance,
+     MSS = m1$parameters$MSS,
+     TAU = m1$parameters$TAU
+   )
+ )
R> 
R> ggplot(df, aes(x = distance)) +
+   theme_bw(8) +
+   geom_line(aes(y = time)) +
+   geom_point(aes(y = pred_time), color = "red") +
+   xlab("Distance (m)") +
+   ylab("Time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1} \end{center}

\end{CodeChunk}

\hypertarget{air-resistance-and-the-calculation-of-force-and-mechanical-power}{%
\subsubsection{Air resistance and the calculation of force and mechanical power}\label{air-resistance-and-the-calculation-of-force-and-mechanical-power}}

To estimate force production at distance or time (using \texttt{shorts::predict\_force\_at\_distance} and \texttt{shorts::predict\_force\_at\_time} functions), as well as power production (using \texttt{shorts::predict\_power\_at\_distance} and \texttt{shorts::predict\_power\_at\_time} functions), one needs to take into account the air resistance. Air resistance (N) is estimated using \texttt{shorts::get\_air\_resistance} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \citet{arsacModelingEnergetics100m2002}, \citet{samozinoSimpleMethodMeasuring2016}, and \citet{vaningenschenauCanCyclePower1991} for more information):

\begin{CodeChunk}
\begin{CodeInput}
R> shorts::get_air_resistance(
+   velocity = 5,
+   bodymass = 80,
+   bodyheight = 1.85,
+   barometric_pressure = 780,
+   air_temperature = 20,
+   wind_velocity = 0.5
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 6.1
\end{CodeOutput}
\end{CodeChunk}

When estimating force and power, the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{shorts::get\_air\_resistance}:

\begin{CodeChunk}
\begin{CodeInput}
R> # To calculate horizontal force produced
R> shorts::predict_force_at_distance(
+   distance = split_distance,
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU,
+   # Additional parameters forwarded to shorts::get_air_resistance
+   # Otherwise, defaults are used
+   bodymass = 80,
+   bodyheight = 1.85,
+   barometric_pressure = 780,
+   air_temperature = 20,
+   wind_velocity = 0.5
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 119.0  58.6  36.9  28.2
\end{CodeOutput}
\begin{CodeInput}
R> 
R> # To calculate power produced
R> shorts::predict_power_at_distance(
+   distance = split_distance,
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU,
+   # Additional parameters forwarded to shorts::get_air_resistance
+   # Otherwise, defaults are used
+   bodymass = 80,
+   bodyheight = 1.85,
+   barometric_pressure = 780,
+   air_temperature = 20,
+   wind_velocity = 0.5
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 868 490 323 251
\end{CodeOutput}
\end{CodeChunk}

The easiest way to get all kinematics for short sprints is to use \texttt{shorts::predict\_kinematics} functions:

\begin{CodeChunk}
\begin{CodeInput}
R> df <- shorts::predict_kinematics(
+   m1,
+   max_time = 6,
+   frequency = 100,
+   # Additional parameters forwarded to shorts::get_air_resistance
+   # Otherwise, defaults are used
+   bodymass = 80,
+   bodyheight = 1.85,
+   barometric_pressure = 780,
+   air_temperature = 20,
+   wind_velocity = 0.5
+ )
R> 
R> head(df)
\end{CodeInput}
\begin{CodeOutput}
#>   time distance velocity acceleration air_resistance force
#> 1 0.00 0.000000   0.0000         6.89        0.07536   551
#> 2 0.01 0.000344   0.0686         6.84        0.05609   547
#> 3 0.02 0.001371   0.1367         6.78        0.03978   543
#> 4 0.03 0.003076   0.2043         6.73        0.02636   539
#> 5 0.04 0.005455   0.2714         6.68        0.01576   534
#> 6 0.05 0.008502   0.3379         6.63        0.00792   530
#>   power relative_power
#> 1   0.0          0.000
#> 2  37.5          0.469
#> 3  74.2          0.928
#> 4 110.0          1.375
#> 5 145.0          1.813
#> 6 179.2          2.241
\end{CodeOutput}
\end{CodeChunk}

Plotting the model predictions can be done once we convert data from wide to long with the help of \pkg{dplyr} \citep{R-dplyr}, \pkg{tidyr} \citep{R-tidyr}, and \pkg{tidyverse} \citep{R-tidyverse} packages:

\begin{CodeChunk}
\begin{CodeInput}
R> require(tidyverse)
R> 
R> df <- pivot_longer(data = df, cols = -2)
R> 
R> ggplot(df, aes(x = distance, y = value)) +
+   theme_bw(8) +
+   facet_wrap(~name, scales = "free_y") +
+   geom_line(alpha = 0.7) +
+   ylab(NULL) +
+   xlab("Distance (m)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}

\end{CodeChunk}

\hypertarget{utility-functions}{%
\subsubsection{Utility functions}\label{utility-functions}}

Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90\% of maximum sprinting speed is reached, or where peak power is within 90\% range. To identify these values, \pkg{shorts} package comes with \texttt{shorts::find\_} family of functions:

\begin{CodeChunk}
\begin{CodeInput}
R> # Finds distance where 90% of maximum sprinting speed is reached
R> shorts::find_velocity_critical_distance(
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU,
+   percent = 0.9
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 16.5
\end{CodeOutput}
\begin{CodeInput}
R> 
R> # Finds maximal power and distance (this time using air resistance)
R> shorts::find_max_power_distance(
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU,
+   # Additional parameters forwarded to shorts::get_air_resistance
+   # Otherwise, defaults are used
+   bodymass = 80,
+   bodyheight = 1.85,
+   barometric_pressure = 780,
+   air_temperature = 20,
+   wind_velocity = 0.5
+ )
\end{CodeInput}
\begin{CodeOutput}
#> $max_power
#> [1] 1264
#> 
#> $distance
#> [1] 2.46
\end{CodeOutput}
\begin{CodeInput}
R> 
R> 
R> # Finds distance over 90% power range
R> shorts::find_power_critical_distance(
+   MSS = m1$parameters$MSS,
+   TAU = m1$parameters$TAU,
+   # Additional parameters forwarded to shorts::get_air_resistance
+   # Otherwise, defaults are used
+   bodymass = 80,
+   bodyheight = 1.85,
+   barometric_pressure = 780,
+   air_temperature = 20,
+   wind_velocity = 0.5
+ )
\end{CodeInput}
\begin{CodeOutput}
#> $lower
#> [1] 0.959
#> 
#> $upper
#> [1] 5.44
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{mixed-effects-model}{%
\subsubsection{Mixed-effects model}\label{mixed-effects-model}}

Sprint performance is often evaluated with a group of athletes (e.g., soccer club) representing a single strata of interest. Sports scientists can estimate individual profiles, or utilize mixed-effects models. To perform mixed-effects models in \pkg{shorts} for split times, one can use \texttt{shorts::mixed\_model\_using\_splits} function. To demonstrate this functionality, we load the \texttt{split\_times} dataset provided in the \pkg{shorts} package:

\begin{CodeChunk}
\begin{CodeInput}
R> data(split_times)
R> 
R> head(split_times)
\end{CodeInput}
\begin{CodeOutput}
#> # A tibble: 6 x 4
#>   athlete bodyweight distance     time
#>   <chr>        <dbl>    <dbl> <I<dbl>>
#> 1 John            75        5     1.20
#> 2 John            75       10     1.97
#> 3 John            75       15     2.66
#> 4 John            75       20     3.31
#> 5 John            75       30     4.59
#> 6 John            75       40     5.85
\end{CodeOutput}
\begin{CodeInput}
R> 
R> # Mixed model
R> m2 <- shorts::mixed_model_using_splits(
+   data = split_times,
+   distance = "distance",
+   time = "time",
+   athlete = "athlete",
+ 
+   # Select random effects
+   # Default is MSS and TAU
+   random = MSS + TAU ~ 1
+ )
R> 
R> m2
\end{CodeInput}
\begin{CodeOutput}
#> Estimated fixed model parameters
#> --------------------------------
#>                 MSS                 TAU                 MAC 
#>               8.065               0.655              12.309 
#>                PMAX     time_correction distance_correction 
#>              24.818               0.000               0.000 
#> 
#> Estimated random model parameters
#> ----------------------------------
#>     athlete  MSS   TAU  MAC PMAX time_correction
#> 1     James 9.69 0.847 11.4 27.7               0
#> 2       Jim 7.83 0.505 15.5 30.4               0
#> 3      John 7.78 0.727 10.7 20.8               0
#> 4 Kimberley 8.57 0.802 10.7 22.9               0
#> 5  Samantha 6.45 0.395 16.3 26.4               0
#>   distance_correction
#> 1                   0
#> 2                   0
#> 3                   0
#> 4                   0
#> 5                   0
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>    0.0260    0.9998   -0.0293    0.0496    0.0496    0.0214 
#>       MAE      MAPE 
#>    0.0172    0.9019
\end{CodeOutput}
\end{CodeChunk}

Additional information about mixed-effects model performed using the \texttt{nlme} package \citep{R-nlme} can be obtained using \texttt{summary}:

\begin{CodeChunk}
\begin{CodeInput}
R> summary(m2)
\end{CodeInput}
\begin{CodeOutput}
#> Nonlinear mixed-effects model fit by maximum likelihood
#>   Model: corrected_time ~ TAU * I(LambertW::W(-exp(1)^(-distance/(MSS *      TAU) - 1))) + distance/MSS + TAU 
#>  Data: train 
#>     AIC   BIC logLik
#>   -75.1 -66.7   43.5
#> 
#> Random effects:
#>  Formula: list(MSS ~ 1, TAU ~ 1)
#>  Level: athlete
#>  Structure: General positive-definite, Log-Cholesky parametrization
#>          StdDev Corr 
#> MSS      1.066  MSS  
#> TAU      0.178  0.877
#> Residual 0.026       
#> 
#> Fixed effects: MSS + TAU ~ 1 
#>     Value Std.Error DF t-value p-value
#> MSS  8.06     0.495 24   16.30       0
#> TAU  0.66     0.084 24    7.82       0
#>  Correlation: 
#>     MSS  
#> TAU 0.874
#> 
#> Standardized Within-Group Residuals:
#>    Min     Q1    Med     Q3    Max 
#> -1.909 -0.605  0.154  0.523  1.129 
#> 
#> Number of Observations: 30
#> Number of Groups: 5
\end{CodeOutput}
\end{CodeChunk}

The following plot contains kinematics for all athletes in \texttt{split\_times} dataset. Please note that power calculation takes default parameters for each individual:

\begin{CodeChunk}
\begin{CodeInput}
R> df <- predict_kinematics(m2, max_time = 10)
R> 
R> df <- pivot_longer(df, cols = c(-1, -3))
R> 
R> ggplot(
+   filter(df, distance < 40),
+   aes(x = distance, y = value, group = athlete, color = athlete)
+ ) +
+   theme_bw(8) +
+   facet_wrap(~name, scales = "free_y") +
+   geom_line(alpha = 0.7) +
+   ylab(NULL) +
+   xlab("Distance (m)") +
+   theme(legend.position = "top", legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}

\end{CodeChunk}

\hypertarget{estimating-short-sprint-parameters-using-radar-gun}{%
\subsection{Estimating short sprint parameters using radar gun}\label{estimating-short-sprint-parameters-using-radar-gun}}

Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as an target or outcome variable. Thus equation \eqref{eq:velocity-time} is used to estimate MSS and TAU.

Let's consider the same example of an athlete with MSS equal to 9\(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92\(ms^-2\) performing 40m sprint with velocity estimated using radar run (in this case with 1Hz sampling rate).

\begin{CodeChunk}
\begin{CodeInput}
R> sprint_time <- seq(0, 6, 1)
R> 
R> sprint_velocity <- c(0.00, 4.83, 7.07, 8.10, 8.59, 8.81, 8.91)
R> 
R> m3 <- shorts::model_using_radar(
+   velocity = sprint_velocity,
+   time = sprint_time
+ )
R> 
R> m3
\end{CodeInput}
\begin{CodeOutput}
#> Estimated model parameters
#> --------------------------
#>                 MSS                 TAU                 MAC 
#>                9.00                1.30                6.92 
#>                PMAX     time_correction distance_correction 
#>               15.58                0.00                0.00 
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>   0.00327   1.00000  -0.00406   0.00532   0.00532   0.00276 
#>       MAE      MAPE 
#>   0.00207       NaN
\end{CodeOutput}
\end{CodeChunk}

Both split and radar gun models allow the use of \emph{weighted} non-linear regression. For example, we can give more weight to shorter distance or faster velocities. Weighted non-linear regression is performed by setting \texttt{weights} parameter:

\begin{CodeChunk}
\begin{CodeInput}
R> m3_weighted <- shorts::model_using_radar(
+   velocity = sprint_velocity,
+   time = sprint_time,
+   weights = 1 / (sprint_velocity + 1)
+ )
R> 
R> m3_weighted
\end{CodeInput}
\begin{CodeOutput}
#> Estimated model parameters
#> --------------------------
#>                 MSS                 TAU                 MAC 
#>                9.00                1.30                6.92 
#>                PMAX     time_correction distance_correction 
#>               15.58                0.00                0.00 
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>   0.00108   1.00000  -0.00406   0.00534   0.00534   0.00276 
#>       MAE      MAPE 
#>   0.00206       NaN
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{mixed-effects-model-1}{%
\subsubsection{Mixed-effects model}\label{mixed-effects-model-1}}

Mixed-effects model using radar data is done using \texttt{shorts::mixed\_model\_using\_radar} function. To perform mixed model, let's load data that comes with \pkg{shorts} package.

\begin{CodeChunk}
\begin{CodeInput}
R> data("radar_gun_data")
R> 
R> head(radar_gun_data)
\end{CodeInput}
\begin{CodeOutput}
#> # A tibble: 6 x 4
#>   athlete bodyweight  time velocity
#>   <chr>        <dbl> <dbl>    <dbl>
#> 1 John            75  0       0    
#> 2 John            75  0.01    0.075
#> 3 John            75  0.02    0.149
#> 4 John            75  0.03    0.222
#> 5 John            75  0.04    0.291
#> 6 John            75  0.05    0.367
\end{CodeOutput}
\begin{CodeInput}
R> 
R> m4 <- shorts::mixed_model_using_radar(
+   radar_gun_data,
+   time = "time",
+   velocity = "velocity",
+   athlete = "athlete"
+ )
R> 
R> m4
\end{CodeInput}
\begin{CodeOutput}
#> Estimated fixed model parameters
#> --------------------------------
#>                 MSS                 TAU                 MAC 
#>                8.30                1.01                8.24 
#>                PMAX     time_correction distance_correction 
#>               17.09                0.00                0.00 
#> 
#> Estimated random model parameters
#> ----------------------------------
#>     athlete   MSS   TAU  MAC PMAX time_correction
#> 1     James 10.00 1.111 9.00 22.5               0
#> 2       Jim  8.00 0.889 9.00 18.0               0
#> 3      John  8.00 1.069 7.48 15.0               0
#> 4 Kimberley  9.01 1.286 7.01 15.8               0
#> 5  Samantha  6.50 0.685 9.50 15.4               0
#>   distance_correction
#> 1                   0
#> 2                   0
#> 3                   0
#> 4                   0
#> 5                   0
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>    0.0516    0.9994   -0.2191    0.1983    0.2191    0.0516 
#>       MAE      MAPE 
#>    0.0395       NaN
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{problems-with-estimation}{%
\section{Problems with estimation}\label{problems-with-estimation}}

There is a challenge when collecting sprint data that could have a substantial impact on modeled outcomes. To ensure accurate parameter outcomes, the initial force production must be synced with start time. Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \pkg{shorts} package.

\hypertarget{problems-with-time-sync-with-radar-gun}{%
\subsection{Problems with time sync with radar gun}\label{problems-with-time-sync-with-radar-gun}}

One source of error in the modeled estimation using radar gun is the time synchronization. In theory, synchronization is ideal when sprint is initiated at \(t=0\) (i.e., \(v(t=0) = 0\)). In practice, this is often not the case. Let's use our athlete and add and deduct 0.5second to simulate an error in synchronization and its effect on estimated MSS and TAU.

\begin{CodeChunk}
\begin{CodeInput}
R> df <- tibble(
+   `true time` = sprint_time,
+   velocity = sprint_velocity,
+   `0.5s added` = `true time` + 0.5,
+   `0.5s deducted` = `true time` - 0.5
+ )
R> 
R> head(df)
\end{CodeInput}
\begin{CodeOutput}
#> # A tibble: 6 x 4
#>   `true time` velocity `0.5s added` `0.5s deducted`
#>         <dbl>    <dbl>        <dbl>           <dbl>
#> 1           0     0             0.5            -0.5
#> 2           1     4.83          1.5             0.5
#> 3           2     7.07          2.5             1.5
#> 4           3     8.1           3.5             2.5
#> 5           4     8.59          4.5             3.5
#> 6           5     8.81          5.5             4.5
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")
R> 
R> ggplot(plot_df, aes(x = value, y = velocity, color = `Sync issue`)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   xlab("Time (s)") +
+   ylab(expression("Velocity (" * ms^-1 * ")"))
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}

\end{CodeChunk}

The following three models estimate MSS and TAU from the three datasets:

\begin{CodeChunk}
\begin{CodeInput}
R> # Without synchronization issues
R> m5 <- shorts::model_using_radar(
+   velocity = df$velocity,
+   time = df$`true time`
+ )
R> 
R> # With time added
R> m6 <- shorts::model_using_radar(
+   velocity = df$velocity,
+   time = df$`0.5s added`
+ )
R> 
R> # With time deducted
R> m7 <- shorts::model_using_radar(
+   velocity = df$velocity,
+   time = df$`0.5s deducted`
+ )
R> 
R> rbind(
+   data.frame(
+     model = "True time",
+     t(coef(m5))
+   ),
+   data.frame(
+     model = "Added 0.5s time",
+     t(coef(m6))
+   ),
+   data.frame(
+     model = "Deducted 0.5s time",
+     t(coef(m7))
+   )
+ )
\end{CodeInput}
\begin{CodeOutput}
#>                model   MSS  TAU  MAC PMAX time_correction
#> 1          True time  9.00 1.30 6.92 15.6               0
#> 2    Added 0.5s time  9.91 2.34 4.23 10.5               0
#> 3 Deducted 0.5s time 10.08 1.86 5.43 13.7               0
#>   distance_correction
#> 1                   0
#> 2                   0
#> 3                   0
\end{CodeOutput}
\end{CodeChunk}

As can be seen from the example, TAU is affected by an error in synchronization of time with velocity. The potential solution incorporated into the \pkg{shorts} package involves estimation of the \emph{time correction} parameter using the following equation:

\begin{equation}
  v(t) = MSS \times (1 - e^{-\frac{t + time \; correction}{TAU}}) \label{eq:velocity-time-correction}
\end{equation}

This model is incorporated in the \texttt{shorts::model\_using\_radar\_with\_time\_correction} function:

\begin{CodeChunk}
\begin{CodeInput}
R> # With time added
R> m8 <- shorts::model_using_radar_with_time_correction(
+   velocity = df$velocity,
+   time = df$`0.5s added`
+ )
R> coef(m8)
\end{CodeInput}
\begin{CodeOutput}
#>                 MSS                 TAU                 MAC 
#>                9.00                1.30                6.92 
#>                PMAX     time_correction distance_correction 
#>               15.58               -0.50                0.00
\end{CodeOutput}
\begin{CodeInput}
R> 
R> # With time deducted
R> m9 <- shorts::model_using_radar_with_time_correction(
+   velocity = df$velocity,
+   time = df$`0.5s deducted`
+ )
R> coef(m9)
\end{CodeInput}
\begin{CodeOutput}
#>                 MSS                 TAU                 MAC 
#>                9.00                1.30                6.92 
#>                PMAX     time_correction distance_correction 
#>               15.58                0.50                0.00
\end{CodeOutput}
\end{CodeChunk}

When using \texttt{shorts::predict\_} family of functions, one can provide estimated time correction to get predictions at original time scale.

\begin{CodeChunk}
\begin{CodeInput}
R> # Using the true time
R> predict_velocity_at_time(
+   time = df$`true time`,
+   MSS = m5$parameters$MSS,
+   TAU = m5$parameters$TAU
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 0.00 4.83 7.07 8.11 8.59 8.81 8.91
\end{CodeOutput}
\begin{CodeInput}
R> 
R> # Using time with sync issues
R> predict_velocity_at_time(
+   time = df$`0.5s added`,
+   MSS = m8$parameters$MSS,
+   TAU = m8$parameters$TAU,
+   time_correction = m8$parameters$time_correction
+ )
\end{CodeInput}
\begin{CodeOutput}
#> [1] 7.82e-05 4.83e+00 7.07e+00 8.11e+00 8.59e+00 8.81e+00
#> [7] 8.91e+00
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{mixed-model-approach}{%
\subsubsection{Mixed-model approach}\label{mixed-model-approach}}

When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the \texttt{shorts::mixed\_model\_using\_radar\_with\_time\_correction} function.

\begin{CodeChunk}
\begin{CodeInput}
R> # Adding 0.5s to radar_gun_data
R> radar_gun_data$time <- radar_gun_data$time + 0.5
R> 
R> # Mixed model with time correction being fixed effect
R> m10 <- shorts::mixed_model_using_radar_with_time_correction(
+   radar_gun_data,
+   time = "time",
+   velocity = "velocity",
+   athlete = "athlete",
+   random = MSS + TAU ~ 1
+ )
R> 
R> m10
\end{CodeInput}
\begin{CodeOutput}
#> Estimated fixed model parameters
#> --------------------------------
#>                 MSS                 TAU                 MAC 
#>                8.30                1.01                8.24 
#>                PMAX     time_correction distance_correction 
#>               17.10               -0.50                0.00 
#> 
#> Estimated random model parameters
#> ----------------------------------
#>     athlete   MSS   TAU  MAC PMAX time_correction
#> 1     James 10.00 1.111 9.00 22.5            -0.5
#> 2       Jim  8.00 0.889 9.00 18.0            -0.5
#> 3      John  8.00 1.069 7.48 15.0            -0.5
#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5
#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5
#>   distance_correction
#> 1                   0
#> 2                   0
#> 3                   0
#> 4                   0
#> 5                   0
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>    0.0516    0.9994   -0.2190    0.1983    0.2190    0.0516 
#>       MAE      MAPE 
#>    0.0395       Inf
\end{CodeOutput}
\begin{CodeInput}
R> 
R> 
R> # Mixed model with time correction being random effect
R> m11 <- shorts::mixed_model_using_radar_with_time_correction(
+   radar_gun_data,
+   time = "time",
+   velocity = "velocity",
+   athlete = "athlete",
+   random = MSS + TAU + time_correction ~ 1
+ )
R> 
R> m11
\end{CodeInput}
\begin{CodeOutput}
#> Estimated fixed model parameters
#> --------------------------------
#>                 MSS                 TAU                 MAC 
#>                8.30                1.01                8.24 
#>                PMAX     time_correction distance_correction 
#>               17.10               -0.50                0.00 
#> 
#> Estimated random model parameters
#> ----------------------------------
#>     athlete   MSS   TAU  MAC PMAX time_correction
#> 1     James 10.00 1.110 9.00 22.5            -0.5
#> 2       Jim  8.00 0.889 9.00 18.0            -0.5
#> 3      John  8.00 1.069 7.48 15.0            -0.5
#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5
#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5
#>   distance_correction
#> 1                   0
#> 2                   0
#> 3                   0
#> 4                   0
#> 5                   0
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>    0.0516    0.9994   -0.2188    0.1982    0.2188    0.0516 
#>       MAE      MAPE 
#>    0.0395       Inf
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{problems-at-the-start-when-using-split-times}{%
\subsection{Problems at the start when using split times}\label{problems-at-the-start-when-using-split-times}}

Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9\(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92\(ms^-2\). Let's call them John and Jack. The are both performing 40m sprint using timing gates set at 5, 10, 20, 30, and 40m. The initial timing gate at the start (i.e., \(d=0m\)) serves to activate the timing system (i.e., when they cross the beam).

John represents the \emph{theoretical model}, in which we assume that the initial force production and the timing initiation are perfectly synchronized. Jack, on the other hand, represents a \emph{practical model}, and decides to move slightly behind the initial timing gate (i.e.~for 0.5m) and use body rocking to initiate the sprint start. In other words, Jack is using a \emph{flying start}, a common scenario when testing field sports athletes. Let's see how their sprints differ.

\begin{CodeChunk}
\begin{CodeInput}
R> MSS <- 9
R> TAU <- 1.3
R> MAC <- MSS / TAU
R> 
R> split_times <- tibble(
+   distance = c(5, 10, 20, 30, 40),
+   john_time = shorts::predict_time_at_distance(distance, MSS, TAU),
+ 
+   # Jack's performance
+   jack_distance = distance + 0.5,
+   jack_true_time = shorts::predict_time_at_distance(jack_distance, MSS, TAU),
+   time_05m = shorts::predict_time_at_distance(0.5, MSS, TAU),
+   jack_time = jack_true_time - time_05m
+ )
R> 
R> split_times
\end{CodeInput}
\begin{CodeOutput}
#> # A tibble: 5 x 6
#>   distance john_time jack_distance jack_true_time time_05m
#>      <dbl>  <I<dbl>>         <dbl>       <I<dbl>> <I<dbl>>
#> 1        5      1.42           5.5           1.50    0.400
#> 2       10      2.17          10.5           2.23    0.400
#> 3       20      3.43          20.5           3.49    0.400
#> 4       30      4.60          30.5           4.65    0.400
#> 5       40      5.73          40.5           5.78    0.400
#> # ... with 1 more variable: jack_time <I<dbl>>
\end{CodeOutput}
\end{CodeChunk}

And here is a graphical representation of the sprint splits:

\begin{CodeChunk}
\begin{CodeInput}
R> plot_df <- split_times %>%
+   select(distance, john_time, jack_time) %>%
+   rename(John = john_time, Jack = jack_time) %>%
+   pivot_longer(cols = -1, names_to = "athlete", values_to = "time") %>%
+   mutate(distance = factor(distance))
R> 
R> ggplot(plot_df, aes(x = distance, y = time, color = athlete, group = athlete)) +
+   theme_bw(8) +
+   geom_point() +
+   geom_line() +
+   xlab("Distance (m)") +
+   ylab("Time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-23-1} \end{center}

\end{CodeChunk}

Using the following code, we can see how different are the estimated MSS and TAU parameters:

\begin{CodeChunk}
\begin{CodeInput}
R> # Since this is a perfect simulation and stats::nls will complain
R> # we need to add very small noise, or measurement error to the times
R> set.seed(1667)
R> rand_noise <- rnorm(nrow(split_times), 0, 10^-5)
R> split_times$john_time <- split_times$john_time + rand_noise
R> split_times$jack_time <- split_times$jack_time + rand_noise
R> 
R> john_profile <- shorts::model_using_splits(
+   distance = split_times$distance,
+   time = split_times$john_time
+ )
R> 
R> jack_profile <- shorts::model_using_splits(
+   distance = split_times$distance,
+   time = split_times$jack_time
+ )
R> 
R> sprint_parameters <- rbind(
+   unlist(john_profile$parameters),
+   unlist(jack_profile$parameters)
+ )
R> 
R> rownames(sprint_parameters) <- c("John", "Jack")
R> 
R> sprint_parameters
\end{CodeInput}
\begin{CodeOutput}
#>       MSS   TAU   MAC PMAX time_correction
#> John 9.00 1.300  6.92 15.6               0
#> Jack 8.49 0.704 12.06 25.6               0
#>      distance_correction
#> John                   0
#> Jack                   0
\end{CodeOutput}
\end{CodeChunk}

As can be seen from the results, flying starts yields biased estimates, particularly for the TAU, MAC and PMAX.

Below is a simulation sprint with 5, 10, 20, 30, 40, and 50m splits, with MSS and MAC varying from 6 to 9 (\(ms^-1\) and \(ms^-2\) respectively), and flying start distance varying from 0 to 1m.

\begin{CodeChunk}
\begin{CodeInput}
R> sim_df <- expand.grid(
+   MSS = c(6, 7, 8, 9),
+   MAC = c(6, 7, 8, 9),
+   flying_start_distance = c(
+     seq(0, 0.001, length.out = 20),
+     seq(0.001, 0.01, length.out = 20),
+     seq(0.01, 0.1, length.out = 20),
+     seq(0.1, 1, length.out = 20)
+   ),
+   distance = c(5, 10, 20, 30, 40, 50)
+ )
R> 
R> sim_df <- sim_df %>%
+   mutate(
+     TAU = MSS / MAC,
+     PMAX = MSS * MAC / 4,
+     true_distance = distance + flying_start_distance,
+     true_time = shorts::predict_time_at_distance(true_distance, MSS, TAU),
+     stolen_time = shorts::predict_time_at_distance(flying_start_distance, MSS, TAU),
+     time = true_time - stolen_time
+   )
R> 
R> # Add small noise to allow model fit
R> set.seed(1667)
R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
R> sim_df$time <- sim_df$time + rand_noise
R> 
R> head(sim_df)
\end{CodeInput}
\begin{CodeOutput}
#>   MSS MAC flying_start_distance distance   TAU PMAX
#> 1   6   6                     0        5 1.000  9.0
#> 2   7   6                     0        5 1.167 10.5
#> 3   8   6                     0        5 1.333 12.0
#> 4   9   6                     0        5 1.500 13.5
#> 5   6   7                     0        5 0.857 10.5
#> 6   7   7                     0        5 1.000 12.2
#>   true_distance    true_time stolen_time         time
#> 1             5 1.639197....           0 1.639356....
#> 2             5 1.579728....           0 1.579675....
#> 3             5 1.537453....           0 1.537537....
#> 4             5 1.505904....           0 1.505900....
#> 5             5 1.549960....           0 1.549860....
#> 6             5 1.488596....           0 1.488548....
\end{CodeOutput}
\end{CodeChunk}

Now when we have a simulation dataset, we can check the model estimates and predictions, given the flying start distance:

\begin{CodeChunk}
\begin{CodeInput}
R> # Prediction wrapper
R> pred_wrapper <- function(data) {
+   model <- shorts::model_using_splits(
+     distance = data$distance,
+     time = data$time
+   )
+ 
+   params <- data.frame(t(unlist(model$parameters)))
+ 
+   predicted_time <- shorts::predict_time_at_distance(
+     distance = data$distance,
+     MSS = model$parameters$MSS,
+     TAU = model$parameters$TAU
+   )
+ 
+   colnames(params) <- c(
+     "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
+     "est_time_correction", "est_distance_correction"
+   )
+ 
+   cbind(data, params, data.frame(predicted_time = as.numeric(predicted_time)))
+ }
R> 
R> # estimated parameters and predited time
R> model_df <- sim_df %>%
+   group_by(MSS, TAU, flying_start_distance) %>%
+   do(pred_wrapper(.)) %>%
+   ungroup()
R> 
R> # Prediction residuals
R> model_df$residuals <- model_df$predicted_time - model_df$time
\end{CodeInput}
\end{CodeChunk}

The following image demonstrates the effect of flying start distance on estimated MSS:

\begin{CodeChunk}
\begin{CodeInput}
R> # Estimates plot
R> df <- model_df %>%
+   group_by(MSS, TAU, flying_start_distance) %>%
+   slice(1) %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
+   )
R> 
R> # MSS
R> ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = MAC_string)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_wrap(~MSS_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated MSS (m/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-27-1} \end{center}

\end{CodeChunk}

As can be seen from the image, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:

\begin{CodeChunk}
\begin{CodeInput}
R> # MAC
R> ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = MSS_string)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_wrap(~MAC_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated MAC (m/s/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-28-1} \end{center}

\end{CodeChunk}

MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as cheat distance increases.

And finally, the following image demonstrates the effect of flying start distance on estimated PMAX:

\begin{CodeChunk}
\begin{CodeInput}
R> # PMAX
R> ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_wrap(~MAC_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated PMAX (W/kg)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-29-1} \end{center}

\end{CodeChunk}

Estimated PMAX is also overestimated as cheat distance increases.

Model residuals are also affected by flying start distance. The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40m versus 5, 15, 30m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simulation:

\begin{CodeChunk}
\begin{CodeInput}
R> # Residuals
R> model_df <- model_df %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
+     group = paste(MSS, MAC, flying_start_distance)
+   )
R> 
R> ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.3) +
+   facet_grid(MSS_string ~ MAC_string) +
+   geom_hline(yintercept = 0, linetype = "dashed") +
+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
+   xlab("Distance (m)") +
+   ylab("Predicted time - observed time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-30-1} \end{center}

\end{CodeChunk}

If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler images conveying issues with residuals when there is a flying start:

\begin{CodeChunk}
\begin{CodeInput}
R> ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.3) +
+   geom_hline(yintercept = 0, linetype = "dashed") +
+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
+   xlab("Distance (m)") +
+   ylab("Predicted time - observed time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-31-1} \end{center}

\end{CodeChunk}

Clearly, any type of flying start where there is a difference between initial force production and start time can result in biased parameters and predictions. Since maximal sprint speed is difficult to improve, the effects of start inconsistencies can mask effects of the training intervention. It is thus crucial to standardize the start when testing and implementing the following techniques when using the \pkg{shorts} package.

\hypertarget{how-to-overcome-missing-the-initial-force-production-when-using-timing-gates}{%
\subsubsection{How to overcome missing the initial force production when using timing gates?}\label{how-to-overcome-missing-the-initial-force-production-when-using-timing-gates}}

A potential solution is to use a correction factor - the recommendation in the literature is +0.5sec \citep{haugenSprintMechanicalProperties2020, haugenSprintMechanicalVariables2019}. Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 sec \citep{haugenDifferenceStartImpact2012}. So, while a timing correction factor is warranted to avoid subsequent errors in estimates of kinetic variables (e.g., overestimate power), a correction factor that is too large will have the opposite effect (e.g., underestimate power).

When implementing \emph{time correction}, equation \eqref{eq:time-distance} gets another form:

\begin{equation}
  t(d) = TAU \times W(-e^{\frac{-d}{MSS \times TAU}} - 1) + \frac{d}{MSS} + TAU - time \; correction \label{eq:time-correction}
\end{equation}

Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides estimation of this parameter from the data provide, together with MSS and TAU. To estimate time correction, we use \texttt{shorts::model\_using\_splits\_with\_time\_correction} function. Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5s) or estimated time correction:

\begin{CodeChunk}
\begin{CodeInput}
R> jack_profile_fixed_time_short <- shorts::model_using_splits(
+   distance = split_times$distance,
+   time = split_times$jack_time,
+   time_correction = 0.3
+ )
R> 
R> jack_profile_fixed_time_long <- shorts::model_using_splits(
+   distance = split_times$distance,
+   time = split_times$jack_time,
+   time_correction = 0.5
+ )
R> 
R> jack_profile_time_estimated <- shorts::model_using_splits_with_time_correction(
+   distance = split_times$distance,
+   time = split_times$jack_time
+ )
R> 
R> jack_parameters <- rbind(
+   unlist(john_profile$parameters),
+   unlist(jack_profile$parameters),
+   unlist(jack_profile_fixed_time_short$parameters),
+   unlist(jack_profile_fixed_time_long$parameters),
+   unlist(jack_profile_time_estimated$parameters)
+ )
R> 
R> rownames(jack_parameters) <- c(
+   "John",
+   "Jack - No corrections",
+   "Jack - Fixed time correction (+0.3s)",
+   "Jack - Fixed time correction (+0.5s)",
+   "Jack - Estimated time correction"
+ )
R> 
R> jack_parameters
\end{CodeInput}
\begin{CodeOutput}
#>                                       MSS   TAU   MAC PMAX
#> John                                 9.00 1.300  6.92 15.6
#> Jack - No corrections                8.49 0.704 12.06 25.6
#> Jack - Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2
#> Jack - Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1
#> Jack - Estimated time correction     8.96 1.216  7.37 16.5
#>                                      time_correction
#> John                                           0.000
#> Jack - No corrections                          0.000
#> Jack - Fixed time correction (+0.3s)           0.300
#> Jack - Fixed time correction (+0.5s)           0.500
#> Jack - Estimated time correction               0.284
#>                                      distance_correction
#> John                                                   0
#> Jack - No corrections                                  0
#> Jack - Fixed time correction (+0.3s)                   0
#> Jack - Fixed time correction (+0.5s)                   0
#> Jack - Estimated time correction                       0
\end{CodeOutput}
\end{CodeChunk}

In Jack's case, both +0.3s fixed time correction and time correction estimation yield parameters closer to John's (i.e.~true parameters).

Another model definition, which is novel approach implemented in the \pkg{shorts} packages, is to ulitize \emph{distance correction}, besides time correction. Thus, equation \eqref{eq:time-distance} gets another form:

\begin{equation}
  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction \label{eq:distance-correction}
\end{equation}

This model is implemented in \texttt{shorts::model\_using\_splits\_with\_corrections} function. Let's check this model estimates:

\begin{CodeChunk}
\begin{CodeInput}
R> jack_profile_distance_correction <- shorts::model_using_splits_with_corrections(
+   distance = split_times$distance,
+   time = split_times$jack_time
+ )
R> 
R> jack_parameters <- rbind(
+   unlist(john_profile$parameters),
+   unlist(jack_profile$parameters),
+   unlist(jack_profile_fixed_time_short$parameters),
+   unlist(jack_profile_fixed_time_long$parameters),
+   unlist(jack_profile_time_estimated$parameters),
+   unlist(jack_profile_distance_correction$parameters)
+ )
R> 
R> rownames(jack_parameters) <- c(
+   "John",
+   "Jack - No corrections",
+   "Jack - Fixed time correction (+0.3s)",
+   "Jack - Fixed time correction (+0.5s)",
+   "Jack - Estimated time correction",
+   "Jack - Estimated distance correction"
+ )
R> 
R> jack_parameters
\end{CodeInput}
\begin{CodeOutput}
#>                                       MSS   TAU   MAC PMAX
#> John                                 9.00 1.300  6.92 15.6
#> Jack - No corrections                8.49 0.704 12.06 25.6
#> Jack - Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2
#> Jack - Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1
#> Jack - Estimated time correction     8.96 1.216  7.37 16.5
#> Jack - Estimated distance correction 9.00 1.301  6.92 15.6
#>                                      time_correction
#> John                                           0.000
#> Jack - No corrections                          0.000
#> Jack - Fixed time correction (+0.3s)           0.300
#> Jack - Fixed time correction (+0.5s)           0.500
#> Jack - Estimated time correction               0.284
#> Jack - Estimated distance correction           0.400
#>                                      distance_correction
#> John                                               0.000
#> Jack - No corrections                              0.000
#> Jack - Fixed time correction (+0.3s)               0.000
#> Jack - Fixed time correction (+0.5s)               0.000
#> Jack - Estimated time correction                   0.000
#> Jack - Estimated distance correction               0.503
\end{CodeOutput}
\end{CodeChunk}

As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters. There are few issues with this model definition. Besides being novel and still not validated with practical data, distance correction model has four parameters to estimate, which implies that one needs at least five sprint splits. This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.

We will get back to these issues later, but let's see how these models perform using simulated data with varying flying start distance. The following code contains the wrapper that performs all four models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction):

\begin{CodeChunk}
\begin{CodeInput}
R> pred_wrapper <- function(data) {
+   no_correction <- shorts::model_using_splits(
+     distance = data$distance,
+     time = data$time
+   )
+ 
+   fixed_correction_short <- shorts::model_using_splits(
+     distance = data$distance,
+     time = data$time,
+     time_correction = 0.3
+   )
+ 
+   fixed_correction_long <- shorts::model_using_splits(
+     distance = data$distance,
+     time = data$time,
+     time_correction = 0.5
+   )
+ 
+   time_correction <- shorts::model_using_splits_with_time_correction(
+     distance = data$distance,
+     time = data$time,
+     control = nls.control(tol = 1)
+   )
+ 
+   time_dist_correction <- shorts::model_using_splits_with_corrections(
+     distance = data$distance,
+     time = data$time,
+     control = nls.control(tol = 1)
+   )
+ 
+ 
+   params <- rbind(
+     data.frame(
+       model = "No correction",
+       t(unlist(no_correction$parameters))
+     ),
+     data.frame(
+       model = "Fixed correction +0.3s",
+       t(unlist(fixed_correction_short$parameters))
+     ),
+     data.frame(
+       model = "Fixed correction +0.5s",
+       t(unlist(fixed_correction_long$parameters))
+     ),
+     data.frame(
+       model = "Time correction",
+       t(unlist(time_correction$parameters))
+     ),
+     data.frame(
+       model = "Time and distance correction",
+       t(unlist(time_dist_correction$parameters))
+     )
+   )
+ 
+   colnames(params) <- c(
+     "model", "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
+     "est_time_correction", "est_distance_correction"
+   )
+ 
+   df <- expand_grid(
+     data,
+     params
+   )
+ 
+   df$predicted_time <- shorts::predict_time_at_distance(
+     distance = df$distance,
+     MSS = df$est_MSS,
+     TAU = df$est_TAU,
+     time_correction = df$est_time_correction,
+     distance_correction = df$est_distance_correction
+   )
+ 
+   df$residuals <- df$predicted_time - df$time
+   return(df)
+ }
R> 
R> # estimated parameters and predited time
R> model_df <- sim_df %>%
+   group_by(MSS, TAU, flying_start_distance) %>%
+   do(pred_wrapper(.)) %>%
+   ungroup()
\end{CodeInput}
\end{CodeChunk}

As can be seen from the next figure, estimated time correction model estimates MSS almost perfectly, while estimated time and distance correction model estimates MSS perfectly.

\begin{CodeChunk}
\begin{CodeInput}
R> model_df$model <- factor(
+   model_df$model,
+   levels = c(
+     "No correction",
+     "Fixed correction +0.3s",
+     "Fixed correction +0.5s",
+     "Time correction",
+     "Time and distance correction"
+   )
+ )
R> # Estimates plot
R> df <- model_df %>%
+   group_by(MSS, TAU, flying_start_distance, model) %>%
+   slice(1) %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
+   )
R> 
R> # MSS
R> ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated MSS (m/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}

\end{CodeChunk}

Same conclusions can be said for the MAC parameter. Time and distance corrections model performs perfectly, while time correction model performs almost as good (although worse than for the MSS parameter).

\begin{CodeChunk}
\begin{CodeInput}
R> # MAC
R> ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated MAC (m/s/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}

\end{CodeChunk}

When it comes to PMAX, same conclusion can be reached.

\begin{CodeChunk}
\begin{CodeInput}
R> # PMAX
R> ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated PMAX (W/kg)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}

\end{CodeChunk}

The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the \emph{stolen time}; indicated by the dashed line on the figure).

\begin{CodeChunk}
\begin{CodeInput}
R> # time_correction
R> ggplot(df, aes(x = flying_start_distance, y = est_time_correction, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   geom_line(aes(y = stolen_time), color = "black", linetype = "dashed") +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated time correction (s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}

\end{CodeChunk}

The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents \emph{identity line} since flying start distance is already on the x-axis).

\begin{CodeChunk}
\begin{CodeInput}
R> # distance_correction
R> ggplot(df, aes(x = flying_start_distance, y = est_distance_correction, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   geom_abline(slope = 1, color = "black", linetype = "dashed") +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Flying start distance (m)") +
+   ylab("estimated distance correction (m)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-39-1} \end{center}

\end{CodeChunk}

The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models:

\begin{CodeChunk}
\begin{CodeInput}
R> # Residuals
R> model_df <- model_df %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
+     group = paste(MSS, MAC, flying_start_distance)
+   )
R> 
R> ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.3) +
+   facet_wrap(~model) +
+   geom_hline(yintercept = 0, linetype = "dashed") +
+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
+   xlab("Distance (m)") +
+   ylab("Predicted time - observed time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-40-1} \end{center}

\end{CodeChunk}

Given the simulation data, time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model in the face of flying start issues. Since time correction model is simpler and demand three parameters to be estimated, it might be practically more useful than time and distance correction model, which demand four parameters estimation and thus more than five timing gates and sprint splits.

Time correction and time and distance corrections are also implemented in the mixed-models using \texttt{shorts::mixed\_model\_using\_splits\_with\_time\_correction} and \texttt{shorts::mixed\_model\_using\_splits\_with\_corrections}. We will showcase their use at the end of this article.

\hypertarget{simulation-of-additional-starting-issues}{%
\subsubsection{Simulation of additional starting issues}\label{simulation-of-additional-starting-issues}}

Starting behind the initial timing gate represent only one issue (i.e.~flying start). In this section, we simulate a few other issues to check how sensitive the presented models are to perturbations that can happen in practical settings during field testing.

\hypertarget{triggering-the-gate-before-the-start}{%
\paragraph{Triggering the gate before the start}\label{triggering-the-gate-before-the-start}}

One issue that might happen in certain timing gates setups is an individual triggering timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely). This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is \emph{reaction time} (RT) involved. Both of these scenarios represent \emph{time lag} that is added to the split times. Let's simulate the effect of this time lag on model estimates and predictions.

\begin{CodeChunk}
\begin{CodeInput}
R> sim_df <- expand.grid(
+   MSS = c(6, 7, 8, 9),
+   MAC = c(6, 7, 8, 9),
+   time_lag = seq(0, 0.5, length.out = 50),
+   distance = c(5, 10, 20, 30, 40, 50)
+ )
R> 
R> sim_df <- sim_df %>%
+   mutate(
+     TAU = MSS / MAC,
+     PMAX = MSS * MAC / 4,
+     true_time = shorts::predict_time_at_distance(distance, MSS, TAU),
+     time = true_time + time_lag
+   )
R> 
R> # Add small noise to allow model fit
R> set.seed(1667)
R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
R> sim_df$time <- sim_df$time + rand_noise
R> 
R> head(sim_df)
\end{CodeInput}
\begin{CodeOutput}
#>   MSS MAC time_lag distance   TAU PMAX    true_time
#> 1   6   6        0        5 1.000  9.0 1.639197....
#> 2   7   6        0        5 1.167 10.5 1.579728....
#> 3   8   6        0        5 1.333 12.0 1.537453....
#> 4   9   6        0        5 1.500 13.5 1.505904....
#> 5   6   7        0        5 0.857 10.5 1.549960....
#> 6   7   7        0        5 1.000 12.2 1.488596....
#>           time
#> 1 1.639356....
#> 2 1.579675....
#> 3 1.537537....
#> 4 1.505900....
#> 5 1.549860....
#> 6 1.488548....
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> # estimated parameters and predicted time
R> model_df <- sim_df %>%
+   group_by(MSS, TAU, time_lag) %>%
+   do(pred_wrapper(.)) %>%
+   ungroup()
\end{CodeInput}
\end{CodeChunk}

From the figure below it can be seen that time lag affects estimated MSS for the the model without correction and fixed correction model. Time correction and time and distance corrections model correctly estimated MSS.

\begin{CodeChunk}
\begin{CodeInput}
R> model_df$model <- factor(
+   model_df$model,
+   levels = c(
+     "No correction",
+     "Fixed correction +0.3s",
+     "Fixed correction +0.5s",
+     "Time correction",
+     "Time and distance correction"
+   )
+ )
R> # Estimates plot
R> df <- model_df %>%
+   group_by(MSS, TAU, time_lag, model) %>%
+   slice(1) %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
+   )
R> 
R> # MSS
R> ggplot(df, aes(x = time_lag, y = est_MSS, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
+   xlab("Time lag (s)") +
+   ylab("estimated MSS (m/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}

\end{CodeChunk}

From the figure below it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models. Time correction and time and distance corrections model correctly estimated MAC.

\begin{CodeChunk}
\begin{CodeInput}
R> # MAC
R> ggplot(df, aes(x = time_lag, y = est_MAC, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Time lag (s)") +
+   ylab("estimated MAC (m/s/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}

\end{CodeChunk}

Two figure below depicts correctly identified time lag (i.e.~using time correction parameter) using time correction and time and distance corrections models.

\begin{CodeChunk}
\begin{CodeInput}
R> # time_correction
R> ggplot(df, aes(x = time_lag, y = est_time_correction, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   geom_abline(slope = -1, color = "black", linetype = "dashed") +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Time lag (s)") +
+   ylab("estimated time correction (s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}

\end{CodeChunk}

The next figure depicts estimated distance correction for the time and distance correction model. The estimated distance correction parameters looks jumpy due to random noise that we have to added to allow model fit.

\begin{CodeChunk}
\begin{CodeInput}
R> # distance_correction
R> ggplot(df, aes(x = time_lag, y = est_distance_correction, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Time lag (s)") +
+   ylab("estimated distance correction (m)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}

\end{CodeChunk}

The following figure depicts residuals (i.e., predicted time minus observed time).

\begin{CodeChunk}
\begin{CodeInput}
R> # Residuals
R> model_df <- model_df %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
+     group = paste(MSS, MAC, time_lag)
+   )
R> 
R> ggplot(model_df, aes(y = residuals, x = distance, color = time_lag, group = group)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.3) +
+   facet_wrap(~model) +
+   geom_hline(yintercept = 0, linetype = "dashed") +
+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
+   xlab("Distance (m)") +
+   ylab("Predicted time - observed time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}

\end{CodeChunk}

Additional scenario can create time lag: if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts). In this case, time lag is a negative number since it reduces the split times. This scenario will not be simulated in the current paper.

\hypertarget{bad-position-of-the-timing-gates}{%
\paragraph{Bad position of the timing gates}\label{bad-position-of-the-timing-gates}}

Another common issue with timing gates in the practical field settings is the bad measurement of the distance and thus bad positions of the timing gates. In this example, we will simulate the scenario where the initial timing gate is not positioned at \(d=0\), but rather with some error (i.e., -0.5 to 0.5m). Although this scenario represents poor procedure implementation, we have included this simulation to depict effects on estimated parameters.

\begin{CodeChunk}
\begin{CodeInput}
R> sim_df <- expand.grid(
+   MSS = c(6, 7, 8, 9),
+   MAC = c(6, 7, 8, 9),
+   displacement = seq(-0.5, 0.5, length.out = 51),
+   distance = c(5, 10, 20, 30, 40, 50)
+ )
R> 
R> sim_df <- sim_df %>%
+   mutate(
+     TAU = MSS / MAC,
+     PMAX = MSS * MAC / 4,
+     true_distance = distance + displacement,
+     time = shorts::predict_time_at_distance(true_distance, MSS, TAU)
+   )
R> 
R> # Add small noise to allow model fit
R> set.seed(1667)
R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
R> sim_df$time <- sim_df$time + rand_noise
R> 
R> head(sim_df, 2)
\end{CodeInput}
\begin{CodeOutput}
#>   MSS MAC displacement distance  TAU PMAX true_distance
#> 1   6   6         -0.5        5 1.00  9.0           4.5
#> 2   7   6         -0.5        5 1.17 10.5           4.5
#>           time
#> 1 1.534434....
#> 2 1.481967....
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> # estimated parameters and predicted time
R> model_df <- sim_df %>%
+   group_by(MSS, TAU, displacement) %>%
+   do(pred_wrapper(.)) %>%
+   ungroup()
\end{CodeInput}
\end{CodeChunk}

Figure below depicts effect of timing gate displacement on the MSS parameter. Time correction and time and distance correction models are almost identical in their correct estimation of the MSS.

\begin{CodeChunk}
\begin{CodeInput}
R> model_df$model <- factor(
+   model_df$model,
+   levels = c(
+     "No correction",
+     "Fixed correction +0.3s",
+     "Fixed correction +0.5s",
+     "Time correction",
+     "Time and distance correction"
+   )
+ )
R> 
R> # Estimates plot
R> df <- model_df %>%
+   group_by(MSS, TAU, displacement, model) %>%
+   slice(1) %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
+   )
R> 
R> # MSS
R> ggplot(df, aes(x = displacement, y = est_MSS, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
+   xlab("Displacement (m)") +
+   ylab("estimated MSS (m/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}

\end{CodeChunk}

Figure below depicts effect of timing gate displacement on the MSS parameter. Time and distance correction model is the only model that estimated MAC correctly.

\begin{CodeChunk}
\begin{CodeInput}
R> # MAC
R> ggplot(df, aes(x = displacement, y = est_MAC, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Displacement (m)") +
+   ylab("estimated MAC (m/s/s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}

\end{CodeChunk}

Figures below depicts estimated time and distance corrections parameters.

\begin{CodeChunk}
\begin{CodeInput}
R> # time_correction
R> ggplot(df, aes(x = displacement, y = est_time_correction, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Displacement (m)") +
+   ylab("estimated time correction (s)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}

\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> # distance_correction
R> ggplot(df, aes(x = displacement, y = est_distance_correction, color = model)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.7) +
+   geom_abline(slope = 1, color = "black", linetype = "dashed") +
+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
+   xlab("Displacement (m)") +
+   ylab("estimated distance correction (m)") +
+   theme(legend.title = element_blank())
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}

\end{CodeChunk}

The following figure depicts residuals (i.e., predicted time minus observed time).

\begin{CodeChunk}
\begin{CodeInput}
R> # Residuals
R> model_df <- model_df %>%
+   mutate(
+     MSS_string = paste("MSS =", MSS),
+     TAU_string = paste("TAU =", TAU),
+     MAC_string = paste("MAC = ", round(MAC, 2)),
+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
+     group = paste(MSS, MAC, displacement)
+   )
R> 
R> ggplot(model_df, aes(y = residuals, x = distance, color = displacement, group = group)) +
+   theme_bw(8) +
+   geom_line(alpha = 0.3) +
+   facet_wrap(~model) +
+   geom_hline(yintercept = 0, linetype = "dashed") +
+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
+   xlab("Distance (m)") +
+   ylab("Predicted time - observed time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}

\end{CodeChunk}

Additional sources of issues with the timing gates involve combination of the previous scenarios, as well as effect of number and distances of the timing gates \citetext{\citealp{haugenPowerForceVelocityProfilingSprinting2020}; \citealp[also see][]{haugenDifferenceStartImpact2012}}. For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence. More elaborate simulation is beyond the scope of the current paper.

\hypertarget{leave-one-out-cross-validation}{%
\section{Leave-one-out Cross-Validation}\label{leave-one-out-cross-validation}}

To estimate parameter stability, model over-fitting, and performance on the unseen data, \pkg{shorts} model function comes with implemented \emph{leave-one-out cross validation} (LOOCV) \citep{jamesIntroductionStatisticalLearning2017, jovanovicBmbstatsBootstrapMagnitudebased2020, kuhnAppliedPredictiveModeling2018}. LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation. This process is repeated for each observations in the model dataset. LOOCV allows one to check estimated parameters stability, and model performance on the unseen data.

Let's perform LOOCV using Jack's data and time correction model:

\begin{CodeChunk}
\begin{CodeInput}
R> jack_LOOCV <- shorts::model_using_splits_with_time_correction(
+   distance = split_times$distance,
+   time = split_times$jack_time,
+   LOOCV = TRUE
+ )
R> 
R> jack_LOOCV
\end{CodeInput}
\begin{CodeOutput}
#> Estimated model parameters
#> --------------------------
#>                 MSS                 TAU                 MAC 
#>               8.958               1.216               7.367 
#>                PMAX     time_correction distance_correction 
#>              16.499               0.284               0.000 
#> 
#> Model fit estimators
#> --------------------
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>   0.00181   1.00000  -0.00109   0.00189   0.00189   0.00114 
#>       MAE      MAPE 
#>   0.00104   0.04981 
#> 
#> 
#> Leave-One-Out Cross-Validation
#> ------------------------------
#> Parameters:
#>    MSS  TAU  MAC PMAX time_correction distance_correction
#> 1 8.98 1.25 7.21 16.2           0.300                   0
#> 2 8.97 1.22 7.35 16.5           0.284                   0
#> 3 8.95 1.21 7.40 16.5           0.282                   0
#> 4 8.96 1.21 7.39 16.5           0.282                   0
#> 5 8.93 1.20 7.45 16.6           0.278                   0
#> 
#> Model fit:
#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
#>        NA   1.00000  -0.00639   0.00510   0.00639   0.00401 
#>       MAE      MAPE 
#>   0.00349   0.18387
\end{CodeOutput}
\end{CodeChunk}

The model print output provides training dataset estimates and model performance, as well as LOOCV estimates and model performance.

Let's plot estimated parameters across LOOCV folds:

\begin{CodeChunk}
\begin{CodeInput}
R> df <- jack_LOOCV$LOOCV$parameters
R> 
R> df <- pivot_longer(df, cols = 1:6, names_to = "parameter")
R> 
R> df$parameter <- factor(
+   df$parameter,
+   levels = c(
+     "MSS",
+     "TAU",
+     "MAC",
+     "PMAX",
+     "time_correction",
+     "distance_correction"
+   )
+ )
R> 
R> ggplot(df, aes(x = value)) +
+   theme_bw(8) +
+   geom_boxplot() +
+   facet_wrap(~parameter, scales = "free_x") +
+   xlab(NULL) +
+   ylab(NULL) +
+   theme(
+     axis.ticks.y = element_blank(),
+     axis.text.y = element_blank()
+   )
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-56-1} \end{center}

\end{CodeChunk}

Here is the plot of the training and LOOCV residuals:

\begin{CodeChunk}
\begin{CodeInput}
R> df <- data.frame(
+   distance = jack_LOOCV$data$distance,
+   time = jack_LOOCV$data$time,
+   pred_time = jack_LOOCV$data$pred_time,
+   LOOCV_time = jack_LOOCV$LOOCV$data$pred_time
+ )
R> 
R> df <- df %>%
+   pivot_longer(cols = c("pred_time", "LOOCV_time"))
R> 
R> df$resid <- df$value - df$time
R> 
R> ggplot(df, aes(x = distance, y = resid, color = name)) +
+   theme_bw(8) +
+   geom_hline(yintercept = 0, linetype = "dashed") +
+   geom_point() +
+   theme(legend.title = element_blank()) +
+   xlab("Distance (m)") +
+   ylab("Predicted - observed time (s)")
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}

\end{CodeChunk}

As expected, model has more issues predicting unseen short or long sprints. Please note, that since LOOCV removes one observation, if model estimates three parameters, one needs to have at least five observations, since we need to make sure that model can be estimated once a single observation is removed. LOOCV can also be implemented with the mixed-effects models in the \pkg{shorts} package.

\hypertarget{example-analysis}{%
\section{Example analysis}\label{example-analysis}}

Let's utilize demonstrated functionalities of the \pkg{shorts} package using real-world data. The first dataset comes from Usain Bolt run from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data set for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package.

\hypertarget{usain-bolts-run-from-london-2017}{%
\subsection{Usain Bolt's run from London 2017}\label{usain-bolts-run-from-london-2017}}

The following dataset represents Usain Bolt's race in the finals at the IAAF World Championship held in London, 2017. Since reaction time enters the splits, we want to see how will that affect the model estimates, and particularly, will estimated time correction model pick reaction time up.

For the sake of this analysis, only 10m splits over 60m are used.

\begin{CodeChunk}
\begin{CodeInput}
R> bolt_reaction_time <- 0.183
R> 
R> bolt_distance <- c(10, 20, 30, 40, 50, 60)
R> bolt_time <- c(1.963, 2.983, 3.883, 4.763, 5.643, 6.493)
R> 
R> # No corrections model
R> bolt_m1 <- shorts::model_using_splits(
+   distance = bolt_distance,
+   time = bolt_time
+ )
R> 
R> # Model with reaction time as fixed time correction
R> bolt_m2 <- shorts::model_using_splits(
+   distance = bolt_distance,
+   time = bolt_time,
+   time_correction = -bolt_reaction_time
+ )
R> 
R> # Model with estimated time correction
R> bolt_m3 <- shorts::model_using_splits_with_time_correction(
+   distance = bolt_distance,
+   time = bolt_time
+ )
R> 
R> # Model with estimated time correction, but deducted reaction time
R> bolt_m4 <- shorts::model_using_splits_with_time_correction(
+   distance = bolt_distance,
+   time = bolt_time - bolt_reaction_time
+ )
R> 
R> # Model with estimated time and distance corrections
R> bolt_m5 <- shorts::model_using_splits_with_corrections(
+   distance = bolt_distance,
+   time = bolt_time
+ )
R> 
R> # Model with estimated time and distance corrections and deducted reaction time
R> bolt_m6 <- shorts::model_using_splits_with_corrections(
+   distance = bolt_distance,
+   time = bolt_time - bolt_reaction_time
+ )
R> 
R> bolt_model <- rbind(
+   data.frame(
+     model = "No correction",
+     t(coef(bolt_m1))
+   ),
+   data.frame(
+     model = "No correction - RT",
+     t(coef(bolt_m2))
+   ),
+   data.frame(
+     model = "Time correction",
+     t(coef(bolt_m3))
+   ),
+   data.frame(
+     model = "Time correction - RT",
+     t(coef(bolt_m4))
+   ),
+   data.frame(
+     model = "Distance correction",
+     t(coef(bolt_m5))
+   ),
+   data.frame(
+     model = "Distance correction - RT",
+     t(coef(bolt_m6))
+   )
+ )
R> 
R> bolt_model
\end{CodeInput}
\begin{CodeOutput}
#>                      model  MSS   TAU   MAC PMAX
#> 1            No correction 12.1 1.564  7.77 23.6
#> 2       No correction - RT 11.7 1.205  9.74 28.6
#> 3          Time correction 11.7 1.202  9.76 28.6
#> 4     Time correction - RT 11.7 1.202  9.76 28.6
#> 5      Distance correction 11.6 0.855 13.56 39.3
#> 6 Distance correction - RT 11.6 0.855 13.56 39.3
#>   time_correction distance_correction
#> 1         0.00000                0.00
#> 2        -0.18300                0.00
#> 3        -0.18483                0.00
#> 4        -0.00183                0.00
#> 5        -0.81151               -3.98
#> 6        -0.62851               -3.98
\end{CodeOutput}
\end{CodeChunk}

\hypertarget{vescovi-data}{%
\subsection{Vescovi data}\label{vescovi-data}}

Vescovi's data set represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players). Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset. Soccer players were older (24.6±3.6 vs.~18.9±2.7 yr, p \textless{} 0.0001), however there were no differences for height (167.3±5.9 vs.~167.0±5.7 cm, p = 0.886), body mass (62.5±5.9 vs.~64.0±9.4 kg, p = 0.500) or any sprint interval time (p \textgreater{} 0.650).

The protocol for assessing linear sprint speed has been described previously \citep{vescoviImpactMaximumSpeed2014, vescoviLocomotorHeartRateMetabolic2016, vescoviSprintSpeedCharacteristics2012} and was identical for each cohort. Briefly, all athletes performed a standardized warm-up that included general exercises such as jogging, shuffling, multi-directional movements, and dynamic stretching exercises. Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, 30, and 35 meters at a height of approximately 1.0 meter. Participants stood with their lead foot positioned approximately 5 cm behind the initial infrared beam (i.e., start line). Only forward movement was permitted (no leaning or rocking backwards) and timing started when the laser of the starting gate was triggered. The best 35 m time, and all associated split times were kept for analysis.

Below is the mixed-effects models analysis of the dataset.

\begin{CodeChunk}
\begin{CodeInput}
R> data("vescovi")
R> 
R> # Convert data to long
R> df <- vescovi %>%
+   select(1:13) %>%
+   # slice(1:10) %>%
+   pivot_longer(cols = 9:13, names_to = "distance", values_to = "time") %>%
+   mutate(
+     distance = as.numeric(str_extract(distance, "^[0-9]+"))
+   )
R> 
R> head(df)
\end{CodeInput}
\begin{CodeOutput}
#> # A tibble: 6 x 10
#>   Team  Surface Athlete   Age Height Bodyweight   BMI   BSA
#>   <chr> <chr>   <fct>   <dbl>  <dbl>      <dbl> <dbl> <dbl>
#> 1 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
#> 2 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
#> 3 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
#> 4 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
#> 5 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
#> 6 FH Sr Hard C~ FMSS-2~    21   172.       63.6  21.5   1.7
#> # ... with 2 more variables: distance <dbl>, time <dbl>
\end{CodeOutput}
\end{CodeChunk}

Here we will utilize the following models: no corrections model, fixed time correction model (using 0.3s heuristic rule of thumb), estimated time correction as a fixed effect model, estimated time correction as a random effect model, estimated distance correction as fixed effect model (and time correction as random effect), and estimated distance correction as random effect model.

\begin{CodeChunk}
\begin{CodeInput}
R> no_corrections <- shorts::mixed_model_using_splits(
+   df,
+   distance = "distance",
+   time = "time",
+   athlete = "Athlete"
+ )
R> 
R> fixed_correction <- shorts::mixed_model_using_splits(
+   df,
+   distance = "distance",
+   time = "time",
+   athlete = "Athlete",
+   time_correction = 0.3
+ )
R> 
R> time_correction_fixed <- shorts::mixed_model_using_splits_with_time_correction(
+   df,
+   distance = "distance",
+   time = "time",
+   athlete = "Athlete",
+   random = MSS + TAU ~ 1
+ )
R> 
R> time_correction_random <- shorts::mixed_model_using_splits_with_time_correction(
+   df,
+   distance = "distance",
+   time = "time",
+   athlete = "Athlete",
+   random = MSS + TAU + time_correction ~ 1
+ )
R> 
R> time_distance_correction_fixed <- shorts::mixed_model_using_splits_with_corrections(
+   df,
+   distance = "distance",
+   time = "time",
+   athlete = "Athlete",
+   random = MSS + TAU + time_correction ~ 1
+ )
R> 
R> time_distance_correction_random <- shorts::mixed_model_using_splits_with_corrections(
+   df,
+   distance = "distance",
+   time = "time",
+   athlete = "Athlete",
+   random = MSS + TAU + time_correction + distance_correction ~ 1
+ )
\end{CodeInput}
\end{CodeChunk}

The following image represents model fit estimator RSE for each model. As can be seen, RSE is reduced the more flexible the model.

\begin{CodeChunk}
\begin{CodeInput}
R> model_fit <- rbind(
+   data.frame(
+     model = "No corrections",
+     t(unlist(no_corrections$model_fit))
+   ),
+   data.frame(
+     model = "Fixed correction +0.3s",
+     t(unlist(fixed_correction$model_fit))
+   ),
+   data.frame(
+     model = "Time correction fixed",
+     t(unlist(time_correction_fixed$model_fit))
+   ),
+   data.frame(
+     model = "Time correction random",
+     t(unlist(time_correction_random$model_fit))
+   ),
+   data.frame(
+     model = "Time and distance correction fixed",
+     t(unlist(time_distance_correction_fixed$model_fit))
+   ),
+   data.frame(
+     model = "Time and distance correction random",
+     t(unlist(time_distance_correction_random$model_fit))
+   )
+ )
R> 
R> model_fit$model <- factor(
+   model_fit$model,
+   levels = rev(c(
+     "No corrections",
+     "Fixed correction +0.3s",
+     "Time correction fixed",
+     "Time correction random",
+     "Time and distance correction fixed",
+     "Time and distance correction random"
+   ))
+ )
R> 
R> ggplot(model_fit, aes(x = RSE, y = model)) +
+   theme_bw(8) +
+   geom_point() +
+   xlab("RSE (s)") +
+   ylab(NULL)
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-61-1} \end{center}

\end{CodeChunk}

The following image depicts estimated parameters for each model:

\begin{CodeChunk}
\begin{CodeInput}
R> est_params <- rbind(
+   data.frame(
+     model = "No corrections",
+     no_corrections$parameters$random
+   ),
+   data.frame(
+     model = "Fixed correction +0.3s",
+     fixed_correction$parameters$random
+   ),
+   data.frame(
+     model = "Time correction fixed",
+     time_correction_fixed$parameters$random
+   ),
+   data.frame(
+     model = "Time correction random",
+     time_correction_random$parameters$random
+   ),
+   data.frame(
+     model = "Time and distance correction fixed",
+     time_distance_correction_fixed$parameters$random
+   ),
+   data.frame(
+     model = "Time and distance correction random",
+     time_distance_correction_random$parameters$random
+   )
+ )
R> 
R> est_params$model <- factor(
+   est_params$model,
+   levels = rev(c(
+     "No corrections",
+     "Fixed correction +0.3s",
+     "Time correction fixed",
+     "Time correction random",
+     "Time and distance correction fixed",
+     "Time and distance correction random"
+   ))
+ )
R> 
R> est_params <- est_params %>%
+   pivot_longer(cols = -(1:2), names_to = "parameter")
R> 
R> est_params$parameter <- factor(
+   est_params$parameter,
+   levels = c(
+     "MSS",
+     "TAU",
+     "MAC",
+     "PMAX",
+     "time_correction",
+     "distance_correction"
+   )
+ )
R> 
R> ggplot(est_params, aes(y = model, x = value)) +
+   theme_bw(8) +
+   geom_boxplot() +
+   facet_wrap(~parameter, scales = "free_x") +
+   xlab(NULL) +
+   ylab(NULL)
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-62-1} \end{center}

\end{CodeChunk}

The following image depicts model residuals across distance splits. To provide practical magnitude of the residuals, we have used between subject observed time SD multiplied with 0.2 and -0.2. This provides practical anchor for the residual magnitude, often referred to as \emph{smallest worthwhile change} (SWC) or \emph{smallest effect size of interest} (SESOI) \citep{jovanovicBmbstatsBootstrapMagnitudebased2020}. If the residuals are within this magnitude band, they have no practical significance, and thus the model is good in making practically useful predictions.

Error bars represent residual bias and \(\pm\) 1SD.

\begin{CodeChunk}
\begin{CodeInput}
R> model_resid <- rbind(
+   data.frame(
+     model = "No corrections",
+     no_corrections$data
+   ),
+   data.frame(
+     model = "Fixed correction +0.3s",
+     fixed_correction$data
+   ),
+   data.frame(
+     model = "Time correction fixed",
+     time_correction_fixed$data
+   ),
+   data.frame(
+     model = "Time correction random",
+     time_correction_random$data
+   ),
+   data.frame(
+     model = "Time and distance correction fixed",
+     time_distance_correction_fixed$data
+   ),
+   data.frame(
+     model = "Time and distance correction random",
+     time_distance_correction_random$data
+   )
+ )
R> 
R> 
R> model_resid$model <- factor(
+   model_resid$model,
+   levels = rev(c(
+     "No corrections",
+     "Fixed correction +0.3s",
+     "Time correction fixed",
+     "Time correction random",
+     "Time and distance correction fixed",
+     "Time and distance correction random"
+   ))
+ )
R> 
R> model_resid$resid <- model_resid$pred_time - model_resid$time
R> 
R> # Create SWC / SESOI band
R> model_SESOI <- model_resid %>%
+   group_by(model, distance) %>%
+   summarise(
+     bias = mean(resid),
+     variance = sd(resid),
+     upper = bias + variance,
+     lower = bias - variance,
+     MAD = mean(abs(resid)),
+     SESOI_upper = sd(time) * 0.2,
+     SESOI_lower = -sd(time) * 0.2
+   )
R> 
R> # Plot
R> ggplot(model_resid, aes(y = model)) +
+   theme_bw(8) +
+   geom_vline(
+     data = model_SESOI,
+     aes(xintercept = SESOI_lower), color = "blue", alpha = 0.5, linetype = "dashed"
+   ) +
+   geom_vline(
+     data = model_SESOI,
+     aes(xintercept = SESOI_upper), color = "blue", alpha = 0.5, linetype = "dashed"
+   ) +
+   geom_vline(xintercept = 0, color = "blue", alpha = 0.5) +
+   geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
+   geom_errorbarh(data = model_SESOI, aes(xmin = lower, xmax = upper), height = 0.1, color = "black") +
+   geom_point(data = model_SESOI, aes(x = bias), color = "black") +
+   facet_wrap(~distance, scales = "free_x") +
+   xlab("Predicted time - observed time (s)") +
+   ylab(NULL)
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-63-1} \end{center}

\end{CodeChunk}

The following figure depicts model residuals estimators (bias, or mean residual; variance, or SD of the residuals, and MAD, or mean absolute difference).

\begin{CodeChunk}
\begin{CodeInput}
R> df <- model_SESOI %>%
+   pivot_longer(cols = -(1:2), names_to = "estimator") %>%
+   filter(estimator %in% c("bias", "variance", "MAD"))
R> 
R> df$model <- factor(
+   df$model,
+   levels = rev(c(
+     "No corrections",
+     "Fixed correction +0.3s",
+     "Time correction fixed",
+     "Time correction random",
+     "Time and distance correction fixed",
+     "Time and distance correction random"
+   ))
+ )
R> 
R> df$estimator <- factor(
+   df$estimator,
+   levels = c("bias", "variance", "MAD")
+ )
R> 
R> ggplot(df, aes(x = value, y = model)) +
+   theme_bw(8) +
+   geom_point() +
+   facet_grid(distance ~ estimator, scales = "free_x") +
+   xlab(NULL) +
+   ylab(NULL)
\end{CodeInput}

\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-64-1} \end{center}

\end{CodeChunk}

Which model should should be used? Although providing a better fit (using RSE as an estimator of model fit), time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction). Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for timing inconsistencies.

We are hoping that the \pkg{shorts} package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.

\renewcommand\refname{References}
\bibliography{references.bib,packages.bib}

\end{document}

### 2020-10-21 paper.Rmd
Vescovi's comments and changes

@@
 abstract: >
-  The abstract of the article.
+  Short sprint performance is one of the most distinguishable and admired physical trait in sports.
+Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
+Short sprint performance data comes from two sources: (1) split times using timing gates and (2) radar gun.
+In this paper, we demonstrate how these data can be modeled using the \pkg{shorts} package and how timing issues (i.e., synchronization issues with the radar gun, or using the flying start with the timing gates), which are common in practical measurement settings, can introduce error in parameter estimates.
+Further, we demonstrated how different model definitions implemented in \pkg{shorts} package can alleviate these errors.
 keywords:
@@
 
-options("width" = 60)
+options("width" = 70)
 
@@
   message = FALSE,
-  width = 60
+  width = 70
 )
@@
 For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)[@brownAssessmentLinearSprinting2004].
-
-Describing individual differences in short sprint performance, quantifying training intervention effects, but more importantly, understanding short sprint *bottle necks* (i.e.
-limiting factors) has been one of the prominent tasks of sports science profession since its establishment.
-The \pkg{shorts} package [@R-shorts], written in the R language [@R-base], represents an open-source tool that aims to help sports scientist in achieving these goals through mathematical modeling.
-
-In the current paper, we will provide explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
+Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
+
+The use of force plates is considered the gold standard for assessing mechanical properties of sprinting; however, there are logistical and financial challenges to capturing the profile of an entire sprint [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+Radar and laser technology are frequently used laboratory-grade methods [@buchheitMechanicalDeterminantsAcceleration2014; @edwardsSprintAccelerationCharacteristics2020; @jimenez-reyesRelationshipVerticalHorizontal2018; @marcote-pequenoAssociationForceVelocity2019] but not normally accessible to practitioners working in sports.
+Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
+The \pkg{shorts} package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+
+In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
 
@@
 
-The Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6s).
+Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6s).
 
@@
 
-If we plot acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
+Plotting acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
 
@@
 
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), and (2) getting a velocity trace using a radar gun.
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), (2) getting a velocity trace using a radar gun.
 Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in \pkg{shorts} package using non-linear least squares regression implemented in the `nls` function in the \pkg{base} R [@R-base] and `nlme` function in the \pkg{nlme} package [@R-nlme] for the mixed-effect models.
@@
 Let's consider an example of an athlete with MSS equal to 9 $ms^-1$, TAU equal to 1.3, and MAC equal to 6.92 $ms^-2$ performing 40m sprint with timing gates positioned at each 10m split.
-When it comes to split times, distance is a predictor, and time is outcome variable, thus the equation \@ref(eq:velocity-time) takes another form:
+For split times, distance is a predictor, and time is the outcome variable, thus the equation \@ref(eq:velocity-time) becomes:
 
@@
 
-Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as an target or outcome variable.
-Thus equation \@ref(eq:velocity-time) is used to estimate MSS and TAU.
+Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as the outcome variable. Thus equation \@ref(eq:velocity-time) is used to estimate MSS and TAU.
 
@@
 There is a challenge when collecting sprint data that could have a substantial impact on modeled outcomes.
-To ensure accurate parameter outcomes, the initial force production must be synced with start time.
+To ensure accurate parameter outcomes, the initial force production must be synced with start time [@haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintMechanicalVariables2019].
 Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \pkg{shorts} package.
@@
 
-One source of error in the modeled estimation using radar gun is the time synchronization.
-In theory, synchronization is ideal when sprint is initiated at $t=0$ (i.e., $v(t=0) = 0$).
+One source of error in the modeled estimation using a radar gun is the time synchronization.
+In theory, synchronization is ideal when a sprint is initiated at $t=0$ (i.e., $v(t=0) = 0$).
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5second to simulate an error in synchronization and its effect on estimated MSS and TAU.
+Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
 
@@
 Let's call them John and Jack.
-The are both performing 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
+They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
 The initial timing gate at the start (i.e., $d=0m$) serves to activate the timing system (i.e., when they cross the beam).
@@
 In other words, Jack is using a *flying start*, a common scenario when testing field sports athletes.
-Let's see how their sprints differ.
+Let's see how their sprint outcomes differ.
 
@@
 
-Using the following code, we can see how different are the estimated MSS and TAU parameters:
+Using the following code, we can see the differences in estimated MSS and TAU parameters:
 
@@
 
-As can be seen from the results, flying starts yields biased estimates, particularly for the TAU, MAC and PMAX.
+As can be seen from the results, a flying start yields biased estimates, particularly for the TAU, MAC and PMAX.
 
@@
 
-MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as cheat distance increases.
+MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
 
@@
 
-Estimated PMAX is also overestimated as cheat distance increases.
+Estimated PMAX is also overestimated as flying start distance increases.
 
@@
 
-When implementing *time correction*, equation \@ref(eq:time-distance) gets another form:
+When implementing *time correction*, equation \@ref(eq:time-distance) becomes:
 
@@
 
-Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides estimation of this parameter from the data provide, together with MSS and TAU.
+Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
 To estimate time correction, we use `shorts::model_using_splits_with_time_correction` function.
@@
 
-Another model definition, which is novel approach implemented in the \pkg{shorts} packages, is to ulitize *distance correction*, besides time correction.
-Thus, equation \@ref(eq:time-distance) gets another form:
+Another model definition, which is a novel approach implemented in the \pkg{shorts} packages, is to utilize *distance correction*, besides time correction.
+Thus, equation \@ref(eq:time-distance) becomes:
 
@@
 
-This model is implemented in `shorts::model_using_splits_with_corrections` function. Let's check this model estimates:
+This model is implemented in `shorts::model_using_splits_with_corrections` function. Below are the model estimates:
 
@@
 As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters.
-There are few issues with this model definition.
-Besides being novel and still not validated with practical data, distance correction model has four parameters to estimate, which implies that one needs at least five sprint splits.
+There are a few issues with this model definition.
+Besides being novel and still not validated with actual data, distance correction model has four parameters to estimate, which implies that at least five sprint splits are needed.
 This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.
 
-We will get back to these issues later, but let's see how these models perform using simulated data with varying flying start distance.
+We will get back to these issues later, but we can examine how these models perform using simulated data with varying flying start distance.
 The following code contains the wrapper that performs all four models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction):
@@
 
-As can be seen from the next figure, estimated time correction model estimates MSS almost perfectly, while estimated time and distance correction model estimates MSS perfectly.
+As can be seen from the next figure, the estimated time correction model estimates MSS almost perfectly, while the estimated time and distance correction model estimates MSS perfectly.
 
@@
 
-Same conclusions can be said for the MAC parameter.
-Time and distance corrections model performs perfectly, while time correction model performs almost as good (although worse than for the MSS parameter).
+Similar outcomes are observed for the MAC parameter. The time and distance corrections model performs perfectly, while the time correction model performs almost as good.
 
@@
 
-When it comes to PMAX, same conclusion can be reached.
+PMAX demonstrates the same properties as MSS and MAC.
 
@@
 
-Given the simulation data, time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model in the face of flying start issues.
-Since time correction model is simpler and demand three parameters to be estimated, it might be practically more useful than time and distance correction model, which demand four parameters estimation and thus more than five timing gates and sprint splits.
+The outcomes from the simulation data clearly demonstrates that the time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model when attempting to overcome the flying start issues.
+Since the time correction model is simpler and requires three parameters to be estimated, it might be practically more useful than the time and distance correction model, which requires four parameters estimation and thus more than five timing gates and sprint splits.
 
@@
 flying start).
-In this section, we simulate a few other issues to check how sensitive the presented models are to perturbations that can happen in practical settings during field testing.
-
-#### Triggering the gate before the start
-
-One issue that might happen in certain timing gates setups is an individual triggering timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
+In this section, we simulate one more issue to check the sensitivity of the presented models to other (less common) perturbations when performing field testing.
+
+One issue that might happen with timing gates is triggering the timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
 This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is *reaction time* (RT) involved.
 Both of these scenarios represent *time lag* that is added to the split times.
-Let's simulate the effect of this time lag on model estimates and predictions.
+Below we simulate the effect of this time lag on model estimates and predictions.
 
@@
 From the figure below it can be seen that time lag affects estimated MSS for the the model without correction and fixed correction model.
-Time correction and time and distance corrections model correctly estimated MSS.
+Time correction and time and distance corrections models correctly estimated MSS.
 
@@
 
-Two figure below depicts correctly identified time lag (i.e. using time correction parameter) using time correction and time and distance corrections models.
+The figure below depicts correctly identified time lag (i.e. using time correction parameter) using time correction and time and distance corrections models.
 
@@
 
-Additional scenario can create time lag: if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
+There are few other starting issues worth mentioning.
+For example, if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
 In this case, time lag is a negative number since it reduces the split times.
-This scenario will not be simulated in the current paper.
-
-#### Bad position of the timing gates
-
 Another common issue with timing gates in the practical field settings is the bad measurement of the distance and thus bad positions of the timing gates.
-In this example, we will simulate the scenario where the initial timing gate is not positioned at $d=0$, but rather with some error (i.e., -0.5 to 0.5m).
-Although this scenario represents poor procedure implementation, we have included this simulation to depict effects on estimated parameters.
-
-```{r}
-sim_df <- expand.grid(
-  MSS = c(6, 7, 8, 9),
-  MAC = c(6, 7, 8, 9),
-  displacement = seq(-0.5, 0.5, length.out = 51),
-  distance = c(5, 10, 20, 30, 40, 50)
-)
-
-sim_df <- sim_df %>%
-  mutate(
-    TAU = MSS / MAC,
-    PMAX = MSS * MAC / 4,
-    true_distance = distance + displacement,
-    time = shorts::predict_time_at_distance(true_distance, MSS, TAU)
-  )
-
-# Add small noise to allow model fit
-set.seed(1667)
-rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
-sim_df$time <- sim_df$time + rand_noise
-
-head(sim_df, 2)
-```
-
-```{r}
-# estimated parameters and predicted time
-model_df <- sim_df %>%
-  group_by(MSS, TAU, displacement) %>%
-  do(pred_wrapper(.)) %>%
-  ungroup()
-```
-
-Figure below depicts effect of timing gate displacement on the MSS parameter. Time correction and time and distance correction models are almost identical in their correct estimation of the MSS.
-
-```{r}
-model_df$model <- factor(
-  model_df$model,
-  levels = c(
-    "No correction",
-    "Fixed correction +0.3s",
-    "Fixed correction +0.5s",
-    "Time correction",
-    "Time and distance correction"
-  )
-)
-
-# Estimates plot
-df <- model_df %>%
-  group_by(MSS, TAU, displacement, model) %>%
-  slice(1) %>%
-  mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2))
-  )
-
-# MSS
-ggplot(df, aes(x = displacement, y = est_MSS, color = model)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-  xlab("Displacement (m)") +
-  ylab("estimated MSS (m/s)") +
-  theme(legend.title = element_blank())
-```
-
-Figure below depicts effect of timing gate displacement on the MSS parameter. Time and distance correction model is the only model that estimated MAC correctly.
-
-```{r}
-# MAC
-ggplot(df, aes(x = displacement, y = est_MAC, color = model)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Displacement (m)") +
-  ylab("estimated MAC (m/s/s)") +
-  theme(legend.title = element_blank())
-```
-
-Figures below depicts estimated time and distance corrections parameters.
-
-```{r}
-# time_correction
-ggplot(df, aes(x = displacement, y = est_time_correction, color = model)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Displacement (m)") +
-  ylab("estimated time correction (s)") +
-  theme(legend.title = element_blank())
-```
-
-```{r}
-# distance_correction
-ggplot(df, aes(x = displacement, y = est_distance_correction, color = model)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  geom_abline(slope = 1, color = "black", linetype = "dashed") +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Displacement (m)") +
-  ylab("estimated distance correction (m)") +
-  theme(legend.title = element_blank())
-```
-
-The following figure depicts residuals (i.e., predicted time minus observed time).
-
-```{r}
-# Residuals
-model_df <- model_df %>%
-  mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-    group = paste(MSS, MAC, displacement)
-  )
-
-ggplot(model_df, aes(y = residuals, x = distance, color = displacement, group = group)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.3) +
-  facet_wrap(~model) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-  xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)")
-```
-
-Additional sources of issues with the timing gates involve combination of the previous scenarios, as well as effect of number and distances of the timing gates [@haugenPowerForceVelocityProfilingSprinting2020; also see @haugenDifferenceStartImpact2012].
+
+The number and distances of the timing gates can also affect the precision of the estimated sprint parameters [@haugenPowerForceVelocityProfilingSprinting2020; @haugenDifferenceStartImpact2012].
+
+In field testing, multiple starting issues can be present.
 For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence.
@@
 
-Let's perform LOOCV using Jack's data and time correction model:
+Let's perform LOOCV using Jack's data and the time correction model:
 
@@
 
-Let's plot estimated parameters across LOOCV folds:
+Next we plot estimated parameters across LOOCV folds:
 
@@
 
-As expected, model has more issues predicting unseen short or long sprints.
-Please note, that since LOOCV removes one observation, if model estimates three parameters, one needs to have at least five observations, since we need to make sure that model can be estimated once a single observation is removed.
+As expected, the model has more issues predicting unseen short or long sprints.
+Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
 LOOCV can also be implemented with the mixed-effects models in the \pkg{shorts} package.
@@
 Let's utilize demonstrated functionalities of the \pkg{shorts} package using real-world data.
-The first dataset comes from Usain Bolt run from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data set for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package.
+The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package (see `?vescovi`).
 
@@
 The following dataset represents Usain Bolt's race in the finals at the IAAF World Championship held in London, 2017.
-Since reaction time enters the splits, we want to see how will that affect the model estimates, and particularly, will estimated time correction model pick reaction time up.
+Since reaction time enters the splits, we want to see how that will affect the model estimates, and particularly, if the estimated time correction model will pick-up reaction time.
 
@@
 
-Vescovi's data set represents a  sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players).
+The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players).
 Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset.
-Soccer players were older (24.6±3.6 vs.
-18.9±2.7 yr, p < 0.0001), however there were no differences for height (167.3±5.9 vs.
-167.0±5.7 cm, p = 0.886), body mass (62.5±5.9 vs.
-64.0±9.4 kg, p = 0.500) or any sprint interval time (p > 0.650).
 
@@
 
-Here we will utilize the following models: no corrections model, fixed time correction model (using 0.3s heuristic rule of thumb), estimated time correction as a fixed effect model, estimated time correction as a random effect model, estimated distance correction as fixed effect model (and time correction as random effect), and estimated distance correction as random effect model.
+The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model.
 
@@
 This provides practical anchor for the residual magnitude, often referred to as *smallest worthwhile change* (SWC) or *smallest effect size of interest* (SESOI) [@jovanovicBmbstatsBootstrapMagnitudebased2020].
-If the residuals are within this magnitude band, they have no practical significance, and thus the model is good in making practically useful predictions.
-
-Error bars represent residual bias and $\pm$ 1SD.
+If the residuals are within this magnitude band, then the model is good in making practically useful predictions.
+
+Error bars represent residual bias $\pm$ 1SD.
 
@@
 Which model should should be used?
-Although providing a better fit (using RSE as an estimator of model fit), time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction).
+Although providing a better fit (using RSE as an estimator of model fit), the time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction).
 Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for timing inconsistencies.

### 2020-10-21 paper.tex
Vescovi's comments and changes

@@
 \Abstract{
-The abstract of the article.
+Short sprint performance is one of the most distinguishable and admired physical trait in sports.
+Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
+Short sprint performance data comes from two sources: (1) split times using timing gates and (2) radar gun.
+In this paper, we demonstrate how these data can be modeled using the \pkg{shorts} package and how timing issues (i.e., synchronization issues with the radar gun, or using the flying start with the timing gates), which are common in practical measurement settings, can introduce error in parameter estimates.
+Further, we demonstrated how different model definitions implemented in \pkg{shorts} package can alleviate these errors.
 }
@@
 For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)\citep{brownAssessmentLinearSprinting2004}.
-
-Describing individual differences in short sprint performance, quantifying training intervention effects, but more importantly, understanding short sprint \emph{bottle necks} (i.e.~limiting factors) has been one of the prominent tasks of sports science profession since its establishment.
-The \pkg{shorts} package \citep{R-shorts}, written in the R language \citep{R-base}, represents an open-source tool that aims to help sports scientist in achieving these goals through mathematical modeling.
-
-In the current paper, we will provide explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
+Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
+
+The use of force plates is considered the gold standard for assessing mechanical properties of sprinting; however, there are logistical and financial challenges to capturing the profile of an entire sprint \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+Radar and laser technology are frequently used laboratory-grade methods \citep{buchheitMechanicalDeterminantsAcceleration2014, edwardsSprintAccelerationCharacteristics2020, jimenez-reyesRelationshipVerticalHorizontal2018, marcote-pequenoAssociationForceVelocity2019} but not normally accessible to practitioners working in sports.
+Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
+The \pkg{shorts} package \citep{R-shorts}, written in the R language \citep{R-base}, represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+
+In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
 
@@
 
-The Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6s).
+Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6s).
 
@@
 
-If we plot acceleration against velocity (Figure \ref{fig:four-athletes-profile}), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
+Plotting acceleration against velocity (Figure \ref{fig:four-athletes-profile}), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
 
@@
 
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), and (2) getting a velocity trace using a radar gun.
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), (2) getting a velocity trace using a radar gun.
 Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \pkg{shorts} package using non-linear least squares regression implemented in the \texttt{nls} function in the \pkg{base} R \citep{R-base} and \texttt{nlme} function in the \pkg{nlme} package \citep{R-nlme} for the mixed-effect models.
@@
 Let's consider an example of an athlete with MSS equal to 9 \(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^-2\) performing 40m sprint with timing gates positioned at each 10m split.
-When it comes to split times, distance is a predictor, and time is outcome variable, thus the equation \eqref{eq:velocity-time} takes another form:
+For split times, distance is a predictor, and time is the outcome variable, thus the equation \eqref{eq:velocity-time} becomes:
 
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>   0.00249   1.00000  -0.00178   0.00265   0.00265   0.00176 
-#>       MAE      MAPE 
-#>   0.00157   0.04742
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>   0.00249   1.00000  -0.00178   0.00265   0.00265   0.00176   0.00157 
+#>      MAPE 
+#>   0.04742
 \end{CodeOutput}
@@
 #> ---
-#> Signif. codes:  
-#> 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+#> Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
 #> 
@@
 \begin{CodeOutput}
-#>   time distance velocity acceleration air_resistance force
-#> 1 0.00 0.000000   0.0000         6.89        0.07536   551
-#> 2 0.01 0.000344   0.0686         6.84        0.05609   547
-#> 3 0.02 0.001371   0.1367         6.78        0.03978   543
-#> 4 0.03 0.003076   0.2043         6.73        0.02636   539
-#> 5 0.04 0.005455   0.2714         6.68        0.01576   534
-#> 6 0.05 0.008502   0.3379         6.63        0.00792   530
-#>   power relative_power
-#> 1   0.0          0.000
-#> 2  37.5          0.469
-#> 3  74.2          0.928
-#> 4 110.0          1.375
-#> 5 145.0          1.813
-#> 6 179.2          2.241
+#>   time distance velocity acceleration air_resistance force power
+#> 1 0.00 0.000000   0.0000         6.89        0.07536   551   0.0
+#> 2 0.01 0.000344   0.0686         6.84        0.05609   547  37.5
+#> 3 0.02 0.001371   0.1367         6.78        0.03978   543  74.2
+#> 4 0.03 0.003076   0.2043         6.73        0.02636   539 110.0
+#> 5 0.04 0.005455   0.2714         6.68        0.01576   534 145.0
+#> 6 0.05 0.008502   0.3379         6.63        0.00792   530 179.2
+#>   relative_power
+#> 1          0.000
+#> 2          0.469
+#> 3          0.928
+#> 4          1.375
+#> 5          1.813
+#> 6          2.241
 \end{CodeOutput}
@@
 #> ----------------------------------
-#>     athlete  MSS   TAU  MAC PMAX time_correction
-#> 1     James 9.69 0.847 11.4 27.7               0
-#> 2       Jim 7.83 0.505 15.5 30.4               0
-#> 3      John 7.78 0.727 10.7 20.8               0
-#> 4 Kimberley 8.57 0.802 10.7 22.9               0
-#> 5  Samantha 6.45 0.395 16.3 26.4               0
-#>   distance_correction
-#> 1                   0
-#> 2                   0
-#> 3                   0
-#> 4                   0
-#> 5                   0
+#>     athlete  MSS   TAU  MAC PMAX time_correction distance_correction
+#> 1     James 9.69 0.847 11.4 27.7               0                   0
+#> 2       Jim 7.83 0.505 15.5 30.4               0                   0
+#> 3      John 7.78 0.727 10.7 20.8               0                   0
+#> 4 Kimberley 8.57 0.802 10.7 22.9               0                   0
+#> 5  Samantha 6.45 0.395 16.3 26.4               0                   0
 #> 
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>    0.0260    0.9998   -0.0293    0.0496    0.0496    0.0214 
-#>       MAE      MAPE 
-#>    0.0172    0.9019
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>    0.0260    0.9998   -0.0293    0.0496    0.0496    0.0214    0.0172 
+#>      MAPE 
+#>    0.9019
 \end{CodeOutput}
@@
 
-Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as an target or outcome variable.
-Thus equation \eqref{eq:velocity-time} is used to estimate MSS and TAU.
+Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as the outcome variable. Thus equation \eqref{eq:velocity-time} is used to estimate MSS and TAU.
 
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>   0.00327   1.00000  -0.00406   0.00532   0.00532   0.00276 
-#>       MAE      MAPE 
-#>   0.00207       NaN
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>   0.00327   1.00000  -0.00406   0.00532   0.00532   0.00276   0.00207 
+#>      MAPE 
+#>       NaN
 \end{CodeOutput}
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>   0.00108   1.00000  -0.00406   0.00534   0.00534   0.00276 
-#>       MAE      MAPE 
-#>   0.00206       NaN
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>   0.00108   1.00000  -0.00406   0.00534   0.00534   0.00276   0.00206 
+#>      MAPE 
+#>       NaN
 \end{CodeOutput}
@@
 #> ----------------------------------
-#>     athlete   MSS   TAU  MAC PMAX time_correction
-#> 1     James 10.00 1.111 9.00 22.5               0
-#> 2       Jim  8.00 0.889 9.00 18.0               0
-#> 3      John  8.00 1.069 7.48 15.0               0
-#> 4 Kimberley  9.01 1.286 7.01 15.8               0
-#> 5  Samantha  6.50 0.685 9.50 15.4               0
-#>   distance_correction
-#> 1                   0
-#> 2                   0
-#> 3                   0
-#> 4                   0
-#> 5                   0
+#>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
+#> 1     James 10.00 1.111 9.00 22.5               0                   0
+#> 2       Jim  8.00 0.889 9.00 18.0               0                   0
+#> 3      John  8.00 1.069 7.48 15.0               0                   0
+#> 4 Kimberley  9.01 1.286 7.01 15.8               0                   0
+#> 5  Samantha  6.50 0.685 9.50 15.4               0                   0
 #> 
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>    0.0516    0.9994   -0.2191    0.1983    0.2191    0.0516 
-#>       MAE      MAPE 
-#>    0.0395       NaN
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>    0.0516    0.9994   -0.2191    0.1983    0.2191    0.0516    0.0395 
+#>      MAPE 
+#>       NaN
 \end{CodeOutput}
@@
 There is a challenge when collecting sprint data that could have a substantial impact on modeled outcomes.
-To ensure accurate parameter outcomes, the initial force production must be synced with start time.
+To ensure accurate parameter outcomes, the initial force production must be synced with start time \citep{haugenPowerForceVelocityProfilingSprinting2020, haugenSprintMechanicalVariables2019}.
 Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \pkg{shorts} package.
@@
 
-One source of error in the modeled estimation using radar gun is the time synchronization.
-In theory, synchronization is ideal when sprint is initiated at \(t=0\) (i.e., \(v(t=0) = 0\)).
+One source of error in the modeled estimation using a radar gun is the time synchronization.
+In theory, synchronization is ideal when a sprint is initiated at \(t=0\) (i.e., \(v(t=0) = 0\)).
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5second to simulate an error in synchronization and its effect on estimated MSS and TAU.
+Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
 
@@
 \begin{CodeOutput}
-#> [1] 7.82e-05 4.83e+00 7.07e+00 8.11e+00 8.59e+00 8.81e+00
-#> [7] 8.91e+00
+#> [1] 7.82e-05 4.83e+00 7.07e+00 8.11e+00 8.59e+00 8.81e+00 8.91e+00
 \end{CodeOutput}
@@
 #> ----------------------------------
-#>     athlete   MSS   TAU  MAC PMAX time_correction
-#> 1     James 10.00 1.111 9.00 22.5            -0.5
-#> 2       Jim  8.00 0.889 9.00 18.0            -0.5
-#> 3      John  8.00 1.069 7.48 15.0            -0.5
-#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5
-#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5
-#>   distance_correction
-#> 1                   0
-#> 2                   0
-#> 3                   0
-#> 4                   0
-#> 5                   0
+#>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
+#> 1     James 10.00 1.111 9.00 22.5            -0.5                   0
+#> 2       Jim  8.00 0.889 9.00 18.0            -0.5                   0
+#> 3      John  8.00 1.069 7.48 15.0            -0.5                   0
+#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5                   0
+#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5                   0
 #> 
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>    0.0516    0.9994   -0.2190    0.1983    0.2190    0.0516 
-#>       MAE      MAPE 
-#>    0.0395       Inf
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>    0.0516    0.9994   -0.2190    0.1983    0.2190    0.0516    0.0395 
+#>      MAPE 
+#>       Inf
 \end{CodeOutput}
@@
 #> ----------------------------------
-#>     athlete   MSS   TAU  MAC PMAX time_correction
-#> 1     James 10.00 1.110 9.00 22.5            -0.5
-#> 2       Jim  8.00 0.889 9.00 18.0            -0.5
-#> 3      John  8.00 1.069 7.48 15.0            -0.5
-#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5
-#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5
-#>   distance_correction
-#> 1                   0
-#> 2                   0
-#> 3                   0
-#> 4                   0
-#> 5                   0
+#>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
+#> 1     James 10.00 1.110 9.00 22.5            -0.5                   0
+#> 2       Jim  8.00 0.889 9.00 18.0            -0.5                   0
+#> 3      John  8.00 1.069 7.48 15.0            -0.5                   0
+#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5                   0
+#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5                   0
 #> 
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>    0.0516    0.9994   -0.2188    0.1982    0.2188    0.0516 
-#>       MAE      MAPE 
-#>    0.0395       Inf
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>    0.0516    0.9994   -0.2188    0.1982    0.2188    0.0516    0.0395 
+#>      MAPE 
+#>       Inf
 \end{CodeOutput}
@@
 Let's call them John and Jack.
-The are both performing 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
+They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
 The initial timing gate at the start (i.e., \(d=0m\)) serves to activate the timing system (i.e., when they cross the beam).
@@
 In other words, Jack is using a \emph{flying start}, a common scenario when testing field sports athletes.
-Let's see how their sprints differ.
+Let's see how their sprint outcomes differ.
 
@@
 #> # A tibble: 5 x 6
-#>   distance john_time jack_distance jack_true_time time_05m
-#>      <dbl>  <I<dbl>>         <dbl>       <I<dbl>> <I<dbl>>
-#> 1        5      1.42           5.5           1.50    0.400
-#> 2       10      2.17          10.5           2.23    0.400
-#> 3       20      3.43          20.5           3.49    0.400
-#> 4       30      4.60          30.5           4.65    0.400
-#> 5       40      5.73          40.5           5.78    0.400
-#> # ... with 1 more variable: jack_time <I<dbl>>
+#>   distance john_time jack_distance jack_true_time time_05m jack_time
+#>      <dbl>  <I<dbl>>         <dbl>       <I<dbl>> <I<dbl>>  <I<dbl>>
+#> 1        5      1.42           5.5           1.50    0.400      1.10
+#> 2       10      2.17          10.5           2.23    0.400      1.83
+#> 3       20      3.43          20.5           3.49    0.400      3.09
+#> 4       30      4.60          30.5           4.65    0.400      4.25
+#> 5       40      5.73          40.5           5.78    0.400      5.39
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-#>       MSS   TAU   MAC PMAX time_correction
-#> John 9.00 1.300  6.92 15.6               0
-#> Jack 8.49 0.704 12.06 25.6               0
-#>      distance_correction
-#> John                   0
-#> Jack                   0
-\end{CodeOutput}
-\end{CodeChunk}
-
-As can be seen from the results, flying starts yields biased estimates, particularly for the TAU, MAC and PMAX.
+#>       MSS   TAU   MAC PMAX time_correction distance_correction
+#> John 9.00 1.300  6.92 15.6               0                   0
+#> Jack 8.49 0.704 12.06 25.6               0                   0
+\end{CodeOutput}
+\end{CodeChunk}
+
+As can be seen from the results, a flying start yields biased estimates, particularly for the TAU, MAC and PMAX.
 
@@
 \begin{CodeOutput}
-#>   MSS MAC flying_start_distance distance   TAU PMAX
-#> 1   6   6                     0        5 1.000  9.0
-#> 2   7   6                     0        5 1.167 10.5
-#> 3   8   6                     0        5 1.333 12.0
-#> 4   9   6                     0        5 1.500 13.5
-#> 5   6   7                     0        5 0.857 10.5
-#> 6   7   7                     0        5 1.000 12.2
-#>   true_distance    true_time stolen_time         time
-#> 1             5 1.639197....           0 1.639356....
-#> 2             5 1.579728....           0 1.579675....
-#> 3             5 1.537453....           0 1.537537....
-#> 4             5 1.505904....           0 1.505900....
-#> 5             5 1.549960....           0 1.549860....
-#> 6             5 1.488596....           0 1.488548....
+#>   MSS MAC flying_start_distance distance   TAU PMAX true_distance
+#> 1   6   6                     0        5 1.000  9.0             5
+#> 2   7   6                     0        5 1.167 10.5             5
+#> 3   8   6                     0        5 1.333 12.0             5
+#> 4   9   6                     0        5 1.500 13.5             5
+#> 5   6   7                     0        5 0.857 10.5             5
+#> 6   7   7                     0        5 1.000 12.2             5
+#>      true_time stolen_time         time
+#> 1 1.639197....           0 1.639356....
+#> 2 1.579728....           0 1.579675....
+#> 3 1.537453....           0 1.537537....
+#> 4 1.505904....           0 1.505900....
+#> 5 1.549960....           0 1.549860....
+#> 6 1.488596....           0 1.488548....
 \end{CodeOutput}
@@
 
-When implementing \emph{time correction}, equation \eqref{eq:time-distance} gets another form:
+When implementing \emph{time correction}, equation \eqref{eq:time-distance} becomes:
 
@@
 
-Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides estimation of this parameter from the data provide, together with MSS and TAU.
+Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
 To estimate time correction, we use \texttt{shorts::model\_using\_splits\_with\_time\_correction} function.
@@
 
-Another model definition, which is novel approach implemented in the \pkg{shorts} packages, is to ulitize \emph{distance correction}, besides time correction.
-Thus, equation \eqref{eq:time-distance} gets another form:
+Another model definition, which is a novel approach implemented in the \pkg{shorts} packages, is to utilize \emph{distance correction}, besides time correction.
+Thus, equation \eqref{eq:time-distance} becomes:
 
@@
 
-This model is implemented in \texttt{shorts::model\_using\_splits\_with\_corrections} function. Let's check this model estimates:
+This model is implemented in \texttt{shorts::model\_using\_splits\_with\_corrections} function. Below are the model estimates:
 
@@
 Starting behind the initial timing gate represent only one issue (i.e.~flying start).
-In this section, we simulate a few other issues to check how sensitive the presented models are to perturbations that can happen in practical settings during field testing.
-
-\hypertarget{triggering-the-gate-before-the-start}{%
-\paragraph{Triggering the gate before the start}\label{triggering-the-gate-before-the-start}}
-
-One issue that might happen in certain timing gates setups is an individual triggering timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
+In this section, we simulate one more issue to check the sensitivity of the presented models to other (less common) perturbations when performing field testing.
+
+One issue that might happen with timing gates is triggering the timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
 This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is \emph{reaction time} (RT) involved.
 Both of these scenarios represent \emph{time lag} that is added to the split times.
-Let's simulate the effect of this time lag on model estimates and predictions.
+Below we simulate the effect of this time lag on model estimates and predictions.
 
@@
 \begin{CodeOutput}
-#>   MSS MAC time_lag distance   TAU PMAX    true_time
-#> 1   6   6        0        5 1.000  9.0 1.639197....
-#> 2   7   6        0        5 1.167 10.5 1.579728....
-#> 3   8   6        0        5 1.333 12.0 1.537453....
-#> 4   9   6        0        5 1.500 13.5 1.505904....
-#> 5   6   7        0        5 0.857 10.5 1.549960....
-#> 6   7   7        0        5 1.000 12.2 1.488596....
-#>           time
-#> 1 1.639356....
-#> 2 1.579675....
-#> 3 1.537537....
-#> 4 1.505900....
-#> 5 1.549860....
-#> 6 1.488548....
+#>   MSS MAC time_lag distance   TAU PMAX    true_time         time
+#> 1   6   6        0        5 1.000  9.0 1.639197.... 1.639356....
+#> 2   7   6        0        5 1.167 10.5 1.579728.... 1.579675....
+#> 3   8   6        0        5 1.333 12.0 1.537453.... 1.537537....
+#> 4   9   6        0        5 1.500 13.5 1.505904.... 1.505900....
+#> 5   6   7        0        5 0.857 10.5 1.549960.... 1.549860....
+#> 6   7   7        0        5 1.000 12.2 1.488596.... 1.488548....
 \end{CodeOutput}
@@
 
-Two figure below depicts correctly identified time lag (i.e.~using time correction parameter) using time correction and time and distance corrections models.
+The figure below depicts correctly identified time lag (i.e.~using time correction parameter) using time correction and time and distance corrections models.
 
@@
 
-Additional scenario can create time lag: if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
+There are few other starting issues worth mentioning.
+For example, if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
 In this case, time lag is a negative number since it reduces the split times.
-This scenario will not be simulated in the current paper.
-
-\hypertarget{bad-position-of-the-timing-gates}{%
-\paragraph{Bad position of the timing gates}\label{bad-position-of-the-timing-gates}}
-
 Another common issue with timing gates in the practical field settings is the bad measurement of the distance and thus bad positions of the timing gates.
-In this example, we will simulate the scenario where the initial timing gate is not positioned at \(d=0\), but rather with some error (i.e., -0.5 to 0.5m).
-Although this scenario represents poor procedure implementation, we have included this simulation to depict effects on estimated parameters.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> sim_df <- expand.grid(
-+   MSS = c(6, 7, 8, 9),
-+   MAC = c(6, 7, 8, 9),
-+   displacement = seq(-0.5, 0.5, length.out = 51),
-+   distance = c(5, 10, 20, 30, 40, 50)
-+ )
-R> 
-R> sim_df <- sim_df %>%
-+   mutate(
-+     TAU = MSS / MAC,
-+     PMAX = MSS * MAC / 4,
-+     true_distance = distance + displacement,
-+     time = shorts::predict_time_at_distance(true_distance, MSS, TAU)
-+   )
-R> 
-R> # Add small noise to allow model fit
-R> set.seed(1667)
-R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
-R> sim_df$time <- sim_df$time + rand_noise
-R> 
-R> head(sim_df, 2)
-\end{CodeInput}
-\begin{CodeOutput}
-#>   MSS MAC displacement distance  TAU PMAX true_distance
-#> 1   6   6         -0.5        5 1.00  9.0           4.5
-#> 2   7   6         -0.5        5 1.17 10.5           4.5
-#>           time
-#> 1 1.534434....
-#> 2 1.481967....
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # estimated parameters and predicted time
-R> model_df <- sim_df %>%
-+   group_by(MSS, TAU, displacement) %>%
-+   do(pred_wrapper(.)) %>%
-+   ungroup()
-\end{CodeInput}
-\end{CodeChunk}
-
-Figure below depicts effect of timing gate displacement on the MSS parameter. Time correction and time and distance correction models are almost identical in their correct estimation of the MSS.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> model_df$model <- factor(
-+   model_df$model,
-+   levels = c(
-+     "No correction",
-+     "Fixed correction +0.3s",
-+     "Fixed correction +0.5s",
-+     "Time correction",
-+     "Time and distance correction"
-+   )
-+ )
-R> 
-R> # Estimates plot
-R> df <- model_df %>%
-+   group_by(MSS, TAU, displacement, model) %>%
-+   slice(1) %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
-+   )
-R> 
-R> # MSS
-R> ggplot(df, aes(x = displacement, y = est_MSS, color = model)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-+   xlab("Displacement (m)") +
-+   ylab("estimated MSS (m/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
-
-\end{CodeChunk}
-
-Figure below depicts effect of timing gate displacement on the MSS parameter. Time and distance correction model is the only model that estimated MAC correctly.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # MAC
-R> ggplot(df, aes(x = displacement, y = est_MAC, color = model)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Displacement (m)") +
-+   ylab("estimated MAC (m/s/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
-
-\end{CodeChunk}
-
-Figures below depicts estimated time and distance corrections parameters.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # time_correction
-R> ggplot(df, aes(x = displacement, y = est_time_correction, color = model)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Displacement (m)") +
-+   ylab("estimated time correction (s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
-
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # distance_correction
-R> ggplot(df, aes(x = displacement, y = est_distance_correction, color = model)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   geom_abline(slope = 1, color = "black", linetype = "dashed") +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Displacement (m)") +
-+   ylab("estimated distance correction (m)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
-
-\end{CodeChunk}
-
-The following figure depicts residuals (i.e., predicted time minus observed time).
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Residuals
-R> model_df <- model_df %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-+     group = paste(MSS, MAC, displacement)
-+   )
-R> 
-R> ggplot(model_df, aes(y = residuals, x = distance, color = displacement, group = group)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.3) +
-+   facet_wrap(~model) +
-+   geom_hline(yintercept = 0, linetype = "dashed") +
-+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-+   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
-
-\end{CodeChunk}
-
-Additional sources of issues with the timing gates involve combination of the previous scenarios, as well as effect of number and distances of the timing gates \citetext{\citealp{haugenPowerForceVelocityProfilingSprinting2020}; \citealp[also see][]{haugenDifferenceStartImpact2012}}.
+
+The number and distances of the timing gates can also affect the precision of the estimated sprint parameters \citep{haugenPowerForceVelocityProfilingSprinting2020, haugenDifferenceStartImpact2012}.
+
+In field testing, multiple starting issues can be present.
 For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence.
@@
 #> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>   0.00181   1.00000  -0.00109   0.00189   0.00189   0.00114 
-#>       MAE      MAPE 
-#>   0.00104   0.04981 
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>   0.00181   1.00000  -0.00109   0.00189   0.00189   0.00114   0.00104 
+#>      MAPE 
+#>   0.04981 
 #> 
@@
 #> Model fit:
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE 
-#>        NA   1.00000  -0.00639   0.00510   0.00639   0.00401 
-#>       MAE      MAPE 
-#>   0.00349   0.18387
+#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
+#>        NA   1.00000  -0.00639   0.00510   0.00639   0.00401   0.00349 
+#>      MAPE 
+#>   0.18387
 \end{CodeOutput}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-56-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-49-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
-
-\end{CodeChunk}
-
-As expected, model has more issues predicting unseen short or long sprints.
-Please note, that since LOOCV removes one observation, if model estimates three parameters, one needs to have at least five observations, since we need to make sure that model can be estimated once a single observation is removed.
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
+
+\end{CodeChunk}
+
+As expected, the model has more issues predicting unseen short or long sprints.
+Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
 LOOCV can also be implemented with the mixed-effects models in the \pkg{shorts} package.
@@
 Let's utilize demonstrated functionalities of the \pkg{shorts} package using real-world data.
-The first dataset comes from Usain Bolt run from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data set for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package.
+The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package (see \texttt{?vescovi}).
 
@@
 \begin{CodeOutput}
-#>                      model  MSS   TAU   MAC PMAX
-#> 1            No correction 12.1 1.564  7.77 23.6
-#> 2       No correction - RT 11.7 1.205  9.74 28.6
-#> 3          Time correction 11.7 1.202  9.76 28.6
-#> 4     Time correction - RT 11.7 1.202  9.76 28.6
-#> 5      Distance correction 11.6 0.855 13.56 39.3
-#> 6 Distance correction - RT 11.6 0.855 13.56 39.3
-#>   time_correction distance_correction
-#> 1         0.00000                0.00
-#> 2        -0.18300                0.00
-#> 3        -0.18483                0.00
-#> 4        -0.00183                0.00
-#> 5        -0.81151               -3.98
-#> 6        -0.62851               -3.98
+#>                      model  MSS   TAU   MAC PMAX time_correction
+#> 1            No correction 12.1 1.564  7.77 23.6         0.00000
+#> 2       No correction - RT 11.7 1.205  9.74 28.6        -0.18300
+#> 3          Time correction 11.7 1.202  9.76 28.6        -0.18483
+#> 4     Time correction - RT 11.7 1.202  9.76 28.6        -0.00183
+#> 5      Distance correction 11.6 0.855 13.56 39.3        -0.81151
+#> 6 Distance correction - RT 11.6 0.855 13.56 39.3        -0.62851
+#>   distance_correction
+#> 1                0.00
+#> 2                0.00
+#> 3                0.00
+#> 4                0.00
+#> 5               -3.98
+#> 6               -3.98
 \end{CodeOutput}
@@
 
-Vescovi's data set represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players).
+The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players).
 Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset.
-Soccer players were older (24.6±3.6 vs.~18.9±2.7 yr, p \textless{} 0.0001), however there were no differences for height (167.3±5.9 vs.~167.0±5.7 cm, p = 0.886), body mass (62.5±5.9 vs.~64.0±9.4 kg, p = 0.500) or any sprint interval time (p \textgreater{} 0.650).
 
@@
 #> # A tibble: 6 x 10
-#>   Team  Surface Athlete   Age Height Bodyweight   BMI   BSA
-#>   <chr> <chr>   <fct>   <dbl>  <dbl>      <dbl> <dbl> <dbl>
-#> 1 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
-#> 2 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
-#> 3 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
-#> 4 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
-#> 5 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6
-#> 6 FH Sr Hard C~ FMSS-2~    21   172.       63.6  21.5   1.7
-#> # ... with 2 more variables: distance <dbl>, time <dbl>
-\end{CodeOutput}
-\end{CodeChunk}
-
-Here we will utilize the following models: no corrections model, fixed time correction model (using 0.3s heuristic rule of thumb), estimated time correction as a fixed effect model, estimated time correction as a random effect model, estimated distance correction as fixed effect model (and time correction as random effect), and estimated distance correction as random effect model.
+#>   Team  Surface Athlete   Age Height Bodyweight   BMI   BSA distance
+#>   <chr> <chr>   <fct>   <dbl>  <dbl>      <dbl> <dbl> <dbl>    <dbl>
+#> 1 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6        5
+#> 2 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       10
+#> 3 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       20
+#> 4 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       30
+#> 5 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       35
+#> 6 FH Sr Hard C~ FMSS-2~    21   172.       63.6  21.5   1.7        5
+#> # ... with 1 more variable: time <dbl>
+\end{CodeOutput}
+\end{CodeChunk}
+
+The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-61-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-62-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-55-1} \end{center}
 
@@
 This provides practical anchor for the residual magnitude, often referred to as \emph{smallest worthwhile change} (SWC) or \emph{smallest effect size of interest} (SESOI) \citep{jovanovicBmbstatsBootstrapMagnitudebased2020}.
-If the residuals are within this magnitude band, they have no practical significance, and thus the model is good in making practically useful predictions.
-
-Error bars represent residual bias and \(\pm\) 1SD.
+If the residuals are within this magnitude band, then the model is good in making practically useful predictions.
+
+Error bars represent residual bias \(\pm\) 1SD.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-63-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-56-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-64-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
 

### 2020-10-21 paper.Rmd
Removed shorts::

@@
 
-$W$ in equation \@ref(eq:velocity-time) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `shorts::model_using_splits` function:
+$W$ in equation \@ref(eq:velocity-time) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `model_using_splits` function:
 
@@
 
-m1 <- shorts::model_using_splits(
+m1 <- model_using_splits(
   distance = split_distance,
@@
 
-Once we have estimated MSS and TAU, we can use `shorts::predict_` family of functions to predict various relationships (i.e. time at distance, acceleration at distance, velocity at time):
+Once we have estimated MSS and TAU, we can use `predict_` family of functions to predict various relationships (i.e. time at distance, acceleration at distance, velocity at time):
 
@@
 # Predict time at distance
-shorts::predict_time_at_distance(
+predict_time_at_distance(
   distance = split_distance,
@@
 # Predict acceleration at time
-shorts::predict_acceleration_at_time(
+predict_acceleration_at_time(
   time = c(0, 1, 2, 3, 4, 5, 6),
@@
   time = split_time,
-  pred_time = shorts::predict_time_at_distance(
+  pred_time = predict_time_at_distance(
     distance = split_distance,
@@
 
-To estimate force production at distance or time (using `shorts::predict_force_at_distance` and `shorts::predict_force_at_time` functions), as well as power production (using `shorts::predict_power_at_distance` and `shorts::predict_power_at_time` functions), one needs to take into account the air resistance.
-Air resistance (N) is estimated using `shorts::get_air_resistance` function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
-
-```{r}
-shorts::get_air_resistance(
+To estimate force production at distance or time (using `predict_force_at_distance` and `predict_force_at_time` functions), as well as power production (using `predict_power_at_distance` and `predict_power_at_time` functions), one needs to take into account the air resistance.
+Air resistance (N) is estimated using `get_air_resistance` function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
+
+```{r}
+get_air_resistance(
   velocity = 5,
@@
 
-When estimating force and power, the air resistance parameters can be set using `"..."`, which are forwarded to the `shorts::get_air_resistance`:
+When estimating force and power, the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance`:
 
@@
 # To calculate horizontal force produced
-shorts::predict_force_at_distance(
+predict_force_at_distance(
   distance = split_distance,
@@
   TAU = m1$parameters$TAU,
-  # Additional parameters forwarded to shorts::get_air_resistance
+  # Additional parameters forwarded to get_air_resistance
   # Otherwise, defaults are used
@@
 # To calculate power produced
-shorts::predict_power_at_distance(
+predict_power_at_distance(
   distance = split_distance,
@@
 
-The easiest way to get all kinematics for short sprints is to use `shorts::predict_kinematics` functions:
-
-```{r}
-df <- shorts::predict_kinematics(
+The easiest way to get all kinematics for short sprints is to use `predict_kinematics` functions:
+
+```{r}
+df <- predict_kinematics(
   m1,
@@
   frequency = 100,
-  # Additional parameters forwarded to shorts::get_air_resistance
+  # Additional parameters forwarded to get_air_resistance
   # Otherwise, defaults are used
@@
 Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
-To identify these values, \pkg{shorts} package comes with `shorts::find_` family of functions:
+To identify these values, \pkg{shorts} package comes with `find_` family of functions:
 
@@
 # Finds distance where 90% of maximum sprinting speed is reached
-shorts::find_velocity_critical_distance(
+find_velocity_critical_distance(
   MSS = m1$parameters$MSS,
@@
 # Finds maximal power and distance (this time using air resistance)
-shorts::find_max_power_distance(
+find_max_power_distance(
   MSS = m1$parameters$MSS,
   TAU = m1$parameters$TAU,
-  # Additional parameters forwarded to shorts::get_air_resistance
+  # Additional parameters forwarded to get_air_resistance
   # Otherwise, defaults are used
@@
 # Finds distance over 90% power range
-shorts::find_power_critical_distance(
+find_power_critical_distance(
   MSS = m1$parameters$MSS,
   TAU = m1$parameters$TAU,
-  # Additional parameters forwarded to shorts::get_air_resistance
+  # Additional parameters forwarded to get_air_resistance
   # Otherwise, defaults are used
@@
 Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in \pkg{shorts} for split times, one can use `shorts::mixed_model_using_splits` function.
+To perform mixed-effects models in \pkg{shorts} for split times, one can use `mixed_model_using_splits` function.
 To demonstrate this functionality, we load the `split_times` dataset provided in the \pkg{shorts} package:
@@
 # Mixed model
-m2 <- shorts::mixed_model_using_splits(
+m2 <- mixed_model_using_splits(
   data = split_times,
@@
 
-m3 <- shorts::model_using_radar(
+m3 <- model_using_radar(
   velocity = sprint_velocity,
@@
 ```{r}
-m3_weighted <- shorts::model_using_radar(
+m3_weighted <- model_using_radar(
   velocity = sprint_velocity,
@@
 
-Mixed-effects model using radar data is done using `shorts::mixed_model_using_radar` function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
+Mixed-effects model using radar data is done using `mixed_model_using_radar` function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
 
@@
 
-m4 <- shorts::mixed_model_using_radar(
+m4 <- mixed_model_using_radar(
   radar_gun_data,
@@
 # Without synchronization issues
-m5 <- shorts::model_using_radar(
+m5 <- model_using_radar(
   velocity = df$velocity,
@@
 # With time added
-m6 <- shorts::model_using_radar(
+m6 <- model_using_radar(
   velocity = df$velocity,
@@
 # With time deducted
-m7 <- shorts::model_using_radar(
+m7 <- model_using_radar(
   velocity = df$velocity,
@@
 
-This model is incorporated in the `shorts::model_using_radar_with_time_correction` function:
+This model is incorporated in the `model_using_radar_with_time_correction` function:
 
@@
 # With time added
-m8 <- shorts::model_using_radar_with_time_correction(
+m8 <- model_using_radar_with_time_correction(
   velocity = df$velocity,
@@
 # With time deducted
-m9 <- shorts::model_using_radar_with_time_correction(
+m9 <- model_using_radar_with_time_correction(
   velocity = df$velocity,
@@
 
-When using `shorts::predict_` family of functions, one can provide estimated time correction to get predictions at original time scale.
+When using `predict_` family of functions, one can provide estimated time correction to get predictions at original time scale.
 
@@
 
-When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the `shorts::mixed_model_using_radar_with_time_correction` function.
+When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the `mixed_model_using_radar_with_time_correction` function.
 
@@
 # Mixed model with time correction being fixed effect
-m10 <- shorts::mixed_model_using_radar_with_time_correction(
+m10 <- mixed_model_using_radar_with_time_correction(
   radar_gun_data,
@@
 # Mixed model with time correction being random effect
-m11 <- shorts::mixed_model_using_radar_with_time_correction(
+m11 <- mixed_model_using_radar_with_time_correction(
   radar_gun_data,
@@
   distance = c(5, 10, 20, 30, 40),
-  john_time = shorts::predict_time_at_distance(distance, MSS, TAU),
+  john_time = predict_time_at_distance(distance, MSS, TAU),
 
@@
   jack_distance = distance + 0.5,
-  jack_true_time = shorts::predict_time_at_distance(jack_distance, MSS, TAU),
-  time_05m = shorts::predict_time_at_distance(0.5, MSS, TAU),
+  jack_true_time = predict_time_at_distance(jack_distance, MSS, TAU),
+  time_05m = predict_time_at_distance(0.5, MSS, TAU),
   jack_time = jack_true_time - time_05m
@@
 
-john_profile <- shorts::model_using_splits(
+john_profile <- model_using_splits(
   distance = split_times$distance,
@@
 
-jack_profile <- shorts::model_using_splits(
+jack_profile <- model_using_splits(
   distance = split_times$distance,
@@
     true_distance = distance + flying_start_distance,
-    true_time = shorts::predict_time_at_distance(true_distance, MSS, TAU),
-    stolen_time = shorts::predict_time_at_distance(flying_start_distance, MSS, TAU),
+    true_time = predict_time_at_distance(true_distance, MSS, TAU),
+    stolen_time = predict_time_at_distance(flying_start_distance, MSS, TAU),
     time = true_time - stolen_time
@@
 pred_wrapper <- function(data) {
-  model <- shorts::model_using_splits(
+  model <- model_using_splits(
     distance = data$distance,
@@
 
-  predicted_time <- shorts::predict_time_at_distance(
+  predicted_time <- predict_time_at_distance(
     distance = data$distance,
@@
 Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
-To estimate time correction, we use `shorts::model_using_splits_with_time_correction` function.
+To estimate time correction, we use `model_using_splits_with_time_correction` function.
 Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5s) or estimated time correction:
@@
 ```{r}
-jack_profile_fixed_time_short <- shorts::model_using_splits(
+jack_profile_fixed_time_short <- model_using_splits(
   distance = split_times$distance,
@@
 
-jack_profile_fixed_time_long <- shorts::model_using_splits(
+jack_profile_fixed_time_long <- model_using_splits(
   distance = split_times$distance,
@@
 
-jack_profile_time_estimated <- shorts::model_using_splits_with_time_correction(
+jack_profile_time_estimated <- model_using_splits_with_time_correction(
   distance = split_times$distance,
@@
 
-This model is implemented in `shorts::model_using_splits_with_corrections` function. Below are the model estimates:
-
-```{r}
-jack_profile_distance_correction <- shorts::model_using_splits_with_corrections(
+This model is implemented in `model_using_splits_with_corrections` function. Below are the model estimates:
+
+```{r}
+jack_profile_distance_correction <- model_using_splits_with_corrections(
   distance = split_times$distance,
@@
 pred_wrapper <- function(data) {
-  no_correction <- shorts::model_using_splits(
+  no_correction <- model_using_splits(
     distance = data$distance,
@@
 
-  fixed_correction_short <- shorts::model_using_splits(
+  fixed_correction_short <- model_using_splits(
     distance = data$distance,
@@
 
-  fixed_correction_long <- shorts::model_using_splits(
+  fixed_correction_long <- model_using_splits(
     distance = data$distance,
@@
 
-  time_correction <- shorts::model_using_splits_with_time_correction(
+  time_correction <- model_using_splits_with_time_correction(
     distance = data$distance,
@@
 
-  time_dist_correction <- shorts::model_using_splits_with_corrections(
+  time_dist_correction <- model_using_splits_with_corrections(
     distance = data$distance,
@@
 
-  df$predicted_time <- shorts::predict_time_at_distance(
+  df$predicted_time <- predict_time_at_distance(
     distance = df$distance,
@@
 
-Time correction and time and distance corrections are also implemented in the mixed-models using `shorts::mixed_model_using_splits_with_time_correction` and `shorts::mixed_model_using_splits_with_corrections`.
+Time correction and time and distance corrections are also implemented in the mixed-models using `mixed_model_using_splits_with_time_correction` and `mixed_model_using_splits_with_corrections`.
 We will showcase their use at the end of this article.
@@
     PMAX = MSS * MAC / 4,
-    true_time = shorts::predict_time_at_distance(distance, MSS, TAU),
+    true_time = predict_time_at_distance(distance, MSS, TAU),
     time = true_time + time_lag
@@
 ```{r}
-jack_LOOCV <- shorts::model_using_splits_with_time_correction(
+jack_LOOCV <- model_using_splits_with_time_correction(
   distance = split_times$distance,
@@
 # No corrections model
-bolt_m1 <- shorts::model_using_splits(
+bolt_m1 <- model_using_splits(
   distance = bolt_distance,
@@
 # Model with reaction time as fixed time correction
-bolt_m2 <- shorts::model_using_splits(
+bolt_m2 <- model_using_splits(
   distance = bolt_distance,
@@
 # Model with estimated time correction
-bolt_m3 <- shorts::model_using_splits_with_time_correction(
+bolt_m3 <- model_using_splits_with_time_correction(
   distance = bolt_distance,
@@
 # Model with estimated time correction, but deducted reaction time
-bolt_m4 <- shorts::model_using_splits_with_time_correction(
+bolt_m4 <- model_using_splits_with_time_correction(
   distance = bolt_distance,
@@
 # Model with estimated time and distance corrections
-bolt_m5 <- shorts::model_using_splits_with_corrections(
+bolt_m5 <- model_using_splits_with_corrections(
   distance = bolt_distance,
@@
 # Model with estimated time and distance corrections and deducted reaction time
-bolt_m6 <- shorts::model_using_splits_with_corrections(
+bolt_m6 <- model_using_splits_with_corrections(
   distance = bolt_distance,
@@
 ```{r}
-no_corrections <- shorts::mixed_model_using_splits(
+no_corrections <- mixed_model_using_splits(
   df,
@@
 
-fixed_correction <- shorts::mixed_model_using_splits(
+fixed_correction <- mixed_model_using_splits(
   df,
@@
 
-time_correction_fixed <- shorts::mixed_model_using_splits_with_time_correction(
+time_correction_fixed <- mixed_model_using_splits_with_time_correction(
   df,
@@
 
-time_correction_random <- shorts::mixed_model_using_splits_with_time_correction(
+time_correction_random <- mixed_model_using_splits_with_time_correction(
   df,
@@
 
-time_distance_correction_fixed <- shorts::mixed_model_using_splits_with_corrections(
+time_distance_correction_fixed <- mixed_model_using_splits_with_corrections(
   df,
@@
 
-time_distance_correction_random <- shorts::mixed_model_using_splits_with_corrections(
+time_distance_correction_random <- mixed_model_using_splits_with_corrections(
   df,

### 2020-10-21 paper.Rmd
Mladen changes

@@
   xlab("Time (s)") +
-  theme(legend.position = "top", legend.title = element_blank())
+  theme(
+    legend.position = "top",
+    legend.title = element_blank())
 
@@
   xlab(expression("Velocity (ms"^-1 * ")")) +
-  theme(legend.position = "top", legend.title = element_blank())
+  theme(
+    legend.position = "top",
+    legend.title = element_blank())
 
@@
 )
-```
-
-Here is a plot of observed and model predicted split times, with the help of the \pkg{ggplot2} package [@R-ggplot2]:
-
-```{r}
-require(ggplot2)
-
-df <- data.frame(
-  distance = split_distance,
-  time = split_time,
-  pred_time = predict_time_at_distance(
-    distance = split_distance,
-    MSS = m1$parameters$MSS,
-    TAU = m1$parameters$TAU
-  )
-)
-
-ggplot(df, aes(x = distance)) +
-  theme_bw(8) +
-  geom_line(aes(y = time)) +
-  geom_point(aes(y = pred_time), color = "red") +
-  xlab("Distance (m)") +
-  ylab("Time (s)")
 ```
@@
 
-Plotting the model predictions can be done once we convert data from wide to long with the help of \pkg{dplyr} [@R-dplyr], \pkg{tidyr} [@R-tidyr], and \pkg{tidyverse} [@R-tidyverse] packages:
+Plotting the model predictions can be done once we convert data from wide to long with the help of   \pkg{ggplot2} [@R-ggplot2], \pkg{dplyr} [@R-dplyr], \pkg{tidyr} [@R-tidyr], and \pkg{tidyverse} [@R-tidyverse] packages:
 
@@
 
-head(split_times)
-
 # Mixed model
@@
 
-The following plot contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:
+The following figure contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:
 
@@
   xlab("Distance (m)") +
-  theme(legend.position = "top", legend.title = element_blank())
+  theme(
+    legend.position = "top",
+    legend.title = element_blank())
 ```
@@
 data("radar_gun_data")
-
-head(radar_gun_data)
 
@@
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
-
-```{r}
+Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU (Figure \@ref(fig:radar-gun-sync)).
+
+```{r radar-gun-sync, include=FALSE, fig.cap="Radar gun synchronization issues"}
 df <- tibble(
@@
 
-head(df)
-```
-
-```{r}
 plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")
 
-ggplot(plot_df, aes(x = value, y = velocity, color = `Sync issue`)) +
+ggplot(
+  plot_df,
+  aes(x = value, y = velocity, color = `Sync issue`)
+) +
   theme_bw(8) +
@@
   xlab("Time (s)") +
-  ylab(expression("Velocity (" * ms^-1 * ")"))
+  ylab(expression("Velocity (" * ms^-1 * ")")) +
+  theme(
+    legend.title = element_blank(), 
+    legend.position = "top")
 ```
@@
 
-ggplot(plot_df, aes(x = distance, y = time, color = athlete, group = athlete)) +
+ggplot(
+  plot_df,
+  aes(x = distance, y = time, color = athlete, group = athlete)
+) +
   theme_bw(8) +
@@
   xlab("Distance (m)") +
-  ylab("Time (s)")
+  ylab("Time (s)") +
+  theme(
+    legend.title = element_blank(), 
+    legend.position = "top")
 ```
@@
 sim_df$time <- sim_df$time + rand_noise
-
-head(sim_df)
 ```
@@
 
-  cbind(data, params, data.frame(predicted_time = as.numeric(predicted_time)))
+  cbind(
+    data,
+    params,
+    data.frame(predicted_time = as.numeric(predicted_time))
+  )
 }
 
-# estimated parameters and predited time
+# estimated parameters and predicted time
 model_df <- sim_df %>%
@@
 # MSS
-ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = MAC_string)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_MSS, color = MAC_string)
+) +
   theme_bw(8) +
@@
   ylab("estimated MSS (m/s)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(), 
+    legend.position = "top")
 ```
@@
 # MAC
-ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = MSS_string)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_MAC, color = MSS_string)
+) +
   theme_bw(8) +
@@
   ylab("estimated MAC (m/s/s)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(), 
+    legend.position = "top")
 ```
@@
 # PMAX
-ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)
+) +
   theme_bw(8) +
@@
   ylab("estimated PMAX (W/kg)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(), 
+    legend.position = "top")
 ```
@@
 
-ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+ggplot(
+  model_df,
+  aes(y = residuals, x = distance, color = flying_start_distance, group = group)
+) +
   theme_bw(8) +
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)")
+  ylab("Predicted time - observed time (s)") +
+  theme(legend.position = "top") + 
+  labs(color = "Flying start distance")
 ```
@@
 ```{r}
-ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+ggplot(
+  model_df,
+  aes(y = residuals, x = distance, color = flying_start_distance, group = group)
+) +
   theme_bw(8) +
@@
 
-# estimated parameters and predited time
+# estimated parameters and predicted time
 model_df <- sim_df %>%
@@
 # MSS
-ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = model)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_MSS, color = model)
+) +
   theme_bw(8) +
@@
   ylab("estimated MSS (m/s)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
 # MAC
-ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = model)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_MAC, color = model)
+) +
   theme_bw(8) +
@@
   ylab("estimated MAC (m/s/s)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
 # PMAX
-ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = model)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_PMAX, color = model)
+) +
   theme_bw(8) +
@@
   ylab("estimated PMAX (W/kg)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
 # time_correction
-ggplot(df, aes(x = flying_start_distance, y = est_time_correction, color = model)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_time_correction, color = model)
+) +
   theme_bw(8) +
@@
   ylab("estimated time correction (s)")  +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
 # distance_correction
-ggplot(df, aes(x = flying_start_distance, y = est_distance_correction, color = model)) +
+ggplot(
+  df,
+  aes(x = flying_start_distance, y = est_distance_correction, color = model)
+) +
   theme_bw(8) +
@@
   ylab("estimated distance correction (m)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
 # time_correction
-ggplot(df, aes(x = time_lag, y = est_time_correction, color = model)) +
+ggplot(
+  df,
+  aes(x = time_lag, y = est_time_correction, color = model)
+) +
   theme_bw(8) +
@@
   ylab("estimated time correction (s)") +
-  theme(legend.title = element_blank())
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
 # distance_correction
-ggplot(df, aes(x = time_lag, y = est_distance_correction, color = model)) +
+ggplot(
+  df,
+  aes(x = time_lag, y = est_distance_correction, color = model)
+) +
   theme_bw(8) +
@@
 
-ggplot(model_df, aes(y = residuals, x = distance, color = time_lag, group = group)) +
+ggplot(
+  model_df,
+  aes(y = residuals, x = distance, color = time_lag, group = group)
+) +
   theme_bw(8) +
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)")
+  ylab("Predicted time - observed time (s)")  +
+  theme(legend.position = "top") + 
+  labs(color = "Time lag")
 ```
@@
   xlab("Distance (m)") +
-  ylab("Predicted - observed time (s)")
+  ylab("Predicted - observed time (s)") +
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top")
 ```
@@
   # slice(1:10) %>%
-  pivot_longer(cols = 9:13, names_to = "distance", values_to = "time") %>%
+  pivot_longer(
+    cols = 9:13,
+    names_to = "distance",
+    values_to = "time"
+  ) %>%
   mutate(
@@
   )
-
-head(df)
 ```
@@
     data = model_SESOI,
-    aes(xintercept = SESOI_lower), color = "blue", alpha = 0.5, linetype = "dashed"
+    aes(xintercept = SESOI_lower),
+    color = "blue", alpha = 0.5, linetype = "dashed"
   ) +
@@
     data = model_SESOI,
-    aes(xintercept = SESOI_upper), color = "blue", alpha = 0.5, linetype = "dashed"
+    aes(xintercept = SESOI_upper),
+    color = "blue", alpha = 0.5, linetype = "dashed"
   ) +
@@
   geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
-  geom_errorbarh(data = model_SESOI, aes(xmin = lower, xmax = upper), height = 0.1, color = "black") +
+  geom_errorbarh(
+    data = model_SESOI,
+    aes(xmin = lower, xmax = upper),
+    height = 0.1, color = "black"
+  ) +
   geom_point(data = model_SESOI, aes(x = bias), color = "black") +

### 2020-10-21 paper.tex
Mladen changes

@@
 
-\(W\) in equation \eqref{eq:velocity-time} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{shorts::model\_using\_splits} function:
+\(W\) in equation \eqref{eq:velocity-time} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{model\_using\_splits} function:
 
@@
 R> 
-R> m1 <- shorts::model_using_splits(
+R> m1 <- model_using_splits(
 +   distance = split_distance,
@@
 
-Once we have estimated MSS and TAU, we can use \texttt{shorts::predict\_} family of functions to predict various relationships (i.e.~time at distance, acceleration at distance, velocity at time):
+Once we have estimated MSS and TAU, we can use \texttt{predict\_} family of functions to predict various relationships (i.e.~time at distance, acceleration at distance, velocity at time):
 
@@
 R> # Predict time at distance
-R> shorts::predict_time_at_distance(
+R> predict_time_at_distance(
 +   distance = split_distance,
@@
 R> # Predict acceleration at time
-R> shorts::predict_acceleration_at_time(
+R> predict_acceleration_at_time(
 +   time = c(0, 1, 2, 3, 4, 5, 6),
@@
 
-Here is a plot of observed and model predicted split times, with the help of the \pkg{ggplot2} package \citep{R-ggplot2}:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> require(ggplot2)
-R> 
-R> df <- data.frame(
-+   distance = split_distance,
-+   time = split_time,
-+   pred_time = shorts::predict_time_at_distance(
-+     distance = split_distance,
-+     MSS = m1$parameters$MSS,
-+     TAU = m1$parameters$TAU
-+   )
-+ )
-R> 
-R> ggplot(df, aes(x = distance)) +
-+   theme_bw(8) +
-+   geom_line(aes(y = time)) +
-+   geom_point(aes(y = pred_time), color = "red") +
-+   xlab("Distance (m)") +
-+   ylab("Time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1} \end{center}
-
-\end{CodeChunk}
-
 \hypertarget{air-resistance-and-the-calculation-of-force-and-mechanical-power}{%
@@
 
-To estimate force production at distance or time (using \texttt{shorts::predict\_force\_at\_distance} and \texttt{shorts::predict\_force\_at\_time} functions), as well as power production (using \texttt{shorts::predict\_power\_at\_distance} and \texttt{shorts::predict\_power\_at\_time} functions), one needs to take into account the air resistance.
-Air resistance (N) is estimated using \texttt{shorts::get\_air\_resistance} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \citet{arsacModelingEnergetics100m2002}, \citet{samozinoSimpleMethodMeasuring2016}, and \citet{vaningenschenauCanCyclePower1991} for more information):
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> shorts::get_air_resistance(
+To estimate force production at distance or time (using \texttt{predict\_force\_at\_distance} and \texttt{predict\_force\_at\_time} functions), as well as power production (using \texttt{predict\_power\_at\_distance} and \texttt{predict\_power\_at\_time} functions), one needs to take into account the air resistance.
+Air resistance (N) is estimated using \texttt{get\_air\_resistance} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \citet{arsacModelingEnergetics100m2002}, \citet{samozinoSimpleMethodMeasuring2016}, and \citet{vaningenschenauCanCyclePower1991} for more information):
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> get_air_resistance(
 +   velocity = 5,
@@
 
-When estimating force and power, the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{shorts::get\_air\_resistance}:
+When estimating force and power, the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{get\_air\_resistance}:
 
@@
 R> # To calculate horizontal force produced
-R> shorts::predict_force_at_distance(
+R> predict_force_at_distance(
 +   distance = split_distance,
@@
 +   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to shorts::get_air_resistance
++   # Additional parameters forwarded to get_air_resistance
 +   # Otherwise, defaults are used
@@
 R> # To calculate power produced
-R> shorts::predict_power_at_distance(
+R> predict_power_at_distance(
 +   distance = split_distance,
@@
 
-The easiest way to get all kinematics for short sprints is to use \texttt{shorts::predict\_kinematics} functions:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- shorts::predict_kinematics(
+The easiest way to get all kinematics for short sprints is to use \texttt{predict\_kinematics} functions:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> df <- predict_kinematics(
 +   m1,
@@
 +   frequency = 100,
-+   # Additional parameters forwarded to shorts::get_air_resistance
++   # Additional parameters forwarded to get_air_resistance
 +   # Otherwise, defaults are used
@@
 
-Plotting the model predictions can be done once we convert data from wide to long with the help of \pkg{dplyr} \citep{R-dplyr}, \pkg{tidyr} \citep{R-tidyr}, and \pkg{tidyverse} \citep{R-tidyverse} packages:
+Plotting the model predictions can be done once we convert data from wide to long with the help of \pkg{ggplot2} \citep{R-ggplot2}, \pkg{dplyr} \citep{R-dplyr}, \pkg{tidyr} \citep{R-tidyr}, and \pkg{tidyverse} \citep{R-tidyverse} packages:
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-8-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
@@
 Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90\% of maximum sprinting speed is reached, or where peak power is within 90\% range.
-To identify these values, \pkg{shorts} package comes with \texttt{shorts::find\_} family of functions:
+To identify these values, \pkg{shorts} package comes with \texttt{find\_} family of functions:
 
@@
 R> # Finds distance where 90% of maximum sprinting speed is reached
-R> shorts::find_velocity_critical_distance(
+R> find_velocity_critical_distance(
 +   MSS = m1$parameters$MSS,
@@
 R> # Finds maximal power and distance (this time using air resistance)
-R> shorts::find_max_power_distance(
+R> find_max_power_distance(
 +   MSS = m1$parameters$MSS,
 +   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to shorts::get_air_resistance
++   # Additional parameters forwarded to get_air_resistance
 +   # Otherwise, defaults are used
@@
 R> 
-R> 
 R> # Finds distance over 90% power range
-R> shorts::find_power_critical_distance(
+R> find_power_critical_distance(
 +   MSS = m1$parameters$MSS,
 +   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to shorts::get_air_resistance
++   # Additional parameters forwarded to get_air_resistance
 +   # Otherwise, defaults are used
@@
 Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in \pkg{shorts} for split times, one can use \texttt{shorts::mixed\_model\_using\_splits} function.
+To perform mixed-effects models in \pkg{shorts} for split times, one can use \texttt{mixed\_model\_using\_splits} function.
 To demonstrate this functionality, we load the \texttt{split\_times} dataset provided in the \pkg{shorts} package:
@@
 R> 
-R> head(split_times)
-\end{CodeInput}
-\begin{CodeOutput}
-#> # A tibble: 6 x 4
-#>   athlete bodyweight distance     time
-#>   <chr>        <dbl>    <dbl> <I<dbl>>
-#> 1 John            75        5     1.20
-#> 2 John            75       10     1.97
-#> 3 John            75       15     2.66
-#> 4 John            75       20     3.31
-#> 5 John            75       30     4.59
-#> 6 John            75       40     5.85
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
 R> # Mixed model
-R> m2 <- shorts::mixed_model_using_splits(
+R> m2 <- mixed_model_using_splits(
 +   data = split_times,
@@
 
-The following plot contains kinematics for all athletes in \texttt{split\_times} dataset. Please note that power calculation takes default parameters for each individual:
+The following figure contains kinematics for all athletes in \texttt{split\_times} dataset. Please note that power calculation takes default parameters for each individual:
 
@@
 +   xlab("Distance (m)") +
-+   theme(legend.position = "top", legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
++   theme(
++     legend.position = "top",
++     legend.title = element_blank())
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-11-1} \end{center}
 
@@
 R> 
-R> m3 <- shorts::model_using_radar(
+R> m3 <- model_using_radar(
 +   velocity = sprint_velocity,
@@
 \begin{CodeInput}
-R> m3_weighted <- shorts::model_using_radar(
+R> m3_weighted <- model_using_radar(
 +   velocity = sprint_velocity,
@@
 
-Mixed-effects model using radar data is done using \texttt{shorts::mixed\_model\_using\_radar} function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
+Mixed-effects model using radar data is done using \texttt{mixed\_model\_using\_radar} function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
 
@@
 R> 
-R> head(radar_gun_data)
-\end{CodeInput}
-\begin{CodeOutput}
-#> # A tibble: 6 x 4
-#>   athlete bodyweight  time velocity
-#>   <chr>        <dbl> <dbl>    <dbl>
-#> 1 John            75  0       0    
-#> 2 John            75  0.01    0.075
-#> 3 John            75  0.02    0.149
-#> 4 John            75  0.03    0.222
-#> 5 John            75  0.04    0.291
-#> 6 John            75  0.05    0.367
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> m4 <- shorts::mixed_model_using_radar(
+R> m4 <- mixed_model_using_radar(
 +   radar_gun_data,
@@
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- tibble(
-+   `true time` = sprint_time,
-+   velocity = sprint_velocity,
-+   `0.5s added` = `true time` + 0.5,
-+   `0.5s deducted` = `true time` - 0.5
-+ )
-R> 
-R> head(df)
-\end{CodeInput}
-\begin{CodeOutput}
-#> # A tibble: 6 x 4
-#>   `true time` velocity `0.5s added` `0.5s deducted`
-#>         <dbl>    <dbl>        <dbl>           <dbl>
-#> 1           0     0             0.5            -0.5
-#> 2           1     4.83          1.5             0.5
-#> 3           2     7.07          2.5             1.5
-#> 4           3     8.1           3.5             2.5
-#> 5           4     8.59          4.5             3.5
-#> 6           5     8.81          5.5             4.5
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")
-R> 
-R> ggplot(plot_df, aes(x = value, y = velocity, color = `Sync issue`)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   xlab("Time (s)") +
-+   ylab(expression("Velocity (" * ms^-1 * ")"))
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
-
-\end{CodeChunk}
+Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU (Figure \ref{fig:radar-gun-sync}).
 
@@
 R> # Without synchronization issues
-R> m5 <- shorts::model_using_radar(
+R> m5 <- model_using_radar(
 +   velocity = df$velocity,
@@
 R> # With time added
-R> m6 <- shorts::model_using_radar(
+R> m6 <- model_using_radar(
 +   velocity = df$velocity,
@@
 R> # With time deducted
-R> m7 <- shorts::model_using_radar(
+R> m7 <- model_using_radar(
 +   velocity = df$velocity,
@@
 
-This model is incorporated in the \texttt{shorts::model\_using\_radar\_with\_time\_correction} function:
+This model is incorporated in the \texttt{model\_using\_radar\_with\_time\_correction} function:
 
@@
 R> # With time added
-R> m8 <- shorts::model_using_radar_with_time_correction(
+R> m8 <- model_using_radar_with_time_correction(
 +   velocity = df$velocity,
@@
 R> # With time deducted
-R> m9 <- shorts::model_using_radar_with_time_correction(
+R> m9 <- model_using_radar_with_time_correction(
 +   velocity = df$velocity,
@@
 
-When using \texttt{shorts::predict\_} family of functions, one can provide estimated time correction to get predictions at original time scale.
+When using \texttt{predict\_} family of functions, one can provide estimated time correction to get predictions at original time scale.
 
@@
 
-When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the \texttt{shorts::mixed\_model\_using\_radar\_with\_time\_correction} function.
+When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the \texttt{mixed\_model\_using\_radar\_with\_time\_correction} function.
 
@@
 R> # Mixed model with time correction being fixed effect
-R> m10 <- shorts::mixed_model_using_radar_with_time_correction(
+R> m10 <- mixed_model_using_radar_with_time_correction(
 +   radar_gun_data,
@@
 R> 
-R> 
 R> # Mixed model with time correction being random effect
-R> m11 <- shorts::mixed_model_using_radar_with_time_correction(
+R> m11 <- mixed_model_using_radar_with_time_correction(
 +   radar_gun_data,
@@
 +   distance = c(5, 10, 20, 30, 40),
-+   john_time = shorts::predict_time_at_distance(distance, MSS, TAU),
++   john_time = predict_time_at_distance(distance, MSS, TAU),
 + 
@@
 +   jack_distance = distance + 0.5,
-+   jack_true_time = shorts::predict_time_at_distance(jack_distance, MSS, TAU),
-+   time_05m = shorts::predict_time_at_distance(0.5, MSS, TAU),
++   jack_true_time = predict_time_at_distance(jack_distance, MSS, TAU),
++   time_05m = predict_time_at_distance(0.5, MSS, TAU),
 +   jack_time = jack_true_time - time_05m
@@
 R> 
-R> ggplot(plot_df, aes(x = distance, y = time, color = athlete, group = athlete)) +
+R> ggplot(
++   plot_df,
++   aes(x = distance, y = time, color = athlete, group = athlete)
++ ) +
 +   theme_bw(8) +
@@
 +   xlab("Distance (m)") +
-+   ylab("Time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-23-1} \end{center}
++   ylab("Time (s)") +
++   theme(
++     legend.title = element_blank(), 
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
 
@@
 R> 
-R> john_profile <- shorts::model_using_splits(
+R> john_profile <- model_using_splits(
 +   distance = split_times$distance,
@@
 R> 
-R> jack_profile <- shorts::model_using_splits(
+R> jack_profile <- model_using_splits(
 +   distance = split_times$distance,
@@
 +     true_distance = distance + flying_start_distance,
-+     true_time = shorts::predict_time_at_distance(true_distance, MSS, TAU),
-+     stolen_time = shorts::predict_time_at_distance(flying_start_distance, MSS, TAU),
++     true_time = predict_time_at_distance(true_distance, MSS, TAU),
++     stolen_time = predict_time_at_distance(flying_start_distance, MSS, TAU),
 +     time = true_time - stolen_time
@@
 R> sim_df$time <- sim_df$time + rand_noise
-R> 
-R> head(sim_df)
-\end{CodeInput}
-\begin{CodeOutput}
-#>   MSS MAC flying_start_distance distance   TAU PMAX true_distance
-#> 1   6   6                     0        5 1.000  9.0             5
-#> 2   7   6                     0        5 1.167 10.5             5
-#> 3   8   6                     0        5 1.333 12.0             5
-#> 4   9   6                     0        5 1.500 13.5             5
-#> 5   6   7                     0        5 0.857 10.5             5
-#> 6   7   7                     0        5 1.000 12.2             5
-#>      true_time stolen_time         time
-#> 1 1.639197....           0 1.639356....
-#> 2 1.579728....           0 1.579675....
-#> 3 1.537453....           0 1.537537....
-#> 4 1.505904....           0 1.505900....
-#> 5 1.549960....           0 1.549860....
-#> 6 1.488596....           0 1.488548....
-\end{CodeOutput}
+\end{CodeInput}
 \end{CodeChunk}
@@
 R> pred_wrapper <- function(data) {
-+   model <- shorts::model_using_splits(
++   model <- model_using_splits(
 +     distance = data$distance,
@@
 + 
-+   predicted_time <- shorts::predict_time_at_distance(
++   predicted_time <- predict_time_at_distance(
 +     distance = data$distance,
@@
 + 
-+   cbind(data, params, data.frame(predicted_time = as.numeric(predicted_time)))
++   cbind(
++     data,
++     params,
++     data.frame(predicted_time = as.numeric(predicted_time))
++   )
 + }
 R> 
-R> # estimated parameters and predited time
+R> # estimated parameters and predicted time
 R> model_df <- sim_df %>%
@@
 R> # MSS
-R> ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = MAC_string)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_MSS, color = MAC_string)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated MSS (m/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-27-1} \end{center}
++   theme(
++     legend.title = element_blank(), 
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-24-1} \end{center}
 
@@
 R> # MAC
-R> ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = MSS_string)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_MAC, color = MSS_string)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated MAC (m/s/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-28-1} \end{center}
++   theme(
++     legend.title = element_blank(), 
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-25-1} \end{center}
 
@@
 R> # PMAX
-R> ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated PMAX (W/kg)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-29-1} \end{center}
++   theme(
++     legend.title = element_blank(), 
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-26-1} \end{center}
 
@@
 R> 
-R> ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+R> ggplot(
++   model_df,
++   aes(y = residuals, x = distance, color = flying_start_distance, group = group)
++ ) +
 +   theme_bw(8) +
@@
 +   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-30-1} \end{center}
++   ylab("Predicted time - observed time (s)") +
++   theme(legend.position = "top") + 
++   labs(color = "Flying start distance")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-27-1} \end{center}
 
@@
 \begin{CodeInput}
-R> ggplot(model_df, aes(y = residuals, x = distance, color = flying_start_distance, group = group)) +
+R> ggplot(
++   model_df,
++   aes(y = residuals, x = distance, color = flying_start_distance, group = group)
++ ) +
 +   theme_bw(8) +
@@
 +   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-31-1} \end{center}
++   ylab("Predicted time - observed time (s)") +
++   theme(legend.position = "top") + 
++   labs(color = "Flying start distance")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-28-1} \end{center}
 
@@
 Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
-To estimate time correction, we use \texttt{shorts::model\_using\_splits\_with\_time\_correction} function.
+To estimate time correction, we use \texttt{model\_using\_splits\_with\_time\_correction} function.
 Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5s) or estimated time correction:
@@
 \begin{CodeInput}
-R> jack_profile_fixed_time_short <- shorts::model_using_splits(
+R> jack_profile_fixed_time_short <- model_using_splits(
 +   distance = split_times$distance,
@@
 R> 
-R> jack_profile_fixed_time_long <- shorts::model_using_splits(
+R> jack_profile_fixed_time_long <- model_using_splits(
 +   distance = split_times$distance,
@@
 R> 
-R> jack_profile_time_estimated <- shorts::model_using_splits_with_time_correction(
+R> jack_profile_time_estimated <- model_using_splits_with_time_correction(
 +   distance = split_times$distance,
@@
 
-This model is implemented in \texttt{shorts::model\_using\_splits\_with\_corrections} function. Below are the model estimates:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> jack_profile_distance_correction <- shorts::model_using_splits_with_corrections(
+This model is implemented in \texttt{model\_using\_splits\_with\_corrections} function. Below are the model estimates:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> jack_profile_distance_correction <- model_using_splits_with_corrections(
 +   distance = split_times$distance,
@@
 R> pred_wrapper <- function(data) {
-+   no_correction <- shorts::model_using_splits(
++   no_correction <- model_using_splits(
 +     distance = data$distance,
@@
 + 
-+   fixed_correction_short <- shorts::model_using_splits(
++   fixed_correction_short <- model_using_splits(
 +     distance = data$distance,
@@
 + 
-+   fixed_correction_long <- shorts::model_using_splits(
++   fixed_correction_long <- model_using_splits(
 +     distance = data$distance,
@@
 + 
-+   time_correction <- shorts::model_using_splits_with_time_correction(
++   time_correction <- model_using_splits_with_time_correction(
 +     distance = data$distance,
@@
 + 
-+   time_dist_correction <- shorts::model_using_splits_with_corrections(
++   time_dist_correction <- model_using_splits_with_corrections(
 +     distance = data$distance,
@@
 + 
-+   df$predicted_time <- shorts::predict_time_at_distance(
++   df$predicted_time <- predict_time_at_distance(
 +     distance = df$distance,
@@
 R> 
-R> # estimated parameters and predited time
+R> # estimated parameters and predicted time
 R> model_df <- sim_df %>%
@@
 R> # MSS
-R> ggplot(df, aes(x = flying_start_distance, y = est_MSS, color = model)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_MSS, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated MSS (m/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-32-1} \end{center}
 
@@
 R> # MAC
-R> ggplot(df, aes(x = flying_start_distance, y = est_MAC, color = model)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_MAC, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated MAC (m/s/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-33-1} \end{center}
 
@@
 R> # PMAX
-R> ggplot(df, aes(x = flying_start_distance, y = est_PMAX, color = model)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_PMAX, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated PMAX (W/kg)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-34-1} \end{center}
 
@@
 R> # time_correction
-R> ggplot(df, aes(x = flying_start_distance, y = est_time_correction, color = model)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_time_correction, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated time correction (s)")  +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
 
@@
 R> # distance_correction
-R> ggplot(df, aes(x = flying_start_distance, y = est_distance_correction, color = model)) +
+R> ggplot(
++   df,
++   aes(x = flying_start_distance, y = est_distance_correction, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated distance correction (m)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-39-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
 
@@
 +   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-40-1} \end{center}
++   ylab("Predicted time - observed time (s)") +
++   theme(legend.position = "top") + 
++   labs(color = "Flying start distance")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
 
@@
 
-Time correction and time and distance corrections are also implemented in the mixed-models using \texttt{shorts::mixed\_model\_using\_splits\_with\_time\_correction} and \texttt{shorts::mixed\_model\_using\_splits\_with\_corrections}.
+Time correction and time and distance corrections are also implemented in the mixed-models using \texttt{mixed\_model\_using\_splits\_with\_time\_correction} and \texttt{mixed\_model\_using\_splits\_with\_corrections}.
 We will showcase their use at the end of this article.
@@
 +     PMAX = MSS * MAC / 4,
-+     true_time = shorts::predict_time_at_distance(distance, MSS, TAU),
++     true_time = predict_time_at_distance(distance, MSS, TAU),
 +     time = true_time + time_lag
@@
 R> sim_df$time <- sim_df$time + rand_noise
-R> 
-R> head(sim_df)
-\end{CodeInput}
-\begin{CodeOutput}
-#>   MSS MAC time_lag distance   TAU PMAX    true_time         time
-#> 1   6   6        0        5 1.000  9.0 1.639197.... 1.639356....
-#> 2   7   6        0        5 1.167 10.5 1.579728.... 1.579675....
-#> 3   8   6        0        5 1.333 12.0 1.537453.... 1.537537....
-#> 4   9   6        0        5 1.500 13.5 1.505904.... 1.505900....
-#> 5   6   7        0        5 0.857 10.5 1.549960.... 1.549860....
-#> 6   7   7        0        5 1.000 12.2 1.488596.... 1.488548....
-\end{CodeOutput}
+\end{CodeInput}
 \end{CodeChunk}
@@
 +   ylab("estimated MSS (m/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-40-1} \end{center}
 
@@
 +   ylab("estimated MAC (m/s/s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
 
@@
 R> # time_correction
-R> ggplot(df, aes(x = time_lag, y = est_time_correction, color = model)) +
+R> ggplot(
++   df,
++   aes(x = time_lag, y = est_time_correction, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated time correction (s)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
 
@@
 R> # distance_correction
-R> ggplot(df, aes(x = time_lag, y = est_distance_correction, color = model)) +
+R> ggplot(
++   df,
++   aes(x = time_lag, y = est_distance_correction, color = model)
++ ) +
 +   theme_bw(8) +
@@
 +   ylab("estimated distance correction (m)") +
-+   theme(legend.title = element_blank())
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
 
@@
 R> 
-R> ggplot(model_df, aes(y = residuals, x = distance, color = time_lag, group = group)) +
+R> ggplot(
++   model_df,
++   aes(y = residuals, x = distance, color = time_lag, group = group)
++ ) +
 +   theme_bw(8) +
@@
 +   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
++   ylab("Predicted time - observed time (s)")  +
++   theme(legend.position = "top") + 
++   labs(color = "Time lag")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
 
@@
 \begin{CodeInput}
-R> jack_LOOCV <- shorts::model_using_splits_with_time_correction(
+R> jack_LOOCV <- model_using_splits_with_time_correction(
 +   distance = split_times$distance,
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-49-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
 
@@
 +   xlab("Distance (m)") +
-+   ylab("Predicted - observed time (s)")
-\end{CodeInput}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
++   ylab("Predicted - observed time (s)") +
++   theme(
++     legend.title = element_blank(),
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
 
@@
 R> # No corrections model
-R> bolt_m1 <- shorts::model_using_splits(
+R> bolt_m1 <- model_using_splits(
 +   distance = bolt_distance,
@@
 R> # Model with reaction time as fixed time correction
-R> bolt_m2 <- shorts::model_using_splits(
+R> bolt_m2 <- model_using_splits(
 +   distance = bolt_distance,
@@
 R> # Model with estimated time correction
-R> bolt_m3 <- shorts::model_using_splits_with_time_correction(
+R> bolt_m3 <- model_using_splits_with_time_correction(
 +   distance = bolt_distance,
@@
 R> # Model with estimated time correction, but deducted reaction time
-R> bolt_m4 <- shorts::model_using_splits_with_time_correction(
+R> bolt_m4 <- model_using_splits_with_time_correction(
 +   distance = bolt_distance,
@@
 R> # Model with estimated time and distance corrections
-R> bolt_m5 <- shorts::model_using_splits_with_corrections(
+R> bolt_m5 <- model_using_splits_with_corrections(
 +   distance = bolt_distance,
@@
 R> # Model with estimated time and distance corrections and deducted reaction time
-R> bolt_m6 <- shorts::model_using_splits_with_corrections(
+R> bolt_m6 <- model_using_splits_with_corrections(
 +   distance = bolt_distance,
@@
 +   # slice(1:10) %>%
-+   pivot_longer(cols = 9:13, names_to = "distance", values_to = "time") %>%
++   pivot_longer(
++     cols = 9:13,
++     names_to = "distance",
++     values_to = "time"
++   ) %>%
 +   mutate(
@@
 +   )
-R> 
-R> head(df)
-\end{CodeInput}
-\begin{CodeOutput}
-#> # A tibble: 6 x 10
-#>   Team  Surface Athlete   Age Height Bodyweight   BMI   BSA distance
-#>   <chr> <chr>   <fct>   <dbl>  <dbl>      <dbl> <dbl> <dbl>    <dbl>
-#> 1 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6        5
-#> 2 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       10
-#> 3 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       20
-#> 4 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       30
-#> 5 W So~ Natura~ FMSS-1~    21   163.       54.8  20.5   1.6       35
-#> 6 FH Sr Hard C~ FMSS-2~    21   172.       63.6  21.5   1.7        5
-#> # ... with 1 more variable: time <dbl>
-\end{CodeOutput}
+\end{CodeInput}
 \end{CodeChunk}
@@
 \begin{CodeInput}
-R> no_corrections <- shorts::mixed_model_using_splits(
+R> no_corrections <- mixed_model_using_splits(
 +   df,
@@
 R> 
-R> fixed_correction <- shorts::mixed_model_using_splits(
+R> fixed_correction <- mixed_model_using_splits(
 +   df,
@@
 R> 
-R> time_correction_fixed <- shorts::mixed_model_using_splits_with_time_correction(
+R> time_correction_fixed <- mixed_model_using_splits_with_time_correction(
 +   df,
@@
 R> 
-R> time_correction_random <- shorts::mixed_model_using_splits_with_time_correction(
+R> time_correction_random <- mixed_model_using_splits_with_time_correction(
 +   df,
@@
 R> 
-R> time_distance_correction_fixed <- shorts::mixed_model_using_splits_with_corrections(
+R> time_distance_correction_fixed <- mixed_model_using_splits_with_corrections(
 +   df,
@@
 R> 
-R> time_distance_correction_random <- shorts::mixed_model_using_splits_with_corrections(
+R> time_distance_correction_random <- mixed_model_using_splits_with_corrections(
 +   df,
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-55-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
 
@@
 +     data = model_SESOI,
-+     aes(xintercept = SESOI_lower), color = "blue", alpha = 0.5, linetype = "dashed"
++     aes(xintercept = SESOI_lower),
++     color = "blue", alpha = 0.5, linetype = "dashed"
 +   ) +
@@
 +     data = model_SESOI,
-+     aes(xintercept = SESOI_upper), color = "blue", alpha = 0.5, linetype = "dashed"
++     aes(xintercept = SESOI_upper),
++     color = "blue", alpha = 0.5, linetype = "dashed"
 +   ) +
@@
 +   geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
-+   geom_errorbarh(data = model_SESOI, aes(xmin = lower, xmax = upper), height = 0.1, color = "black") +
++   geom_errorbarh(
++     data = model_SESOI,
++     aes(xmin = lower, xmax = upper),
++     height = 0.1, color = "black"
++   ) +
 +   geom_point(data = model_SESOI, aes(x = bias), color = "black") +
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-56-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
 

### 2020-10-21 paper.Rmd
Draft ready

@@
 Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
-Short sprint performance data comes from two sources: (1) split times using timing gates and (2) radar gun.
-In this paper, we demonstrate how these data can be modeled using the \pkg{shorts} package and how timing issues (i.e., synchronization issues with the radar gun, or using the flying start with the timing gates), which are common in practical measurement settings, can introduce error in parameter estimates.
-Further, we demonstrated how different model definitions implemented in \pkg{shorts} package can alleviate these errors.
+The most common methods to assess short sprint performance are with a radar gun or timing gates.
+In this paper, we: 1) provide the \pkg{shorts} package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the \pkg{shorts} package to help alleviate errors within the subsequent parameter outcomes.
 keywords:
-  formatted: [sprints, sports, modeling]
+  formatted: [sprinting, sports, modeling]
 preamble: >
@@
 bibliography: [references.bib, packages.bib]
-biblio-style: apalike
 link-citations: yes
@@
 # automatically create a bib database for R packages
-knitr::write_bib(c(
-  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
-), "packages.bib")
+#knitr::write_bib(c(
+#  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
+#), "packages.bib")
 
@@
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (<5 sec) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the type of sport.
-For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)[@brownAssessmentLinearSprinting2004].
+Peak anaerobic power is achieved within the first few seconds (<5s) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40m)[@brownAssessmentLinearSprinting2004].
 Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
@@
 Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
-Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20m), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.

### 2020-10-21 paper.tex
Draft ready

@@
 Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
-Short sprint performance data comes from two sources: (1) split times using timing gates and (2) radar gun.
-In this paper, we demonstrate how these data can be modeled using the \pkg{shorts} package and how timing issues (i.e., synchronization issues with the radar gun, or using the flying start with the timing gates), which are common in practical measurement settings, can introduce error in parameter estimates.
-Further, we demonstrated how different model definitions implemented in \pkg{shorts} package can alleviate these errors.
+The most common methods to assess short sprint performance are with a radar gun or timing gates.
+In this paper, we: 1) provide the \pkg{shorts} package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the \pkg{shorts} package to help alleviate errors within the subsequent parameter outcomes.
 }
 
-\Keywords{sprints, sports, modeling}
+\Keywords{sprinting, sports, modeling}
 
@@
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (\textless5 sec) of maximal efforts \citep{mangineSpeedForcePower2014}, whereas the ability to achieve maximal sprint speed varies based on the type of sport.
-For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)\citep{brownAssessmentLinearSprinting2004}.
+Peak anaerobic power is achieved within the first few seconds (\textless5s) of maximal efforts \citep{mangineSpeedForcePower2014}, whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40m)\citep{brownAssessmentLinearSprinting2004}.
 Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
@@
 Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
-Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20m), which can now be incorporated into the method for determining sprint mechanical properties \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.

### 2020-10-21 code.R

@@ -1,0 +1,1538 @@
+## ----setup, include=FALSE-------------------------------------------
+require(knitr)
+require(shorts)
+require(tidyverse)
+require(bookdown)
+
+my_random_seed <- 1667
+set.seed(my_random_seed)
+
+options("width" = 70)
+
+knitr::opts_chunk$set(
+  comment = "#>",
+  collapse = TRUE,
+  cache = FALSE,
+  fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
+  dpi = 600,
+  out.width = "100%",
+  fig.align = "center",
+  fig.width = 6,
+  fig.height = 6 * 0.618, # 1 / phi
+  fig.show = "hold",
+  echo = TRUE,
+  warning = FALSE,
+  message = FALSE,
+  width = 70
+)
+
+# automatically create a bib database for R packages
+#knitr::write_bib(c(
+#  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
+#), "packages.bib")
+
+
+# Set rounding
+op <- options()
+options(digits = 3)
+
+
+## ----four-athletes-table, echo=FALSE--------------------------------
+athletes <- tribble(
+  ~Athlete, ~MSS, ~MAC,
+  "Athlete A", 12, 10,
+  "Athlete B", 12, 6,
+  "Athlete C", 8, 10,
+  "Athlete D", 8, 6
+)
+
+athletes <- athletes %>%
+  mutate(TAU = MSS / MAC)
+
+knitr::kable(
+  athletes,
+  caption = "Four athletes with different MSS and MAC parameters.",
+  digits = 2
+)
+
+
+## ----four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds."----
+kinematics <- expand_grid(
+  athletes,
+  time = seq(0, 6, length.out = 1000)
+) %>%
+  mutate(
+    distance = predict_distance_at_time(time, MSS, TAU),
+    velocity = predict_velocity_at_time(time, MSS, TAU),
+    acceleration = predict_acceleration_at_time(time, MSS, TAU)
+  )
+
+kinematics_long <- kinematics %>%
+  pivot_longer(
+    cols = c("distance", "velocity", "acceleration"),
+    names_to = "variable"
+  ) %>%
+  mutate(
+    variable = factor(
+      variable,
[…]

### 2020-10-21 paper.Rmd
Draft preparation

@@
   fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
-  dpi = 600,
+  dpi = 300,
   out.width = "100%",

### 2020-10-21 code.R

@@ -1,4 +1,4 @@
-## ----setup, include=FALSE-------------------------------------------
+## ----setup, include=FALSE---------------------------------------------------------------------------------------------------------
 require(knitr)
 require(shorts)
 require(tidyverse)
@@ -14,7 +14,7 @@ knitr::opts_chunk$set(
   collapse = TRUE,
   cache = FALSE,
   fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
-  dpi = 600,
+  dpi = 300,
   out.width = "100%",
   fig.align = "center",
   fig.width = 6,
@@ -37,7 +37,7 @@ op <- options()
 options(digits = 3)
 
 
-## ----four-athletes-table, echo=FALSE--------------------------------
+## ----four-athletes-table, echo=FALSE----------------------------------------------------------------------------------------------
 athletes <- tribble(
   ~Athlete, ~MSS, ~MAC,
   "Athlete A", 12, 10,
@@ -120,7 +120,7 @@ gg <- ggplot(
 plot(gg)
 
 
-## -------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 require(shorts)
 
 split_distance <- c(10, 20, 30, 40)
@@ -135,11 +135,11 @@ m1 <- model_using_splits(
 m1
 
 
-## -------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 summary(m1$model)
 
 
-## -------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 # Predict time at distance
 predict_time_at_distance(
   distance = split_distance,
@@ -155,7 +155,7 @@ predict_acceleration_at_time(
 )
 
 
-## -------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 get_air_resistance(
   velocity = 5,
   bodymass = 80,
@@ -166,7 +166,7 @@ get_air_resistance(
 )
 
 
-## -------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
 # To calculate horizontal force produced
 predict_force_at_distance(
   distance = split_distance,
@@ -196,7 +196,7 @@ predict_power_at_distance(
 )
 
 
-## -------------------------------------------------------------------
+## ---------------------------------------------------------------------------------------------------------------------------------
[…]

### 2020-10-26 paper.Rmd
Added calculation of 99% MSS for Bolt's example

@@
 
+Here is the model estimate of the time and distance it takes for Bolt to reach 99% of MSS:
+
+```{r}
+bolt_model <- bolt_model %>%
+  group_by(model) %>%
+  mutate(
+    dist_95_MSS = find_velocity_critical_distance(
+      MSS = MSS, TAU = TAU, time_correction = time_correction, 
+      distance_correction = distance_correction, percent = 0.99
+    ),
+   time_95_MSS = find_velocity_critical_time(
+      MSS = MSS, TAU = TAU, time_correction = time_correction, 
+      percent = 0.99
+    )
+  )
+
+bolt_model[c(1, 8, 9)]
+```
+
 ## Vescovi data

### 2020-10-26 paper.tex
Added calculation of 99% MSS for Bolt's example

@@
 
+Here is the model estimate of the time and distance it takes for Bolt to reach 99\% of MSS:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> bolt_model <- bolt_model %>%
++   group_by(model) %>%
++   mutate(
++     dist_95_MSS = find_velocity_critical_distance(
++       MSS = MSS, TAU = TAU, time_correction = time_correction, 
++       distance_correction = distance_correction, percent = 0.99
++     ),
++    time_95_MSS = find_velocity_critical_time(
++       MSS = MSS, TAU = TAU, time_correction = time_correction, 
++       percent = 0.99
++     )
++   )
+R> 
+R> bolt_model[c(1, 8, 9)]
+\end{CodeInput}
+\begin{CodeOutput}
+#> # A tibble: 6 x 3
+#> # Groups:   model [6]
+#>   model                    dist_95_MSS time_95_MSS
+#>   <chr>                          <dbl>       <dbl>
+#> 1 No correction                   68.7        7.20
+#> 2 No correction - RT              51.1        5.73
+#> 3 Time correction                 51.0        5.72
+#> 4 Time correction - RT            51.0        5.54
+#> 5 Distance correction             39.8        4.75
+#> 6 Distance correction - RT        39.8        4.57
+\end{CodeOutput}
+\end{CodeChunk}
+
 \hypertarget{vescovi-data}{%
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-55-1} \end{center}
 

### 2020-10-26 code.R

@@ -1,4 +1,4 @@
-## ----setup, include=FALSE---------------------------------------------------------------------------------------------------------
+## ----setup, include=FALSE-------------------------------------------
 require(knitr)
 require(shorts)
 require(tidyverse)
@@ -37,7 +37,7 @@ op <- options()
 options(digits = 3)
 
 
-## ----four-athletes-table, echo=FALSE----------------------------------------------------------------------------------------------
+## ----four-athletes-table, echo=FALSE--------------------------------
 athletes <- tribble(
   ~Athlete, ~MSS, ~MAC,
   "Athlete A", 12, 10,
@@ -120,7 +120,7 @@ gg <- ggplot(
 plot(gg)
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 require(shorts)
 
 split_distance <- c(10, 20, 30, 40)
@@ -135,11 +135,11 @@ m1 <- model_using_splits(
 m1
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 summary(m1$model)
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 # Predict time at distance
 predict_time_at_distance(
   distance = split_distance,
@@ -155,7 +155,7 @@ predict_acceleration_at_time(
 )
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 get_air_resistance(
   velocity = 5,
   bodymass = 80,
@@ -166,7 +166,7 @@ get_air_resistance(
 )
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 # To calculate horizontal force produced
 predict_force_at_distance(
   distance = split_distance,
@@ -196,7 +196,7 @@ predict_power_at_distance(
 )
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 df <- predict_kinematics(
   m1,
   max_time = 6,
@@ -213,7 +213,7 @@ df <- predict_kinematics(
 head(df)
 
 
-## ---------------------------------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 require(tidyverse)
 
 df <- pivot_longer(data = df, cols = -2)
[…]

### 2020-10-29 paper.Rmd
Fixed Bolt's critical distance and missing graph

--- 2020-10-21 paper.Rmd
@@
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU (Figure \@ref(fig:radar-gun-sync)).
-
-```{r radar-gun-sync, include=FALSE, fig.cap="Radar gun synchronization issues"}
+Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
+
+```{r}
 df <- tibble(
@@
 bolt_model
+```
+
+Here is the model estimate of the time and distance it takes for Bolt to reach 99% of MSS.
+Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start, not the measurement scale.
+
+```{r}
+bolt_model <- bolt_model %>%
+  group_by(model) %>%
+  mutate(
+    dist_95_MSS = find_velocity_critical_distance(
+      MSS = MSS, TAU = TAU, 
+      #time_correction = time_correction, 
+      #distance_correction = distance_correction,
+      percent = 0.99
+    ),
+   time_95_MSS = find_velocity_critical_time(
+      MSS = MSS, TAU = TAU, 
+      time_correction = time_correction, 
+      percent = 0.99
+    )
+  )
+
+bolt_model[c(1, 8, 9)]
 ```

### 2020-10-29 paper.tex
Fixed Bolt's critical distance and missing graph

--- 2020-10-21 paper.tex
@@
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU (Figure \ref{fig:radar-gun-sync}).
+Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> df <- tibble(
++   `true time` = sprint_time,
++   velocity = sprint_velocity,
++   `0.5s added` = `true time` + 0.5,
++   `0.5s deducted` = `true time` - 0.5
++ )
+R> 
+R> plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")
+R> 
+R> ggplot(
++   plot_df,
++   aes(x = value, y = velocity, color = `Sync issue`)
++ ) +
++   theme_bw(8) +
++   geom_line(alpha = 0.7) +
++   xlab("Time (s)") +
++   ylab(expression("Velocity (" * ms^-1 * ")")) +
++   theme(
++     legend.title = element_blank(), 
++     legend.position = "top")
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
+
+\end{CodeChunk}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-21-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-24-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-25-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-25-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-26-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-26-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-27-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-27-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-28-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-28-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-29-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-32-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-33-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-33-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-34-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-34-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-40-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-48-1} \end{center}
 
@@
 
+Here is the model estimate of the time and distance it takes for Bolt to reach 99\% of MSS.
+Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start, not the measurement scale.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> bolt_model <- bolt_model %>%
++   group_by(model) %>%
++   mutate(
++     dist_95_MSS = find_velocity_critical_distance(
++       MSS = MSS, TAU = TAU, 
++       #time_correction = time_correction, 
++       #distance_correction = distance_correction,
++       percent = 0.99
++     ),
++    time_95_MSS = find_velocity_critical_time(
++       MSS = MSS, TAU = TAU, 
++       time_correction = time_correction, 
++       percent = 0.99
++     )
++   )
+R> 
+R> bolt_model[c(1, 8, 9)]
+\end{CodeInput}
+\begin{CodeOutput}
+#> # A tibble: 6 x 3
+#> # Groups:   model [6]
+#>   model                    dist_95_MSS time_95_MSS
+#>   <chr>                          <dbl>       <dbl>
+#> 1 No correction                   68.7        7.20
+#> 2 No correction - RT              51.1        5.73
+#> 3 Time correction                 51.0        5.72
+#> 4 Time correction - RT            51.0        5.54
+#> 5 Distance correction             35.8        4.75
+#> 6 Distance correction - RT        35.8        4.57
+\end{CodeOutput}
+\end{CodeChunk}
+
 \hypertarget{vescovi-data}{%
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-55-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-56-1} \end{center}
 

### 2020-10-29 code.R

@@ -1,4 +1,4 @@
-## ----setup, include=FALSE-------------------------------------------
+## ----setup, include=FALSE-----------------------------------------------------------------------------------
 require(knitr)
 require(shorts)
 require(tidyverse)
@@ -37,7 +37,7 @@ op <- options()
 options(digits = 3)
 
 
-## ----four-athletes-table, echo=FALSE--------------------------------
+## ----four-athletes-table, echo=FALSE------------------------------------------------------------------------
 athletes <- tribble(
   ~Athlete, ~MSS, ~MAC,
   "Athlete A", 12, 10,
@@ -120,7 +120,7 @@ gg <- ggplot(
 plot(gg)
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 require(shorts)
 
 split_distance <- c(10, 20, 30, 40)
@@ -135,11 +135,11 @@ m1 <- model_using_splits(
 m1
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 summary(m1$model)
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 # Predict time at distance
 predict_time_at_distance(
   distance = split_distance,
@@ -155,7 +155,7 @@ predict_acceleration_at_time(
 )
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 get_air_resistance(
   velocity = 5,
   bodymass = 80,
@@ -166,7 +166,7 @@ get_air_resistance(
 )
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 # To calculate horizontal force produced
 predict_force_at_distance(
   distance = split_distance,
@@ -196,7 +196,7 @@ predict_power_at_distance(
 )
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 df <- predict_kinematics(
   m1,
   max_time = 6,
@@ -213,7 +213,7 @@ df <- predict_kinematics(
 head(df)
 
 
-## -------------------------------------------------------------------
+## -----------------------------------------------------------------------------------------------------------
 require(tidyverse)
 
 df <- pivot_longer(data = df, cols = -2)
@@ -226,7 +226,7 @@ ggplot(df, aes(x = distance, y = value)) +
   xlab("Distance (m)")
 
 
[…]

### 2020-11-09 paper.Rmd
recompiled and fixed 0.95 to 0.99 in Bolts

@@
 
-$W$ in equation \@ref(eq:velocity-time) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `model_using_splits` function:
+$W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `model_using_splits` function:
 
@@
   mutate(
-    dist_95_MSS = find_velocity_critical_distance(
+    dist_99_MSS = find_velocity_critical_distance(
       MSS = MSS, TAU = TAU, 
@@
     ),
-   time_95_MSS = find_velocity_critical_time(
+   time_99_MSS = find_velocity_critical_time(
       MSS = MSS, TAU = TAU, 
-      time_correction = time_correction, 
+      #time_correction = time_correction, 
       percent = 0.99

### 2020-11-09 paper.tex
recompiled and fixed 0.95 to 0.99 in Bolts

@@
 
-\(W\) in equation \eqref{eq:velocity-time} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{model\_using\_splits} function:
+\(W\) in equation \eqref{eq:time-distance} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{model\_using\_splits} function:
 
@@
 +   mutate(
-+     dist_95_MSS = find_velocity_critical_distance(
++     dist_99_MSS = find_velocity_critical_distance(
 +       MSS = MSS, TAU = TAU, 
@@
 +     ),
-+    time_95_MSS = find_velocity_critical_time(
++    time_99_MSS = find_velocity_critical_time(
 +       MSS = MSS, TAU = TAU, 
-+       time_correction = time_correction, 
++       #time_correction = time_correction, 
 +       percent = 0.99
@@
 #> # Groups:   model [6]
-#>   model                    dist_95_MSS time_95_MSS
+#>   model                    dist_99_MSS time_99_MSS
 #>   <chr>                          <dbl>       <dbl>
 #> 1 No correction                   68.7        7.20
-#> 2 No correction - RT              51.1        5.73
-#> 3 Time correction                 51.0        5.72
+#> 2 No correction - RT              51.1        5.55
+#> 3 Time correction                 51.0        5.54
 #> 4 Time correction - RT            51.0        5.54
-#> 5 Distance correction             35.8        4.75
-#> 6 Distance correction - RT        35.8        4.57
+#> 5 Distance correction             35.8        3.94
+#> 6 Distance correction - RT        35.8        3.94
 \end{CodeOutput}

### 2020-12-30 paper.Rmd
Preprint version

@@
   formatted: "\\pkg{shorts}: An R Package for Modeling Short Sprints"
+DOI: "10.31236/osf.io/4jw62"
+URL: https://osf.io/preprints/sportrxiv/4jw62/
 abstract: >
@@
     base_format: rticles::jss_article
+    extra_dependencies: ["flafter"]
 citation_package: natbib
@@
 link-citations: yes
+header-includes: >
+  \usepackage[htt]{hyphenat}
+  \usepackage{booktabs}
+  \usepackage{microtype}
 editor_options: 
@@
 require(bookdown)
+require(magick)
+
+pdf2png <- function(path) {
+  # only do the conversion for non-LaTeX output
+  if (knitr::is_latex_output()) 
+    return(path)
+  path2 <- xfun::with_ext(path, "png")
+  img <- magick::image_read_pdf(path)
+  magick::image_write(img, path2, format = "png")
+  path2
+}
 
@@
 set.seed(my_random_seed)
-
-options("width" = 70)
 
@@
   fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
-  dpi = 300,
+  dpi = 600,
+  dev = "png",
+  fig.process = pdf2png,
   out.width = "100%",
@@
 op <- options()
-options(digits = 3)
+
+options(
+  digits = 3,
+  "width" = 70,
+  scipen = 999)
 ```
@@
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (<5s) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the type of sport.
-For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40m)[@brownAssessmentLinearSprinting2004].
+Peak anaerobic power is achieved within the first few seconds (<5 s) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)[@brownAssessmentLinearSprinting2004].
 Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
@@
 Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
-Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20m), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
@@
   caption = "Four athletes with different MSS and MAC parameters.",
-  digits = 2
-)
-```
-
-Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6s).
-
-```{r four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds."}
+  digits = 2,
+  booktabs = TRUE
+)
+```
+
+Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6 s).
+
+```{r four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds."}
 kinematics <- expand_grid(
@@
 
-Plotting acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
+Plotting acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get *Acceleration-Velocity Profile*, which is linear, according to the mathematical model.
+If the athlete's body mass (kg) is known, as well as additional air resistance parameters (see [Air resistance and the calculation of force and mechanical power] section of this paper), *Force-Velocity Profile* can be estimated (see [Force-Velocity profile] section of this paper).
 
@@
 
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in \pkg{shorts} package using non-linear least squares regression implemented in the `nls` function in the \pkg{base} R [@R-base] and `nlme` function in the \pkg{nlme} package [@R-nlme] for the mixed-effect models.
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
+Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in \pkg{shorts} package using non-linear least squares regression implemented in the `nls()` function in the \pkg{base} R [@R-base] and `nlme()` function in the \pkg{nlme} package [@R-nlme] for the mixed-effect models.
 
@@
 
-Let's consider an example of an athlete with MSS equal to 9 $ms^-1$, TAU equal to 1.3, and MAC equal to 6.92 $ms^-2$ performing 40m sprint with timing gates positioned at each 10m split.
+Let's consider an example of an athlete with MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$ performing 40m sprint with timing gates positioned at each 10m split.
 For split times, distance is a predictor, and time is the outcome variable, thus the equation \@ref(eq:velocity-time) becomes:
@@
 
-$W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `model_using_splits` function:
+$W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `model_using_splits()` function:
 
@@
 Maximal relative power (PMAX) from the output is estimated using $\frac{MSS \times MAC}{4}$, which disregards the air resistance.
-`time_correction` and `distance_corection` parameters will be covered later in the article.
+`time_correction` and `distance_corection` parameters will be covered later in the paper.
 
 Besides providing *residual standard error* (RSE), \pkg{shorts} functions provide additional model fit estimators.
-Additional information can be gained by exploring the returned object, particularly object returned from the `nls` function:
-
-```{r}
-summary(m1$model)
-```
-
-Once we have estimated MSS and TAU, we can use `predict_` family of functions to predict various relationships (i.e. time at distance, acceleration at distance, velocity at time):
+Additional information can be gained by exploring the returned object, particularly object returned from the `nls()` function:
+
+```{r}
+summary(m1)
+```
+
+To extract estimated model parameters, use S3 `coef()` method:
+
+```{r}
+coef(m1)
+```
+
+To create a simple plot of the model, use S3 `plot()` method, which returns \pkg{ggplot2} [@R-ggplot2] object:
+
+```{r}
+plot(m1) + theme_bw(8)
+```
+
+Once we have estimated MSS and TAU, we can use `predict_XXX()` family of functions to predict various relationships (i.e., time at distance, acceleration at distance, velocity at time, etc.):
 
@@
 
-To estimate force production at distance or time (using `predict_force_at_distance` and `predict_force_at_time` functions), as well as power production (using `predict_power_at_distance` and `predict_power_at_time` functions), one needs to take into account the air resistance.
-Air resistance (N) is estimated using `get_air_resistance` function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
+To estimate force production at distance or time (using `predict_force_at_distance()` and `predict_force_at_time()` functions), as well as power production (using `predict_power_at_distance()` and `predict_power_at_time()` functions), one needs to take into account the air resistance.
+Air resistance (N) is estimated using `get_air_resistance()` function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
 
@@
 
-When estimating force and power, the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance`:
+When estimating force and power, the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance()`:
 
@@
 
-The easiest way to get all kinematics for short sprints is to use `predict_kinematics` functions:
+The easiest way to get all kinematics and kinetics for short sprints is to use `predict_kinematics()` function:
 
@@
 
-df <- pivot_longer(data = df, cols = -2)
+variable_names <- colnames(df)
+
+df <- pivot_longer(data = df, cols = -2) %>%
+  mutate(name = factor(name, levels = variable_names))
 
@@
 
+These kinematic and kinetic variables are utilized in [Force-Velocity profile] estimation, which is covered later in this paper.
+
 ### Utility functions
@@
 Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
-To identify these values, \pkg{shorts} package comes with `find_` family of functions:
+To identify these values, \pkg{shorts} package comes with `find_XXX()` family of functions:
 
@@
 Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in \pkg{shorts} for split times, one can use `mixed_model_using_splits` function.
+To perform mixed-effects models in \pkg{shorts} for split times, one can use `mixed_model_using_splits()` function.
 To demonstrate this functionality, we load the `split_times` dataset provided in the \pkg{shorts} package:
@@
 
-Additional information about mixed-effects model performed using the `nlme` package [@R-nlme] can be obtained using `summary`:
+Additional information about mixed-effects model performed using the `nlme` package [@R-nlme] can be obtained using `summary()` function:
 
@@
 
+S3 method `coef()` when applied on mixed-model result will return both the fixed and random effects:
+
+```{r}
+coef(m2)
+```
+
+To create a simple plot of the model, use S3 `plot()` method:
+
+```{r}
+plot(m2) + theme_bw(8)
+```
+
 The following figure contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:
@@
 
-df <- pivot_longer(df, cols = c(-1, -3))
+variable_names <- colnames(df)
+
+df <- pivot_longer(df, cols = c(-1, -3)) %>%
+  mutate(name = factor(name, levels = variable_names))
 
@@
 
-Let's consider the same example of an athlete with MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$ms^-2$ performing 40m sprint with velocity estimated using radar run (in this case with 1Hz sampling rate).
+Let's consider the same example of an athlete with MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$ performing 40m sprint with velocity estimated using radar run (in this case with 1 Hz sampling rate).
 
@@
 m3
+```
+
+```{r}
+plot(m3) + theme_bw(8)
 ```
@@
 
-Mixed-effects model using radar data is done using `mixed_model_using_radar` function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
+Mixed-effects model using radar data is done using `mixed_model_using_radar()` function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
 
@@
 
+```{r}
+plot(m4) + theme_bw(8)
+```
+
+## Force-Velocity profile
+
+To create *Force-Velocity Profile* (FVP) using single athlete estimated sprint model parameters (i.e., TAU and MSS), you can use `get_FV_profile()` function.
+When estimating FVP, athlete body mass (kg) can be set using `bodymass` parameter, while the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance()` function.
+Details of the FVP method implemented in the \pkg{shorts} package, as well as the interpretation from a sprint training perspective, are covered elsewhere [@haugenPowerForceVelocityProfilingSprinting2020; @morinInterpretingPowerForceVelocityProfiles2016; @morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+
+```{r}
+# To create Force-Velocity Profile
+fvp <- get_FV_profile(
+  MSS = m1$parameters$MSS,
+  TAU = m1$parameters$TAU,
+  bodymass = 80,
+  # Additional parameters forwarded to get_air_resistance
+  # Otherwise, defaults are used
+  bodyheight = 1.85,
+  barometric_pressure = 780,
+  air_temperature = 20,
+  wind_velocity = 0.5
+)
+
+fvp
+```
+
+To plot FVP kinematics and kinetics (which are exactly the same as generated by the `predict_kinematics()` function), use S3 `plot()` function.
+By default, FVP estimated kinetics are plotted against velocity (on x-axis).
+
+```{r}
+plot(fvp) + theme_bw(8)
+```
+
+To plot FVP estimated kinetics against time, use `type = "time"` parameter:
+
+```{r}
+plot(fvp, "time") + theme_bw(8)
+```
+
 # Problems with estimation
@@
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5s to simulate an error in synchronization and its effect on estimated MSS and TAU.
+Let's use our athlete and add and deduct 0.5 s to simulate an error in synchronization and its effect on estimated MSS and TAU.
 
@@
 
-As can be seen from the example, TAU is affected by an error in synchronization of time with velocity.
+As can be seen from the example, all estimated parameters are affected by an error in synchronization of time with velocity (with MSS being the least affected in this example).
 The potential solution incorporated into the \pkg{shorts} package involves estimation of the *time correction* parameter using the following equation:
@@
 
-This model is incorporated in the `model_using_radar_with_time_correction` function:
+This model is incorporated in the `model_using_radar_with_time_correction()` function:
 
@@
 
-When using `predict_` family of functions, one can provide estimated time correction to get predictions at original time scale.
+When using `predict_XXX()` family of functions, one can provide estimated time correction to get predictions at original time scale.
 
@@
 
-When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the `mixed_model_using_radar_with_time_correction` function.
+When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the `mixed_model_using_radar_with_time_correction()` function.
 
@@
 
-Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9$ms^-1$, TAU equal to 1.3, and MAC equal to 6.92$ms^-2$.
+Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$.
 Let's call them John and Jack.
-They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
-The initial timing gate at the start (i.e., $d=0m$) serves to activate the timing system (i.e., when they cross the beam).
+They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40 m.
+The initial timing gate at the start (i.e., $d=0$ m) serves to activate the timing system (i.e., when they cross the beam).
 
@@
 Jack, on the other hand, represents a *practical model*, and decides to move slightly behind the initial timing gate (i.e.
-for 0.5m) and use body rocking to initiate the sprint start.
+for 0.5 m) and use body rocking to initiate the sprint start.
 In other words, Jack is using a *flying start*, a common scenario when testing field sports athletes.
@@
 sprint_parameters <- rbind(
-  unlist(john_profile$parameters),
-  unlist(jack_profile$parameters)
+  coef(john_profile),
+  coef(jack_profile)
 )
@@
 
-Below is a simulation sprint with 5, 10, 20, 30, 40, and 50m splits, with MSS and MAC varying from 6 to 9 ($ms^-1$ and $ms^-2$ respectively), and flying start distance varying from 0 to 1m.
+Below is a simulation sprint with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 ($ms^{-1}$ and $ms^{-2}$ respectively), and flying start distance varying from 0 to 1 m.
 
@@
 set.seed(1667)
-rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
+rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
 sim_df$time <- sim_df$time + rand_noise
@@
 
-  params <- data.frame(t(unlist(model$parameters)))
+  params <- data.frame(t(coef(model)))
 
@@
 
-The following image demonstrates the effect of flying start distance on estimated MSS:
+The following figure demonstrates the effect of flying start distance on estimated MSS:
 
@@
 
-As can be seen from the image, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
+As can be seen from the figure, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
 
@@
 Model residuals are also affected by flying start distance.
-The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40m versus 5, 15, 30m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simulation:
+The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40 m versus 5, 15, 30 m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simulation:
 
@@
   model_df,
-  aes(y = residuals, x = distance, color = flying_start_distance, group = group)
+  aes(
+    y = residuals,
+    x = distance,
+    color = flying_start_distance,
+    group = group)
 ) +
@@
 
-If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler images conveying issues with residuals when there is a flying start:
+If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler figure conveying issues with residuals when there is a flying start:
 
@@
 
-A potential solution is to use a correction factor - the recommendation in the literature is +0.5sec [@haugenSprintMechanicalProperties2020; @haugenSprintMechanicalVariables2019].
-Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 sec [@haugenDifferenceStartImpact2012].
+A potential solution is to use a correction factor - the recommendation in the literature is +0.5 s [@haugenSprintMechanicalProperties2020; @haugenSprintMechanicalVariables2019].
+Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 s [@haugenDifferenceStartImpact2012].
 So, while a timing correction factor is warranted to avoid subsequent errors in estimates of kinetic variables (e.g., overestimate power), a correction factor that is too large will have the opposite effect (e.g., underestimate power).
 
-When implementing *time correction*, equation \@ref(eq:time-distance) becomes:
+Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
+Exactly the same method is suggested by @stenrothForcevelocityProfilingIce2020, named *time shift method*, and the estimated parameter named *time shift parameter*.
+We have named this parameter *time correction* to be in agreement with the parameter introduced in [Problems with time sync with radar gun] section of this paper, as well as the available literature.
+
+When implementing time correction, equation \@ref(eq:time-distance) becomes:
 
@@
 
-Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
-To estimate time correction, we use `model_using_splits_with_time_correction` function.
-Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5s) or estimated time correction:
+To estimate time correction parameter, we use `model_using_splits_with_time_correction()` function.
+Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5 s) or estimated time correction:
 
@@
 jack_parameters <- rbind(
-  unlist(john_profile$parameters),
-  unlist(jack_profile$parameters),
-  unlist(jack_profile_fixed_time_short$parameters),
-  unlist(jack_profile_fixed_time_long$parameters),
-  unlist(jack_profile_time_estimated$parameters)
+  coef(john_profile),
+  coef(jack_profile),
+  coef(jack_profile_fixed_time_short),
+  coef(jack_profile_fixed_time_long),
+  coef(jack_profile_time_estimated)
 )
@@
 
-In Jack's case, both +0.3s fixed time correction and time correction estimation yield parameters closer to John's (i.e. true parameters).
-
-Another model definition, which is a novel approach implemented in the \pkg{shorts} packages, is to utilize *distance correction*, besides time correction.
+In Jack's case, both +0.3 s fixed time correction and time correction estimation yield parameters closer to John's (i.e. true parameters).
+
+Another model definition, which is a novel approach implemented in the \pkg{shorts} package, is to utilize *distance correction*, besides time correction.
 Thus, equation \@ref(eq:time-distance) becomes:
@@
 
-This model is implemented in `model_using_splits_with_corrections` function. Below are the model estimates:
+This model is implemented in `model_using_splits_with_corrections()` function. Below are the model estimates:
 
@@
 jack_parameters <- rbind(
-  unlist(john_profile$parameters),
-  unlist(jack_profile$parameters),
-  unlist(jack_profile_fixed_time_short$parameters),
-  unlist(jack_profile_fixed_time_long$parameters),
-  unlist(jack_profile_time_estimated$parameters),
-  unlist(jack_profile_distance_correction$parameters)
+  coef(john_profile),
+  coef(jack_profile),
+  coef(jack_profile_fixed_time_short),
+  coef(jack_profile_fixed_time_long),
+  coef(jack_profile_time_estimated),
+  coef(jack_profile_distance_correction)
 )
@@
 This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.
+One strategy that is sometimes implemented is adding zeros to the sample (i.e., $t=0$ and $d=0$), which increase the number of observations.
+Unfortunately, this strategy should not be implemented, as explained later in the [Should we add zero to the sample?] section of this paper.
 
@@
       model = "No correction",
-      t(unlist(no_correction$parameters))
+      t(coef(no_correction))
     ),
@@
       model = "Fixed correction +0.3s",
-      t(unlist(fixed_correction_short$parameters))
+      t(coef(fixed_correction_short))
     ),
@@
       model = "Fixed correction +0.5s",
-      t(unlist(fixed_correction_long$parameters))
+      t(coef(fixed_correction_long))
     ),
@@
       model = "Time correction",
-      t(unlist(time_correction$parameters))
+      t(coef(time_correction))
     ),
@@
       model = "Time and distance correction",
-      t(unlist(time_dist_correction$parameters))
+      t(coef(time_dist_correction))
     )
@@
 
-Time correction and time and distance corrections are also implemented in the mixed-models using `mixed_model_using_splits_with_time_correction` and `mixed_model_using_splits_with_corrections`.
-We will showcase their use at the end of this article.
+Time correction and time and distance corrections are also implemented in the mixed-models using `mixed_model_using_splits_with_time_correction()` and `mixed_model_using_splits_with_corrections()`.
+We will showcase their use at the end of this paper.
 
@@
 
-Starting behind the initial timing gate represent only one issue (i.e.
-flying start).
+Starting behind the initial timing gate represent only one issue (i.e., flying start).
 In this section, we simulate one more issue to check the sensitivity of the presented models to other (less common) perturbations when performing field testing.
@@
 One issue that might happen with timing gates is triggering the timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
-This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is *reaction time* (RT) involved.
+This is very similar to the situation when timing starts on a signal (i.e., gun during 100 m sprint race) and there is *reaction time* (RT) involved.
 Both of these scenarios represent *time lag* that is added to the split times.
@@
 The next figure depicts estimated distance correction for the time and distance correction model.
-The estimated distance correction parameters looks jumpy due to random noise that we have to added to allow model fit.
+The estimated distance correction parameters looks jumpy due to random noise that we have to added to allow model fit, as well as the model estimation error.
 
@@
 
+## Should we add zero to the sample?
+
+Fellow sports scientists are often considering adding $t=0$ s at $d=0$ m to the collected split times with the aim of increasing the number of observations.
+The question is whether this strategy is sound and if it should be employed.
+The short answer is no, it shouldn't, particularly if there are flying start issues.
+In the following example, we are demonstrating the issue when zeros are added to the sample when fitting multiple models with the known true parameters:
+
+```{r}
+# Create split times from known MSS and TAU
+df <- tibble(
+  distance = c(5, 10, 20, 30, 40),
+  time = predict_time_at_distance(
+    distance,
+    MSS = 9,
+    TAU = 1.3
+  )
+) %>%
+  mutate(
+    # Add random noise to time
+    time = time + rnorm(n(), 0, 10^-5),
+    gate_time = time - 0.1
+  )
+
+# Model without time correction
+m_no_correction <- model_using_splits(
+  distance = df$distance, 
+  time = df$gate_time
+)
+
+# Model without time correction, but with zeros added
+m_no_correction_zero <- model_using_splits(
+  distance = c(0, df$distance), 
+  time = c(0, df$gate_time)
+)
+
+# Model without adding zeros for the start
+m_no_zero <- model_using_splits_with_time_correction(
+  distance = df$distance, 
+  time = df$gate_time
+)
+
+# Model with added zeros for the start (d=0 and t=0)
+m_with_zero <- model_using_splits_with_time_correction(
+  distance = c(0, df$distance), 
+  time = c(0, df$gate_time)
+)
+
+# Print results
+data.frame(
+  model = c(
+    "Without time correction",
+    "Without time correction with zeros added", 
+    "With time correction",
+    "With time correction with zeros added"),
+  rbind(
+    coef(m_no_correction),
+    coef(m_no_correction_zero),
+    coef(m_no_zero),
+    coef(m_with_zero))
+)
+```
+
+As can be seen from the output, only the model with time correction is able to correctly recover true sprint parameters, but adding zeros to the sample in this case, results in MSS and TAU estimates very close to the estimates of the model without time corrections.
+Thus, adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
+
 # Leave-one-out Cross-Validation
@@
 LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation.
-This process is repeated for each observations in the model dataset.
+This process is repeated for each observation in the model dataset.
 LOOCV allows one to check estimated parameters stability, and model performance on the unseen data.
@@
 
-As expected, the model has more issues predicting unseen short or long sprints.
+As expected, the model has more issues predicting unseen split times for both short or long distances.
 Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
@@
 
-For the sake of this analysis, only 10m splits over 60m are used.
+For the sake of this analysis, only 10 m splits over 60 m race distance are used.
 
@@
 Briefly, all athletes performed a standardized warm-up that included general exercises such as jogging, shuffling, multi-directional movements, and dynamic stretching exercises.
-Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, 30, and 35 meters at a height of approximately 1.0 meter.
+Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, 30, and 35 m at a height of approximately 1.0 m.
 Participants stood with their lead foot positioned approximately 5 cm behind the initial infrared beam (i.e., start line).
@@
 
-Error bars represent residual bias $\pm$ 1SD.
+Error bars represent residual bias $\pm$ 1 SD.
 

### 2020-12-30 paper.tex
Preprint version

@@
+% To remove when JSS.cls is fixed
+% https://github.com/rstudio/rticles/issues/329
+\RequirePackage[2020/02/02]{latexrelease}
 \documentclass[
@@
 
+% Pandoc citation processing
+
 % Pandoc header
+\usepackage[htt]{hyphenat} \usepackage{booktabs} \usepackage{microtype}
+
+\usepackage{flafter}
 
@@
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (\textless5s) of maximal efforts \citep{mangineSpeedForcePower2014}, whereas the ability to achieve maximal sprint speed varies based on the type of sport.
-For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40m)\citep{brownAssessmentLinearSprinting2004}.
+Peak anaerobic power is achieved within the first few seconds (\textless5 s) of maximal efforts \citep{mangineSpeedForcePower2014}, whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)\citep{brownAssessmentLinearSprinting2004}.
 Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
@@
 Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
-Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20m), which can now be incorporated into the method for determining sprint mechanical properties \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
@@
 \centering
-\begin{tabular}[t]{l|r|r|r}
-\hline
+\begin{tabular}[t]{lrrr}
+\toprule
 Athlete & MSS & MAC & TAU\\
-\hline
+\midrule
 Athlete A & 12 & 10 & 1.20\\
-\hline
 Athlete B & 12 & 6 & 2.00\\
-\hline
 Athlete C & 8 & 10 & 0.80\\
-\hline
 Athlete D & 8 & 6 & 1.33\\
-\hline
+\bottomrule
 \end{tabular}
@@
 
-Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6s).
+Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6 s).
 
@@
 
-\caption[Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds]{Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds.}\label{fig:four-athletes-kinematics}
+\caption[Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds]{Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds.}\label{fig:four-athletes-kinematics}
 \end{figure}
@@
 
-Plotting acceleration against velocity (Figure \ref{fig:four-athletes-profile}), we will get Acceleration-Velocity profile, which is linear, according to the mathematical model.
+Plotting acceleration against velocity (Figure \ref{fig:four-athletes-profile}), we will get \emph{Acceleration-Velocity Profile}, which is linear, according to the mathematical model.
+If the athlete's body mass (kg) is known, as well as additional air resistance parameters (see \protect\hyperlink{air-resistance-and-the-calculation-of-force-and-mechanical-power}{Air resistance and the calculation of force and mechanical power} section of this paper), \emph{Force-Velocity Profile} can be estimated (see \protect\hyperlink{force-velocity-profile}{Force-Velocity profile} section of this paper).
 
@@
 
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40m), (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \pkg{shorts} package using non-linear least squares regression implemented in the \texttt{nls} function in the \pkg{base} R \citep{R-base} and \texttt{nlme} function in the \pkg{nlme} package \citep{R-nlme} for the mixed-effect models.
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
+Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \pkg{shorts} package using non-linear least squares regression implemented in the \texttt{nls()} function in the \pkg{base} R \citep{R-base} and \texttt{nlme()} function in the \pkg{nlme} package \citep{R-nlme} for the mixed-effect models.
 
@@
 
-Let's consider an example of an athlete with MSS equal to 9 \(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^-2\) performing 40m sprint with timing gates positioned at each 10m split.
+Let's consider an example of an athlete with MSS equal to 9 \(ms^{-1}\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^{-2}\) performing 40m sprint with timing gates positioned at each 10m split.
 For split times, distance is a predictor, and time is the outcome variable, thus the equation \eqref{eq:velocity-time} becomes:
@@
 
-\(W\) in equation \eqref{eq:time-distance} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{model\_using\_splits} function:
+\(W\) in equation \eqref{eq:time-distance} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{model\_using\_splits()} function:
 
@@
 Maximal relative power (PMAX) from the output is estimated using \(\frac{MSS \times MAC}{4}\), which disregards the air resistance.
-\texttt{time\_correction} and \texttt{distance\_corection} parameters will be covered later in the article.
+\texttt{time\_correction} and \texttt{distance\_corection} parameters will be covered later in the paper.
 
 Besides providing \emph{residual standard error} (RSE), \pkg{shorts} functions provide additional model fit estimators.
-Additional information can be gained by exploring the returned object, particularly object returned from the \texttt{nls} function:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> summary(m1$model)
+Additional information can be gained by exploring the returned object, particularly object returned from the \texttt{nls()} function:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> summary(m1)
 \end{CodeInput}
@@
 #>     Estimate Std. Error t value Pr(>|t|)    
-#> MSS   9.0121     0.0155     581  3.0e-06 ***
-#> TAU   1.3083     0.0066     198  2.5e-05 ***
+#> MSS   9.0121     0.0155     581 0.000003 ***
+#> TAU   1.3083     0.0066     198 0.000025 ***
 #> ---
@@
 #> Number of iterations to convergence: 4 
-#> Achieved convergence tolerance: 3.11e-06
-\end{CodeOutput}
-\end{CodeChunk}
-
-Once we have estimated MSS and TAU, we can use \texttt{predict\_} family of functions to predict various relationships (i.e.~time at distance, acceleration at distance, velocity at time):
+#> Achieved convergence tolerance: 0.00000311
+\end{CodeOutput}
+\end{CodeChunk}
+
+To extract estimated model parameters, use S3 \texttt{coef()} method:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> coef(m1)
+\end{CodeInput}
+\begin{CodeOutput}
+#>                 MSS                 TAU                 MAC 
+#>                9.01                1.31                6.89 
+#>                PMAX     time_correction distance_correction 
+#>               15.52                0.00                0.00
+\end{CodeOutput}
+\end{CodeChunk}
+
+To create a simple plot of the model, use S3 \texttt{plot()} method, which returns \pkg{ggplot2} \citep{R-ggplot2} object:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(m1) + theme_bw(8)
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1} \end{center}
+
+\end{CodeChunk}
+
+Once we have estimated MSS and TAU, we can use \texttt{predict\_XXX()} family of functions to predict various relationships (i.e., time at distance, acceleration at distance, velocity at time, etc.):
 
@@
 
-To estimate force production at distance or time (using \texttt{predict\_force\_at\_distance} and \texttt{predict\_force\_at\_time} functions), as well as power production (using \texttt{predict\_power\_at\_distance} and \texttt{predict\_power\_at\_time} functions), one needs to take into account the air resistance.
-Air resistance (N) is estimated using \texttt{get\_air\_resistance} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \citet{arsacModelingEnergetics100m2002}, \citet{samozinoSimpleMethodMeasuring2016}, and \citet{vaningenschenauCanCyclePower1991} for more information):
+To estimate force production at distance or time (using \texttt{predict\_force\_at\_distance()} and \texttt{predict\_force\_at\_time()} functions), as well as power production (using \texttt{predict\_power\_at\_distance()} and \texttt{predict\_power\_at\_time()} functions), one needs to take into account the air resistance.
+Air resistance (N) is estimated using \texttt{get\_air\_resistance()} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \citet{arsacModelingEnergetics100m2002}, \citet{samozinoSimpleMethodMeasuring2016}, and \citet{vaningenschenauCanCyclePower1991} for more information):
 
@@
 
-When estimating force and power, the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{get\_air\_resistance}:
+When estimating force and power, the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{get\_air\_resistance()}:
 
@@
 
-The easiest way to get all kinematics for short sprints is to use \texttt{predict\_kinematics} functions:
+The easiest way to get all kinematics and kinetics for short sprints is to use \texttt{predict\_kinematics()} function:
 
@@
 \begin{CodeOutput}
-#>   time distance velocity acceleration air_resistance force power
-#> 1 0.00 0.000000   0.0000         6.89        0.07536   551   0.0
-#> 2 0.01 0.000344   0.0686         6.84        0.05609   547  37.5
-#> 3 0.02 0.001371   0.1367         6.78        0.03978   543  74.2
-#> 4 0.03 0.003076   0.2043         6.73        0.02636   539 110.0
-#> 5 0.04 0.005455   0.2714         6.68        0.01576   534 145.0
-#> 6 0.05 0.008502   0.3379         6.63        0.00792   530 179.2
-#>   relative_power
-#> 1          0.000
-#> 2          0.469
-#> 3          0.928
-#> 4          1.375
-#> 5          1.813
-#> 6          2.241
+#>   time distance velocity acceleration bodymass net_horizontal_force
+#> 1 0.00 0.000000   0.0000         6.89       80                  551
+#> 2 0.01 0.000344   0.0686         6.84       80                  547
+#> 3 0.02 0.001371   0.1367         6.78       80                  543
+#> 4 0.03 0.003076   0.2043         6.73       80                  539
+#> 5 0.04 0.005455   0.2714         6.68       80                  534
+#> 6 0.05 0.008502   0.3379         6.63       80                  530
+#>   air_resistance horizontal_force horizontal_force_relative
+#> 1        0.07536              551                      6.89
+#> 2        0.05609              547                      6.84
+#> 3        0.03978              543                      6.78
+#> 4        0.02636              539                      6.73
+#> 5        0.01576              534                      6.68
+#> 6        0.00792              530                      6.63
+#>   vertical_force resultant_force resultant_force_relative power
+#> 1            785             959                     12.0   0.0
+#> 2            785             957                     12.0  37.5
+#> 3            785             954                     11.9  74.2
+#> 4            785             952                     11.9 110.0
+#> 5            785             950                     11.9 145.0
+#> 6            785             947                     11.8 179.2
+#>   relative_power    RF force_angle
+#> 1          0.000 0.575        54.9
+#> 2          0.469 0.572        55.1
+#> 3          0.928 0.569        55.3
+#> 4          1.375 0.566        55.5
+#> 5          1.813 0.563        55.7
+#> 6          2.241 0.560        55.9
 \end{CodeOutput}
@@
 R> 
-R> df <- pivot_longer(data = df, cols = -2)
+R> variable_names <- colnames(df)
+R> 
+R> df <- pivot_longer(data = df, cols = -2) %>%
++   mutate(name = factor(name, levels = variable_names))
 R> 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
-
-\end{CodeChunk}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
+
+\end{CodeChunk}
+
+These kinematic and kinetic variables are utilized in \protect\hyperlink{force-velocity-profile}{Force-Velocity profile} estimation, which is covered later in this paper.
 
@@
 Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90\% of maximum sprinting speed is reached, or where peak power is within 90\% range.
-To identify these values, \pkg{shorts} package comes with \texttt{find\_} family of functions:
+To identify these values, \pkg{shorts} package comes with \texttt{find\_XXX()} family of functions:
 
@@
 Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in \pkg{shorts} for split times, one can use \texttt{mixed\_model\_using\_splits} function.
+To perform mixed-effects models in \pkg{shorts} for split times, one can use \texttt{mixed\_model\_using\_splits()} function.
 To demonstrate this functionality, we load the \texttt{split\_times} dataset provided in the \pkg{shorts} package:
@@
 #> Estimated random model parameters
-#> ----------------------------------
+#> ---------------------------------
 #>     athlete  MSS   TAU  MAC PMAX time_correction distance_correction
@@
 
-Additional information about mixed-effects model performed using the \texttt{nlme} package \citep{R-nlme} can be obtained using \texttt{summary}:
+Additional information about mixed-effects model performed using the \texttt{nlme} package \citep{R-nlme} can be obtained using \texttt{summary()} function:
 
@@
 
+S3 method \texttt{coef()} when applied on mixed-model result will return both the fixed and random effects:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> coef(m2)
+\end{CodeInput}
+\begin{CodeOutput}
+#> $fixed
+#>                 MSS                 TAU                 MAC 
+#>               8.065               0.655              12.309 
+#>                PMAX     time_correction distance_correction 
+#>              24.818               0.000               0.000 
+#> 
+#> $random
+#>     athlete  MSS   TAU  MAC PMAX time_correction distance_correction
+#> 1     James 9.69 0.847 11.4 27.7               0                   0
+#> 2       Jim 7.83 0.505 15.5 30.4               0                   0
+#> 3      John 7.78 0.727 10.7 20.8               0                   0
+#> 4 Kimberley 8.57 0.802 10.7 22.9               0                   0
+#> 5  Samantha 6.45 0.395 16.3 26.4               0                   0
+\end{CodeOutput}
+\end{CodeChunk}
+
+To create a simple plot of the model, use S3 \texttt{plot()} method:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(m2) + theme_bw(8)
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-14-1} \end{center}
+
+\end{CodeChunk}
+
 The following figure contains kinematics for all athletes in \texttt{split\_times} dataset. Please note that power calculation takes default parameters for each individual:
@@
 R> 
-R> df <- pivot_longer(df, cols = c(-1, -3))
+R> variable_names <- colnames(df)
+R> 
+R> df <- pivot_longer(df, cols = c(-1, -3)) %>%
++   mutate(name = factor(name, levels = variable_names))
 R> 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-11-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
 
@@
 
-Let's consider the same example of an athlete with MSS equal to 9\(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92\(ms^-2\) performing 40m sprint with velocity estimated using radar run (in this case with 1Hz sampling rate).
+Let's consider the same example of an athlete with MSS equal to 9 \(ms^{-1}\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^{-2}\) performing 40m sprint with velocity estimated using radar run (in this case with 1 Hz sampling rate).
 
@@
 
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(m3) + theme_bw(8)
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
+
+\end{CodeChunk}
+
 Both split and radar gun models allow the use of \emph{weighted} non-linear regression.
@@
 
-Mixed-effects model using radar data is done using \texttt{mixed\_model\_using\_radar} function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
+Mixed-effects model using radar data is done using \texttt{mixed\_model\_using\_radar()} function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
 
@@
 #> Estimated random model parameters
-#> ----------------------------------
+#> ---------------------------------
 #>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
@@
 
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(m4) + theme_bw(8)
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
+
+\end{CodeChunk}
+
+\hypertarget{force-velocity-profile}{%
+\subsection{Force-Velocity profile}\label{force-velocity-profile}}
+
+To create \emph{Force-Velocity Profile} (FVP) using single athlete estimated sprint model parameters (i.e., TAU and MSS), you can use \texttt{get\_FV\_profile()} function.
+When estimating FVP, athlete body mass (kg) can be set using \texttt{bodymass} parameter, while the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{get\_air\_resistance()} function.
+Details of the FVP method implemented in the \pkg{shorts} package, as well as the interpretation from a sprint training perspective, are covered elsewhere \citep{haugenPowerForceVelocityProfilingSprinting2020, morinInterpretingPowerForceVelocityProfiles2016, morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # To create Force-Velocity Profile
+R> fvp <- get_FV_profile(
++   MSS = m1$parameters$MSS,
++   TAU = m1$parameters$TAU,
++   bodymass = 80,
++   # Additional parameters forwarded to get_air_resistance
++   # Otherwise, defaults are used
++   bodyheight = 1.85,
++   barometric_pressure = 780,
++   air_temperature = 20,
++   wind_velocity = 0.5
++ )
+R> 
+R> fvp
+\end{CodeInput}
+\begin{CodeOutput}
+#> Estimated Force-Velocity Profile
+#> --------------------------------
+#>      bodymass            F0        F0_rel            V0          Pmax 
+#>      80.00000     544.51032       6.80638       9.36184    1274.40402 
+#> Pmax_relative      FV_slope  RFmax_cutoff         RFmax           Drf 
+#>      15.93005      -0.72703       0.30000       0.48779      -0.06676 
+#>        RSE_FV       RSE_Drf 
+#>       1.54814       0.00447
+\end{CodeOutput}
+\end{CodeChunk}
+
+To plot FVP kinematics and kinetics (which are exactly the same as generated by the \texttt{predict\_kinematics()} function), use S3 \texttt{plot()} function.
+By default, FVP estimated kinetics are plotted against velocity (on x-axis).
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(fvp) + theme_bw(8)
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-22-1} \end{center}
+
+\end{CodeChunk}
+
+To plot FVP estimated kinetics against time, use \texttt{type\ =\ "time"} parameter:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> plot(fvp, "time") + theme_bw(8)
+\end{CodeInput}
+
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-23-1} \end{center}
+
+\end{CodeChunk}
+
 \hypertarget{problems-with-estimation}{%
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-24-1} \end{center}
 
@@
 
-As can be seen from the example, TAU is affected by an error in synchronization of time with velocity.
+As can be seen from the example, all estimated parameters are affected by an error in synchronization of time with velocity (with MSS being the least affected in this example).
 The potential solution incorporated into the \pkg{shorts} package involves estimation of the \emph{time correction} parameter using the following equation:
@@
 
-This model is incorporated in the \texttt{model\_using\_radar\_with\_time\_correction} function:
+This model is incorporated in the \texttt{model\_using\_radar\_with\_time\_correction()} function:
 
@@
 
-When using \texttt{predict\_} family of functions, one can provide estimated time correction to get predictions at original time scale.
+When using \texttt{predict\_XXX()} family of functions, one can provide estimated time correction to get predictions at original time scale.
 
@@
 \begin{CodeOutput}
-#> [1] 7.82e-05 4.83e+00 7.07e+00 8.11e+00 8.59e+00 8.81e+00 8.91e+00
+#> [1] 0.0000782 4.8299729 7.0681475 8.1053182 8.5859434 8.8086652
+#> [7] 8.9118746
 \end{CodeOutput}
@@
 
-When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the \texttt{mixed\_model\_using\_radar\_with\_time\_correction} function.
+When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the \texttt{mixed\_model\_using\_radar\_with\_time\_correction()} function.
 
@@
 
-Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9\(ms^-1\), TAU equal to 1.3, and MAC equal to 6.92\(ms^-2\).
+Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9 \(ms^{-1}\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^{-2}\).
 Let's call them John and Jack.
-They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40m.
-The initial timing gate at the start (i.e., \(d=0m\)) serves to activate the timing system (i.e., when they cross the beam).
+They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40 m.
+The initial timing gate at the start (i.e., \(d=0\) m) serves to activate the timing system (i.e., when they cross the beam).
 
 John represents the \emph{theoretical model}, in which we assume that the initial force production and the timing initiation are perfectly synchronized.
-Jack, on the other hand, represents a \emph{practical model}, and decides to move slightly behind the initial timing gate (i.e.~for 0.5m) and use body rocking to initiate the sprint start.
+Jack, on the other hand, represents a \emph{practical model}, and decides to move slightly behind the initial timing gate (i.e.~for 0.5 m) and use body rocking to initiate the sprint start.
 In other words, Jack is using a \emph{flying start}, a common scenario when testing field sports athletes.
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-21-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-30-1} \end{center}
 
@@
 R> sprint_parameters <- rbind(
-+   unlist(john_profile$parameters),
-+   unlist(jack_profile$parameters)
++   coef(john_profile),
++   coef(jack_profile)
 + )
@@
 
-Below is a simulation sprint with 5, 10, 20, 30, 40, and 50m splits, with MSS and MAC varying from 6 to 9 (\(ms^-1\) and \(ms^-2\) respectively), and flying start distance varying from 0 to 1m.
+Below is a simulation sprint with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 (\(ms^{-1}\) and \(ms^{-2}\) respectively), and flying start distance varying from 0 to 1 m.
 
@@
 R> set.seed(1667)
-R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
+R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
 R> sim_df$time <- sim_df$time + rand_noise
@@
 + 
-+   params <- data.frame(t(unlist(model$parameters)))
++   params <- data.frame(t(coef(model)))
 + 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-25-1} \end{center}
-
-\end{CodeChunk}
-
-As can be seen from the image, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-34-1} \end{center}
+
+\end{CodeChunk}
+
+As can be seen from the figure, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-26-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-27-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
 
@@
 +   model_df,
-+   aes(y = residuals, x = distance, color = flying_start_distance, group = group)
++   aes(
++     y = residuals,
++     x = distance,
++     color = flying_start_distance,
++     group = group)
 + ) +
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-28-1} \end{center}
-
-\end{CodeChunk}
-
-If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler images conveying issues with residuals when there is a flying start:
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
+
+\end{CodeChunk}
+
+If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler figure conveying issues with residuals when there is a flying start:
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-29-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
 
@@
 
-A potential solution is to use a correction factor - the recommendation in the literature is +0.5sec \citep{haugenSprintMechanicalProperties2020, haugenSprintMechanicalVariables2019}.
-Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 sec \citep{haugenDifferenceStartImpact2012}.
+A potential solution is to use a correction factor - the recommendation in the literature is +0.5 s \citep{haugenSprintMechanicalProperties2020, haugenSprintMechanicalVariables2019}.
+Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 s \citep{haugenDifferenceStartImpact2012}.
 So, while a timing correction factor is warranted to avoid subsequent errors in estimates of kinetic variables (e.g., overestimate power), a correction factor that is too large will have the opposite effect (e.g., underestimate power).
 
-When implementing \emph{time correction}, equation \eqref{eq:time-distance} becomes:
+Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
+Exactly the same method is suggested by \citet{stenrothForcevelocityProfilingIce2020}, named \emph{time shift method}, and the estimated parameter named \emph{time shift parameter}.
+We have named this parameter \emph{time correction} to be in agreement with the parameter introduced in \protect\hyperlink{problems-with-time-sync-with-radar-gun}{Problems with time sync with radar gun} section of this paper, as well as the available literature.
+
+When implementing time correction, equation \eqref{eq:time-distance} becomes:
 
@@
 
-Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
-To estimate time correction, we use \texttt{model\_using\_splits\_with\_time\_correction} function.
-Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5s) or estimated time correction:
+To estimate time correction parameter, we use \texttt{model\_using\_splits\_with\_time\_correction()} function.
+Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5 s) or estimated time correction:
 
@@
 R> jack_parameters <- rbind(
-+   unlist(john_profile$parameters),
-+   unlist(jack_profile$parameters),
-+   unlist(jack_profile_fixed_time_short$parameters),
-+   unlist(jack_profile_fixed_time_long$parameters),
-+   unlist(jack_profile_time_estimated$parameters)
++   coef(john_profile),
++   coef(jack_profile),
++   coef(jack_profile_fixed_time_short),
++   coef(jack_profile_fixed_time_long),
++   coef(jack_profile_time_estimated)
 + )
@@
 
-In Jack's case, both +0.3s fixed time correction and time correction estimation yield parameters closer to John's (i.e.~true parameters).
-
-Another model definition, which is a novel approach implemented in the \pkg{shorts} packages, is to utilize \emph{distance correction}, besides time correction.
+In Jack's case, both +0.3 s fixed time correction and time correction estimation yield parameters closer to John's (i.e.~true parameters).
+
+Another model definition, which is a novel approach implemented in the \pkg{shorts} package, is to utilize \emph{distance correction}, besides time correction.
 Thus, equation \eqref{eq:time-distance} becomes:
@@
 
-This model is implemented in \texttt{model\_using\_splits\_with\_corrections} function. Below are the model estimates:
+This model is implemented in \texttt{model\_using\_splits\_with\_corrections()} function. Below are the model estimates:
 
@@
 R> jack_parameters <- rbind(
-+   unlist(john_profile$parameters),
-+   unlist(jack_profile$parameters),
-+   unlist(jack_profile_fixed_time_short$parameters),
-+   unlist(jack_profile_fixed_time_long$parameters),
-+   unlist(jack_profile_time_estimated$parameters),
-+   unlist(jack_profile_distance_correction$parameters)
++   coef(john_profile),
++   coef(jack_profile),
++   coef(jack_profile_fixed_time_short),
++   coef(jack_profile_fixed_time_long),
++   coef(jack_profile_time_estimated),
++   coef(jack_profile_distance_correction)
 + )
@@
 This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.
+One strategy that is sometimes implemented is adding zeros to the sample (i.e., \(t=0\) and \(d=0\)), which increase the number of observations.
+Unfortunately, this strategy should not be implemented, as explained later in the \protect\hyperlink{should-we-add-zero-to-the-sample}{Should we add zero to the sample?} section of this paper.
 
@@
 +       model = "No correction",
-+       t(unlist(no_correction$parameters))
++       t(coef(no_correction))
 +     ),
@@
 +       model = "Fixed correction +0.3s",
-+       t(unlist(fixed_correction_short$parameters))
++       t(coef(fixed_correction_short))
 +     ),
@@
 +       model = "Fixed correction +0.5s",
-+       t(unlist(fixed_correction_long$parameters))
++       t(coef(fixed_correction_long))
 +     ),
@@
 +       model = "Time correction",
-+       t(unlist(time_correction$parameters))
++       t(coef(time_correction))
 +     ),
@@
 +       model = "Time and distance correction",
-+       t(unlist(time_dist_correction$parameters))
++       t(coef(time_dist_correction))
 +     )
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-33-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-34-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
 
@@
 
-Time correction and time and distance corrections are also implemented in the mixed-models using \texttt{mixed\_model\_using\_splits\_with\_time\_correction} and \texttt{mixed\_model\_using\_splits\_with\_corrections}.
-We will showcase their use at the end of this article.
+Time correction and time and distance corrections are also implemented in the mixed-models using \texttt{mixed\_model\_using\_splits\_with\_time\_correction()} and \texttt{mixed\_model\_using\_splits\_with\_corrections()}.
+We will showcase their use at the end of this paper.
 
@@
 
-Starting behind the initial timing gate represent only one issue (i.e.~flying start).
+Starting behind the initial timing gate represent only one issue (i.e., flying start).
 In this section, we simulate one more issue to check the sensitivity of the presented models to other (less common) perturbations when performing field testing.
@@
 One issue that might happen with timing gates is triggering the timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
-This is very similar to the situation when timing starts on a signal (i.e., gun during 100m sprint race) and there is \emph{reaction time} (RT) involved.
+This is very similar to the situation when timing starts on a signal (i.e., gun during 100 m sprint race) and there is \emph{reaction time} (RT) involved.
 Both of these scenarios represent \emph{time lag} that is added to the split times.
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
 
@@
 
+\hypertarget{should-we-add-zero-to-the-sample}{%
+\subsection{Should we add zero to the sample?}\label{should-we-add-zero-to-the-sample}}
+
+Fellow sports scientists are often considering adding \(t=0\) s at \(d=0\) m to the collected split times with the aim of increasing the number of observations.
+The question is whether this strategy is sound and if it should be employed.
+The short answer is no, it shouldn't, particularly if there are flying start issues.
+In the following example, we are demonstrating the issue when zeros are added to the sample when fitting multiple models with the known true parameters:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> # Create split times from known MSS and TAU
+R> df <- tibble(
++   distance = c(5, 10, 20, 30, 40),
++   time = predict_time_at_distance(
++     distance,
++     MSS = 9,
++     TAU = 1.3
++   )
++ ) %>%
++   mutate(
++     # Add random noise to time
++     time = time + rnorm(n(), 0, 10^-5),
++     gate_time = time - 0.1
++   )
+R> 
+R> # Model without time correction
+R> m_no_correction <- model_using_splits(
++   distance = df$distance, 
++   time = df$gate_time
++ )
+R> 
+R> # Model without time correction, but with zeros added
+R> m_no_correction_zero <- model_using_splits(
++   distance = c(0, df$distance), 
++   time = c(0, df$gate_time)
++ )
+R> 
+R> # Model without adding zeros for the start
+R> m_no_zero <- model_using_splits_with_time_correction(
++   distance = df$distance, 
++   time = df$gate_time
++ )
+R> 
+R> # Model with added zeros for the start (d=0 and t=0)
+R> m_with_zero <- model_using_splits_with_time_correction(
++   distance = c(0, df$distance), 
++   time = c(0, df$gate_time)
++ )
+R> 
+R> # Print results
+R> data.frame(
++   model = c(
++     "Without time correction",
++     "Without time correction with zeros added", 
++     "With time correction",
++     "With time correction with zeros added"),
++   rbind(
++     coef(m_no_correction),
++     coef(m_no_correction_zero),
++     coef(m_no_zero),
++     coef(m_with_zero))
++ )
+\end{CodeInput}
+\begin{CodeOutput}
+#>                                      model  MSS  TAU  MAC PMAX
+#> 1                  Without time correction 8.78 1.09 8.06 17.7
+#> 2 Without time correction with zeros added 8.78 1.09 8.06 17.7
+#> 3                     With time correction 9.00 1.30 6.92 15.6
+#> 4    With time correction with zeros added 8.79 1.10 7.97 17.5
+#>   time_correction distance_correction
+#> 1         0.00000                   0
+#> 2         0.00000                   0
+#> 3         0.09994                   0
+#> 4         0.00739                   0
+\end{CodeOutput}
+\end{CodeChunk}
+
+As can be seen from the output, only the model with time correction is able to correctly recover true sprint parameters, but adding zeros to the sample in this case, results in MSS and TAU estimates very close to the estimates of the model without time corrections.
+Thus, adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
+
 \hypertarget{leave-one-out-cross-validation}{%
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-48-1} \end{center}
-
-\end{CodeChunk}
-
-As expected, the model has more issues predicting unseen short or long sprints.
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-58-1} \end{center}
+
+\end{CodeChunk}
+
+As expected, the model has more issues predicting unseen split times for both short or long distances.
 Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-63-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-64-1} \end{center}
 
@@
 
-Error bars represent residual bias \(\pm\) 1SD.
+Error bars represent residual bias \(\pm\) 1 SD.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-55-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-65-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-56-1} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-66-1} \end{center}
 

### 2020-12-31 code.Rmd
JSS submission

```{r setup, include=FALSE}
require(knitr)
require(shorts)
require(tidyverse)
require(bookdown)

my_random_seed <- 1667
set.seed(my_random_seed)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
  dpi = 300,
  out.width = "100%",
  fig.align = "center",
  fig.width = 6,
  fig.height = 6 * 0.618, # 1 / phi
  fig.show = "hold",
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  width = 70
)

# automatically create a bib database for R packages
#knitr::write_bib(c(
#  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
#), "packages.bib")

# Set rounding
op <- options()

options(
  digits = 3,
  "width" = 70,
  scipen = 999)

```
```{r four-athletes-table, echo=FALSE}
athletes <- tribble(
  ~Athlete, ~MSS, ~MAC,
  "Athlete A", 12, 10,
  "Athlete B", 12, 6,
  "Athlete C", 8, 10,
  "Athlete D", 8, 6
)

athletes <- athletes %>%
  mutate(TAU = MSS / MAC)

knitr::kable(
  athletes,
  caption = "Four athletes with different MSS and MAC parameters.",
  digits = 2,
  booktabs = TRUE
)

```
```{r four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds."}
kinematics <- expand_grid(
  athletes,
  time = seq(0, 6, length.out = 1000)
) %>%
  mutate(
    distance = predict_distance_at_time(time, MSS, TAU),
    velocity = predict_velocity_at_time(time, MSS, TAU),
    acceleration = predict_acceleration_at_time(time, MSS, TAU)
  )

kinematics_long <- kinematics %>%
  pivot_longer(
    cols = c("distance", "velocity", "acceleration"),
    names_to = "variable"
  ) %>%
  mutate(
    variable = factor(
      variable,
      levels = c("distance", "velocity", "acceleration"),
      labels = c(
        expression("Distance (m" * ")"),
        expression("Velocity (ms"^-1 * ")"),
        expression("Acceleration (ms"^-2 * ")")
      )
    )
  )

gg <- ggplot(
  kinematics_long,
  aes(x = time, y = value, color = Athlete)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_wrap(
    ~variable,
    scales = "free_y",
    labeller = label_parsed
  ) +
  ylab(NULL) +
  xlab("Time (s)") +
  theme(
    legend.position = "top",
    legend.title = element_blank())

plot(gg)

```
```{r four-athletes-profile, echo=FALSE, fig.cap="Acceleration-Velocity profile of four athletes with different MSS and MAC parameters."}
gg <- ggplot(
  kinematics,
  aes(x = velocity, y = acceleration, color = Athlete)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  ylab(expression("Acceleration (ms"^-2 * ")")) +
  xlab(expression("Velocity (ms"^-1 * ")")) +
  theme(
    legend.position = "top",
    legend.title = element_blank())

plot(gg)

```
```{r }
require(shorts)

split_distance <- c(10, 20, 30, 40)

split_time <- c(2.17, 3.43, 4.60, 5.73)

m1 <- model_using_splits(
  distance = split_distance,
  time = split_time
)

m1

```
```{r }
summary(m1)

```
```{r }
coef(m1)

```
```{r }
plot(m1) + theme_bw(8)

```
```{r }
# Predict time at distance
predict_time_at_distance(
  distance = split_distance,
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU
)

# Predict acceleration at time
predict_acceleration_at_time(
  time = c(0, 1, 2, 3, 4, 5, 6),
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU
)

```
```{r }
get_air_resistance(
  velocity = 5,
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

```
```{r }
# To calculate horizontal force produced
predict_force_at_distance(
  distance = split_distance,
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

# To calculate power produced
predict_power_at_distance(
  distance = split_distance,
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

```
```{r }
df <- predict_kinematics(
  m1,
  max_time = 6,
  frequency = 100,
  # Additional parameters forwarded to get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

head(df)

```
```{r }
require(tidyverse)

variable_names <- colnames(df)

df <- pivot_longer(data = df, cols = -2) %>%
  mutate(name = factor(name, levels = variable_names))

ggplot(df, aes(x = distance, y = value)) +
  theme_bw(8) +
  facet_wrap(~name, scales = "free_y") +
  geom_line(alpha = 0.7) +
  ylab(NULL) +
  xlab("Distance (m)")

```
```{r }
# Finds distance where 90% of maximum sprinting speed is reached
find_velocity_critical_distance(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  percent = 0.9
)

# Finds maximal power and distance (this time using air resistance)
find_max_power_distance(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

# Finds distance over 90% power range
find_power_critical_distance(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  # Additional parameters forwarded to get_air_resistance
  # Otherwise, defaults are used
  bodymass = 80,
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

```
```{r }
data(split_times)

# Mixed model
m2 <- mixed_model_using_splits(
  data = split_times,
  distance = "distance",
  time = "time",
  athlete = "athlete",

  # Select random effects
  # Default is MSS and TAU
  random = MSS + TAU ~ 1
)

m2

```
```{r }
summary(m2)

```
```{r }
coef(m2)

```
```{r }
plot(m2) + theme_bw(8)

```
```{r }
df <- predict_kinematics(m2, max_time = 10)

variable_names <- colnames(df)

df <- pivot_longer(df, cols = c(-1, -3)) %>%
  mutate(name = factor(name, levels = variable_names))

ggplot(
  filter(df, distance < 40),
  aes(x = distance, y = value, group = athlete, color = athlete)
) +
  theme_bw(8) +
  facet_wrap(~name, scales = "free_y") +
  geom_line(alpha = 0.7) +
  ylab(NULL) +
  xlab("Distance (m)") +
  theme(
    legend.position = "top",
    legend.title = element_blank())

```
```{r }
sprint_time <- seq(0, 6, 1)

sprint_velocity <- c(0.00, 4.83, 7.07, 8.10, 8.59, 8.81, 8.91)

m3 <- model_using_radar(
  velocity = sprint_velocity,
  time = sprint_time
)

m3

```
```{r }
plot(m3) + theme_bw(8)

```
```{r }
m3_weighted <- model_using_radar(
  velocity = sprint_velocity,
  time = sprint_time,
  weights = 1 / (sprint_velocity + 1)
)

m3_weighted

```
```{r }
data("radar_gun_data")

m4 <- mixed_model_using_radar(
  radar_gun_data,
  time = "time",
  velocity = "velocity",
  athlete = "athlete"
)

m4

```
```{r }
plot(m4) + theme_bw(8)

```
```{r }
# To create Force-Velocity Profile
fvp <- get_FV_profile(
  MSS = m1$parameters$MSS,
  TAU = m1$parameters$TAU,
  bodymass = 80,
  # Additional parameters forwarded to get_air_resistance
  # Otherwise, defaults are used
  bodyheight = 1.85,
  barometric_pressure = 780,
  air_temperature = 20,
  wind_velocity = 0.5
)

fvp

```
```{r }
plot(fvp) + theme_bw(8)

```
```{r }
plot(fvp, "time") + theme_bw(8)

```
```{r }
df <- tibble(
  `true time` = sprint_time,
  velocity = sprint_velocity,
  `0.5s added` = `true time` + 0.5,
  `0.5s deducted` = `true time` - 0.5
)

plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")

ggplot(
  plot_df,
  aes(x = value, y = velocity, color = `Sync issue`)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  xlab("Time (s)") +
  ylab(expression("Velocity (" * ms^-1 * ")")) +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# Without synchronization issues
m5 <- model_using_radar(
  velocity = df$velocity,
  time = df$`true time`
)

# With time added
m6 <- model_using_radar(
  velocity = df$velocity,
  time = df$`0.5s added`
)

# With time deducted
m7 <- model_using_radar(
  velocity = df$velocity,
  time = df$`0.5s deducted`
)

rbind(
  data.frame(
    model = "True time",
    t(coef(m5))
  ),
  data.frame(
    model = "Added 0.5s time",
    t(coef(m6))
  ),
  data.frame(
    model = "Deducted 0.5s time",
    t(coef(m7))
  )
)

```
```{r }
# With time added
m8 <- model_using_radar_with_time_correction(
  velocity = df$velocity,
  time = df$`0.5s added`
)
coef(m8)

# With time deducted
m9 <- model_using_radar_with_time_correction(
  velocity = df$velocity,
  time = df$`0.5s deducted`
)
coef(m9)

```
```{r }
# Using the true time
predict_velocity_at_time(
  time = df$`true time`,
  MSS = m5$parameters$MSS,
  TAU = m5$parameters$TAU
)

# Using time with sync issues
predict_velocity_at_time(
  time = df$`0.5s added`,
  MSS = m8$parameters$MSS,
  TAU = m8$parameters$TAU,
  time_correction = m8$parameters$time_correction
)

```
```{r }
# Adding 0.5s to radar_gun_data
radar_gun_data$time <- radar_gun_data$time + 0.5

# Mixed model with time correction being fixed effect
m10 <- mixed_model_using_radar_with_time_correction(
  radar_gun_data,
  time = "time",
  velocity = "velocity",
  athlete = "athlete",
  random = MSS + TAU ~ 1
)

m10

# Mixed model with time correction being random effect
m11 <- mixed_model_using_radar_with_time_correction(
  radar_gun_data,
  time = "time",
  velocity = "velocity",
  athlete = "athlete",
  random = MSS + TAU + time_correction ~ 1
)

m11

```
```{r }
MSS <- 9
TAU <- 1.3
MAC <- MSS / TAU

split_times <- tibble(
  distance = c(5, 10, 20, 30, 40),
  john_time = predict_time_at_distance(distance, MSS, TAU),

  # Jack's performance
  jack_distance = distance + 0.5,
  jack_true_time = predict_time_at_distance(jack_distance, MSS, TAU),
  time_05m = predict_time_at_distance(0.5, MSS, TAU),
  jack_time = jack_true_time - time_05m
)

split_times

```
```{r }
plot_df <- split_times %>%
  select(distance, john_time, jack_time) %>%
  rename(John = john_time, Jack = jack_time) %>%
  pivot_longer(cols = -1, names_to = "athlete", values_to = "time") %>%
  mutate(distance = factor(distance))

ggplot(
  plot_df,
  aes(x = distance, y = time, color = athlete, group = athlete)
) +
  theme_bw(8) +
  geom_point() +
  geom_line() +
  xlab("Distance (m)") +
  ylab("Time (s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# Since this is a perfect simulation and stats::nls will complain
# we need to add very small noise, or measurement error to the times
set.seed(1667)
rand_noise <- rnorm(nrow(split_times), 0, 10^-5)
split_times$john_time <- split_times$john_time + rand_noise
split_times$jack_time <- split_times$jack_time + rand_noise

john_profile <- model_using_splits(
  distance = split_times$distance,
  time = split_times$john_time
)

jack_profile <- model_using_splits(
  distance = split_times$distance,
  time = split_times$jack_time
)

sprint_parameters <- rbind(
  coef(john_profile),
  coef(jack_profile)
)

rownames(sprint_parameters) <- c("John", "Jack")

sprint_parameters

```
```{r }
sim_df <- expand.grid(
  MSS = c(6, 7, 8, 9),
  MAC = c(6, 7, 8, 9),
  flying_start_distance = c(
    seq(0, 0.001, length.out = 20),
    seq(0.001, 0.01, length.out = 20),
    seq(0.01, 0.1, length.out = 20),
    seq(0.1, 1, length.out = 20)
  ),
  distance = c(5, 10, 20, 30, 40, 50)
)

sim_df <- sim_df %>%
  mutate(
    TAU = MSS / MAC,
    PMAX = MSS * MAC / 4,
    true_distance = distance + flying_start_distance,
    true_time = predict_time_at_distance(true_distance, MSS, TAU),
    stolen_time = predict_time_at_distance(flying_start_distance, MSS, TAU),
    time = true_time - stolen_time
  )

# Add small noise to allow model fit
set.seed(1667)
rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
sim_df$time <- sim_df$time + rand_noise

```
```{r }
# Prediction wrapper
pred_wrapper <- function(data) {
  model <- model_using_splits(
    distance = data$distance,
    time = data$time
  )

  params <- data.frame(t(coef(model)))

  predicted_time <- predict_time_at_distance(
    distance = data$distance,
    MSS = model$parameters$MSS,
    TAU = model$parameters$TAU
  )

  colnames(params) <- c(
    "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
    "est_time_correction", "est_distance_correction"
  )

  cbind(
    data,
    params,
    data.frame(predicted_time = as.numeric(predicted_time))
  )
}

# estimated parameters and predicted time
model_df <- sim_df %>%
  group_by(MSS, TAU, flying_start_distance) %>%
  do(pred_wrapper(.)) %>%
  ungroup()

# Prediction residuals
model_df$residuals <- model_df$predicted_time - model_df$time

```
```{r }
# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, flying_start_distance) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(
  df,
  aes(x = flying_start_distance, y = est_MSS, color = MAC_string)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_wrap(~MSS_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated MSS (m/s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# MAC
ggplot(
  df,
  aes(x = flying_start_distance, y = est_MAC, color = MSS_string)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_wrap(~MAC_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated MAC (m/s/s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# PMAX
ggplot(
  df,
  aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_wrap(~MAC_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated PMAX (W/kg)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, flying_start_distance)
  )

ggplot(
  model_df,
  aes(
    y = residuals,
    x = distance,
    color = flying_start_distance,
    group = group)
) +
  theme_bw(8) +
  geom_line(alpha = 0.3) +
  facet_grid(MSS_string ~ MAC_string) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (s)") +
  theme(legend.position = "top") +
  labs(color = "Flying start distance")

```
```{r }
ggplot(
  model_df,
  aes(
    y = residuals,
    x = distance,
    color = flying_start_distance,
    group = group)
) +
  theme_bw(8) +
  geom_line(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (s)") +
  theme(legend.position = "top") +
  labs(color = "Flying start distance")

```
```{r }
jack_profile_fixed_time_short <- model_using_splits(
  distance = split_times$distance,
  time = split_times$jack_time,
  time_correction = 0.3
)

jack_profile_fixed_time_long <- model_using_splits(
  distance = split_times$distance,
  time = split_times$jack_time,
  time_correction = 0.5
)

jack_profile_time_estimated <- model_using_splits_with_time_correction(
  distance = split_times$distance,
  time = split_times$jack_time
)

jack_parameters <- rbind(
  coef(john_profile),
  coef(jack_profile),
  coef(jack_profile_fixed_time_short),
  coef(jack_profile_fixed_time_long),
  coef(jack_profile_time_estimated)
)

rownames(jack_parameters) <- c(
  "John",
  "Jack - No corrections",
  "Jack - Fixed time correction (+0.3s)",
  "Jack - Fixed time correction (+0.5s)",
  "Jack - Estimated time correction"
)

jack_parameters

```
```{r }
jack_profile_distance_correction <- model_using_splits_with_corrections(
  distance = split_times$distance,
  time = split_times$jack_time
)

jack_parameters <- rbind(
  coef(john_profile),
  coef(jack_profile),
  coef(jack_profile_fixed_time_short),
  coef(jack_profile_fixed_time_long),
  coef(jack_profile_time_estimated),
  coef(jack_profile_distance_correction)
)

rownames(jack_parameters) <- c(
  "John",
  "Jack - No corrections",
  "Jack - Fixed time correction (+0.3s)",
  "Jack - Fixed time correction (+0.5s)",
  "Jack - Estimated time correction",
  "Jack - Estimated distance correction"
)

jack_parameters

```
```{r }
pred_wrapper <- function(data) {
  no_correction <- model_using_splits(
    distance = data$distance,
    time = data$time
  )

  fixed_correction_short <- model_using_splits(
    distance = data$distance,
    time = data$time,
    time_correction = 0.3
  )

  fixed_correction_long <- model_using_splits(
    distance = data$distance,
    time = data$time,
    time_correction = 0.5
  )

  time_correction <- model_using_splits_with_time_correction(
    distance = data$distance,
    time = data$time,
    control = nls.control(tol = 1)
  )

  time_dist_correction <- model_using_splits_with_corrections(
    distance = data$distance,
    time = data$time,
    control = nls.control(tol = 1)
  )

  params <- rbind(
    data.frame(
      model = "No correction",
      t(coef(no_correction))
    ),
    data.frame(
      model = "Fixed correction +0.3s",
      t(coef(fixed_correction_short))
    ),
    data.frame(
      model = "Fixed correction +0.5s",
      t(coef(fixed_correction_long))
    ),
    data.frame(
      model = "Time correction",
      t(coef(time_correction))
    ),
    data.frame(
      model = "Time and distance correction",
      t(coef(time_dist_correction))
    )
  )

  colnames(params) <- c(
    "model", "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
    "est_time_correction", "est_distance_correction"
  )

  df <- expand_grid(
    data,
    params
  )

  df$predicted_time <- predict_time_at_distance(
    distance = df$distance,
    MSS = df$est_MSS,
    TAU = df$est_TAU,
    time_correction = df$est_time_correction,
    distance_correction = df$est_distance_correction
  )

  df$residuals <- df$predicted_time - df$time
  return(df)
}

# estimated parameters and predicted time
model_df <- sim_df %>%
  group_by(MSS, TAU, flying_start_distance) %>%
  do(pred_wrapper(.)) %>%
  ungroup()

```
```{r }
model_df$model <- factor(
  model_df$model,
  levels = c(
    "No correction",
    "Fixed correction +0.3s",
    "Fixed correction +0.5s",
    "Time correction",
    "Time and distance correction"
  )
)
# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, flying_start_distance, model) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(
  df,
  aes(x = flying_start_distance, y = est_MSS, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated MSS (m/s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# MAC
ggplot(
  df,
  aes(x = flying_start_distance, y = est_MAC, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated MAC (m/s/s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# PMAX
ggplot(
  df,
  aes(x = flying_start_distance, y = est_PMAX, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated PMAX (W/kg)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# time_correction
ggplot(
  df,
  aes(x = flying_start_distance, y = est_time_correction, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_line(aes(y = stolen_time), color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated time correction (s)")  +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# distance_correction
ggplot(
  df,
  aes(x = flying_start_distance, y = est_distance_correction, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_abline(slope = 1, color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Flying start distance (m)") +
  ylab("estimated distance correction (m)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, flying_start_distance)
  )

ggplot(
  model_df,
  aes(
    y = residuals,
    x = distance,
    color = flying_start_distance,
    group = group)
) +
  theme_bw(8) +
  geom_line(alpha = 0.3) +
  facet_wrap(~model) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (s)") +
  theme(legend.position = "top") +
  labs(color = "Flying start distance")

```
```{r }
sim_df <- expand.grid(
  MSS = c(6, 7, 8, 9),
  MAC = c(6, 7, 8, 9),
  time_lag = seq(0, 0.5, length.out = 50),
  distance = c(5, 10, 20, 30, 40, 50)
)

sim_df <- sim_df %>%
  mutate(
    TAU = MSS / MAC,
    PMAX = MSS * MAC / 4,
    true_time = predict_time_at_distance(distance, MSS, TAU),
    time = true_time + time_lag
  )

# Add small noise to allow model fit
set.seed(1667)
rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
sim_df$time <- sim_df$time + rand_noise

```
```{r }
# estimated parameters and predicted time
model_df <- sim_df %>%
  group_by(MSS, TAU, time_lag) %>%
  do(pred_wrapper(.)) %>%
  ungroup()

```
```{r }
model_df$model <- factor(
  model_df$model,
  levels = c(
    "No correction",
    "Fixed correction +0.3s",
    "Fixed correction +0.5s",
    "Time correction",
    "Time and distance correction"
  )
)
# Estimates plot
df <- model_df %>%
  group_by(MSS, TAU, time_lag, model) %>%
  slice(1) %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2))
  )

# MSS
ggplot(df, aes(x = time_lag, y = est_MSS, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
  xlab("Time lag (s)") +
  ylab("estimated MSS (m/s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# MAC
ggplot(df, aes(x = time_lag, y = est_MAC, color = model)) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Time lag (s)") +
  ylab("estimated MAC (m/s/s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# time_correction
ggplot(
  df,
  aes(x = time_lag, y = est_time_correction, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  geom_abline(slope = -1, color = "black", linetype = "dashed") +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Time lag (s)") +
  ylab("estimated time correction (s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# distance_correction
ggplot(
  df,
  aes(x = time_lag, y = est_distance_correction, color = model)
) +
  theme_bw(8) +
  geom_line(alpha = 0.7) +
  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
  xlab("Time lag (s)") +
  ylab("estimated distance correction (m)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
# Residuals
model_df <- model_df %>%
  mutate(
    MSS_string = paste("MSS =", MSS),
    TAU_string = paste("TAU =", TAU),
    MAC_string = paste("MAC = ", round(MAC, 2)),
    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
    group = paste(MSS, MAC, time_lag)
  )

ggplot(
  model_df,
  aes(y = residuals, x = distance, color = time_lag, group = group)
) +
  theme_bw(8) +
  geom_line(alpha = 0.3) +
  facet_wrap(~model) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
  xlab("Distance (m)") +
  ylab("Predicted time - observed time (s)")  +
  theme(legend.position = "top") +
  labs(color = "Time lag")

```
```{r }
# Create split times from known MSS and TAU
df <- tibble(
  distance = c(5, 10, 20, 30, 40),
  time = predict_time_at_distance(
    distance,
    MSS = 9,
    TAU = 1.3
  )
) %>%
  mutate(
    # Add random noise to time
    time = time + rnorm(n(), 0, 10^-5),
    gate_time = time - 0.1
  )

# Model without time correction
m_no_correction <- model_using_splits(
  distance = df$distance,
  time = df$gate_time
)

# Model without time correction, but with zeros added
m_no_correction_zero <- model_using_splits(
  distance = c(0, df$distance),
  time = c(0, df$gate_time)
)

# Model without adding zeros for the start
m_no_zero <- model_using_splits_with_time_correction(
  distance = df$distance,
  time = df$gate_time
)

# Model with added zeros for the start (d=0 and t=0)
m_with_zero <- model_using_splits_with_time_correction(
  distance = c(0, df$distance),
  time = c(0, df$gate_time)
)

# Print results
data.frame(
  model = c(
    "Without time correction",
    "Without time correction with zeros added",
    "With time correction",
    "With time correction with zeros added"),
  rbind(
    coef(m_no_correction),
    coef(m_no_correction_zero),
    coef(m_no_zero),
    coef(m_with_zero))
)

```
```{r }
jack_LOOCV <- model_using_splits_with_time_correction(
  distance = split_times$distance,
  time = split_times$jack_time,
  LOOCV = TRUE
)

jack_LOOCV

```
```{r }
df <- jack_LOOCV$LOOCV$parameters

df <- pivot_longer(df, cols = 1:6, names_to = "parameter")

df$parameter <- factor(
  df$parameter,
  levels = c(
    "MSS",
    "TAU",
    "MAC",
    "PMAX",
    "time_correction",
    "distance_correction"
  )
)

ggplot(df, aes(x = value)) +
  theme_bw(8) +
  geom_boxplot() +
  facet_wrap(~parameter, scales = "free_x") +
  xlab(NULL) +
  ylab(NULL) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )

```
```{r }
df <- data.frame(
  distance = jack_LOOCV$data$distance,
  time = jack_LOOCV$data$time,
  pred_time = jack_LOOCV$data$pred_time,
  LOOCV_time = jack_LOOCV$LOOCV$data$pred_time
)

df <- df %>%
  pivot_longer(cols = c("pred_time", "LOOCV_time"))

df$resid <- df$value - df$time

ggplot(df, aes(x = distance, y = resid, color = name)) +
  theme_bw(8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point() +
  theme(legend.title = element_blank()) +
  xlab("Distance (m)") +
  ylab("Predicted - observed time (s)") +
  theme(
    legend.title = element_blank(),
    legend.position = "top")

```
```{r }
bolt_reaction_time <- 0.183

bolt_distance <- c(10, 20, 30, 40, 50, 60)
bolt_time <- c(1.963, 2.983, 3.883, 4.763, 5.643, 6.493)

# No corrections model
bolt_m1 <- model_using_splits(
  distance = bolt_distance,
  time = bolt_time
)

# Model with reaction time as fixed time correction
bolt_m2 <- model_using_splits(
  distance = bolt_distance,
  time = bolt_time,
  time_correction = -bolt_reaction_time
)

# Model with estimated time correction
bolt_m3 <- model_using_splits_with_time_correction(
  distance = bolt_distance,
  time = bolt_time
)

# Model with estimated time correction, but deducted reaction time
bolt_m4 <- model_using_splits_with_time_correction(
  distance = bolt_distance,
  time = bolt_time - bolt_reaction_time
)

# Model with estimated time and distance corrections
bolt_m5 <- model_using_splits_with_corrections(
  distance = bolt_distance,
  time = bolt_time
)

# Model with estimated time and distance corrections and deducted reaction time
bolt_m6 <- model_using_splits_with_corrections(
  distance = bolt_distance,
  time = bolt_time - bolt_reaction_time
)

bolt_model <- rbind(
  data.frame(
    model = "No correction",
    t(coef(bolt_m1))
  ),
  data.frame(
    model = "No correction - RT",
    t(coef(bolt_m2))
  ),
  data.frame(
    model = "Time correction",
    t(coef(bolt_m3))
  ),
  data.frame(
    model = "Time correction - RT",
    t(coef(bolt_m4))
  ),
  data.frame(
    model = "Distance correction",
    t(coef(bolt_m5))
  ),
  data.frame(
    model = "Distance correction - RT",
    t(coef(bolt_m6))
  )
)

bolt_model

```
```{r }
bolt_model <- bolt_model %>%
  group_by(model) %>%
  mutate(
    dist_99_MSS = find_velocity_critical_distance(
      MSS = MSS, TAU = TAU,
      #time_correction = time_correction,
      #distance_correction = distance_correction,
      percent = 0.99
    ),
   time_99_MSS = find_velocity_critical_time(
      MSS = MSS, TAU = TAU,
      #time_correction = time_correction,
      percent = 0.99
    )
  )

bolt_model[c(1, 8, 9)]

```
```{r }
data("vescovi")

# Convert data to long
df <- vescovi %>%
  select(1:13) %>%
  # slice(1:10) %>%
  pivot_longer(
    cols = 9:13,
    names_to = "distance",
    values_to = "time"
  ) %>%
  mutate(
    distance = as.numeric(str_extract(distance, "^[0-9]+"))
  )

```
```{r }
no_corrections <- mixed_model_using_splits(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete"
)

fixed_correction <- mixed_model_using_splits(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  time_correction = 0.3
)

time_correction_fixed <- mixed_model_using_splits_with_time_correction(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU ~ 1
)

time_correction_random <- mixed_model_using_splits_with_time_correction(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU + time_correction ~ 1
)

time_distance_correction_fixed <- mixed_model_using_splits_with_corrections(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU + time_correction ~ 1
)

time_distance_correction_random <- mixed_model_using_splits_with_corrections(
  df,
  distance = "distance",
  time = "time",
  athlete = "Athlete",
  random = MSS + TAU + time_correction + distance_correction ~ 1
)

```
```{r }
model_fit <- rbind(
  data.frame(
    model = "No corrections",
    t(unlist(no_corrections$model_fit))
  ),
  data.frame(
    model = "Fixed correction +0.3s",
    t(unlist(fixed_correction$model_fit))
  ),
  data.frame(
    model = "Time correction fixed",
    t(unlist(time_correction_fixed$model_fit))
  ),
  data.frame(
    model = "Time correction random",
    t(unlist(time_correction_random$model_fit))
  ),
  data.frame(
    model = "Time and distance correction fixed",
    t(unlist(time_distance_correction_fixed$model_fit))
  ),
  data.frame(
    model = "Time and distance correction random",
    t(unlist(time_distance_correction_random$model_fit))
  )
)

model_fit$model <- factor(
  model_fit$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction +0.3s",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

ggplot(model_fit, aes(x = RSE, y = model)) +
  theme_bw(8) +
  geom_point() +
  xlab("RSE (s)") +
  ylab(NULL)

```
```{r }
est_params <- rbind(
  data.frame(
    model = "No corrections",
    no_corrections$parameters$random
  ),
  data.frame(
    model = "Fixed correction +0.3s",
    fixed_correction$parameters$random
  ),
  data.frame(
    model = "Time correction fixed",
    time_correction_fixed$parameters$random
  ),
  data.frame(
    model = "Time correction random",
    time_correction_random$parameters$random
  ),
  data.frame(
    model = "Time and distance correction fixed",
    time_distance_correction_fixed$parameters$random
  ),
  data.frame(
    model = "Time and distance correction random",
    time_distance_correction_random$parameters$random
  )
)

est_params$model <- factor(
  est_params$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction +0.3s",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

est_params <- est_params %>%
  pivot_longer(cols = -(1:2), names_to = "parameter")

est_params$parameter <- factor(
  est_params$parameter,
  levels = c(
    "MSS",
    "TAU",
    "MAC",
    "PMAX",
    "time_correction",
    "distance_correction"
  )
)

ggplot(est_params, aes(y = model, x = value)) +
  theme_bw(8) +
  geom_boxplot() +
  facet_wrap(~parameter, scales = "free_x") +
  xlab(NULL) +
  ylab(NULL)

```
```{r }
model_resid <- rbind(
  data.frame(
    model = "No corrections",
    no_corrections$data
  ),
  data.frame(
    model = "Fixed correction +0.3s",
    fixed_correction$data
  ),
  data.frame(
    model = "Time correction fixed",
    time_correction_fixed$data
  ),
  data.frame(
    model = "Time correction random",
    time_correction_random$data
  ),
  data.frame(
    model = "Time and distance correction fixed",
    time_distance_correction_fixed$data
  ),
  data.frame(
    model = "Time and distance correction random",
    time_distance_correction_random$data
  )
)

model_resid$model <- factor(
  model_resid$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction +0.3s",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

model_resid$resid <- model_resid$pred_time - model_resid$time

# Create SWC / SESOI band
model_SESOI <- model_resid %>%
  group_by(model, distance) %>%
  summarise(
    bias = mean(resid),
    variance = sd(resid),
    upper = bias + variance,
    lower = bias - variance,
    MAD = mean(abs(resid)),
    SESOI_upper = sd(time) * 0.2,
    SESOI_lower = -sd(time) * 0.2
  )

# Plot
ggplot(model_resid, aes(y = model)) +
  theme_bw(8) +
  geom_vline(
    data = model_SESOI,
    aes(xintercept = SESOI_lower),
    color = "blue", alpha = 0.5, linetype = "dashed"
  ) +
  geom_vline(
    data = model_SESOI,
    aes(xintercept = SESOI_upper),
    color = "blue", alpha = 0.5, linetype = "dashed"
  ) +
  geom_vline(xintercept = 0, color = "blue", alpha = 0.5) +
  geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
  geom_errorbarh(
    data = model_SESOI,
    aes(xmin = lower, xmax = upper),
    height = 0.1, color = "black"
  ) +
  geom_point(data = model_SESOI, aes(x = bias), color = "black") +
  facet_wrap(~distance, scales = "free_x") +
  xlab("Predicted time - observed time (s)") +
  ylab(NULL)

```
```{r }
df <- model_SESOI %>%
  pivot_longer(cols = -(1:2), names_to = "estimator") %>%
  filter(estimator %in% c("bias", "variance", "MAD"))

df$model <- factor(
  df$model,
  levels = rev(c(
    "No corrections",
    "Fixed correction +0.3s",
    "Time correction fixed",
    "Time correction random",
    "Time and distance correction fixed",
    "Time and distance correction random"
  ))
)

df$estimator <- factor(
  df$estimator,
  levels = c("bias", "variance", "MAD")
)

ggplot(df, aes(x = value, y = model)) +
  theme_bw(8) +
  geom_point() +
  facet_grid(distance ~ estimator, scales = "free_x") +
  xlab(NULL) +
  ylab(NULL)

```
```{r include=FALSE}
# Return original options
options(op)

sessionInfo()
```

### 2020-12-31 paper.Rmd
JSS submission

@@
 require(bookdown)
-require(magick)
-
-pdf2png <- function(path) {
-  # only do the conversion for non-LaTeX output
-  if (knitr::is_latex_output()) 
-    return(path)
-  path2 <- xfun::with_ext(path, "png")
-  img <- magick::image_read_pdf(path)
-  magick::image_write(img, path2, format = "png")
-  path2
-}
 
@@
   fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
-  dpi = 600,
-  dev = "png",
-  fig.process = pdf2png,
+  dpi = 300,
   out.width = "100%",

### 2020-12-31 code.R

@@ -1,4 +1,4 @@
-## ----setup, include=FALSE-----------------------------------------------------------------------------------
+## ----setup, include=FALSE-------------------------------------------
 require(knitr)
 require(shorts)
 require(tidyverse)
@@ -6,8 +6,6 @@ require(bookdown)
 
 my_random_seed <- 1667
 set.seed(my_random_seed)
-
-options("width" = 70)
 
 knitr::opts_chunk$set(
   comment = "#>",
@@ -34,10 +32,14 @@ knitr::opts_chunk$set(
 
 # Set rounding
 op <- options()
-options(digits = 3)
+
+options(
+  digits = 3,
+  "width" = 70,
+  scipen = 999)
 
 
-## ----four-athletes-table, echo=FALSE------------------------------------------------------------------------
+## ----four-athletes-table, echo=FALSE--------------------------------
 athletes <- tribble(
   ~Athlete, ~MSS, ~MAC,
   "Athlete A", 12, 10,
@@ -52,11 +54,12 @@ athletes <- athletes %>%
 knitr::kable(
   athletes,
   caption = "Four athletes with different MSS and MAC parameters.",
-  digits = 2
+  digits = 2,
+  booktabs = TRUE
 )
 
 
-## ----four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6seconds."----
+## ----four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds."----
 kinematics <- expand_grid(
   athletes,
   time = seq(0, 6, length.out = 1000)
@@ -120,7 +123,7 @@ gg <- ggplot(
 plot(gg)
 
 
-## -----------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
 require(shorts)
 
 split_distance <- c(10, 20, 30, 40)
@@ -135,11 +138,19 @@ m1 <- model_using_splits(
 m1
 
 
-## -----------------------------------------------------------------------------------------------------------
-summary(m1$model)
+## -------------------------------------------------------------------
+summary(m1)
 
 
-## -----------------------------------------------------------------------------------------------------------
+## -------------------------------------------------------------------
+coef(m1)
+
+
+## -------------------------------------------------------------------
+plot(m1) + theme_bw(8)
+
+
[…]

### 2021-04-04 paper.Rmd
PeerJ version of the manuscript

@@
 ---
-documentclass: jss
+title: "shorts: An R Package for Modeling Short Sprints"
+preprint: false
+date: "`r Sys.Date()`"
 author: 
   - name: Mladen Jovanović
-    affiliation: |
-      | University of Belgrade
-    address: |
-      | Faculty of Sport and Physical Education
-      | Belgrade, Serbia
-    email: \email{coach.mladen.jovanovic@gmail.com}
+    affiliation: 1
+    corresponding: true
+    email: coach.mladen.jovanovic@gmail.com
   - name: Jason D. Vescovi
-    affiliation: |
-      | University of Toronto
-    address: |
-      | Faculty of Kinesiology and Physical Education
-      | Graduate School of Exercise Science
-      | Toronto, ON Canada
-    email: \email{vescovij@gmail.com}
-title:
-  formatted: "\\pkg{shorts}: An R Package for Modeling Short Sprints"
-DOI: "10.31236/osf.io/4jw62"
-URL: https://osf.io/preprints/sportrxiv/4jw62/
+    affiliation: 2
+    corresponding: false
+    email: vescovij@gmail.com
+affiliation:
+  - code: 1
+    address: Faculty of Sport and Physical Education, University of Belgrade, Serbia
+  - code: 2
+    address: Faculty of Kinesiology and Physical Education, Graduate School of Exercise Science, Toronto, ON Canada
 abstract: >
-  Short sprint performance is one of the most distinguishable and admired physical trait in sports.
+  Short sprint performance is one of the most distinguishable and admired physical traits in sports.
 Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
 The most common methods to assess short sprint performance are with a radar gun or timing gates.
-In this paper, we: 1) provide the \pkg{shorts} package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the \pkg{shorts} package to help alleviate errors within the subsequent parameter outcomes.
-keywords:
-  formatted: [sprinting, sports, modeling]
-preamble: >
-  \usepackage{amsmath}
-output:
-  bookdown::pdf_book:
-    base_format: rticles::jss_article
-    extra_dependencies: ["flafter"]
-citation_package: natbib
-bibliography: [references.bib, packages.bib]
-link-citations: yes
+In this paper, we: 1) provide the **shorts** package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the **shorts** package to help alleviate errors within the subsequent parameter outcomes.
 header-includes: >
+  \usepackage{lipsum}
   \usepackage[htt]{hyphenat}
@@
   \usepackage{microtype}
+preamble: >
+  \usepackage{amsmath}
+citation_package: natbib
+bibliography: [references.bib, packages.bib]
+output:
+  bookdown::pdf_book:
+    base_format: rticles::peerj_article # for using bookdown features like \@ref()
+    extra_dependencies: ["flafter"]
+  rticles::peerj_article: default
+link-citations: yes
 editor_options:
@@
   digits = 3,
-  "width" = 70,
+  "width" = 65,
   scipen = 999)
@@
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
-The \pkg{shorts} package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
-
-In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
+The **shorts** package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+
+In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the **shorts** package, issues that can arise during measurement and estimation, and potential solutions to those problems.
 
@@
 
-The parameters of the equation \@ref(eq:velocity-time) are \textit{maximum sprinting speed} (MSS; expressed in $ms^{-1}$) and \textit{relative acceleration} (TAU).
-Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in $ms^{-2}$) \@ref(eq:maximal-acceleration).
+The parameters of the equation \@ref(eq:velocity-time) are *maximum sprinting speed* (MSS; expressed in $ms^{-1}$) and *relative acceleration* (TAU).
+Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; *maximal acceleration*, expressed in $ms^{-2}$) \@ref(eq:maximal-acceleration).
 
@@
 Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in \pkg{shorts} package using non-linear least squares regression implemented in the `nls()` function in the \pkg{base} R [@R-base] and `nlme()` function in the \pkg{nlme} package [@R-nlme] for the mixed-effect models.
+Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **shorts** package using non-linear least squares regression implemented in the `nls()` function in the **base R** [@R-base] and `nlme()` function in the **nlme** package [@R-nlme] for the mixed-effect models.
 
@@
 
-Besides providing *residual standard error* (RSE), \pkg{shorts} functions provide additional model fit estimators.
+Besides providing *residual standard error* (RSE), **shorts** functions provide additional model fit estimators.
 Additional information can be gained by exploring the returned object, particularly object returned from the `nls()` function:
@@
 
-To create a simple plot of the model, use S3 `plot()` method, which returns \pkg{ggplot2} [@R-ggplot2] object:
+To create a simple plot of the model, use S3 `plot()` method, which returns **ggplot2** [@R-ggplot2] object:
 
@@
 
-Plotting the model predictions can be done once we convert data from wide to long with the help of   \pkg{ggplot2} [@R-ggplot2], \pkg{dplyr} [@R-dplyr], \pkg{tidyr} [@R-tidyr], and \pkg{tidyverse} [@R-tidyverse] packages:
+Plotting the model predictions can be done once we convert data from wide to long with the help of   **ggplot2** [@R-ggplot2], **dplyr** [@R-dplyr], **tidyr** [@R-tidyr], and **tidyverse** [@R-tidyverse] packages:
 
@@
 Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
-To identify these values, \pkg{shorts} package comes with `find_XXX()` family of functions:
+To identify these values, **shorts** package comes with `find_XXX()` family of functions:
 
@@
 Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in \pkg{shorts} for split times, one can use `mixed_model_using_splits()` function.
-To demonstrate this functionality, we load the `split_times` dataset provided in the \pkg{shorts} package:
+To perform mixed-effects models in **shorts** for split times, one can use `mixed_model_using_splits()` function.
+To demonstrate this functionality, we load the `split_times` dataset provided in the **shorts** package:
 
@@
 
-Mixed-effects model using radar data is done using `mixed_model_using_radar()` function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
+Mixed-effects model using radar data is done using `mixed_model_using_radar()` function. To perform mixed model, let's load data that comes with **shorts** package.
 
@@
 When estimating FVP, athlete body mass (kg) can be set using `bodymass` parameter, while the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance()` function.
-Details of the FVP method implemented in the \pkg{shorts} package, as well as the interpretation from a sprint training perspective, are covered elsewhere [@haugenPowerForceVelocityProfilingSprinting2020; @morinInterpretingPowerForceVelocityProfiles2016; @morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+Details of the FVP method implemented in the **shorts** package, as well as the interpretation from a sprint training perspective, are covered elsewhere [@haugenPowerForceVelocityProfilingSprinting2020; @morinInterpretingPowerForceVelocityProfiles2016; @morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
 
@@
 To ensure accurate parameter outcomes, the initial force production must be synced with start time [@haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintMechanicalVariables2019].
-Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \pkg{shorts} package.
+Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the **shorts** package.
 
@@
 As can be seen from the example, all estimated parameters are affected by an error in synchronization of time with velocity (with MSS being the least affected in this example).
-The potential solution incorporated into the \pkg{shorts} package involves estimation of the *time correction* parameter using the following equation:
+The potential solution incorporated into the **shorts** package involves estimation of the *time correction* parameter using the following equation:
 
@@
     true_time = predict_time_at_distance(true_distance, MSS, TAU),
-    stolen_time = predict_time_at_distance(flying_start_distance, MSS, TAU),
+    stolen_time = predict_time_at_distance(
+      flying_start_distance, MSS, TAU),
     time = true_time - stolen_time
@@
 Since maximal sprint speed is difficult to improve, the effects of start inconsistencies can mask effects of the training intervention.
-It is thus crucial to standardize the start when testing and implementing the following techniques when using the \pkg{shorts} package.
+It is thus crucial to standardize the start when testing and implementing the following techniques when using the **shorts** package.
 
@@
 
-Rather than providing *apriori* time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
+Rather than providing *apriori* time correction from the literature, **shorts** package provides an estimation of this parameter from the data provided, together with MSS and TAU.
 Exactly the same method is suggested by @stenrothForcevelocityProfilingIce2020, named *time shift method*, and the estimated parameter named *time shift parameter*.
@@
 
-Another model definition, which is a novel approach implemented in the \pkg{shorts} package, is to utilize *distance correction*, besides time correction.
-Thus, equation \@ref(eq:time-distance) becomes:
+Another model definition, which is a novel approach implemented in the **shorts** package, is to utilize *distance correction*, besides time correction. Thus, equation \@ref(eq:time-distance) becomes:
 
@@
   df,
-  aes(x = flying_start_distance, y = est_time_correction, color = model)
-) +
+  aes(
+    x = flying_start_distance,
+    y = est_time_correction,
+    color = model)) +
   theme_bw(8) +
   geom_line(alpha = 0.7) +
-  geom_line(aes(y = stolen_time), color = "black", linetype = "dashed") +
+  geom_line(
+    aes(y = stolen_time), color = "black", linetype = "dashed") +
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
@@
   df,
-  aes(x = flying_start_distance, y = est_distance_correction, color = model)
-) +
+  aes(
+    x = flying_start_distance,
+    y = est_distance_correction, 
+    color = model)) +
   theme_bw(8) +
@@
 
-To estimate parameter stability, model over-fitting, and performance on the unseen data, \pkg{shorts} model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
+To estimate parameter stability, model over-fitting, and performance on the unseen data, **shorts** model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
 LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation.
@@
 Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
-LOOCV can also be implemented with the mixed-effects models in the \pkg{shorts} package.
+LOOCV can also be implemented with the mixed-effects models in the **shorts** package.
 
@@
 
-Let's utilize demonstrated functionalities of the \pkg{shorts} package using real-world data.
-The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package (see `?vescovi`).
+Let's utilize demonstrated functionalities of the **shorts** package using real-world data.
+The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the **shorts** package (see `?vescovi`).
 
@@
 
-# Model with estimated time and distance corrections and deducted reaction time
+# Model with estimated time and distance corrections and
+#  deducted reaction time
 bolt_m6 <- model_using_splits_with_corrections(
@@
 
-time_correction_fixed <- mixed_model_using_splits_with_time_correction(
+time_correction_fixed <- 
+  mixed_model_using_splits_with_time_correction(
     df,
@@
 
-time_correction_random <- mixed_model_using_splits_with_time_correction(
+time_correction_random <- 
+  mixed_model_using_splits_with_time_correction(
     df,
@@
 
-time_distance_correction_fixed <- mixed_model_using_splits_with_corrections(
+time_distance_correction_fixed <- 
+  mixed_model_using_splits_with_corrections(
     df,
@@
 
-time_distance_correction_random <- mixed_model_using_splits_with_corrections(
+time_distance_correction_random <- 
+  mixed_model_using_splits_with_corrections(
     df,
@@
 
-We are hoping that the \pkg{shorts} package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
+We are hoping that the **shorts** package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
 

### 2021-04-04 paper.tex
PeerJ version of the manuscript

@@
-% To remove when JSS.cls is fixed
-% https://github.com/rstudio/rticles/issues/329
-\RequirePackage[2020/02/02]{latexrelease}
-\documentclass[
-]{jss}
-
-\usepackage[utf8]{inputenc}
-
-\providecommand{\tightlist}{%
+%% Submissions for peer-review must enable line-numbering
+%% using the lineno option in the \documentclass command.
+%%
+%% Preprints and camera-ready submissions do not need
+%% line numbers, and should have this option removed.
+%%
+%% Please note that the line numbering option requires
+%% version 1.1 or newer of the wlpeerj.cls file, and
+%% the corresponding author info requires v1.2
+
+\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
+
+% ZNK -- Adding headers for pandoc
+
+\setlength{\emergencystretch}{3em}
+\providecommand{\tightlist}{
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\author{
-Mladen Jovanović\\University of Belgrade \And Jason D. Vescovi\\University of Toronto
-}
-\title{\pkg{shorts}: An R Package for Modeling Short Sprints}
-
-\Plainauthor{Mladen Jovanović, Jason D. Vescovi}
-
-\Abstract{
-Short sprint performance is one of the most distinguishable and admired physical trait in sports.
+\usepackage{lipsum}
+\usepackage[unicode=true]{hyperref}
+\usepackage{longtable}
+
+% Pandoc syntax highlighting
+% See https://github.com/rstudio/rticles/issues/182
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
+% Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc} % for calculating minipage widths
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% Pandoc Header
+\usepackage{lipsum} \usepackage[htt]{hyphenat} \usepackage{booktabs} \usepackage{microtype}
+
+\usepackage{flafter}
+
+\title{shorts: An R Package for Modeling Short Sprints}
+
+\author[1]{Mladen Jovanović}
+
+\corrauthor[1]{Mladen Jovanović}{\href{mailto:coach.mladen.jovanovic@gmail.com}{\nolinkurl{coach.mladen.jovanovic@gmail.com}}}
+\author[2]{Jason D. Vescovi}
+
+\affil[1]{Faculty of Sport and Physical Education, University of Belgrade, Serbia}
+\affil[2]{Faculty of Kinesiology and Physical Education, Graduate School of Exercise Science, Toronto, ON Canada}
+
+%
+% \author[1]{First Author}
+% \author[2]{Second Author}
+% \affil[1]{Address of first author}
+% \affil[2]{Address of second author}
+% \corrauthor[1]{First Author}{f.author@email.com}
+
+%
+
+\begin{abstract}
+Short sprint performance is one of the most distinguishable and admired physical traits in sports.
 Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
 The most common methods to assess short sprint performance are with a radar gun or timing gates.
-In this paper, we: 1) provide the \pkg{shorts} package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the \pkg{shorts} package to help alleviate errors within the subsequent parameter outcomes.
-}
-
-\Keywords{sprinting, sports, modeling}
-
-%% publication information
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{}
-%% \Acceptdate{2012-06-04}
-
-\Address{
-    Mladen Jovanović\\
-    University of Belgrade\\
-    Faculty of Sport and Physical Education\\
-Belgrade, Serbia\\
-  E-mail: \email{coach.mladen.jovanovic@gmail.com}\\
-
-      Jason D. Vescovi\\
-    University of Toronto\\
-    Faculty of Kinesiology and Physical Education\\
-Graduate School of Exercise Science\\
-Toronto, ON Canada\\
-  E-mail: \email{vescovij@gmail.com}\\
-
-  }
-
-% Pandoc citation processing
-
-% Pandoc header
-\usepackage[htt]{hyphenat} \usepackage{booktabs} \usepackage{microtype}
-
-\usepackage{flafter}
-
-\usepackage{amsmath}
+In this paper, we: 1) provide the \textbf{shorts} package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the \textbf{shorts} package to help alleviate errors within the subsequent parameter outcomes.
+% Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+\end{abstract}
 
 \begin{document}
+
+\flushbottom
+\maketitle
+\thispagestyle{empty}
 
@@
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (\textless5 s) of maximal efforts \citep{mangineSpeedForcePower2014}, whereas the ability to achieve maximal sprint speed varies based on the type of sport.
-For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) \citep{ward-smithEnergyConversionStrategies2001}, but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)\citep{brownAssessmentLinearSprinting2004}.
+Peak anaerobic power is achieved within the first few seconds (\textless5 s) of maximal efforts (\protect\hyperlink{ref-mangineSpeedForcePower2014}{Mangine et al.
+2014}), whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) (\protect\hyperlink{ref-ward-smithEnergyConversionStrategies2001}{Ward-Smith 2001}), but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)(\protect\hyperlink{ref-brownAssessmentLinearSprinting2004}{Brown, Vescovi, and Vanheest 2004}).
 Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
 
-The use of force plates is considered the gold standard for assessing mechanical properties of sprinting; however, there are logistical and financial challenges to capturing the profile of an entire sprint \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
-Radar and laser technology are frequently used laboratory-grade methods \citep{buchheitMechanicalDeterminantsAcceleration2014, edwardsSprintAccelerationCharacteristics2020, jimenez-reyesRelationshipVerticalHorizontal2018, marcote-pequenoAssociationForceVelocity2019} but not normally accessible to practitioners working in sports.
+The use of force plates is considered the gold standard for assessing mechanical properties of sprinting; however, there are logistical and financial challenges to capturing the profile of an entire sprint (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
+2016}).
+Radar and laser technology are frequently used laboratory-grade methods (\protect\hyperlink{ref-buchheitMechanicalDeterminantsAcceleration2014}{Buchheit et al.
+2014}; \protect\hyperlink{ref-edwardsSprintAccelerationCharacteristics2020}{Edwards et al.
+2020}; \protect\hyperlink{ref-jimenez-reyesRelationshipVerticalHorizontal2018}{Jiménez-Reyes et al.
+2018}; \protect\hyperlink{ref-marcote-pequenoAssociationForceVelocity2019}{Marcote-Pequeño et al.
+2019}) but not normally accessible to practitioners working in sports.
 Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
-Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
+2016}).
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
-The \pkg{shorts} package \citep{R-shorts}, written in the R language \citep{R-base}, represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling \citep{morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
-
-In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \pkg{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
+The \textbf{shorts} package (\protect\hyperlink{ref-R-shorts}{Jovanovic 2020}), written in the R language (\protect\hyperlink{ref-R-base}{R Core Team 2020}), represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
+2016}).
+
+In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the \textbf{shorts} package, issues that can arise during measurement and estimation, and potential solutions to those problems.
 
@@
 
-Short sprints have been modeled using the mono-exponential equation \eqref{eq:velocity-time} originally proposed by \citet{doi:10.1098/rspb.1927.0035}, and more recently popularized by \citet{clarkNFLCombine40Yard2017}, and \citet{samozinoSimpleMethodMeasuring2016}.
+Short sprints have been modeled using the mono-exponential equation \eqref{eq:velocity-time} originally proposed by \protect\hyperlink{ref-doi:10.1098ux2frspb.1927.0035}{Furusawa, Hill, and Parkinson} (\protect\hyperlink{ref-doi:10.1098ux2frspb.1927.0035}{1927}), and more recently popularized by \protect\hyperlink{ref-clarkNFLCombine40Yard2017}{Clark et al.} (\protect\hyperlink{ref-clarkNFLCombine40Yard2017}{2017}), and \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.} (\protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{2016}).
 Equation \eqref{eq:velocity-time} represents function for instantaneous horizontal velocity \(v\) given the time \(t\) and two model parameters:
@@
 
-The parameters of the equation \eqref{eq:velocity-time} are \textit{maximum sprinting speed} (MSS; expressed in \(ms^{-1}\)) and \textit{relative acceleration} (TAU).
-Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \textit{maximal acceleration}, expressed in \(ms^{-2}\)) \eqref{eq:maximal-acceleration}.
+The parameters of the equation \eqref{eq:velocity-time} are \emph{maximum sprinting speed} (MSS; expressed in \(ms^{-1}\)) and \emph{relative acceleration} (TAU).
+Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; \emph{maximal acceleration}, expressed in \(ms^{-2}\)) \eqref{eq:maximal-acceleration}.
 
@@
 
-\begin{CodeChunk}
 \begin{table}
@@
 
-\end{CodeChunk}
-
 Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6 s).
 
-\begin{CodeChunk}
 \begin{figure}
@@
 
-\caption[Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds]{Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds.}\label{fig:four-athletes-kinematics}
+\caption{Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds.}\label{fig:four-athletes-kinematics}
 \end{figure}
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
 \begin{figure}
@@
 
-\caption[Acceleration-Velocity profile of four athletes with different MSS and MAC parameters]{Acceleration-Velocity profile of four athletes with different MSS and MAC parameters.}\label{fig:four-athletes-profile}
+\caption{Acceleration-Velocity profile of four athletes with different MSS and MAC parameters.}\label{fig:four-athletes-profile}
 \end{figure}
-\end{CodeChunk}
 
@@
 Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \pkg{shorts} package using non-linear least squares regression implemented in the \texttt{nls()} function in the \pkg{base} R \citep{R-base} and \texttt{nlme()} function in the \pkg{nlme} package \citep{R-nlme} for the mixed-effect models.
+Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \textbf{shorts} package using non-linear least squares regression implemented in the \texttt{nls()} function in the \textbf{base R} (\protect\hyperlink{ref-R-base}{R Core Team 2020}) and \texttt{nlme()} function in the \textbf{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2020}) for the mixed-effect models.
 
@@
 
-\(W\) in equation \eqref{eq:time-distance} represents Lambert's W function \citep{R-LambertW}. MSS and TAU parameters are estimated using \texttt{model\_using\_splits()} function:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> require(shorts)
-R> 
-R> split_distance <- c(10, 20, 30, 40)
-R> 
-R> split_time <- c(2.17, 3.43, 4.60, 5.73)
-R> 
-R> m1 <- model_using_splits(
-+   distance = split_distance,
-+   time = split_time
-+ )
-R> 
-R> m1
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated model parameters
-#> --------------------------
-#>                 MSS                 TAU                 MAC 
-#>                9.01                1.31                6.89 
-#>                PMAX     time_correction distance_correction 
-#>               15.52                0.00                0.00 
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>   0.00249   1.00000  -0.00178   0.00265   0.00265   0.00176   0.00157 
-#>      MAPE 
-#>   0.04742
-\end{CodeOutput}
-\end{CodeChunk}
+\(W\) in equation \eqref{eq:time-distance} represents Lambert's W function (\protect\hyperlink{ref-R-LambertW}{Goerg 2020}).
+MSS and TAU parameters are estimated using \texttt{model\_using\_splits()} function:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{require}\NormalTok{(shorts)}
+
+\NormalTok{split\_distance }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{)}
+
+\NormalTok{split\_time }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{2.17}\NormalTok{, }\FloatTok{3.43}\NormalTok{, }\FloatTok{4.60}\NormalTok{, }\FloatTok{5.73}\NormalTok{)}
+
+\NormalTok{m1 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_distance,}
+  \AttributeTok{time =}\NormalTok{ split\_time}
+\NormalTok{)}
+
+\NormalTok{m1}
+\CommentTok{\#\textgreater{} Estimated model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                9.01                1.31                6.89 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               15.52                0.00                0.00 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}   0.00249   1.00000  {-}0.00178   0.00265   0.00265   0.00176 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}   0.00157   0.04742}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-Besides providing \emph{residual standard error} (RSE), \pkg{shorts} functions provide additional model fit estimators.
+Besides providing \emph{residual standard error} (RSE), \textbf{shorts} functions provide additional model fit estimators.
 Additional information can be gained by exploring the returned object, particularly object returned from the \texttt{nls()} function:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> summary(m1)
-\end{CodeInput}
-\begin{CodeOutput}
-#> 
-#> Formula: corrected_time ~ TAU * I(LambertW::W(-exp(1)^(-distance/(MSS * 
-#>     TAU) - 1))) + distance/MSS + TAU
-#> 
-#> Parameters:
-#>     Estimate Std. Error t value Pr(>|t|)    
-#> MSS   9.0121     0.0155     581 0.000003 ***
-#> TAU   1.3083     0.0066     198 0.000025 ***
-#> ---
-#> Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-#> 
-#> Residual standard error: 0.00249 on 2 degrees of freedom
-#> 
-#> Number of iterations to convergence: 4 
-#> Achieved convergence tolerance: 0.00000311
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{summary}\NormalTok{(m1)}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Formula: corrected\_time \textasciitilde{} TAU * I(LambertW::W({-}exp(1)\^{}({-}distance/(MSS * }
+\CommentTok{\#\textgreater{}     TAU) {-} 1))) + distance/MSS + TAU}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Parameters:}
+\CommentTok{\#\textgreater{}     Estimate Std. Error t value Pr(\textgreater{}|t|)    }
+\CommentTok{\#\textgreater{} MSS   9.0121     0.0155     581 0.000003 ***}
+\CommentTok{\#\textgreater{} TAU   1.3083     0.0066     198 0.000025 ***}
+\CommentTok{\#\textgreater{} {-}{-}{-}}
+\CommentTok{\#\textgreater{} Signif.
+codes:  0 \textquotesingle{}***\textquotesingle{} 0.001 \textquotesingle{}**\textquotesingle{} 0.01 \textquotesingle{}*\textquotesingle{} 0.05 \textquotesingle{}.\textquotesingle{} 0.1 \textquotesingle{} \textquotesingle{} 1}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Residual standard error: 0.00249 on 2 degrees of freedom}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Number of iterations to convergence: 4 }
+\CommentTok{\#\textgreater{} Achieved convergence tolerance: 0.00000311}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> coef(m1)
-\end{CodeInput}
-\begin{CodeOutput}
-#>                 MSS                 TAU                 MAC 
-#>                9.01                1.31                6.89 
-#>                PMAX     time_correction distance_correction 
-#>               15.52                0.00                0.00
-\end{CodeOutput}
-\end{CodeChunk}
-
-To create a simple plot of the model, use S3 \texttt{plot()} method, which returns \pkg{ggplot2} \citep{R-ggplot2} object:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot(m1) + theme_bw(8)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{coef}\NormalTok{(m1)}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                9.01                1.31                6.89 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               15.52                0.00                0.00}
+\end{Highlighting}
+\end{Shaded}
+
+To create a simple plot of the model, use S3 \texttt{plot()} method, which returns \textbf{ggplot2} (\protect\hyperlink{ref-R-ggplot2}{Wickham, Chang, et al. 2020}) object:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(m1) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 Once we have estimated MSS and TAU, we can use \texttt{predict\_XXX()} family of functions to predict various relationships (i.e., time at distance, acceleration at distance, velocity at time, etc.):
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Predict time at distance
-R> predict_time_at_distance(
-+   distance = split_distance,
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 2.17 3.43 4.60 5.73
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # Predict acceleration at time
-R> predict_acceleration_at_time(
-+   time = c(0, 1, 2, 3, 4, 5, 6),
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 6.8884 3.2075 1.4935 0.6954 0.3238 0.1508 0.0702
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Predict time at distance}
+\FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_distance,}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 2.17 3.43 4.60 5.73}
+
+\CommentTok{\# Predict acceleration at time}
+\FunctionTok{predict\_acceleration\_at\_time}\NormalTok{(}
+  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{4}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{6}\NormalTok{),}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 6.8884 3.2075 1.4935 0.6954 0.3238 0.1508 0.0702}
+\end{Highlighting}
+\end{Shaded}
 
@@
 To estimate force production at distance or time (using \texttt{predict\_force\_at\_distance()} and \texttt{predict\_force\_at\_time()} functions), as well as power production (using \texttt{predict\_power\_at\_distance()} and \texttt{predict\_power\_at\_time()} functions), one needs to take into account the air resistance.
-Air resistance (N) is estimated using \texttt{get\_air\_resistance()} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \citet{arsacModelingEnergetics100m2002}, \citet{samozinoSimpleMethodMeasuring2016}, and \citet{vaningenschenauCanCyclePower1991} for more information):
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> get_air_resistance(
-+   velocity = 5,
-+   bodymass = 80,
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 6.1
-\end{CodeOutput}
-\end{CodeChunk}
+Air resistance (N) is estimated using \texttt{get\_air\_resistance()} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \protect\hyperlink{ref-arsacModelingEnergetics100m2002}{Arsac and Locatelli} (\protect\hyperlink{ref-arsacModelingEnergetics100m2002}{2002}), \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.} (\protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{2016}), and \protect\hyperlink{ref-vaningenschenauCanCyclePower1991}{van Ingen Schenau, Jacobs, and de Koning} (\protect\hyperlink{ref-vaningenschenauCanCyclePower1991}{1991}) for more information):
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{get\_air\_resistance}\NormalTok{(}
+  \AttributeTok{velocity =} \DecValTok{5}\NormalTok{,}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 6.1}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # To calculate horizontal force produced
-R> predict_force_at_distance(
-+   distance = split_distance,
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to get_air_resistance
-+   # Otherwise, defaults are used
-+   bodymass = 80,
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 119.0  58.6  36.9  28.2
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # To calculate power produced
-R> predict_power_at_distance(
-+   distance = split_distance,
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to get_air_resistance
-+   # Otherwise, defaults are used
-+   bodymass = 80,
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 868 490 323 251
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# To calculate horizontal force produced}
+\FunctionTok{predict\_force\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_distance,}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
+  \CommentTok{\# Otherwise, defaults are used}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 119.0  58.6  36.9  28.2}
+
+\CommentTok{\# To calculate power produced}
+\FunctionTok{predict\_power\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_distance,}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
+  \CommentTok{\# Otherwise, defaults are used}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 868 490 323 251}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- predict_kinematics(
-+   m1,
-+   max_time = 6,
-+   frequency = 100,
-+   # Additional parameters forwarded to get_air_resistance
-+   # Otherwise, defaults are used
-+   bodymass = 80,
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-R> 
-R> head(df)
-\end{CodeInput}
-\begin{CodeOutput}
-#>   time distance velocity acceleration bodymass net_horizontal_force
-#> 1 0.00 0.000000   0.0000         6.89       80                  551
-#> 2 0.01 0.000344   0.0686         6.84       80                  547
-#> 3 0.02 0.001371   0.1367         6.78       80                  543
-#> 4 0.03 0.003076   0.2043         6.73       80                  539
-#> 5 0.04 0.005455   0.2714         6.68       80                  534
-#> 6 0.05 0.008502   0.3379         6.63       80                  530
-#>   air_resistance horizontal_force horizontal_force_relative
-#> 1        0.07536              551                      6.89
-#> 2        0.05609              547                      6.84
-#> 3        0.03978              543                      6.78
-#> 4        0.02636              539                      6.73
-#> 5        0.01576              534                      6.68
-#> 6        0.00792              530                      6.63
-#>   vertical_force resultant_force resultant_force_relative power
-#> 1            785             959                     12.0   0.0
-#> 2            785             957                     12.0  37.5
-#> 3            785             954                     11.9  74.2
-#> 4            785             952                     11.9 110.0
-#> 5            785             950                     11.9 145.0
-#> 6            785             947                     11.8 179.2
-#>   relative_power    RF force_angle
-#> 1          0.000 0.575        54.9
-#> 2          0.469 0.572        55.1
-#> 3          0.928 0.569        55.3
-#> 4          1.375 0.566        55.5
-#> 5          1.813 0.563        55.7
-#> 6          2.241 0.560        55.9
-\end{CodeOutput}
-\end{CodeChunk}
-
-Plotting the model predictions can be done once we convert data from wide to long with the help of \pkg{ggplot2} \citep{R-ggplot2}, \pkg{dplyr} \citep{R-dplyr}, \pkg{tidyr} \citep{R-tidyr}, and \pkg{tidyverse} \citep{R-tidyverse} packages:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> require(tidyverse)
-R> 
-R> variable_names <- colnames(df)
-R> 
-R> df <- pivot_longer(data = df, cols = -2) %>%
-+   mutate(name = factor(name, levels = variable_names))
-R> 
-R> ggplot(df, aes(x = distance, y = value)) +
-+   theme_bw(8) +
-+   facet_wrap(~name, scales = "free_y") +
-+   geom_line(alpha = 0.7) +
-+   ylab(NULL) +
-+   xlab("Distance (m)")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{predict\_kinematics}\NormalTok{(}
+\NormalTok{  m1,}
+  \AttributeTok{max\_time =} \DecValTok{6}\NormalTok{,}
+  \AttributeTok{frequency =} \DecValTok{100}\NormalTok{,}
+  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
+  \CommentTok{\# Otherwise, defaults are used}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+
+\FunctionTok{head}\NormalTok{(df)}
+\CommentTok{\#\textgreater{}   time distance velocity acceleration bodymass}
+\CommentTok{\#\textgreater{} 1 0.00 0.000000   0.0000         6.89       80}
+\CommentTok{\#\textgreater{} 2 0.01 0.000344   0.0686         6.84       80}
+\CommentTok{\#\textgreater{} 3 0.02 0.001371   0.1367         6.78       80}
+\CommentTok{\#\textgreater{} 4 0.03 0.003076   0.2043         6.73       80}
+\CommentTok{\#\textgreater{} 5 0.04 0.005455   0.2714         6.68       80}
+\CommentTok{\#\textgreater{} 6 0.05 0.008502   0.3379         6.63       80}
+\CommentTok{\#\textgreater{}   net\_horizontal\_force air\_resistance horizontal\_force}
+\CommentTok{\#\textgreater{} 1                  551        0.07536              551}
+\CommentTok{\#\textgreater{} 2                  547        0.05609              547}
+\CommentTok{\#\textgreater{} 3                  543        0.03978              543}
+\CommentTok{\#\textgreater{} 4                  539        0.02636              539}
+\CommentTok{\#\textgreater{} 5                  534        0.01576              534}
+\CommentTok{\#\textgreater{} 6                  530        0.00792              530}
+\CommentTok{\#\textgreater{}   horizontal\_force\_relative vertical\_force resultant\_force}
+\CommentTok{\#\textgreater{} 1                      6.89            785             959}
+\CommentTok{\#\textgreater{} 2                      6.84            785             957}
+\CommentTok{\#\textgreater{} 3                      6.78            785             954}
+\CommentTok{\#\textgreater{} 4                      6.73            785             952}
+\CommentTok{\#\textgreater{} 5                      6.68            785             950}
+\CommentTok{\#\textgreater{} 6                      6.63            785             947}
+\CommentTok{\#\textgreater{}   resultant\_force\_relative power relative\_power    RF}
+\CommentTok{\#\textgreater{} 1                     12.0   0.0          0.000 0.575}
+\CommentTok{\#\textgreater{} 2                     12.0  37.5          0.469 0.572}
+\CommentTok{\#\textgreater{} 3                     11.9  74.2          0.928 0.569}
+\CommentTok{\#\textgreater{} 4                     11.9 110.0          1.375 0.566}
+\CommentTok{\#\textgreater{} 5                     11.9 145.0          1.813 0.563}
+\CommentTok{\#\textgreater{} 6                     11.8 179.2          2.241 0.560}
+\CommentTok{\#\textgreater{}   force\_angle}
+\CommentTok{\#\textgreater{} 1        54.9}
+\CommentTok{\#\textgreater{} 2        55.1}
+\CommentTok{\#\textgreater{} 3        55.3}
+\CommentTok{\#\textgreater{} 4        55.5}
+\CommentTok{\#\textgreater{} 5        55.7}
+\CommentTok{\#\textgreater{} 6        55.9}
+\end{Highlighting}
+\end{Shaded}
+
+Plotting the model predictions can be done once we convert data from wide to long with the help of \textbf{ggplot2} (\protect\hyperlink{ref-R-ggplot2}{Wickham, Chang, et al.
+2020}), \textbf{dplyr} (\protect\hyperlink{ref-R-dplyr}{Wickham, François, et al.
+2020}), \textbf{tidyr} (\protect\hyperlink{ref-R-tidyr}{Wickham 2020}), and \textbf{tidyverse} (\protect\hyperlink{ref-R-tidyverse}{Wickham 2019}) packages:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{require}\NormalTok{(tidyverse)}
+
+\NormalTok{variable\_names }\OtherTok{\textless{}{-}} \FunctionTok{colnames}\NormalTok{(df)}
+
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{data =}\NormalTok{ df, }\AttributeTok{cols =} \SpecialCharTok{{-}}\DecValTok{2}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{name =} \FunctionTok{factor}\NormalTok{(name, }\AttributeTok{levels =}\NormalTok{ variable\_names))}
+
+\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{y =}\NormalTok{ value)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{name, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
-
-\end{CodeChunk}
 
@@
 Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90\% of maximum sprinting speed is reached, or where peak power is within 90\% range.
-To identify these values, \pkg{shorts} package comes with \texttt{find\_XXX()} family of functions:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Finds distance where 90% of maximum sprinting speed is reached
-R> find_velocity_critical_distance(
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU,
-+   percent = 0.9
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 16.5
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # Finds maximal power and distance (this time using air resistance)
-R> find_max_power_distance(
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to get_air_resistance
-+   # Otherwise, defaults are used
-+   bodymass = 80,
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> $max_power
-#> [1] 1264
-#> 
-#> $distance
-#> [1] 2.46
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # Finds distance over 90% power range
-R> find_power_critical_distance(
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU,
-+   # Additional parameters forwarded to get_air_resistance
-+   # Otherwise, defaults are used
-+   bodymass = 80,
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> $lower
-#> [1] 0.959
-#> 
-#> $upper
-#> [1] 5.44
-\end{CodeOutput}
-\end{CodeChunk}
+To identify these values, \textbf{shorts} package comes with \texttt{find\_XXX()} family of functions:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Finds distance where 90\% of maximum sprinting speed is reached}
+\FunctionTok{find\_velocity\_critical\_distance}\NormalTok{(}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \AttributeTok{percent =} \FloatTok{0.9}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 16.5}
+
+\CommentTok{\# Finds maximal power and distance (this time using air resistance)}
+\FunctionTok{find\_max\_power\_distance}\NormalTok{(}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
+  \CommentTok{\# Otherwise, defaults are used}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} $max\_power}
+\CommentTok{\#\textgreater{} [1] 1264}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} $distance}
+\CommentTok{\#\textgreater{} [1] 2.46}
+
+\CommentTok{\# Finds distance over 90\% power range}
+\FunctionTok{find\_power\_critical\_distance}\NormalTok{(}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
+  \CommentTok{\# Otherwise, defaults are used}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} $lower}
+\CommentTok{\#\textgreater{} [1] 0.959}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} $upper}
+\CommentTok{\#\textgreater{} [1] 5.44}
+\end{Highlighting}
+\end{Shaded}
 
@@
 Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in \pkg{shorts} for split times, one can use \texttt{mixed\_model\_using\_splits()} function.
-To demonstrate this functionality, we load the \texttt{split\_times} dataset provided in the \pkg{shorts} package:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> data(split_times)
-R> 
-R> # Mixed model
-R> m2 <- mixed_model_using_splits(
-+   data = split_times,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "athlete",
-+ 
-+   # Select random effects
-+   # Default is MSS and TAU
-+   random = MSS + TAU ~ 1
-+ )
-R> 
-R> m2
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated fixed model parameters
-#> --------------------------------
-#>                 MSS                 TAU                 MAC 
-#>               8.065               0.655              12.309 
-#>                PMAX     time_correction distance_correction 
-#>              24.818               0.000               0.000 
-#> 
-#> Estimated random model parameters
-#> ---------------------------------
-#>     athlete  MSS   TAU  MAC PMAX time_correction distance_correction
-#> 1     James 9.69 0.847 11.4 27.7               0                   0
-#> 2       Jim 7.83 0.505 15.5 30.4               0                   0
-#> 3      John 7.78 0.727 10.7 20.8               0                   0
-#> 4 Kimberley 8.57 0.802 10.7 22.9               0                   0
-#> 5  Samantha 6.45 0.395 16.3 26.4               0                   0
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>    0.0260    0.9998   -0.0293    0.0496    0.0496    0.0214    0.0172 
-#>      MAPE 
-#>    0.9019
-\end{CodeOutput}
-\end{CodeChunk}
-
-Additional information about mixed-effects model performed using the \texttt{nlme} package \citep{R-nlme} can be obtained using \texttt{summary()} function:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> summary(m2)
-\end{CodeInput}
-\begin{CodeOutput}
-#> Nonlinear mixed-effects model fit by maximum likelihood
-#>   Model: corrected_time ~ TAU * I(LambertW::W(-exp(1)^(-distance/(MSS *      TAU) - 1))) + distance/MSS + TAU 
-#>   Data: train 
-#>     AIC   BIC logLik
-#>   -75.1 -66.7   43.5
-#> 
-#> Random effects:
-#>  Formula: list(MSS ~ 1, TAU ~ 1)
-#>  Level: athlete
-#>  Structure: General positive-definite, Log-Cholesky parametrization
-#>          StdDev Corr 
-#> MSS      1.066  MSS  
-#> TAU      0.178  0.877
-#> Residual 0.026       
-#> 
-#> Fixed effects:  MSS + TAU ~ 1 
-#>     Value Std.Error DF t-value p-value
-#> MSS  8.06     0.495 24   16.30       0
-#> TAU  0.66     0.084 24    7.82       0
-#>  Correlation: 
-#>     MSS  
-#> TAU 0.874
-#> 
-#> Standardized Within-Group Residuals:
-#>    Min     Q1    Med     Q3    Max 
-#> -1.909 -0.605  0.154  0.523  1.129 
-#> 
-#> Number of Observations: 30
-#> Number of Groups: 5
-\end{CodeOutput}
-\end{CodeChunk}
+To perform mixed-effects models in \textbf{shorts} for split times, one can use \texttt{mixed\_model\_using\_splits()} function.
+To demonstrate this functionality, we load the \texttt{split\_times} dataset provided in the \textbf{shorts} package:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{data}\NormalTok{(split\_times)}
+
+\CommentTok{\# Mixed model}
+\NormalTok{m2 }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_splits}\NormalTok{(}
+  \AttributeTok{data =}\NormalTok{ split\_times,}
+  \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+  \AttributeTok{athlete =} \StringTok{"athlete"}\NormalTok{,}
+
+  \CommentTok{\# Select random effects}
+  \CommentTok{\# Default is MSS and TAU}
+  \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{)}
+
+\NormalTok{m2}
+\CommentTok{\#\textgreater{} Estimated fixed model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}               8.065               0.655              12.309 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}              24.818               0.000               0.000 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Estimated random model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}     athlete  MSS   TAU  MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1     James 9.69 0.847 11.4 27.7               0}
+\CommentTok{\#\textgreater{} 2       Jim 7.83 0.505 15.5 30.4               0}
+\CommentTok{\#\textgreater{} 3      John 7.78 0.727 10.7 20.8               0}
+\CommentTok{\#\textgreater{} 4 Kimberley 8.57 0.802 10.7 22.9               0}
+\CommentTok{\#\textgreater{} 5  Samantha 6.45 0.395 16.3 26.4               0}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                   0}
+\CommentTok{\#\textgreater{} 2                   0}
+\CommentTok{\#\textgreater{} 3                   0}
+\CommentTok{\#\textgreater{} 4                   0}
+\CommentTok{\#\textgreater{} 5                   0}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}    0.0260    0.9998   {-}0.0293    0.0496    0.0496    0.0214 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}    0.0172    0.9019}
+\end{Highlighting}
+\end{Shaded}
+
+Additional information about mixed-effects model performed using the \texttt{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2020}) can be obtained using \texttt{summary()} function:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{summary}\NormalTok{(m2)}
+\CommentTok{\#\textgreater{} Nonlinear mixed{-}effects model fit by maximum likelihood}
+\CommentTok{\#\textgreater{}   Model: corrected\_time \textasciitilde{} TAU * I(LambertW::W({-}exp(1)\^{}({-}distance/(MSS *      TAU) {-} 1))) + distance/MSS + TAU }
+\CommentTok{\#\textgreater{}   Data: train }
+\CommentTok{\#\textgreater{}     AIC   BIC logLik}
+\CommentTok{\#\textgreater{}   {-}75.1 {-}66.7   43.5}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Random effects:}
+\CommentTok{\#\textgreater{}  Formula: list(MSS \textasciitilde{} 1, TAU \textasciitilde{} 1)}
+\CommentTok{\#\textgreater{}  Level: athlete}
+\CommentTok{\#\textgreater{}  Structure: General positive{-}definite, Log{-}Cholesky parametrization}
+\CommentTok{\#\textgreater{}          StdDev Corr }
+\CommentTok{\#\textgreater{} MSS      1.066  MSS  }
+\CommentTok{\#\textgreater{} TAU      0.178  0.877}
+\CommentTok{\#\textgreater{} Residual 0.026       }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Fixed effects:  MSS + TAU \textasciitilde{} 1 }
+\CommentTok{\#\textgreater{}     Value Std.Error DF t{-}value p{-}value}
+\CommentTok{\#\textgreater{} MSS  8.06     0.495 24   16.30       0}
+\CommentTok{\#\textgreater{} TAU  0.66     0.084 24    7.82       0}
+\CommentTok{\#\textgreater{}  Correlation: }
+\CommentTok{\#\textgreater{}     MSS  }
+\CommentTok{\#\textgreater{} TAU 0.874}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Standardized Within{-}Group Residuals:}
+\CommentTok{\#\textgreater{}    Min     Q1    Med     Q3    Max }
+\CommentTok{\#\textgreater{} {-}1.909 {-}0.605  0.154  0.523  1.129 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Number of Observations: 30}
+\CommentTok{\#\textgreater{} Number of Groups: 5}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> coef(m2)
-\end{CodeInput}
-\begin{CodeOutput}
-#> $fixed
-#>                 MSS                 TAU                 MAC 
-#>               8.065               0.655              12.309 
-#>                PMAX     time_correction distance_correction 
-#>              24.818               0.000               0.000 
-#> 
-#> $random
-#>     athlete  MSS   TAU  MAC PMAX time_correction distance_correction
-#> 1     James 9.69 0.847 11.4 27.7               0                   0
-#> 2       Jim 7.83 0.505 15.5 30.4               0                   0
-#> 3      John 7.78 0.727 10.7 20.8               0                   0
-#> 4 Kimberley 8.57 0.802 10.7 22.9               0                   0
-#> 5  Samantha 6.45 0.395 16.3 26.4               0                   0
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{coef}\NormalTok{(m2)}
+\CommentTok{\#\textgreater{} $fixed}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}               8.065               0.655              12.309 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}              24.818               0.000               0.000 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} $random}
+\CommentTok{\#\textgreater{}     athlete  MSS   TAU  MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1     James 9.69 0.847 11.4 27.7               0}
+\CommentTok{\#\textgreater{} 2       Jim 7.83 0.505 15.5 30.4               0}
+\CommentTok{\#\textgreater{} 3      John 7.78 0.727 10.7 20.8               0}
+\CommentTok{\#\textgreater{} 4 Kimberley 8.57 0.802 10.7 22.9               0}
+\CommentTok{\#\textgreater{} 5  Samantha 6.45 0.395 16.3 26.4               0}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                   0}
+\CommentTok{\#\textgreater{} 2                   0}
+\CommentTok{\#\textgreater{} 3                   0}
+\CommentTok{\#\textgreater{} 4                   0}
+\CommentTok{\#\textgreater{} 5                   0}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot(m2) + theme_bw(8)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(m2) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following figure contains kinematics for all athletes in \texttt{split\_times} dataset. Please note that power calculation takes default parameters for each individual:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- predict_kinematics(m2, max_time = 10)
-R> 
-R> variable_names <- colnames(df)
-R> 
-R> df <- pivot_longer(df, cols = c(-1, -3)) %>%
-+   mutate(name = factor(name, levels = variable_names))
-R> 
-R> ggplot(
-+   filter(df, distance < 40),
-+   aes(x = distance, y = value, group = athlete, color = athlete)
-+ ) +
-+   theme_bw(8) +
-+   facet_wrap(~name, scales = "free_y") +
-+   geom_line(alpha = 0.7) +
-+   ylab(NULL) +
-+   xlab("Distance (m)") +
-+   theme(
-+     legend.position = "top",
-+     legend.title = element_blank())
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{predict\_kinematics}\NormalTok{(m2, }\AttributeTok{max\_time =} \DecValTok{10}\NormalTok{)}
+
+\NormalTok{variable\_names }\OtherTok{\textless{}{-}} \FunctionTok{colnames}\NormalTok{(df)}
+
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{pivot\_longer}\NormalTok{(df, }\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{name =} \FunctionTok{factor}\NormalTok{(name, }\AttributeTok{levels =}\NormalTok{ variable\_names))}
+
+\FunctionTok{ggplot}\NormalTok{(}
+  \FunctionTok{filter}\NormalTok{(df, distance }\SpecialCharTok{\textless{}} \DecValTok{40}\NormalTok{),}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{y =}\NormalTok{ value, }\AttributeTok{group =}\NormalTok{ athlete, }\AttributeTok{color =}\NormalTok{ athlete)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{name, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{,}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{())}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> sprint_time <- seq(0, 6, 1)
-R> 
-R> sprint_velocity <- c(0.00, 4.83, 7.07, 8.10, 8.59, 8.81, 8.91)
-R> 
-R> m3 <- model_using_radar(
-+   velocity = sprint_velocity,
-+   time = sprint_time
-+ )
-R> 
-R> m3
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated model parameters
-#> --------------------------
-#>                 MSS                 TAU                 MAC 
-#>                9.00                1.30                6.92 
-#>                PMAX     time_correction distance_correction 
-#>               15.58                0.00                0.00 
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>   0.00327   1.00000  -0.00406   0.00532   0.00532   0.00276   0.00207 
-#>      MAPE 
-#>       NaN
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot(m3) + theme_bw(8)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sprint\_time }\OtherTok{\textless{}{-}} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{1}\NormalTok{)}
+
+\NormalTok{sprint\_velocity }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{0.00}\NormalTok{, }\FloatTok{4.83}\NormalTok{, }\FloatTok{7.07}\NormalTok{, }\FloatTok{8.10}\NormalTok{, }\FloatTok{8.59}\NormalTok{, }\FloatTok{8.81}\NormalTok{, }\FloatTok{8.91}\NormalTok{)}
+
+\NormalTok{m3 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ sprint\_velocity,}
+  \AttributeTok{time =}\NormalTok{ sprint\_time}
+\NormalTok{)}
+
+\NormalTok{m3}
+\CommentTok{\#\textgreater{} Estimated model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                9.00                1.30                6.92 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               15.58                0.00                0.00 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}   0.00327   1.00000  {-}0.00406   0.00532   0.00532   0.00276 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}   0.00207       NaN}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(m3) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> m3_weighted <- model_using_radar(
-+   velocity = sprint_velocity,
-+   time = sprint_time,
-+   weights = 1 / (sprint_velocity + 1)
-+ )
-R> 
-R> m3_weighted
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated model parameters
-#> --------------------------
-#>                 MSS                 TAU                 MAC 
-#>                9.00                1.30                6.92 
-#>                PMAX     time_correction distance_correction 
-#>               15.58                0.00                0.00 
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>   0.00108   1.00000  -0.00406   0.00534   0.00534   0.00276   0.00206 
-#>      MAPE 
-#>       NaN
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{m3\_weighted }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ sprint\_velocity,}
+  \AttributeTok{time =}\NormalTok{ sprint\_time,}
+  \AttributeTok{weights =} \DecValTok{1} \SpecialCharTok{/}\NormalTok{ (sprint\_velocity }\SpecialCharTok{+} \DecValTok{1}\NormalTok{)}
+\NormalTok{)}
+
+\NormalTok{m3\_weighted}
+\CommentTok{\#\textgreater{} Estimated model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                9.00                1.30                6.92 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               15.58                0.00                0.00 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}   0.00108   1.00000  {-}0.00406   0.00534   0.00534   0.00276 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}   0.00206       NaN}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-Mixed-effects model using radar data is done using \texttt{mixed\_model\_using\_radar()} function. To perform mixed model, let's load data that comes with \pkg{shorts} package.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> data("radar_gun_data")
-R> 
-R> m4 <- mixed_model_using_radar(
-+   radar_gun_data,
-+   time = "time",
-+   velocity = "velocity",
-+   athlete = "athlete"
-+ )
-R> 
-R> m4
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated fixed model parameters
-#> --------------------------------
-#>                 MSS                 TAU                 MAC 
-#>                8.30                1.01                8.24 
-#>                PMAX     time_correction distance_correction 
-#>               17.09                0.00                0.00 
-#> 
-#> Estimated random model parameters
-#> ---------------------------------
-#>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
-#> 1     James 10.00 1.111 9.00 22.5               0                   0
-#> 2       Jim  8.00 0.889 9.00 18.0               0                   0
-#> 3      John  8.00 1.069 7.48 15.0               0                   0
-#> 4 Kimberley  9.01 1.286 7.01 15.8               0                   0
-#> 5  Samantha  6.50 0.685 9.50 15.4               0                   0
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>    0.0516    0.9994   -0.2191    0.1983    0.2191    0.0516    0.0395 
-#>      MAPE 
-#>       NaN
-\end{CodeOutput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot(m4) + theme_bw(8)
-\end{CodeInput}
+Mixed-effects model using radar data is done using \texttt{mixed\_model\_using\_radar()} function. To perform mixed model, let's load data that comes with \textbf{shorts} package.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{data}\NormalTok{(}\StringTok{"radar\_gun\_data"}\NormalTok{)}
+
+\NormalTok{m4 }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_radar}\NormalTok{(}
+\NormalTok{  radar\_gun\_data,}
+  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+  \AttributeTok{velocity =} \StringTok{"velocity"}\NormalTok{,}
+  \AttributeTok{athlete =} \StringTok{"athlete"}
+\NormalTok{)}
+
+\NormalTok{m4}
+\CommentTok{\#\textgreater{} Estimated fixed model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                8.30                1.01                8.24 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               17.09                0.00                0.00 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Estimated random model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}     athlete   MSS   TAU  MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1     James 10.00 1.111 9.00 22.5               0}
+\CommentTok{\#\textgreater{} 2       Jim  8.00 0.889 9.00 18.0               0}
+\CommentTok{\#\textgreater{} 3      John  8.00 1.069 7.48 15.0               0}
+\CommentTok{\#\textgreater{} 4 Kimberley  9.01 1.286 7.01 15.8               0}
+\CommentTok{\#\textgreater{} 5  Samantha  6.50 0.685 9.50 15.4               0}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                   0}
+\CommentTok{\#\textgreater{} 2                   0}
+\CommentTok{\#\textgreater{} 3                   0}
+\CommentTok{\#\textgreater{} 4                   0}
+\CommentTok{\#\textgreater{} 5                   0}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}    0.0516    0.9994   {-}0.2191    0.1983    0.2191    0.0516 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}    0.0395       NaN}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(m4) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
-
-\end{CodeChunk}
 
@@
 When estimating FVP, athlete body mass (kg) can be set using \texttt{bodymass} parameter, while the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{get\_air\_resistance()} function.
-Details of the FVP method implemented in the \pkg{shorts} package, as well as the interpretation from a sprint training perspective, are covered elsewhere \citep{haugenPowerForceVelocityProfilingSprinting2020, morinInterpretingPowerForceVelocityProfiles2016, morinSimpleMethodComputing2019, samozinoSimpleMethodMeasuring2016}.
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # To create Force-Velocity Profile
-R> fvp <- get_FV_profile(
-+   MSS = m1$parameters$MSS,
-+   TAU = m1$parameters$TAU,
-+   bodymass = 80,
-+   # Additional parameters forwarded to get_air_resistance
-+   # Otherwise, defaults are used
-+   bodyheight = 1.85,
-+   barometric_pressure = 780,
-+   air_temperature = 20,
-+   wind_velocity = 0.5
-+ )
-R> 
-R> fvp
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated Force-Velocity Profile
-#> --------------------------------
-#>      bodymass            F0        F0_rel            V0          Pmax 
-#>      80.00000     544.51032       6.80638       9.36184    1274.40402 
-#> Pmax_relative      FV_slope  RFmax_cutoff         RFmax           Drf 
-#>      15.93005      -0.72703       0.30000       0.48779      -0.06676 
-#>        RSE_FV       RSE_Drf 
-#>       1.54814       0.00447
-\end{CodeOutput}
-\end{CodeChunk}
+Details of the FVP method implemented in the \textbf{shorts} package, as well as the interpretation from a sprint training perspective, are covered elsewhere (\protect\hyperlink{ref-haugenPowerForceVelocityProfilingSprinting2020}{Thomas A.
+Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-morinInterpretingPowerForceVelocityProfiles2016}{Jean-Benoît Morin and Samozino 2016}; \protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
+2016}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# To create Force{-}Velocity Profile}
+\NormalTok{fvp }\OtherTok{\textless{}{-}} \FunctionTok{get\_FV\_profile}\NormalTok{(}
+  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
+  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
+  \CommentTok{\# Otherwise, defaults are used}
+  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
+  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
+  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
+  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
+\NormalTok{)}
+
+\NormalTok{fvp}
+\CommentTok{\#\textgreater{} Estimated Force{-}Velocity Profile}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}      bodymass            F0        F0\_rel            V0 }
+\CommentTok{\#\textgreater{}      80.00000     544.51032       6.80638       9.36184 }
+\CommentTok{\#\textgreater{}          Pmax Pmax\_relative      FV\_slope  RFmax\_cutoff }
+\CommentTok{\#\textgreater{}    1274.40402      15.93005      {-}0.72703       0.30000 }
+\CommentTok{\#\textgreater{}         RFmax           Drf        RSE\_FV       RSE\_Drf }
+\CommentTok{\#\textgreater{}       0.48779      {-}0.06676       1.54814       0.00447}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot(fvp) + theme_bw(8)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(fvp) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 To plot FVP estimated kinetics against time, use \texttt{type\ =\ "time"} parameter:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot(fvp, "time") + theme_bw(8)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(fvp, }\StringTok{"time"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-23-1} \end{center}
-
-\end{CodeChunk}
 
@@
 There is a challenge when collecting sprint data that could have a substantial impact on modeled outcomes.
-To ensure accurate parameter outcomes, the initial force production must be synced with start time \citep{haugenPowerForceVelocityProfilingSprinting2020, haugenSprintMechanicalVariables2019}.
-Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \pkg{shorts} package.
+To ensure accurate parameter outcomes, the initial force production must be synced with start time (\protect\hyperlink{ref-haugenPowerForceVelocityProfilingSprinting2020}{Thomas A.
+Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-haugenSprintMechanicalVariables2019}{Thomas A.
+Haugen, Breitschädel, and Seiler 2019}).
+Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the \textbf{shorts} package.
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- tibble(
-+   `true time` = sprint_time,
-+   velocity = sprint_velocity,
-+   `0.5s added` = `true time` + 0.5,
-+   `0.5s deducted` = `true time` - 0.5
-+ )
-R> 
-R> plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")
-R> 
-R> ggplot(
-+   plot_df,
-+   aes(x = value, y = velocity, color = `Sync issue`)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   xlab("Time (s)") +
-+   ylab(expression("Velocity (" * ms^-1 * ")")) +
-+   theme(
-+     legend.title = element_blank(), 
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
+  \StringTok{\textasciigrave{}}\AttributeTok{true time}\StringTok{\textasciigrave{}} \OtherTok{=}\NormalTok{ sprint\_time,}
+  \AttributeTok{velocity =}\NormalTok{ sprint\_velocity,}
+  \StringTok{\textasciigrave{}}\AttributeTok{0.5s added}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{\textasciigrave{}}\AttributeTok{true time}\StringTok{\textasciigrave{}} \SpecialCharTok{+} \FloatTok{0.5}\NormalTok{,}
+  \StringTok{\textasciigrave{}}\AttributeTok{0.5s deducted}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{\textasciigrave{}}\AttributeTok{true time}\StringTok{\textasciigrave{}} \SpecialCharTok{{-}} \FloatTok{0.5}
+\NormalTok{)}
+
+\NormalTok{plot\_df }\OtherTok{\textless{}{-}} \FunctionTok{pivot\_longer}\NormalTok{(df, }\AttributeTok{cols =} \SpecialCharTok{{-}}\DecValTok{2}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"Sync issue"}\NormalTok{)}
+
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  plot\_df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ value, }\AttributeTok{y =}\NormalTok{ velocity, }\AttributeTok{color =} \StringTok{\textasciigrave{}}\AttributeTok{Sync issue}\StringTok{\textasciigrave{}}\NormalTok{)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\FunctionTok{expression}\NormalTok{(}\StringTok{"Velocity ("} \SpecialCharTok{*}\NormalTok{ ms}\SpecialCharTok{\^{}{-}}\DecValTok{1} \SpecialCharTok{*} \StringTok{")"}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following three models estimate MSS and TAU from the three datasets:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Without synchronization issues
-R> m5 <- model_using_radar(
-+   velocity = df$velocity,
-+   time = df$`true time`
-+ )
-R> 
-R> # With time added
-R> m6 <- model_using_radar(
-+   velocity = df$velocity,
-+   time = df$`0.5s added`
-+ )
-R> 
-R> # With time deducted
-R> m7 <- model_using_radar(
-+   velocity = df$velocity,
-+   time = df$`0.5s deducted`
-+ )
-R> 
-R> rbind(
-+   data.frame(
-+     model = "True time",
-+     t(coef(m5))
-+   ),
-+   data.frame(
-+     model = "Added 0.5s time",
-+     t(coef(m6))
-+   ),
-+   data.frame(
-+     model = "Deducted 0.5s time",
-+     t(coef(m7))
-+   )
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#>                model   MSS  TAU  MAC PMAX time_correction
-#> 1          True time  9.00 1.30 6.92 15.6               0
-#> 2    Added 0.5s time  9.91 2.34 4.23 10.5               0
-#> 3 Deducted 0.5s time 10.08 1.86 5.43 13.7               0
-#>   distance_correction
-#> 1                   0
-#> 2                   0
-#> 3                   0
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Without synchronization issues}
+\NormalTok{m5 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{velocity,}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{true time}\StringTok{\textasciigrave{}}
+\NormalTok{)}
+
+\CommentTok{\# With time added}
+\NormalTok{m6 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{velocity,}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{0.5s added}\StringTok{\textasciigrave{}}
+\NormalTok{)}
+
+\CommentTok{\# With time deducted}
+\NormalTok{m7 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{velocity,}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{0.5s deducted}\StringTok{\textasciigrave{}}
+\NormalTok{)}
+
+\FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"True time"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(m5))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Added 0.5s time"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(m6))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Deducted 0.5s time"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(m7))}
+\NormalTok{  )}
+\NormalTok{)}
+\CommentTok{\#\textgreater{}                model   MSS  TAU  MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1          True time  9.00 1.30 6.92 15.6               0}
+\CommentTok{\#\textgreater{} 2    Added 0.5s time  9.91 2.34 4.23 10.5               0}
+\CommentTok{\#\textgreater{} 3 Deducted 0.5s time 10.08 1.86 5.43 13.7               0}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                   0}
+\CommentTok{\#\textgreater{} 2                   0}
+\CommentTok{\#\textgreater{} 3                   0}
+\end{Highlighting}
+\end{Shaded}
 
 As can be seen from the example, all estimated parameters are affected by an error in synchronization of time with velocity (with MSS being the least affected in this example).
-The potential solution incorporated into the \pkg{shorts} package involves estimation of the \emph{time correction} parameter using the following equation:
+The potential solution incorporated into the \textbf{shorts} package involves estimation of the \emph{time correction} parameter using the following equation:
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # With time added
-R> m8 <- model_using_radar_with_time_correction(
-+   velocity = df$velocity,
-+   time = df$`0.5s added`
-+ )
-R> coef(m8)
-\end{CodeInput}
-\begin{CodeOutput}
-#>                 MSS                 TAU                 MAC 
-#>                9.00                1.30                6.92 
-#>                PMAX     time_correction distance_correction 
-#>               15.58               -0.50                0.00
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # With time deducted
-R> m9 <- model_using_radar_with_time_correction(
-+   velocity = df$velocity,
-+   time = df$`0.5s deducted`
-+ )
-R> coef(m9)
-\end{CodeInput}
-\begin{CodeOutput}
-#>                 MSS                 TAU                 MAC 
-#>                9.00                1.30                6.92 
-#>                PMAX     time_correction distance_correction 
-#>               15.58                0.50                0.00
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# With time added}
+\NormalTok{m8 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{velocity,}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{0.5s added}\StringTok{\textasciigrave{}}
+\NormalTok{)}
+\FunctionTok{coef}\NormalTok{(m8)}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                9.00                1.30                6.92 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               15.58               {-}0.50                0.00}
+
+\CommentTok{\# With time deducted}
+\NormalTok{m9 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_radar\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{velocity =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{velocity,}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{0.5s deducted}\StringTok{\textasciigrave{}}
+\NormalTok{)}
+\FunctionTok{coef}\NormalTok{(m9)}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                9.00                1.30                6.92 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               15.58                0.50                0.00}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Using the true time
-R> predict_velocity_at_time(
-+   time = df$`true time`,
-+   MSS = m5$parameters$MSS,
-+   TAU = m5$parameters$TAU
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 0.00 4.83 7.07 8.11 8.59 8.81 8.91
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # Using time with sync issues
-R> predict_velocity_at_time(
-+   time = df$`0.5s added`,
-+   MSS = m8$parameters$MSS,
-+   TAU = m8$parameters$TAU,
-+   time_correction = m8$parameters$time_correction
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#> [1] 0.0000782 4.8299729 7.0681475 8.1053182 8.5859434 8.8086652
-#> [7] 8.9118746
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Using the true time}
+\FunctionTok{predict\_velocity\_at\_time}\NormalTok{(}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{true time}\StringTok{\textasciigrave{}}\NormalTok{,}
+  \AttributeTok{MSS =}\NormalTok{ m5}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m5}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 0.00 4.83 7.07 8.11 8.59 8.81 8.91}
+
+\CommentTok{\# Using time with sync issues}
+\FunctionTok{predict\_velocity\_at\_time}\NormalTok{(}
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\StringTok{\textasciigrave{}}\AttributeTok{0.5s added}\StringTok{\textasciigrave{}}\NormalTok{,}
+  \AttributeTok{MSS =}\NormalTok{ m8}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ m8}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \AttributeTok{time\_correction =}\NormalTok{ m8}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{time\_correction}
+\NormalTok{)}
+\CommentTok{\#\textgreater{} [1] 0.0000782 4.8299729 7.0681475 8.1053182 8.5859434 8.8086652}
+\CommentTok{\#\textgreater{} [7] 8.9118746}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Adding 0.5s to radar_gun_data
-R> radar_gun_data$time <- radar_gun_data$time + 0.5
-R> 
-R> # Mixed model with time correction being fixed effect
-R> m10 <- mixed_model_using_radar_with_time_correction(
-+   radar_gun_data,
-+   time = "time",
-+   velocity = "velocity",
-+   athlete = "athlete",
-+   random = MSS + TAU ~ 1
-+ )
-R> 
-R> m10
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated fixed model parameters
-#> --------------------------------
-#>                 MSS                 TAU                 MAC 
-#>                8.30                1.01                8.24 
-#>                PMAX     time_correction distance_correction 
-#>               17.10               -0.50                0.00 
-#> 
-#> Estimated random model parameters
-#> ---------------------------------
-#>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
-#> 1     James 10.00 1.111 9.00 22.5            -0.5                   0
-#> 2       Jim  8.00 0.889 9.00 18.0            -0.5                   0
-#> 3      John  8.00 1.069 7.48 15.0            -0.5                   0
-#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5                   0
-#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5                   0
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>    0.0516    0.9994   -0.2190    0.1983    0.2190    0.0516    0.0395 
-#>      MAPE 
-#>       Inf
-\end{CodeOutput}
-\begin{CodeInput}
-R> 
-R> # Mixed model with time correction being random effect
-R> m11 <- mixed_model_using_radar_with_time_correction(
-+   radar_gun_data,
-+   time = "time",
-+   velocity = "velocity",
-+   athlete = "athlete",
-+   random = MSS + TAU + time_correction ~ 1
-+ )
-R> 
-R> m11
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated fixed model parameters
-#> --------------------------------
-#>                 MSS                 TAU                 MAC 
-#>                8.30                1.01                8.24 
-#>                PMAX     time_correction distance_correction 
-#>               17.10               -0.50                0.00 
-#> 
-#> Estimated random model parameters
-#> ---------------------------------
-#>     athlete   MSS   TAU  MAC PMAX time_correction distance_correction
-#> 1     James 10.00 1.110 9.00 22.5            -0.5                   0
-#> 2       Jim  8.00 0.889 9.00 18.0            -0.5                   0
-#> 3      John  8.00 1.069 7.48 15.0            -0.5                   0
-#> 4 Kimberley  9.01 1.285 7.01 15.8            -0.5                   0
-#> 5  Samantha  6.50 0.685 9.50 15.4            -0.5                   0
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>    0.0516    0.9994   -0.2188    0.1982    0.2188    0.0516    0.0395 
-#>      MAPE 
-#>       Inf
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Adding 0.5s to radar\_gun\_data}
+\NormalTok{radar\_gun\_data}\SpecialCharTok{$}\NormalTok{time }\OtherTok{\textless{}{-}}\NormalTok{ radar\_gun\_data}\SpecialCharTok{$}\NormalTok{time }\SpecialCharTok{+} \FloatTok{0.5}
+
+\CommentTok{\# Mixed model with time correction being fixed effect}
+\NormalTok{m10 }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_radar\_with\_time\_correction}\NormalTok{(}
+\NormalTok{  radar\_gun\_data,}
+  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+  \AttributeTok{velocity =} \StringTok{"velocity"}\NormalTok{,}
+  \AttributeTok{athlete =} \StringTok{"athlete"}\NormalTok{,}
+  \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{)}
+
+\NormalTok{m10}
+\CommentTok{\#\textgreater{} Estimated fixed model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                8.30                1.01                8.24 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               17.10               {-}0.50                0.00 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Estimated random model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}     athlete   MSS   TAU  MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1     James 10.00 1.111 9.00 22.5            {-}0.5}
+\CommentTok{\#\textgreater{} 2       Jim  8.00 0.889 9.00 18.0            {-}0.5}
+\CommentTok{\#\textgreater{} 3      John  8.00 1.069 7.48 15.0            {-}0.5}
+\CommentTok{\#\textgreater{} 4 Kimberley  9.01 1.285 7.01 15.8            {-}0.5}
+\CommentTok{\#\textgreater{} 5  Samantha  6.50 0.685 9.50 15.4            {-}0.5}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                   0}
+\CommentTok{\#\textgreater{} 2                   0}
+\CommentTok{\#\textgreater{} 3                   0}
+\CommentTok{\#\textgreater{} 4                   0}
+\CommentTok{\#\textgreater{} 5                   0}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}    0.0516    0.9994   {-}0.2190    0.1983    0.2190    0.0516 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}    0.0395       Inf}
+
+\CommentTok{\# Mixed model with time correction being random effect}
+\NormalTok{m11 }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_radar\_with\_time\_correction}\NormalTok{(}
+\NormalTok{  radar\_gun\_data,}
+  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+  \AttributeTok{velocity =} \StringTok{"velocity"}\NormalTok{,}
+  \AttributeTok{athlete =} \StringTok{"athlete"}\NormalTok{,}
+  \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{)}
+
+\NormalTok{m11}
+\CommentTok{\#\textgreater{} Estimated fixed model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}                8.30                1.01                8.24 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}               17.10               {-}0.50                0.00 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Estimated random model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}     athlete   MSS   TAU  MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1     James 10.00 1.110 9.00 22.5            {-}0.5}
+\CommentTok{\#\textgreater{} 2       Jim  8.00 0.889 9.00 18.0            {-}0.5}
+\CommentTok{\#\textgreater{} 3      John  8.00 1.069 7.48 15.0            {-}0.5}
+\CommentTok{\#\textgreater{} 4 Kimberley  9.01 1.285 7.01 15.8            {-}0.5}
+\CommentTok{\#\textgreater{} 5  Samantha  6.50 0.685 9.50 15.4            {-}0.5}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                   0}
+\CommentTok{\#\textgreater{} 2                   0}
+\CommentTok{\#\textgreater{} 3                   0}
+\CommentTok{\#\textgreater{} 4                   0}
+\CommentTok{\#\textgreater{} 5                   0}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}    0.0516    0.9994   {-}0.2188    0.1982    0.2188    0.0516 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}    0.0395       Inf}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> MSS <- 9
-R> TAU <- 1.3
-R> MAC <- MSS / TAU
-R> 
-R> split_times <- tibble(
-+   distance = c(5, 10, 20, 30, 40),
-+   john_time = predict_time_at_distance(distance, MSS, TAU),
-+ 
-+   # Jack's performance
-+   jack_distance = distance + 0.5,
-+   jack_true_time = predict_time_at_distance(jack_distance, MSS, TAU),
-+   time_05m = predict_time_at_distance(0.5, MSS, TAU),
-+   jack_time = jack_true_time - time_05m
-+ )
-R> 
-R> split_times
-\end{CodeInput}
-\begin{CodeOutput}
-#> # A tibble: 5 x 6
-#>   distance john_time jack_distance jack_true_time time_05m jack_time
-#>      <dbl>  <I<dbl>>         <dbl>       <I<dbl>> <I<dbl>>  <I<dbl>>
-#> 1        5      1.42           5.5           1.50    0.400      1.10
-#> 2       10      2.17          10.5           2.23    0.400      1.83
-#> 3       20      3.43          20.5           3.49    0.400      3.09
-#> 4       30      4.60          30.5           4.65    0.400      4.25
-#> 5       40      5.73          40.5           5.78    0.400      5.39
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{MSS }\OtherTok{\textless{}{-}} \DecValTok{9}
+\NormalTok{TAU }\OtherTok{\textless{}{-}} \FloatTok{1.3}
+\NormalTok{MAC }\OtherTok{\textless{}{-}}\NormalTok{ MSS }\SpecialCharTok{/}\NormalTok{ TAU}
+
+\NormalTok{split\_times }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{),}
+  \AttributeTok{john\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(distance, MSS, TAU),}
+
+  \CommentTok{\# Jack\textquotesingle{}s performance}
+  \AttributeTok{jack\_distance =}\NormalTok{ distance }\SpecialCharTok{+} \FloatTok{0.5}\NormalTok{,}
+  \AttributeTok{jack\_true\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(jack\_distance, MSS, TAU),}
+  \AttributeTok{time\_05m =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}\FloatTok{0.5}\NormalTok{, MSS, TAU),}
+  \AttributeTok{jack\_time =}\NormalTok{ jack\_true\_time }\SpecialCharTok{{-}}\NormalTok{ time\_05m}
+\NormalTok{)}
+
+\NormalTok{split\_times}
+\CommentTok{\#\textgreater{} \# A tibble: 5 x 6}
+\CommentTok{\#\textgreater{}   distance john\_time jack\_distance jack\_true\_time time\_05m}
+\CommentTok{\#\textgreater{}      \textless{}dbl\textgreater{}  \textless{}I\textless{}dbl\textgreater{}\textgreater{}         \textless{}dbl\textgreater{}       \textless{}I\textless{}dbl\textgreater{}\textgreater{} \textless{}I\textless{}dbl\textgreater{}\textgreater{}}
+\CommentTok{\#\textgreater{} 1        5      1.42           5.5           1.50    0.400}
+\CommentTok{\#\textgreater{} 2       10      2.17          10.5           2.23    0.400}
+\CommentTok{\#\textgreater{} 3       20      3.43          20.5           3.49    0.400}
+\CommentTok{\#\textgreater{} 4       30      4.60          30.5           4.65    0.400}
+\CommentTok{\#\textgreater{} 5       40      5.73          40.5           5.78    0.400}
+\CommentTok{\#\textgreater{} \# ... with 1 more variable: jack\_time \textless{}I\textless{}dbl\textgreater{}\textgreater{}}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> plot_df <- split_times %>%
-+   select(distance, john_time, jack_time) %>%
-+   rename(John = john_time, Jack = jack_time) %>%
-+   pivot_longer(cols = -1, names_to = "athlete", values_to = "time") %>%
-+   mutate(distance = factor(distance))
-R> 
-R> ggplot(
-+   plot_df,
-+   aes(x = distance, y = time, color = athlete, group = athlete)
-+ ) +
-+   theme_bw(8) +
-+   geom_point() +
-+   geom_line() +
-+   xlab("Distance (m)") +
-+   ylab("Time (s)") +
-+   theme(
-+     legend.title = element_blank(), 
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot\_df }\OtherTok{\textless{}{-}}\NormalTok{ split\_times }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(distance, john\_time, jack\_time) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{rename}\NormalTok{(}\AttributeTok{John =}\NormalTok{ john\_time, }\AttributeTok{Jack =}\NormalTok{ jack\_time) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"athlete"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"time"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{distance =} \FunctionTok{factor}\NormalTok{(distance))}
+
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  plot\_df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{y =}\NormalTok{ time, }\AttributeTok{color =}\NormalTok{ athlete, }\AttributeTok{group =}\NormalTok{ athlete)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 Using the following code, we can see the differences in estimated MSS and TAU parameters:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Since this is a perfect simulation and stats::nls will complain
-R> # we need to add very small noise, or measurement error to the times
-R> set.seed(1667)
-R> rand_noise <- rnorm(nrow(split_times), 0, 10^-5)
-R> split_times$john_time <- split_times$john_time + rand_noise
-R> split_times$jack_time <- split_times$jack_time + rand_noise
-R> 
-R> john_profile <- model_using_splits(
-+   distance = split_times$distance,
-+   time = split_times$john_time
-+ )
-R> 
-R> jack_profile <- model_using_splits(
-+   distance = split_times$distance,
-+   time = split_times$jack_time
-+ )
-R> 
-R> sprint_parameters <- rbind(
-+   coef(john_profile),
-+   coef(jack_profile)
-+ )
-R> 
-R> rownames(sprint_parameters) <- c("John", "Jack")
-R> 
-R> sprint_parameters
-\end{CodeInput}
-\begin{CodeOutput}
-#>       MSS   TAU   MAC PMAX time_correction distance_correction
-#> John 9.00 1.300  6.92 15.6               0                   0
-#> Jack 8.49 0.704 12.06 25.6               0                   0
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Since this is a perfect simulation and stats::nls will complain}
+\CommentTok{\# we need to add very small noise, or measurement error to the times}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1667}\NormalTok{)}
+\NormalTok{rand\_noise }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(split\_times), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{5}\NormalTok{)}
+\NormalTok{split\_times}\SpecialCharTok{$}\NormalTok{john\_time }\OtherTok{\textless{}{-}}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{john\_time }\SpecialCharTok{+}\NormalTok{ rand\_noise}
+\NormalTok{split\_times}\SpecialCharTok{$}\NormalTok{jack\_time }\OtherTok{\textless{}{-}}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time }\SpecialCharTok{+}\NormalTok{ rand\_noise}
+
+\NormalTok{john\_profile }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{john\_time}
+\NormalTok{)}
+
+\NormalTok{jack\_profile }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time}
+\NormalTok{)}
+
+\NormalTok{sprint\_parameters }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{coef}\NormalTok{(john\_profile),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile)}
+\NormalTok{)}
+
+\FunctionTok{rownames}\NormalTok{(sprint\_parameters) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"John"}\NormalTok{, }\StringTok{"Jack"}\NormalTok{)}
+
+\NormalTok{sprint\_parameters}
+\CommentTok{\#\textgreater{}       MSS   TAU   MAC PMAX time\_correction distance\_correction}
+\CommentTok{\#\textgreater{} John 9.00 1.300  6.92 15.6               0                   0}
+\CommentTok{\#\textgreater{} Jack 8.49 0.704 12.06 25.6               0                   0}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> sim_df <- expand.grid(
-+   MSS = c(6, 7, 8, 9),
-+   MAC = c(6, 7, 8, 9),
-+   flying_start_distance = c(
-+     seq(0, 0.001, length.out = 20),
-+     seq(0.001, 0.01, length.out = 20),
-+     seq(0.01, 0.1, length.out = 20),
-+     seq(0.1, 1, length.out = 20)
-+   ),
-+   distance = c(5, 10, 20, 30, 40, 50)
-+ )
-R> 
-R> sim_df <- sim_df %>%
-+   mutate(
-+     TAU = MSS / MAC,
-+     PMAX = MSS * MAC / 4,
-+     true_distance = distance + flying_start_distance,
-+     true_time = predict_time_at_distance(true_distance, MSS, TAU),
-+     stolen_time = predict_time_at_distance(flying_start_distance, MSS, TAU),
-+     time = true_time - stolen_time
-+   )
-R> 
-R> # Add small noise to allow model fit
-R> set.seed(1667)
-R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
-R> sim_df$time <- sim_df$time + rand_noise
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sim\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand.grid}\NormalTok{(}
+  \AttributeTok{MSS =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
+  \AttributeTok{MAC =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
+  \AttributeTok{flying\_start\_distance =} \FunctionTok{c}\NormalTok{(}
+    \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.001}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{),}
+    \FunctionTok{seq}\NormalTok{(}\FloatTok{0.001}\NormalTok{, }\FloatTok{0.01}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{),}
+    \FunctionTok{seq}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{),}
+    \FunctionTok{seq}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{)}
+\NormalTok{  ),}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{50}\NormalTok{)}
+\NormalTok{)}
+
+\NormalTok{sim\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{TAU =}\NormalTok{ MSS }\SpecialCharTok{/}\NormalTok{ MAC,}
+    \AttributeTok{PMAX =}\NormalTok{ MSS }\SpecialCharTok{*}\NormalTok{ MAC }\SpecialCharTok{/} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{true\_distance =}\NormalTok{ distance }\SpecialCharTok{+}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{true\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(true\_distance, MSS, TAU),}
+    \AttributeTok{stolen\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+\NormalTok{      flying\_start\_distance, MSS, TAU),}
+    \AttributeTok{time =}\NormalTok{ true\_time }\SpecialCharTok{{-}}\NormalTok{ stolen\_time}
+\NormalTok{  )}
+
+\CommentTok{\# Add small noise to allow model fit}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1667}\NormalTok{)}
+\NormalTok{rand\_noise }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(sim\_df), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{5}\NormalTok{)}
+\NormalTok{sim\_df}\SpecialCharTok{$}\NormalTok{time }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df}\SpecialCharTok{$}\NormalTok{time }\SpecialCharTok{+}\NormalTok{ rand\_noise}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Prediction wrapper
-R> pred_wrapper <- function(data) {
-+   model <- model_using_splits(
-+     distance = data$distance,
-+     time = data$time
-+   )
-+ 
-+   params <- data.frame(t(coef(model)))
-+ 
-+   predicted_time <- predict_time_at_distance(
-+     distance = data$distance,
-+     MSS = model$parameters$MSS,
-+     TAU = model$parameters$TAU
-+   )
-+ 
-+   colnames(params) <- c(
-+     "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
-+     "est_time_correction", "est_distance_correction"
-+   )
-+ 
-+   cbind(
-+     data,
-+     params,
-+     data.frame(predicted_time = as.numeric(predicted_time))
-+   )
-+ }
-R> 
-R> # estimated parameters and predicted time
-R> model_df <- sim_df %>%
-+   group_by(MSS, TAU, flying_start_distance) %>%
-+   do(pred_wrapper(.)) %>%
-+   ungroup()
-R> 
-R> # Prediction residuals
-R> model_df$residuals <- model_df$predicted_time - model_df$time
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Prediction wrapper}
+\NormalTok{pred\_wrapper }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data) \{}
+\NormalTok{  model }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time}
+\NormalTok{  )}
+
+\NormalTok{  params }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(model)))}
+
+\NormalTok{  predicted\_time }\OtherTok{\textless{}{-}} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{MSS =}\NormalTok{ model}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+    \AttributeTok{TAU =}\NormalTok{ model}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU}
+\NormalTok{  )}
+
+  \FunctionTok{colnames}\NormalTok{(params) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"est\_MSS"}\NormalTok{, }\StringTok{"est\_TAU"}\NormalTok{, }\StringTok{"est\_MAC"}\NormalTok{, }\StringTok{"est\_PMAX"}\NormalTok{,}
+    \StringTok{"est\_time\_correction"}\NormalTok{, }\StringTok{"est\_distance\_correction"}
+\NormalTok{  )}
+
+  \FunctionTok{cbind}\NormalTok{(}
+\NormalTok{    data,}
+\NormalTok{    params,}
+    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{predicted\_time =} \FunctionTok{as.numeric}\NormalTok{(predicted\_time))}
+\NormalTok{  )}
+\NormalTok{\}}
+
+\CommentTok{\# estimated parameters and predicted time}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{do}\NormalTok{(}\FunctionTok{pred\_wrapper}\NormalTok{(.)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{()}
+
+\CommentTok{\# Prediction residuals}
+\NormalTok{model\_df}\SpecialCharTok{$}\NormalTok{residuals }\OtherTok{\textless{}{-}}\NormalTok{ model\_df}\SpecialCharTok{$}\NormalTok{predicted\_time }\SpecialCharTok{{-}}\NormalTok{ model\_df}\SpecialCharTok{$}\NormalTok{time}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Estimates plot
-R> df <- model_df %>%
-+   group_by(MSS, TAU, flying_start_distance) %>%
-+   slice(1) %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
-+   )
-R> 
-R> # MSS
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_MSS, color = MAC_string)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_wrap(~MSS_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated MSS (m/s)") +
-+   theme(
-+     legend.title = element_blank(), 
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Estimates plot}
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{slice}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
+    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
+    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{))}
+\NormalTok{  )}
+
+\CommentTok{\# MSS}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MSS, }\AttributeTok{color =}\NormalTok{ MAC\_string)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MSS (m/s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 As can be seen from the figure, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # MAC
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_MAC, color = MSS_string)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_wrap(~MAC_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated MAC (m/s/s)") +
-+   theme(
-+     legend.title = element_blank(), 
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# MAC}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MAC, }\AttributeTok{color =}\NormalTok{ MSS\_string)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MAC (m/s/s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # PMAX
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_wrap(~MAC_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated PMAX (W/kg)") +
-+   theme(
-+     legend.title = element_blank(), 
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# PMAX}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_PMAX, }\AttributeTok{color =}\NormalTok{ MSS\_string)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated PMAX (W/kg)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Residuals
-R> model_df <- model_df %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-+     group = paste(MSS, MAC, flying_start_distance)
-+   )
-R> 
-R> ggplot(
-+   model_df,
-+   aes(
-+     y = residuals,
-+     x = distance,
-+     color = flying_start_distance,
-+     group = group)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.3) +
-+   facet_grid(MSS_string ~ MAC_string) +
-+   geom_hline(yintercept = 0, linetype = "dashed") +
-+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-+   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)") +
-+   theme(legend.position = "top") + 
-+   labs(color = "Flying start distance")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Residuals}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
+    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
+    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{group =} \FunctionTok{paste}\NormalTok{(MSS, MAC, flying\_start\_distance)}
+\NormalTok{  )}
+
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  model\_df,}
+  \FunctionTok{aes}\NormalTok{(}
+    \AttributeTok{y =}\NormalTok{ residuals,}
+    \AttributeTok{x =}\NormalTok{ distance,}
+    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{group =}\NormalTok{ group)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MSS\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MAC\_string) }\SpecialCharTok{+}
+  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler figure conveying issues with residuals when there is a flying start:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> ggplot(
-+   model_df,
-+   aes(
-+     y = residuals,
-+     x = distance,
-+     color = flying_start_distance,
-+     group = group)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.3) +
-+   geom_hline(yintercept = 0, linetype = "dashed") +
-+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-+   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)") +
-+   theme(legend.position = "top") + 
-+   labs(color = "Flying start distance")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  model\_df,}
+  \FunctionTok{aes}\NormalTok{(}
+    \AttributeTok{y =}\NormalTok{ residuals,}
+    \AttributeTok{x =}\NormalTok{ distance,}
+    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{group =}\NormalTok{ group)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
-
-\end{CodeChunk}
 
@@
 Since maximal sprint speed is difficult to improve, the effects of start inconsistencies can mask effects of the training intervention.
-It is thus crucial to standardize the start when testing and implementing the following techniques when using the \pkg{shorts} package.
+It is thus crucial to standardize the start when testing and implementing the following techniques when using the \textbf{shorts} package.
 
@@
 
-A potential solution is to use a correction factor - the recommendation in the literature is +0.5 s \citep{haugenSprintMechanicalProperties2020, haugenSprintMechanicalVariables2019}.
-Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 s \citep{haugenDifferenceStartImpact2012}.
+A potential solution is to use a correction factor - the recommendation in the literature is +0.5 s (\protect\hyperlink{ref-haugenSprintMechanicalProperties2020}{Thomas A.
+Haugen, Breitschädel, and Seiler 2020}, \protect\hyperlink{ref-haugenSprintMechanicalVariables2019}{2019}).
+Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 s (\protect\hyperlink{ref-haugenDifferenceStartImpact2012}{Thomas A.
+Haugen, Tønnessen, and Seiler 2012}).
 So, while a timing correction factor is warranted to avoid subsequent errors in estimates of kinetic variables (e.g., overestimate power), a correction factor that is too large will have the opposite effect (e.g., underestimate power).
 
-Rather than providing \emph{apriori} time correction from the literature, \pkg{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
-Exactly the same method is suggested by \citet{stenrothForcevelocityProfilingIce2020}, named \emph{time shift method}, and the estimated parameter named \emph{time shift parameter}.
+Rather than providing \emph{apriori} time correction from the literature, \textbf{shorts} package provides an estimation of this parameter from the data provided, together with MSS and TAU.
+Exactly the same method is suggested by \protect\hyperlink{ref-stenrothForcevelocityProfilingIce2020}{Stenroth, Vartiainen, and Karjalainen} (\protect\hyperlink{ref-stenrothForcevelocityProfilingIce2020}{2020}), named \emph{time shift method}, and the estimated parameter named \emph{time shift parameter}.
 We have named this parameter \emph{time correction} to be in agreement with the parameter introduced in \protect\hyperlink{problems-with-time-sync-with-radar-gun}{Problems with time sync with radar gun} section of this paper, as well as the available literature.
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> jack_profile_fixed_time_short <- model_using_splits(
-+   distance = split_times$distance,
-+   time = split_times$jack_time,
-+   time_correction = 0.3
-+ )
-R> 
-R> jack_profile_fixed_time_long <- model_using_splits(
-+   distance = split_times$distance,
-+   time = split_times$jack_time,
-+   time_correction = 0.5
-+ )
-R> 
-R> jack_profile_time_estimated <- model_using_splits_with_time_correction(
-+   distance = split_times$distance,
-+   time = split_times$jack_time
-+ )
-R> 
-R> jack_parameters <- rbind(
-+   coef(john_profile),
-+   coef(jack_profile),
-+   coef(jack_profile_fixed_time_short),
-+   coef(jack_profile_fixed_time_long),
-+   coef(jack_profile_time_estimated)
-+ )
-R> 
-R> rownames(jack_parameters) <- c(
-+   "John",
-+   "Jack - No corrections",
-+   "Jack - Fixed time correction (+0.3s)",
-+   "Jack - Fixed time correction (+0.5s)",
-+   "Jack - Estimated time correction"
-+ )
-R> 
-R> jack_parameters
-\end{CodeInput}
-\begin{CodeOutput}
-#>                                       MSS   TAU   MAC PMAX
-#> John                                 9.00 1.300  6.92 15.6
-#> Jack - No corrections                8.49 0.704 12.06 25.6
-#> Jack - Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2
-#> Jack - Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1
-#> Jack - Estimated time correction     8.96 1.216  7.37 16.5
-#>                                      time_correction
-#> John                                           0.000
-#> Jack - No corrections                          0.000
-#> Jack - Fixed time correction (+0.3s)           0.300
-#> Jack - Fixed time correction (+0.5s)           0.500
-#> Jack - Estimated time correction               0.284
-#>                                      distance_correction
-#> John                                                   0
-#> Jack - No corrections                                  0
-#> Jack - Fixed time correction (+0.3s)                   0
-#> Jack - Fixed time correction (+0.5s)                   0
-#> Jack - Estimated time correction                       0
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jack\_profile\_fixed\_time\_short }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time,}
+  \AttributeTok{time\_correction =} \FloatTok{0.3}
+\NormalTok{)}
+
+\NormalTok{jack\_profile\_fixed\_time\_long }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time,}
+  \AttributeTok{time\_correction =} \FloatTok{0.5}
+\NormalTok{)}
+
+\NormalTok{jack\_profile\_time\_estimated }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time}
+\NormalTok{)}
+
+\NormalTok{jack\_parameters }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{coef}\NormalTok{(john\_profile),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_fixed\_time\_short),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_fixed\_time\_long),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_time\_estimated)}
+\NormalTok{)}
+
+\FunctionTok{rownames}\NormalTok{(jack\_parameters) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+  \StringTok{"John"}\NormalTok{,}
+  \StringTok{"Jack {-} No corrections"}\NormalTok{,}
+  \StringTok{"Jack {-} Fixed time correction (+0.3s)"}\NormalTok{,}
+  \StringTok{"Jack {-} Fixed time correction (+0.5s)"}\NormalTok{,}
+  \StringTok{"Jack {-} Estimated time correction"}
+\NormalTok{)}
+
+\NormalTok{jack\_parameters}
+\CommentTok{\#\textgreater{}                                       MSS   TAU   MAC PMAX}
+\CommentTok{\#\textgreater{} John                                 9.00 1.300  6.92 15.6}
+\CommentTok{\#\textgreater{} Jack {-} No corrections                8.49 0.704 12.06 25.6}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1}
+\CommentTok{\#\textgreater{} Jack {-} Estimated time correction     8.96 1.216  7.37 16.5}
+\CommentTok{\#\textgreater{}                                      time\_correction}
+\CommentTok{\#\textgreater{} John                                           0.000}
+\CommentTok{\#\textgreater{} Jack {-} No corrections                          0.000}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)           0.300}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)           0.500}
+\CommentTok{\#\textgreater{} Jack {-} Estimated time correction               0.284}
+\CommentTok{\#\textgreater{}                                      distance\_correction}
+\CommentTok{\#\textgreater{} John                                                   0}
+\CommentTok{\#\textgreater{} Jack {-} No corrections                                  0}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)                   0}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)                   0}
+\CommentTok{\#\textgreater{} Jack {-} Estimated time correction                       0}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-Another model definition, which is a novel approach implemented in the \pkg{shorts} package, is to utilize \emph{distance correction}, besides time correction.
+Another model definition, which is a novel approach implemented in the \textbf{shorts} package, is to utilize \emph{distance correction}, besides time correction.
 Thus, equation \eqref{eq:time-distance} becomes:
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> jack_profile_distance_correction <- model_using_splits_with_corrections(
-+   distance = split_times$distance,
-+   time = split_times$jack_time
-+ )
-R> 
-R> jack_parameters <- rbind(
-+   coef(john_profile),
-+   coef(jack_profile),
-+   coef(jack_profile_fixed_time_short),
-+   coef(jack_profile_fixed_time_long),
-+   coef(jack_profile_time_estimated),
-+   coef(jack_profile_distance_correction)
-+ )
-R> 
-R> rownames(jack_parameters) <- c(
-+   "John",
-+   "Jack - No corrections",
-+   "Jack - Fixed time correction (+0.3s)",
-+   "Jack - Fixed time correction (+0.5s)",
-+   "Jack - Estimated time correction",
-+   "Jack - Estimated distance correction"
-+ )
-R> 
-R> jack_parameters
-\end{CodeInput}
-\begin{CodeOutput}
-#>                                       MSS   TAU   MAC PMAX
-#> John                                 9.00 1.300  6.92 15.6
-#> Jack - No corrections                8.49 0.704 12.06 25.6
-#> Jack - Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2
-#> Jack - Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1
-#> Jack - Estimated time correction     8.96 1.216  7.37 16.5
-#> Jack - Estimated distance correction 9.00 1.301  6.92 15.6
-#>                                      time_correction
-#> John                                           0.000
-#> Jack - No corrections                          0.000
-#> Jack - Fixed time correction (+0.3s)           0.300
-#> Jack - Fixed time correction (+0.5s)           0.500
-#> Jack - Estimated time correction               0.284
-#> Jack - Estimated distance correction           0.400
-#>                                      distance_correction
-#> John                                               0.000
-#> Jack - No corrections                              0.000
-#> Jack - Fixed time correction (+0.3s)               0.000
-#> Jack - Fixed time correction (+0.5s)               0.000
-#> Jack - Estimated time correction                   0.000
-#> Jack - Estimated distance correction               0.503
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jack\_profile\_distance\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time}
+\NormalTok{)}
+
+\NormalTok{jack\_parameters }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{coef}\NormalTok{(john\_profile),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_fixed\_time\_short),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_fixed\_time\_long),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_time\_estimated),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_distance\_correction)}
+\NormalTok{)}
+
+\FunctionTok{rownames}\NormalTok{(jack\_parameters) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+  \StringTok{"John"}\NormalTok{,}
+  \StringTok{"Jack {-} No corrections"}\NormalTok{,}
+  \StringTok{"Jack {-} Fixed time correction (+0.3s)"}\NormalTok{,}
+  \StringTok{"Jack {-} Fixed time correction (+0.5s)"}\NormalTok{,}
+  \StringTok{"Jack {-} Estimated time correction"}\NormalTok{,}
+  \StringTok{"Jack {-} Estimated distance correction"}
+\NormalTok{)}
+
+\NormalTok{jack\_parameters}
+\CommentTok{\#\textgreater{}                                       MSS   TAU   MAC PMAX}
+\CommentTok{\#\textgreater{} John                                 9.00 1.300  6.92 15.6}
+\CommentTok{\#\textgreater{} Jack {-} No corrections                8.49 0.704 12.06 25.6}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1}
+\CommentTok{\#\textgreater{} Jack {-} Estimated time correction     8.96 1.216  7.37 16.5}
+\CommentTok{\#\textgreater{} Jack {-} Estimated distance correction 9.00 1.301  6.92 15.6}
+\CommentTok{\#\textgreater{}                                      time\_correction}
+\CommentTok{\#\textgreater{} John                                           0.000}
+\CommentTok{\#\textgreater{} Jack {-} No corrections                          0.000}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)           0.300}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)           0.500}
+\CommentTok{\#\textgreater{} Jack {-} Estimated time correction               0.284}
+\CommentTok{\#\textgreater{} Jack {-} Estimated distance correction           0.400}
+\CommentTok{\#\textgreater{}                                      distance\_correction}
+\CommentTok{\#\textgreater{} John                                               0.000}
+\CommentTok{\#\textgreater{} Jack {-} No corrections                              0.000}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)               0.000}
+\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)               0.000}
+\CommentTok{\#\textgreater{} Jack {-} Estimated time correction                   0.000}
+\CommentTok{\#\textgreater{} Jack {-} Estimated distance correction               0.503}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> pred_wrapper <- function(data) {
-+   no_correction <- model_using_splits(
-+     distance = data$distance,
-+     time = data$time
-+   )
-+ 
-+   fixed_correction_short <- model_using_splits(
-+     distance = data$distance,
-+     time = data$time,
-+     time_correction = 0.3
-+   )
-+ 
-+   fixed_correction_long <- model_using_splits(
-+     distance = data$distance,
-+     time = data$time,
-+     time_correction = 0.5
-+   )
-+ 
-+   time_correction <- model_using_splits_with_time_correction(
-+     distance = data$distance,
-+     time = data$time,
-+     control = nls.control(tol = 1)
-+   )
-+ 
-+   time_dist_correction <- model_using_splits_with_corrections(
-+     distance = data$distance,
-+     time = data$time,
-+     control = nls.control(tol = 1)
-+   )
-+ 
-+ 
-+   params <- rbind(
-+     data.frame(
-+       model = "No correction",
-+       t(coef(no_correction))
-+     ),
-+     data.frame(
-+       model = "Fixed correction +0.3s",
-+       t(coef(fixed_correction_short))
-+     ),
-+     data.frame(
-+       model = "Fixed correction +0.5s",
-+       t(coef(fixed_correction_long))
-+     ),
-+     data.frame(
-+       model = "Time correction",
-+       t(coef(time_correction))
-+     ),
-+     data.frame(
-+       model = "Time and distance correction",
-+       t(coef(time_dist_correction))
-+     )
-+   )
-+ 
-+   colnames(params) <- c(
-+     "model", "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
-+     "est_time_correction", "est_distance_correction"
-+   )
-+ 
-+   df <- expand_grid(
-+     data,
-+     params
-+   )
-+ 
-+   df$predicted_time <- predict_time_at_distance(
-+     distance = df$distance,
-+     MSS = df$est_MSS,
-+     TAU = df$est_TAU,
-+     time_correction = df$est_time_correction,
-+     distance_correction = df$est_distance_correction
-+   )
-+ 
-+   df$residuals <- df$predicted_time - df$time
-+   return(df)
-+ }
-R> 
-R> # estimated parameters and predicted time
-R> model_df <- sim_df %>%
-+   group_by(MSS, TAU, flying_start_distance) %>%
-+   do(pred_wrapper(.)) %>%
-+   ungroup()
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pred\_wrapper }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data) \{}
+\NormalTok{  no\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time}
+\NormalTok{  )}
+
+\NormalTok{  fixed\_correction\_short }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
+    \AttributeTok{time\_correction =} \FloatTok{0.3}
+\NormalTok{  )}
+
+\NormalTok{  fixed\_correction\_long }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
+    \AttributeTok{time\_correction =} \FloatTok{0.5}
+\NormalTok{  )}
+
+\NormalTok{  time\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
+    \AttributeTok{control =} \FunctionTok{nls.control}\NormalTok{(}\AttributeTok{tol =} \DecValTok{1}\NormalTok{)}
+\NormalTok{  )}
+
+\NormalTok{  time\_dist\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
+    \AttributeTok{control =} \FunctionTok{nls.control}\NormalTok{(}\AttributeTok{tol =} \DecValTok{1}\NormalTok{)}
+\NormalTok{  )}
+
+\NormalTok{  params }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+    \FunctionTok{data.frame}\NormalTok{(}
+      \AttributeTok{model =} \StringTok{"No correction"}\NormalTok{,}
+      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(no\_correction))}
+\NormalTok{    ),}
+    \FunctionTok{data.frame}\NormalTok{(}
+      \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(fixed\_correction\_short))}
+\NormalTok{    ),}
+    \FunctionTok{data.frame}\NormalTok{(}
+      \AttributeTok{model =} \StringTok{"Fixed correction +0.5s"}\NormalTok{,}
+      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(fixed\_correction\_long))}
+\NormalTok{    ),}
+    \FunctionTok{data.frame}\NormalTok{(}
+      \AttributeTok{model =} \StringTok{"Time correction"}\NormalTok{,}
+      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(time\_correction))}
+\NormalTok{    ),}
+    \FunctionTok{data.frame}\NormalTok{(}
+      \AttributeTok{model =} \StringTok{"Time and distance correction"}\NormalTok{,}
+      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(time\_dist\_correction))}
+\NormalTok{    )}
+\NormalTok{  )}
+
+  \FunctionTok{colnames}\NormalTok{(params) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"model"}\NormalTok{, }\StringTok{"est\_MSS"}\NormalTok{, }\StringTok{"est\_TAU"}\NormalTok{, }\StringTok{"est\_MAC"}\NormalTok{, }\StringTok{"est\_PMAX"}\NormalTok{,}
+    \StringTok{"est\_time\_correction"}\NormalTok{, }\StringTok{"est\_distance\_correction"}
+\NormalTok{  )}
+
+\NormalTok{  df }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}
+\NormalTok{    data,}
+\NormalTok{    params}
+\NormalTok{  )}
+
+\NormalTok{  df}\SpecialCharTok{$}\NormalTok{predicted\_time }\OtherTok{\textless{}{-}} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+    \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance,}
+    \AttributeTok{MSS =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_MSS,}
+    \AttributeTok{TAU =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_TAU,}
+    \AttributeTok{time\_correction =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_time\_correction,}
+    \AttributeTok{distance\_correction =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_distance\_correction}
+\NormalTok{  )}
+
+\NormalTok{  df}\SpecialCharTok{$}\NormalTok{residuals }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{predicted\_time }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{time}
+  \FunctionTok{return}\NormalTok{(df)}
+\NormalTok{\}}
+
+\CommentTok{\# estimated parameters and predicted time}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{do}\NormalTok{(}\FunctionTok{pred\_wrapper}\NormalTok{(.)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> model_df$model <- factor(
-+   model_df$model,
-+   levels = c(
-+     "No correction",
-+     "Fixed correction +0.3s",
-+     "Fixed correction +0.5s",
-+     "Time correction",
-+     "Time and distance correction"
-+   )
-+ )
-R> # Estimates plot
-R> df <- model_df %>%
-+   group_by(MSS, TAU, flying_start_distance, model) %>%
-+   slice(1) %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
-+   )
-R> 
-R> # MSS
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_MSS, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated MSS (m/s)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_df}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  model\_df}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"No correction"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.5s"}\NormalTok{,}
+    \StringTok{"Time correction"}\NormalTok{,}
+    \StringTok{"Time and distance correction"}
+\NormalTok{  )}
+\NormalTok{)}
+\CommentTok{\# Estimates plot}
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance, model) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{slice}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
+    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
+    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{))}
+\NormalTok{  )}
+
+\CommentTok{\# MSS}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MSS, }\AttributeTok{color =}\NormalTok{ model)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MSS\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MSS (m/s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 Similar outcomes are observed for the MAC parameter. The time and distance corrections model performs perfectly, while the time correction model performs almost as good.
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # MAC
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_MAC, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated MAC (m/s/s)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# MAC}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MAC, }\AttributeTok{color =}\NormalTok{ model)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MAC (m/s/s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 PMAX demonstrates the same properties as MSS and MAC.
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # PMAX
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_PMAX, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated PMAX (W/kg)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# PMAX}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_PMAX, }\AttributeTok{color =}\NormalTok{ model)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated PMAX (W/kg)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the \emph{stolen time}; indicated by the dashed line on the figure).
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # time_correction
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_time_correction, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   geom_line(aes(y = stolen_time), color = "black", linetype = "dashed") +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated time correction (s)")  +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# time\_correction}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}
+    \AttributeTok{x =}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{y =}\NormalTok{ est\_time\_correction,}
+    \AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}
+    \FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ stolen\_time), }\AttributeTok{color =} \StringTok{"black"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated time correction (s)"}\NormalTok{)  }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents \emph{identity line} since flying start distance is already on the x-axis).
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # distance_correction
-R> ggplot(
-+   df,
-+   aes(x = flying_start_distance, y = est_distance_correction, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   geom_abline(slope = 1, color = "black", linetype = "dashed") +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Flying start distance (m)") +
-+   ylab("estimated distance correction (m)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# distance\_correction}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}
+    \AttributeTok{x =}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{y =}\NormalTok{ est\_distance\_correction, }
+    \AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_abline}\NormalTok{(}\AttributeTok{slope =} \DecValTok{1}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated distance correction (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Residuals
-R> model_df <- model_df %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-+     group = paste(MSS, MAC, flying_start_distance)
-+   )
-R> 
-R> ggplot(
-+   model_df,
-+   aes(
-+     y = residuals,
-+     x = distance,
-+     color = flying_start_distance,
-+     group = group)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.3) +
-+   facet_wrap(~model) +
-+   geom_hline(yintercept = 0, linetype = "dashed") +
-+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-+   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)") +
-+   theme(legend.position = "top") + 
-+   labs(color = "Flying start distance")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Residuals}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
+    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
+    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{group =} \FunctionTok{paste}\NormalTok{(MSS, MAC, flying\_start\_distance)}
+\NormalTok{  )}
+
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  model\_df,}
+  \FunctionTok{aes}\NormalTok{(}
+    \AttributeTok{y =}\NormalTok{ residuals,}
+    \AttributeTok{x =}\NormalTok{ distance,}
+    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{group =}\NormalTok{ group)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{model) }\SpecialCharTok{+}
+  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> sim_df <- expand.grid(
-+   MSS = c(6, 7, 8, 9),
-+   MAC = c(6, 7, 8, 9),
-+   time_lag = seq(0, 0.5, length.out = 50),
-+   distance = c(5, 10, 20, 30, 40, 50)
-+ )
-R> 
-R> sim_df <- sim_df %>%
-+   mutate(
-+     TAU = MSS / MAC,
-+     PMAX = MSS * MAC / 4,
-+     true_time = predict_time_at_distance(distance, MSS, TAU),
-+     time = true_time + time_lag
-+   )
-R> 
-R> # Add small noise to allow model fit
-R> set.seed(1667)
-R> rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
-R> sim_df$time <- sim_df$time + rand_noise
-\end{CodeInput}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # estimated parameters and predicted time
-R> model_df <- sim_df %>%
-+   group_by(MSS, TAU, time_lag) %>%
-+   do(pred_wrapper(.)) %>%
-+   ungroup()
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{sim\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand.grid}\NormalTok{(}
+  \AttributeTok{MSS =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
+  \AttributeTok{MAC =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
+  \AttributeTok{time\_lag =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{50}\NormalTok{),}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{50}\NormalTok{)}
+\NormalTok{)}
+
+\NormalTok{sim\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{TAU =}\NormalTok{ MSS }\SpecialCharTok{/}\NormalTok{ MAC,}
+    \AttributeTok{PMAX =}\NormalTok{ MSS }\SpecialCharTok{*}\NormalTok{ MAC }\SpecialCharTok{/} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{true\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(distance, MSS, TAU),}
+    \AttributeTok{time =}\NormalTok{ true\_time }\SpecialCharTok{+}\NormalTok{ time\_lag}
+\NormalTok{  )}
+
+\CommentTok{\# Add small noise to allow model fit}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1667}\NormalTok{)}
+\NormalTok{rand\_noise }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(sim\_df), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{4}\NormalTok{)}
+\NormalTok{sim\_df}\SpecialCharTok{$}\NormalTok{time }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df}\SpecialCharTok{$}\NormalTok{time }\SpecialCharTok{+}\NormalTok{ rand\_noise}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# estimated parameters and predicted time}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, time\_lag) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{do}\NormalTok{(}\FunctionTok{pred\_wrapper}\NormalTok{(.)) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{ungroup}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> model_df$model <- factor(
-+   model_df$model,
-+   levels = c(
-+     "No correction",
-+     "Fixed correction +0.3s",
-+     "Fixed correction +0.5s",
-+     "Time correction",
-+     "Time and distance correction"
-+   )
-+ )
-R> # Estimates plot
-R> df <- model_df %>%
-+   group_by(MSS, TAU, time_lag, model) %>%
-+   slice(1) %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2))
-+   )
-R> 
-R> # MSS
-R> ggplot(df, aes(x = time_lag, y = est_MSS, color = model)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-+   xlab("Time lag (s)") +
-+   ylab("estimated MSS (m/s)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_df}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  model\_df}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"No correction"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.5s"}\NormalTok{,}
+    \StringTok{"Time correction"}\NormalTok{,}
+    \StringTok{"Time and distance correction"}
+\NormalTok{  )}
+\NormalTok{)}
+\CommentTok{\# Estimates plot}
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, time\_lag, model) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{slice}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
+    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
+    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{))}
+\NormalTok{  )}
+
+\CommentTok{\# MSS}
+\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_MSS, }\AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MSS\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MSS (m/s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # MAC
-R> ggplot(df, aes(x = time_lag, y = est_MAC, color = model)) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Time lag (s)") +
-+   ylab("estimated MAC (m/s/s)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# MAC}
+\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_MAC, }\AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MAC (m/s/s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The figure below depicts correctly identified time lag (i.e.~using time correction parameter) using time correction and time and distance corrections models.
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # time_correction
-R> ggplot(
-+   df,
-+   aes(x = time_lag, y = est_time_correction, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   geom_abline(slope = -1, color = "black", linetype = "dashed") +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Time lag (s)") +
-+   ylab("estimated time correction (s)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# time\_correction}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_time\_correction, }\AttributeTok{color =}\NormalTok{ model)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_abline}\NormalTok{(}\AttributeTok{slope =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated time correction (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # distance_correction
-R> ggplot(
-+   df,
-+   aes(x = time_lag, y = est_distance_correction, color = model)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.7) +
-+   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-+   xlab("Time lag (s)") +
-+   ylab("estimated distance correction (m)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# distance\_correction}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_distance\_correction, }\AttributeTok{color =}\NormalTok{ model)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated distance correction (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following figure depicts residuals (i.e., predicted time minus observed time).
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Residuals
-R> model_df <- model_df %>%
-+   mutate(
-+     MSS_string = paste("MSS =", MSS),
-+     TAU_string = paste("TAU =", TAU),
-+     MAC_string = paste("MAC = ", round(MAC, 2)),
-+     PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-+     group = paste(MSS, MAC, time_lag)
-+   )
-R> 
-R> ggplot(
-+   model_df,
-+   aes(y = residuals, x = distance, color = time_lag, group = group)
-+ ) +
-+   theme_bw(8) +
-+   geom_line(alpha = 0.3) +
-+   facet_wrap(~model) +
-+   geom_hline(yintercept = 0, linetype = "dashed") +
-+   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-+   xlab("Distance (m)") +
-+   ylab("Predicted time - observed time (s)")  +
-+   theme(legend.position = "top") + 
-+   labs(color = "Time lag")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Residuals}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
+    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
+    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{)),}
+    \AttributeTok{group =} \FunctionTok{paste}\NormalTok{(MSS, MAC, time\_lag)}
+\NormalTok{  )}
+
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  model\_df,}
+  \FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ residuals, }\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{color =}\NormalTok{ time\_lag, }\AttributeTok{group =}\NormalTok{ group)}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{model) }\SpecialCharTok{+}
+  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{)  }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Time lag"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-The number and distances of the timing gates can also affect the precision of the estimated sprint parameters \citep{haugenPowerForceVelocityProfilingSprinting2020, haugenDifferenceStartImpact2012}.
+The number and distances of the timing gates can also affect the precision of the estimated sprint parameters (\protect\hyperlink{ref-haugenPowerForceVelocityProfilingSprinting2020}{Thomas A.
+Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-haugenDifferenceStartImpact2012}{Thomas A.
+Haugen, Tønnessen, and Seiler 2012}).
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> # Create split times from known MSS and TAU
-R> df <- tibble(
-+   distance = c(5, 10, 20, 30, 40),
-+   time = predict_time_at_distance(
-+     distance,
-+     MSS = 9,
-+     TAU = 1.3
-+   )
-+ ) %>%
-+   mutate(
-+     # Add random noise to time
-+     time = time + rnorm(n(), 0, 10^-5),
-+     gate_time = time - 0.1
-+   )
-R> 
-R> # Model without time correction
-R> m_no_correction <- model_using_splits(
-+   distance = df$distance, 
-+   time = df$gate_time
-+ )
-R> 
-R> # Model without time correction, but with zeros added
-R> m_no_correction_zero <- model_using_splits(
-+   distance = c(0, df$distance), 
-+   time = c(0, df$gate_time)
-+ )
-R> 
-R> # Model without adding zeros for the start
-R> m_no_zero <- model_using_splits_with_time_correction(
-+   distance = df$distance, 
-+   time = df$gate_time
-+ )
-R> 
-R> # Model with added zeros for the start (d=0 and t=0)
-R> m_with_zero <- model_using_splits_with_time_correction(
-+   distance = c(0, df$distance), 
-+   time = c(0, df$gate_time)
-+ )
-R> 
-R> # Print results
-R> data.frame(
-+   model = c(
-+     "Without time correction",
-+     "Without time correction with zeros added", 
-+     "With time correction",
-+     "With time correction with zeros added"),
-+   rbind(
-+     coef(m_no_correction),
-+     coef(m_no_correction_zero),
-+     coef(m_no_zero),
-+     coef(m_with_zero))
-+ )
-\end{CodeInput}
-\begin{CodeOutput}
-#>                                      model  MSS  TAU  MAC PMAX
-#> 1                  Without time correction 8.78 1.09 8.06 17.7
-#> 2 Without time correction with zeros added 8.78 1.09 8.06 17.7
-#> 3                     With time correction 9.00 1.30 6.92 15.6
-#> 4    With time correction with zeros added 8.79 1.10 7.97 17.5
-#>   time_correction distance_correction
-#> 1         0.00000                   0
-#> 2         0.00000                   0
-#> 3         0.09994                   0
-#> 4         0.00739                   0
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Create split times from known MSS and TAU}
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{),}
+  \AttributeTok{time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+\NormalTok{    distance,}
+    \AttributeTok{MSS =} \DecValTok{9}\NormalTok{,}
+    \AttributeTok{TAU =} \FloatTok{1.3}
+\NormalTok{  )}
+\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \CommentTok{\# Add random noise to time}
+    \AttributeTok{time =}\NormalTok{ time }\SpecialCharTok{+} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{5}\NormalTok{),}
+    \AttributeTok{gate\_time =}\NormalTok{ time }\SpecialCharTok{{-}} \FloatTok{0.1}
+\NormalTok{  )}
+
+\CommentTok{\# Model without time correction}
+\NormalTok{m\_no\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{gate\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model without time correction, but with zeros added}
+\NormalTok{m\_no\_correction\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance), }
+  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{gate\_time)}
+\NormalTok{)}
+
+\CommentTok{\# Model without adding zeros for the start}
+\NormalTok{m\_no\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{gate\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model with added zeros for the start (d=0 and t=0)}
+\NormalTok{m\_with\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance), }
+  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{gate\_time)}
+\NormalTok{)}
+
+\CommentTok{\# Print results}
+\FunctionTok{data.frame}\NormalTok{(}
+  \AttributeTok{model =} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"Without time correction"}\NormalTok{,}
+    \StringTok{"Without time correction with zeros added"}\NormalTok{, }
+    \StringTok{"With time correction"}\NormalTok{,}
+    \StringTok{"With time correction with zeros added"}\NormalTok{),}
+  \FunctionTok{rbind}\NormalTok{(}
+    \FunctionTok{coef}\NormalTok{(m\_no\_correction),}
+    \FunctionTok{coef}\NormalTok{(m\_no\_correction\_zero),}
+    \FunctionTok{coef}\NormalTok{(m\_no\_zero),}
+    \FunctionTok{coef}\NormalTok{(m\_with\_zero))}
+\NormalTok{)}
+\CommentTok{\#\textgreater{}                                      model  MSS  TAU  MAC PMAX}
+\CommentTok{\#\textgreater{} 1                  Without time correction 8.78 1.09 8.06 17.7}
+\CommentTok{\#\textgreater{} 2 Without time correction with zeros added 8.78 1.09 8.06 17.7}
+\CommentTok{\#\textgreater{} 3                     With time correction 9.00 1.30 6.92 15.6}
+\CommentTok{\#\textgreater{} 4    With time correction with zeros added 8.79 1.10 7.97 17.5}
+\CommentTok{\#\textgreater{}   time\_correction distance\_correction}
+\CommentTok{\#\textgreater{} 1         0.00000                   0}
+\CommentTok{\#\textgreater{} 2         0.00000                   0}
+\CommentTok{\#\textgreater{} 3         0.09994                   0}
+\CommentTok{\#\textgreater{} 4         0.00739                   0}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-To estimate parameter stability, model over-fitting, and performance on the unseen data, \pkg{shorts} model function comes with implemented \emph{leave-one-out cross validation} (LOOCV) \citep{jamesIntroductionStatisticalLearning2017, jovanovicBmbstatsBootstrapMagnitudebased2020, kuhnAppliedPredictiveModeling2018}.
+To estimate parameter stability, model over-fitting, and performance on the unseen data, \textbf{shorts} model function comes with implemented \emph{leave-one-out cross validation} (LOOCV) (\protect\hyperlink{ref-jamesIntroductionStatisticalLearning2017}{James et al.
+2017}; \protect\hyperlink{ref-jovanovicBmbstatsBootstrapMagnitudebased2020}{Jovanović 2020}; \protect\hyperlink{ref-kuhnAppliedPredictiveModeling2018}{Kuhn and Johnson 2018}).
 LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation.
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> jack_LOOCV <- model_using_splits_with_time_correction(
-+   distance = split_times$distance,
-+   time = split_times$jack_time,
-+   LOOCV = TRUE
-+ )
-R> 
-R> jack_LOOCV
-\end{CodeInput}
-\begin{CodeOutput}
-#> Estimated model parameters
-#> --------------------------
-#>                 MSS                 TAU                 MAC 
-#>               8.958               1.216               7.367 
-#>                PMAX     time_correction distance_correction 
-#>              16.499               0.284               0.000 
-#> 
-#> Model fit estimators
-#> --------------------
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>   0.00181   1.00000  -0.00109   0.00189   0.00189   0.00114   0.00104 
-#>      MAPE 
-#>   0.04981 
-#> 
-#> 
-#> Leave-One-Out Cross-Validation
-#> ------------------------------
-#> Parameters:
-#>    MSS  TAU  MAC PMAX time_correction distance_correction
-#> 1 8.98 1.25 7.21 16.2           0.300                   0
-#> 2 8.97 1.22 7.35 16.5           0.284                   0
-#> 3 8.95 1.21 7.40 16.5           0.282                   0
-#> 4 8.96 1.21 7.39 16.5           0.282                   0
-#> 5 8.93 1.20 7.45 16.6           0.278                   0
-#> 
-#> Model fit:
-#>       RSE R_squared    minErr    maxErr maxAbsErr      RMSE       MAE 
-#>        NA   1.00000  -0.00639   0.00510   0.00639   0.00401   0.00349 
-#>      MAPE 
-#>   0.18387
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jack\_LOOCV }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time,}
+  \AttributeTok{LOOCV =} \ConstantTok{TRUE}
+\NormalTok{)}
+
+\NormalTok{jack\_LOOCV}
+\CommentTok{\#\textgreater{} Estimated model parameters}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
+\CommentTok{\#\textgreater{}               8.958               1.216               7.367 }
+\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
+\CommentTok{\#\textgreater{}              16.499               0.284               0.000 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit estimators}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}   0.00181   1.00000  {-}0.00109   0.00189   0.00189   0.00114 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}   0.00104   0.04981 }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Leave{-}One{-}Out Cross{-}Validation}
+\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+\CommentTok{\#\textgreater{} Parameters:}
+\CommentTok{\#\textgreater{}    MSS  TAU  MAC PMAX time\_correction distance\_correction}
+\CommentTok{\#\textgreater{} 1 8.98 1.25 7.21 16.2           0.300                   0}
+\CommentTok{\#\textgreater{} 2 8.97 1.22 7.35 16.5           0.284                   0}
+\CommentTok{\#\textgreater{} 3 8.95 1.21 7.40 16.5           0.282                   0}
+\CommentTok{\#\textgreater{} 4 8.96 1.21 7.39 16.5           0.282                   0}
+\CommentTok{\#\textgreater{} 5 8.93 1.20 7.45 16.6           0.278                   0}
+\CommentTok{\#\textgreater{} }
+\CommentTok{\#\textgreater{} Model fit:}
+\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
+\CommentTok{\#\textgreater{}        NA   1.00000  {-}0.00639   0.00510   0.00639   0.00401 }
+\CommentTok{\#\textgreater{}       MAE      MAPE }
+\CommentTok{\#\textgreater{}   0.00349   0.18387}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- jack_LOOCV$LOOCV$parameters
-R> 
-R> df <- pivot_longer(df, cols = 1:6, names_to = "parameter")
-R> 
-R> df$parameter <- factor(
-+   df$parameter,
-+   levels = c(
-+     "MSS",
-+     "TAU",
-+     "MAC",
-+     "PMAX",
-+     "time_correction",
-+     "distance_correction"
-+   )
-+ )
-R> 
-R> ggplot(df, aes(x = value)) +
-+   theme_bw(8) +
-+   geom_boxplot() +
-+   facet_wrap(~parameter, scales = "free_x") +
-+   xlab(NULL) +
-+   ylab(NULL) +
-+   theme(
-+     axis.ticks.y = element_blank(),
-+     axis.text.y = element_blank()
-+   )
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{LOOCV}\SpecialCharTok{$}\NormalTok{parameters}
+
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{pivot\_longer}\NormalTok{(df, }\AttributeTok{cols =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"parameter"}\NormalTok{)}
+
+\NormalTok{df}\SpecialCharTok{$}\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  df}\SpecialCharTok{$}\NormalTok{parameter,}
+  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"MSS"}\NormalTok{,}
+    \StringTok{"TAU"}\NormalTok{,}
+    \StringTok{"MAC"}\NormalTok{,}
+    \StringTok{"PMAX"}\NormalTok{,}
+    \StringTok{"time\_correction"}\NormalTok{,}
+    \StringTok{"distance\_correction"}
+\NormalTok{  )}
+\NormalTok{)}
+
+\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ value)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_boxplot}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{parameter, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{axis.ticks.y =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{axis.text.y =} \FunctionTok{element\_blank}\NormalTok{()}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 Here is the plot of the training and LOOCV residuals:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- data.frame(
-+   distance = jack_LOOCV$data$distance,
-+   time = jack_LOOCV$data$time,
-+   pred_time = jack_LOOCV$data$pred_time,
-+   LOOCV_time = jack_LOOCV$LOOCV$data$pred_time
-+ )
-R> 
-R> df <- df %>%
-+   pivot_longer(cols = c("pred_time", "LOOCV_time"))
-R> 
-R> df$resid <- df$value - df$time
-R> 
-R> ggplot(df, aes(x = distance, y = resid, color = name)) +
-+   theme_bw(8) +
-+   geom_hline(yintercept = 0, linetype = "dashed") +
-+   geom_point() +
-+   theme(legend.title = element_blank()) +
-+   xlab("Distance (m)") +
-+   ylab("Predicted - observed time (s)") +
-+   theme(
-+     legend.title = element_blank(),
-+     legend.position = "top")
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{time,}
+  \AttributeTok{pred\_time =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{pred\_time,}
+  \AttributeTok{LOOCV\_time =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{pred\_time}
+\NormalTok{)}
+
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ df }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(}\StringTok{"pred\_time"}\NormalTok{, }\StringTok{"LOOCV\_time"}\NormalTok{))}
+
+\NormalTok{df}\SpecialCharTok{$}\NormalTok{resid }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{value }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{time}
+
+\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{y =}\NormalTok{ resid, }\AttributeTok{color =}\NormalTok{ name)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{()) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-58-1} \end{center}
-
-\end{CodeChunk}
 
@@
 Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
-LOOCV can also be implemented with the mixed-effects models in the \pkg{shorts} package.
+LOOCV can also be implemented with the mixed-effects models in the \textbf{shorts} package.
 
@@
 
-Let's utilize demonstrated functionalities of the \pkg{shorts} package using real-world data.
-The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the \pkg{shorts} package (see \texttt{?vescovi}).
+Let's utilize demonstrated functionalities of the \textbf{shorts} package using real-world data.
+The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the \textbf{shorts} package (see \texttt{?vescovi}).
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> bolt_reaction_time <- 0.183
-R> 
-R> bolt_distance <- c(10, 20, 30, 40, 50, 60)
-R> bolt_time <- c(1.963, 2.983, 3.883, 4.763, 5.643, 6.493)
-R> 
-R> # No corrections model
-R> bolt_m1 <- model_using_splits(
-+   distance = bolt_distance,
-+   time = bolt_time
-+ )
-R> 
-R> # Model with reaction time as fixed time correction
-R> bolt_m2 <- model_using_splits(
-+   distance = bolt_distance,
-+   time = bolt_time,
-+   time_correction = -bolt_reaction_time
-+ )
-R> 
-R> # Model with estimated time correction
-R> bolt_m3 <- model_using_splits_with_time_correction(
-+   distance = bolt_distance,
-+   time = bolt_time
-+ )
-R> 
-R> # Model with estimated time correction, but deducted reaction time
-R> bolt_m4 <- model_using_splits_with_time_correction(
-+   distance = bolt_distance,
-+   time = bolt_time - bolt_reaction_time
-+ )
-R> 
-R> # Model with estimated time and distance corrections
-R> bolt_m5 <- model_using_splits_with_corrections(
-+   distance = bolt_distance,
-+   time = bolt_time
-+ )
-R> 
-R> # Model with estimated time and distance corrections and deducted reaction time
-R> bolt_m6 <- model_using_splits_with_corrections(
-+   distance = bolt_distance,
-+   time = bolt_time - bolt_reaction_time
-+ )
-R> 
-R> bolt_model <- rbind(
-+   data.frame(
-+     model = "No correction",
-+     t(coef(bolt_m1))
-+   ),
-+   data.frame(
-+     model = "No correction - RT",
-+     t(coef(bolt_m2))
-+   ),
-+   data.frame(
-+     model = "Time correction",
-+     t(coef(bolt_m3))
-+   ),
-+   data.frame(
-+     model = "Time correction - RT",
-+     t(coef(bolt_m4))
-+   ),
-+   data.frame(
-+     model = "Distance correction",
-+     t(coef(bolt_m5))
-+   ),
-+   data.frame(
-+     model = "Distance correction - RT",
-+     t(coef(bolt_m6))
-+   )
-+ )
-R> 
-R> bolt_model
-\end{CodeInput}
-\begin{CodeOutput}
-#>                      model  MSS   TAU   MAC PMAX time_correction
-#> 1            No correction 12.1 1.564  7.77 23.6         0.00000
-#> 2       No correction - RT 11.7 1.205  9.74 28.6        -0.18300
-#> 3          Time correction 11.7 1.202  9.76 28.6        -0.18483
-#> 4     Time correction - RT 11.7 1.202  9.76 28.6        -0.00183
-#> 5      Distance correction 11.6 0.855 13.56 39.3        -0.81151
-#> 6 Distance correction - RT 11.6 0.855 13.56 39.3        -0.62851
-#>   distance_correction
-#> 1                0.00
-#> 2                0.00
-#> 3                0.00
-#> 4                0.00
-#> 5               -3.98
-#> 6               -3.98
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{bolt\_reaction\_time }\OtherTok{\textless{}{-}} \FloatTok{0.183}
+
+\NormalTok{bolt\_distance }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{50}\NormalTok{, }\DecValTok{60}\NormalTok{)}
+\NormalTok{bolt\_time }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{1.963}\NormalTok{, }\FloatTok{2.983}\NormalTok{, }\FloatTok{3.883}\NormalTok{, }\FloatTok{4.763}\NormalTok{, }\FloatTok{5.643}\NormalTok{, }\FloatTok{6.493}\NormalTok{)}
+
+\CommentTok{\# No corrections model}
+\NormalTok{bolt\_m1 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ bolt\_distance,}
+  \AttributeTok{time =}\NormalTok{ bolt\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model with reaction time as fixed time correction}
+\NormalTok{bolt\_m2 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ bolt\_distance,}
+  \AttributeTok{time =}\NormalTok{ bolt\_time,}
+  \AttributeTok{time\_correction =} \SpecialCharTok{{-}}\NormalTok{bolt\_reaction\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model with estimated time correction}
+\NormalTok{bolt\_m3 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ bolt\_distance,}
+  \AttributeTok{time =}\NormalTok{ bolt\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model with estimated time correction, but deducted reaction time}
+\NormalTok{bolt\_m4 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ bolt\_distance,}
+  \AttributeTok{time =}\NormalTok{ bolt\_time }\SpecialCharTok{{-}}\NormalTok{ bolt\_reaction\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model with estimated time and distance corrections}
+\NormalTok{bolt\_m5 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ bolt\_distance,}
+  \AttributeTok{time =}\NormalTok{ bolt\_time}
+\NormalTok{)}
+
+\CommentTok{\# Model with estimated time and distance corrections and}
+\CommentTok{\#  deducted reaction time}
+\NormalTok{bolt\_m6 }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ bolt\_distance,}
+  \AttributeTok{time =}\NormalTok{ bolt\_time }\SpecialCharTok{{-}}\NormalTok{ bolt\_reaction\_time}
+\NormalTok{)}
+
+\NormalTok{bolt\_model }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"No correction"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(bolt\_m1))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"No correction {-} RT"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(bolt\_m2))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(bolt\_m3))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction {-} RT"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(bolt\_m4))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Distance correction"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(bolt\_m5))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Distance correction {-} RT"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(bolt\_m6))}
+\NormalTok{  )}
+\NormalTok{)}
+
+\NormalTok{bolt\_model}
+\CommentTok{\#\textgreater{}                      model  MSS   TAU   MAC PMAX time\_correction}
+\CommentTok{\#\textgreater{} 1            No correction 12.1 1.564  7.77 23.6         0.00000}
+\CommentTok{\#\textgreater{} 2       No correction {-} RT 11.7 1.205  9.74 28.6        {-}0.18300}
+\CommentTok{\#\textgreater{} 3          Time correction 11.7 1.202  9.76 28.6        {-}0.18483}
+\CommentTok{\#\textgreater{} 4     Time correction {-} RT 11.7 1.202  9.76 28.6        {-}0.00183}
+\CommentTok{\#\textgreater{} 5      Distance correction 11.6 0.855 13.56 39.3        {-}0.81151}
+\CommentTok{\#\textgreater{} 6 Distance correction {-} RT 11.6 0.855 13.56 39.3        {-}0.62851}
+\CommentTok{\#\textgreater{}   distance\_correction}
+\CommentTok{\#\textgreater{} 1                0.00}
+\CommentTok{\#\textgreater{} 2                0.00}
+\CommentTok{\#\textgreater{} 3                0.00}
+\CommentTok{\#\textgreater{} 4                0.00}
+\CommentTok{\#\textgreater{} 5               {-}3.98}
+\CommentTok{\#\textgreater{} 6               {-}3.98}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> bolt_model <- bolt_model %>%
-+   group_by(model) %>%
-+   mutate(
-+     dist_99_MSS = find_velocity_critical_distance(
-+       MSS = MSS, TAU = TAU, 
-+       #time_correction = time_correction, 
-+       #distance_correction = distance_correction,
-+       percent = 0.99
-+     ),
-+    time_99_MSS = find_velocity_critical_time(
-+       MSS = MSS, TAU = TAU, 
-+       #time_correction = time_correction, 
-+       percent = 0.99
-+     )
-+   )
-R> 
-R> bolt_model[c(1, 8, 9)]
-\end{CodeInput}
-\begin{CodeOutput}
-#> # A tibble: 6 x 3
-#> # Groups:   model [6]
-#>   model                    dist_99_MSS time_99_MSS
-#>   <chr>                          <dbl>       <dbl>
-#> 1 No correction                   68.7        7.20
-#> 2 No correction - RT              51.1        5.55
-#> 3 Time correction                 51.0        5.54
-#> 4 Time correction - RT            51.0        5.54
-#> 5 Distance correction             35.8        3.94
-#> 6 Distance correction - RT        35.8        3.94
-\end{CodeOutput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{bolt\_model }\OtherTok{\textless{}{-}}\NormalTok{ bolt\_model }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(model) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{dist\_99\_MSS =} \FunctionTok{find\_velocity\_critical\_distance}\NormalTok{(}
+      \AttributeTok{MSS =}\NormalTok{ MSS, }\AttributeTok{TAU =}\NormalTok{ TAU, }
+      \CommentTok{\#time\_correction = time\_correction, }
+      \CommentTok{\#distance\_correction = distance\_correction,}
+      \AttributeTok{percent =} \FloatTok{0.99}
+\NormalTok{    ),}
+   \AttributeTok{time\_99\_MSS =} \FunctionTok{find\_velocity\_critical\_time}\NormalTok{(}
+      \AttributeTok{MSS =}\NormalTok{ MSS, }\AttributeTok{TAU =}\NormalTok{ TAU, }
+      \CommentTok{\#time\_correction = time\_correction, }
+      \AttributeTok{percent =} \FloatTok{0.99}
+\NormalTok{    )}
+\NormalTok{  )}
+
+\NormalTok{bolt\_model[}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{)]}
+\CommentTok{\#\textgreater{} \# A tibble: 6 x 3}
+\CommentTok{\#\textgreater{} \# Groups:   model [6]}
+\CommentTok{\#\textgreater{}   model                    dist\_99\_MSS time\_99\_MSS}
+\CommentTok{\#\textgreater{}   \textless{}chr\textgreater{}                          \textless{}dbl\textgreater{}       \textless{}dbl\textgreater{}}
+\CommentTok{\#\textgreater{} 1 No correction                   68.7        7.20}
+\CommentTok{\#\textgreater{} 2 No correction {-} RT              51.1        5.55}
+\CommentTok{\#\textgreater{} 3 Time correction                 51.0        5.54}
+\CommentTok{\#\textgreater{} 4 Time correction {-} RT            51.0        5.54}
+\CommentTok{\#\textgreater{} 5 Distance correction             35.8        3.94}
+\CommentTok{\#\textgreater{} 6 Distance correction {-} RT        35.8        3.94}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-The protocol for assessing linear sprint speed has been described previously \citep{vescoviImpactMaximumSpeed2014, vescoviLocomotorHeartRateMetabolic2016, vescoviSprintSpeedCharacteristics2012} and was identical for each cohort.
+The protocol for assessing linear sprint speed has been described previously (\protect\hyperlink{ref-vescoviImpactMaximumSpeed2014}{Vescovi 2014}, \protect\hyperlink{ref-vescoviLocomotorHeartRateMetabolic2016}{2016}, \protect\hyperlink{ref-vescoviSprintSpeedCharacteristics2012}{2012}) and was identical for each cohort.
 Briefly, all athletes performed a standardized warm-up that included general exercises such as jogging, shuffling, multi-directional movements, and dynamic stretching exercises.
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> data("vescovi")
-R> 
-R> # Convert data to long
-R> df <- vescovi %>%
-+   select(1:13) %>%
-+   # slice(1:10) %>%
-+   pivot_longer(
-+     cols = 9:13,
-+     names_to = "distance",
-+     values_to = "time"
-+   ) %>%
-+   mutate(
-+     distance = as.numeric(str_extract(distance, "^[0-9]+"))
-+   )
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{data}\NormalTok{(}\StringTok{"vescovi"}\NormalTok{)}
+
+\CommentTok{\# Convert data to long}
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ vescovi }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{select}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{13}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\# slice(1:10) \%\textgreater{}\%}
+  \FunctionTok{pivot\_longer}\NormalTok{(}
+    \AttributeTok{cols =} \DecValTok{9}\SpecialCharTok{:}\DecValTok{13}\NormalTok{,}
+    \AttributeTok{names\_to =} \StringTok{"distance"}\NormalTok{,}
+    \AttributeTok{values\_to =} \StringTok{"time"}
+\NormalTok{  ) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{distance =} \FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{str\_extract}\NormalTok{(distance, }\StringTok{"\^{}[0{-}9]+"}\NormalTok{))}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> no_corrections <- mixed_model_using_splits(
-+   df,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "Athlete"
-+ )
-R> 
-R> fixed_correction <- mixed_model_using_splits(
-+   df,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "Athlete",
-+   time_correction = 0.3
-+ )
-R> 
-R> time_correction_fixed <- mixed_model_using_splits_with_time_correction(
-+   df,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "Athlete",
-+   random = MSS + TAU ~ 1
-+ )
-R> 
-R> time_correction_random <- mixed_model_using_splits_with_time_correction(
-+   df,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "Athlete",
-+   random = MSS + TAU + time_correction ~ 1
-+ )
-R> 
-R> time_distance_correction_fixed <- mixed_model_using_splits_with_corrections(
-+   df,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "Athlete",
-+   random = MSS + TAU + time_correction ~ 1
-+ )
-R> 
-R> time_distance_correction_random <- mixed_model_using_splits_with_corrections(
-+   df,
-+   distance = "distance",
-+   time = "time",
-+   athlete = "Athlete",
-+   random = MSS + TAU + time_correction + distance_correction ~ 1
-+ )
-\end{CodeInput}
-\end{CodeChunk}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{no\_corrections }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_splits}\NormalTok{(}
+\NormalTok{  df,}
+  \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+  \AttributeTok{athlete =} \StringTok{"Athlete"}
+\NormalTok{)}
+
+\NormalTok{fixed\_correction }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_splits}\NormalTok{(}
+\NormalTok{  df,}
+  \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+  \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
+  \AttributeTok{time\_correction =} \FloatTok{0.3}
+\NormalTok{)}
+
+\NormalTok{time\_correction\_fixed }\OtherTok{\textless{}{-}} 
+  \FunctionTok{mixed\_model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+\NormalTok{    df,}
+    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
+    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{  )}
+
+\NormalTok{time\_correction\_random }\OtherTok{\textless{}{-}} 
+  \FunctionTok{mixed\_model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+\NormalTok{    df,}
+    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
+    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{  )}
+
+\NormalTok{time\_distance\_correction\_fixed }\OtherTok{\textless{}{-}} 
+  \FunctionTok{mixed\_model\_using\_splits\_with\_corrections}\NormalTok{(}
+\NormalTok{    df,}
+    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
+    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{  )}
+
+\NormalTok{time\_distance\_correction\_random }\OtherTok{\textless{}{-}} 
+  \FunctionTok{mixed\_model\_using\_splits\_with\_corrections}\NormalTok{(}
+\NormalTok{    df,}
+    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
+    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
+    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
+    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{+}\NormalTok{ distance\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> model_fit <- rbind(
-+   data.frame(
-+     model = "No corrections",
-+     t(unlist(no_corrections$model_fit))
-+   ),
-+   data.frame(
-+     model = "Fixed correction +0.3s",
-+     t(unlist(fixed_correction$model_fit))
-+   ),
-+   data.frame(
-+     model = "Time correction fixed",
-+     t(unlist(time_correction_fixed$model_fit))
-+   ),
-+   data.frame(
-+     model = "Time correction random",
-+     t(unlist(time_correction_random$model_fit))
-+   ),
-+   data.frame(
-+     model = "Time and distance correction fixed",
-+     t(unlist(time_distance_correction_fixed$model_fit))
-+   ),
-+   data.frame(
-+     model = "Time and distance correction random",
-+     t(unlist(time_distance_correction_random$model_fit))
-+   )
-+ )
-R> 
-R> model_fit$model <- factor(
-+   model_fit$model,
-+   levels = rev(c(
-+     "No corrections",
-+     "Fixed correction +0.3s",
-+     "Time correction fixed",
-+     "Time correction random",
-+     "Time and distance correction fixed",
-+     "Time and distance correction random"
-+   ))
-+ )
-R> 
-R> ggplot(model_fit, aes(x = RSE, y = model)) +
-+   theme_bw(8) +
-+   geom_point() +
-+   xlab("RSE (s)") +
-+   ylab(NULL)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_fit }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"No corrections"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(no\_corrections}\SpecialCharTok{$}\NormalTok{model\_fit))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(fixed\_correction}\SpecialCharTok{$}\NormalTok{model\_fit))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction fixed"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_correction\_fixed}\SpecialCharTok{$}\NormalTok{model\_fit))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction random"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_correction\_random}\SpecialCharTok{$}\NormalTok{model\_fit))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_distance\_correction\_fixed}\SpecialCharTok{$}\NormalTok{model\_fit))}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time and distance correction random"}\NormalTok{,}
+    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_distance\_correction\_random}\SpecialCharTok{$}\NormalTok{model\_fit))}
+\NormalTok{  )}
+\NormalTok{)}
+
+\NormalTok{model\_fit}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  model\_fit}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
+    \StringTok{"No corrections"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \StringTok{"Time correction fixed"}\NormalTok{,}
+    \StringTok{"Time correction random"}\NormalTok{,}
+    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+    \StringTok{"Time and distance correction random"}
+\NormalTok{  ))}
+\NormalTok{)}
+
+\FunctionTok{ggplot}\NormalTok{(model\_fit, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ RSE, }\AttributeTok{y =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"RSE (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following image depicts estimated parameters for each model:
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> est_params <- rbind(
-+   data.frame(
-+     model = "No corrections",
-+     no_corrections$parameters$random
-+   ),
-+   data.frame(
-+     model = "Fixed correction +0.3s",
-+     fixed_correction$parameters$random
-+   ),
-+   data.frame(
-+     model = "Time correction fixed",
-+     time_correction_fixed$parameters$random
-+   ),
-+   data.frame(
-+     model = "Time correction random",
-+     time_correction_random$parameters$random
-+   ),
-+   data.frame(
-+     model = "Time and distance correction fixed",
-+     time_distance_correction_fixed$parameters$random
-+   ),
-+   data.frame(
-+     model = "Time and distance correction random",
-+     time_distance_correction_random$parameters$random
-+   )
-+ )
-R> 
-R> est_params$model <- factor(
-+   est_params$model,
-+   levels = rev(c(
-+     "No corrections",
-+     "Fixed correction +0.3s",
-+     "Time correction fixed",
-+     "Time correction random",
-+     "Time and distance correction fixed",
-+     "Time and distance correction random"
-+   ))
-+ )
-R> 
-R> est_params <- est_params %>%
-+   pivot_longer(cols = -(1:2), names_to = "parameter")
-R> 
-R> est_params$parameter <- factor(
-+   est_params$parameter,
-+   levels = c(
-+     "MSS",
-+     "TAU",
-+     "MAC",
-+     "PMAX",
-+     "time_correction",
-+     "distance_correction"
-+   )
-+ )
-R> 
-R> ggplot(est_params, aes(y = model, x = value)) +
-+   theme_bw(8) +
-+   geom_boxplot() +
-+   facet_wrap(~parameter, scales = "free_x") +
-+   xlab(NULL) +
-+   ylab(NULL)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{est\_params }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"No corrections"}\NormalTok{,}
+\NormalTok{    no\_corrections}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+\NormalTok{    fixed\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction fixed"}\NormalTok{,}
+\NormalTok{    time\_correction\_fixed}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction random"}\NormalTok{,}
+\NormalTok{    time\_correction\_random}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+\NormalTok{    time\_distance\_correction\_fixed}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time and distance correction random"}\NormalTok{,}
+\NormalTok{    time\_distance\_correction\_random}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
+\NormalTok{  )}
+\NormalTok{)}
+
+\NormalTok{est\_params}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  est\_params}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
+    \StringTok{"No corrections"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \StringTok{"Time correction fixed"}\NormalTok{,}
+    \StringTok{"Time correction random"}\NormalTok{,}
+    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+    \StringTok{"Time and distance correction random"}
+\NormalTok{  ))}
+\NormalTok{)}
+
+\NormalTok{est\_params }\OtherTok{\textless{}{-}}\NormalTok{ est\_params }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{), }\AttributeTok{names\_to =} \StringTok{"parameter"}\NormalTok{)}
+
+\NormalTok{est\_params}\SpecialCharTok{$}\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  est\_params}\SpecialCharTok{$}\NormalTok{parameter,}
+  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
+    \StringTok{"MSS"}\NormalTok{,}
+    \StringTok{"TAU"}\NormalTok{,}
+    \StringTok{"MAC"}\NormalTok{,}
+    \StringTok{"PMAX"}\NormalTok{,}
+    \StringTok{"time\_correction"}\NormalTok{,}
+    \StringTok{"distance\_correction"}
+\NormalTok{  )}
+\NormalTok{)}
+
+\FunctionTok{ggplot}\NormalTok{(est\_params, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ model, }\AttributeTok{x =}\NormalTok{ value)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_boxplot}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{parameter, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-64-1} \end{center}
-
-\end{CodeChunk}
 
@@
 To provide practical magnitude of the residuals, we have used between subject observed time SD multiplied with 0.2 and -0.2.
-This provides practical anchor for the residual magnitude, often referred to as \emph{smallest worthwhile change} (SWC) or \emph{smallest effect size of interest} (SESOI) \citep{jovanovicBmbstatsBootstrapMagnitudebased2020}.
+This provides practical anchor for the residual magnitude, often referred to as \emph{smallest worthwhile change} (SWC) or \emph{smallest effect size of interest} (SESOI) (\protect\hyperlink{ref-jovanovicBmbstatsBootstrapMagnitudebased2020}{Jovanović 2020}).
 If the residuals are within this magnitude band, then the model is good in making practically useful predictions.
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> model_resid <- rbind(
-+   data.frame(
-+     model = "No corrections",
-+     no_corrections$data
-+   ),
-+   data.frame(
-+     model = "Fixed correction +0.3s",
-+     fixed_correction$data
-+   ),
-+   data.frame(
-+     model = "Time correction fixed",
-+     time_correction_fixed$data
-+   ),
-+   data.frame(
-+     model = "Time correction random",
-+     time_correction_random$data
-+   ),
-+   data.frame(
-+     model = "Time and distance correction fixed",
-+     time_distance_correction_fixed$data
-+   ),
-+   data.frame(
-+     model = "Time and distance correction random",
-+     time_distance_correction_random$data
-+   )
-+ )
-R> 
-R> 
-R> model_resid$model <- factor(
-+   model_resid$model,
-+   levels = rev(c(
-+     "No corrections",
-+     "Fixed correction +0.3s",
-+     "Time correction fixed",
-+     "Time correction random",
-+     "Time and distance correction fixed",
-+     "Time and distance correction random"
-+   ))
-+ )
-R> 
-R> model_resid$resid <- model_resid$pred_time - model_resid$time
-R> 
-R> # Create SWC / SESOI band
-R> model_SESOI <- model_resid %>%
-+   group_by(model, distance) %>%
-+   summarise(
-+     bias = mean(resid),
-+     variance = sd(resid),
-+     upper = bias + variance,
-+     lower = bias - variance,
-+     MAD = mean(abs(resid)),
-+     SESOI_upper = sd(time) * 0.2,
-+     SESOI_lower = -sd(time) * 0.2
-+   )
-R> 
-R> # Plot
-R> ggplot(model_resid, aes(y = model)) +
-+   theme_bw(8) +
-+   geom_vline(
-+     data = model_SESOI,
-+     aes(xintercept = SESOI_lower),
-+     color = "blue", alpha = 0.5, linetype = "dashed"
-+   ) +
-+   geom_vline(
-+     data = model_SESOI,
-+     aes(xintercept = SESOI_upper),
-+     color = "blue", alpha = 0.5, linetype = "dashed"
-+   ) +
-+   geom_vline(xintercept = 0, color = "blue", alpha = 0.5) +
-+   geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
-+   geom_errorbarh(
-+     data = model_SESOI,
-+     aes(xmin = lower, xmax = upper),
-+     height = 0.1, color = "black"
-+   ) +
-+   geom_point(data = model_SESOI, aes(x = bias), color = "black") +
-+   facet_wrap(~distance, scales = "free_x") +
-+   xlab("Predicted time - observed time (s)") +
-+   ylab(NULL)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{model\_resid }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"No corrections"}\NormalTok{,}
+\NormalTok{    no\_corrections}\SpecialCharTok{$}\NormalTok{data}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+\NormalTok{    fixed\_correction}\SpecialCharTok{$}\NormalTok{data}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction fixed"}\NormalTok{,}
+\NormalTok{    time\_correction\_fixed}\SpecialCharTok{$}\NormalTok{data}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time correction random"}\NormalTok{,}
+\NormalTok{    time\_correction\_random}\SpecialCharTok{$}\NormalTok{data}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+\NormalTok{    time\_distance\_correction\_fixed}\SpecialCharTok{$}\NormalTok{data}
+\NormalTok{  ),}
+  \FunctionTok{data.frame}\NormalTok{(}
+    \AttributeTok{model =} \StringTok{"Time and distance correction random"}\NormalTok{,}
+\NormalTok{    time\_distance\_correction\_random}\SpecialCharTok{$}\NormalTok{data}
+\NormalTok{  )}
+\NormalTok{)}
+
+\NormalTok{model\_resid}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  model\_resid}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
+    \StringTok{"No corrections"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \StringTok{"Time correction fixed"}\NormalTok{,}
+    \StringTok{"Time correction random"}\NormalTok{,}
+    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+    \StringTok{"Time and distance correction random"}
+\NormalTok{  ))}
+\NormalTok{)}
+
+\NormalTok{model\_resid}\SpecialCharTok{$}\NormalTok{resid }\OtherTok{\textless{}{-}}\NormalTok{ model\_resid}\SpecialCharTok{$}\NormalTok{pred\_time }\SpecialCharTok{{-}}\NormalTok{ model\_resid}\SpecialCharTok{$}\NormalTok{time}
+
+\CommentTok{\# Create SWC / SESOI band}
+\NormalTok{model\_SESOI }\OtherTok{\textless{}{-}}\NormalTok{ model\_resid }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{group\_by}\NormalTok{(model, distance) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{summarise}\NormalTok{(}
+    \AttributeTok{bias =} \FunctionTok{mean}\NormalTok{(resid),}
+    \AttributeTok{variance =} \FunctionTok{sd}\NormalTok{(resid),}
+    \AttributeTok{upper =}\NormalTok{ bias }\SpecialCharTok{+}\NormalTok{ variance,}
+    \AttributeTok{lower =}\NormalTok{ bias }\SpecialCharTok{{-}}\NormalTok{ variance,}
+    \AttributeTok{MAD =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{abs}\NormalTok{(resid)),}
+    \AttributeTok{SESOI\_upper =} \FunctionTok{sd}\NormalTok{(time) }\SpecialCharTok{*} \FloatTok{0.2}\NormalTok{,}
+    \AttributeTok{SESOI\_lower =} \SpecialCharTok{{-}}\FunctionTok{sd}\NormalTok{(time) }\SpecialCharTok{*} \FloatTok{0.2}
+\NormalTok{  )}
+
+\CommentTok{\# Plot}
+\FunctionTok{ggplot}\NormalTok{(model\_resid, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_vline}\NormalTok{(}
+    \AttributeTok{data =}\NormalTok{ model\_SESOI,}
+    \FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ SESOI\_lower),}
+    \AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}
+\NormalTok{  ) }\SpecialCharTok{+}
+  \FunctionTok{geom\_vline}\NormalTok{(}
+    \AttributeTok{data =}\NormalTok{ model\_SESOI,}
+    \FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ SESOI\_upper),}
+    \AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}
+\NormalTok{  ) }\SpecialCharTok{+}
+  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_jitter}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ resid), }\AttributeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{height =} \FloatTok{0.25}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_errorbarh}\NormalTok{(}
+    \AttributeTok{data =}\NormalTok{ model\_SESOI,}
+    \FunctionTok{aes}\NormalTok{(}\AttributeTok{xmin =}\NormalTok{ lower, }\AttributeTok{xmax =}\NormalTok{ upper),}
+    \AttributeTok{height =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}
+\NormalTok{  ) }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{data =}\NormalTok{ model\_SESOI, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ bias), }\AttributeTok{color =} \StringTok{"black"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{distance, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\end{CodeChunk}
-
 The following figure depicts model residuals estimators (bias, or mean residual; variance, or SD of the residuals, and MAD, or mean absolute difference).
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> df <- model_SESOI %>%
-+   pivot_longer(cols = -(1:2), names_to = "estimator") %>%
-+   filter(estimator %in% c("bias", "variance", "MAD"))
-R> 
-R> df$model <- factor(
-+   df$model,
-+   levels = rev(c(
-+     "No corrections",
-+     "Fixed correction +0.3s",
-+     "Time correction fixed",
-+     "Time correction random",
-+     "Time and distance correction fixed",
-+     "Time and distance correction random"
-+   ))
-+ )
-R> 
-R> df$estimator <- factor(
-+   df$estimator,
-+   levels = c("bias", "variance", "MAD")
-+ )
-R> 
-R> ggplot(df, aes(x = value, y = model)) +
-+   theme_bw(8) +
-+   geom_point() +
-+   facet_grid(distance ~ estimator, scales = "free_x") +
-+   xlab(NULL) +
-+   ylab(NULL)
-\end{CodeInput}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_SESOI }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{), }\AttributeTok{names\_to =} \StringTok{"estimator"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{filter}\NormalTok{(estimator }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"bias"}\NormalTok{, }\StringTok{"variance"}\NormalTok{, }\StringTok{"MAD"}\NormalTok{))}
+
+\NormalTok{df}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  df}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
+    \StringTok{"No corrections"}\NormalTok{,}
+    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
+    \StringTok{"Time correction fixed"}\NormalTok{,}
+    \StringTok{"Time correction random"}\NormalTok{,}
+    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
+    \StringTok{"Time and distance correction random"}
+\NormalTok{  ))}
+\NormalTok{)}
+
+\NormalTok{df}\SpecialCharTok{$}\NormalTok{estimator }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
+\NormalTok{  df}\SpecialCharTok{$}\NormalTok{estimator,}
+  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"bias"}\NormalTok{, }\StringTok{"variance"}\NormalTok{, }\StringTok{"MAD"}\NormalTok{)}
+\NormalTok{)}
+
+\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ value, }\AttributeTok{y =}\NormalTok{ model)) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(distance }\SpecialCharTok{\textasciitilde{}}\NormalTok{ estimator, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
 \begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-66-1} \end{center}
-
-\end{CodeChunk}
 
@@
 
-We are hoping that the \pkg{shorts} package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
-
-\renewcommand\refname{References}
-\bibliography{references.bib,packages.bib}
+We are hoping that the \textbf{shorts} package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-arsacModelingEnergetics100m2002}{}%
+Arsac, Laurent M., and Elio Locatelli.
+2002.
+{``Modeling the Energetics of 100-m Running by Using Speed Curves of World Champions.''} \emph{Journal of Applied Physiology} 92 (5): 1781--88.
+\url{https://doi.org/10.1152/japplphysiol.00754.2001}.
+
+\leavevmode\hypertarget{ref-brownAssessmentLinearSprinting2004}{}%
+Brown, Todd D., Jason D. Vescovi, and Jaci L. Vanheest. 2004. {``Assessment of Linear Sprinting Performance: A Theoretical Paradigm.''} \emph{Journal of Sports Science \& Medicine} 3 (4): 203--10.
+
+\leavevmode\hypertarget{ref-buchheitMechanicalDeterminantsAcceleration2014}{}%
+Buchheit, Martin, Pierre Samozino, Jonathan Alexander Glynn, Ben Simpson Michael, Hani Al Haddad, Alberto Mendez-Villanueva, and Jean Benoit Morin.
+2014.
+{``Mechanical Determinants of Acceleration and Maximal Sprinting Speed in Highly Trained Young Soccer Players.''} \emph{Journal of Sports Sciences} 32 (20): 1906--13.
+\url{https://doi.org/10.1080/02640414.2014.965191}.
+
+\leavevmode\hypertarget{ref-clarkNFLCombine40Yard2017}{}%
+Clark, Kenneth P., Randall H.
+Rieger, Richard F.
+Bruno, and David J.
+Stearne.
+2017.
+{``The {NFL Combine} 40-{Yard Dash}: {How Important} Is {Maximum Velocity}?''} \emph{Journal of Strength and Conditioning Research}, June, 1.
+\url{https://doi.org/10.1519/JSC.0000000000002081}.
+
+\leavevmode\hypertarget{ref-edwardsSprintAccelerationCharacteristics2020}{}%
+Edwards, Toby, Benjamin Piggott, Harry G.
+Banyard, G.
+Gregory Haff, and Christopher Joyce.
+2020.
+{``Sprint Acceleration Characteristics Across the {Australian} Football Participation Pathway.''} \emph{Sports Biomechanics}, August, 1--13.
+\url{https://doi.org/10.1080/14763141.2020.1790641}.
+
+\leavevmode\hypertarget{ref-doi:10.1098ux2frspb.1927.0035}{}%
+Furusawa, K., Archibald Vivian Hill, and J.
+L.
+Parkinson.
+1927.
+{``The Dynamics of "Sprint" Running.''} \emph{Proceedings of the Royal Society of London.
+Series B, Containing Papers of a Biological Character} 102 (713): 29--42.
+\url{https://doi.org/10.1098/rspb.1927.0035}.
+
+\leavevmode\hypertarget{ref-R-LambertW}{}%
+Goerg, Georg M. 2020. \emph{LambertW: Probabilistic Models to Analyze and Gaussianize Heavy-Tailed, Skewed Data}. \url{https://CRAN.R-project.org/package=LambertW}.
+
+\leavevmode\hypertarget{ref-haugenPowerForceVelocityProfilingSprinting2020}{}%
+Haugen, Thomas A., Felix Breitschädel, and Pierre Samozino.
+2020.
+{``Power-{Force}-{Velocity Profiling} of {Sprinting Athletes}: {Methodological} and {Practical Considerations When Using Timing Gates}.''} \emph{Journal of Strength and Conditioning Research} 34 (6): 1769--73.
+\url{https://doi.org/10.1519/JSC.0000000000002890}.
+
+\leavevmode\hypertarget{ref-haugenSprintMechanicalVariables2019}{}%
+Haugen, Thomas A., Felix Breitschädel, and Stephen Seiler.
+2019.
+{``Sprint Mechanical Variables in Elite Athletes: {Are} Force-Velocity Profiles Sport Specific or Individual?''} Edited by Leonardo A.
+Peyré-Tartaruga.
+\emph{PLOS ONE} 14 (7): e0215551.
+\url{https://doi.org/10.1371/journal.pone.0215551}.
+
+\leavevmode\hypertarget{ref-haugenSprintMechanicalProperties2020}{}%
+---------.
+2020.
+{``Sprint Mechanical Properties in Soccer Players According to Playing Standard, Position, Age and Sex.''} \emph{Journal of Sports Sciences} 38 (9): 1070--76.
+\url{https://doi.org/10.1080/02640414.2020.1741955}.
+
+\leavevmode\hypertarget{ref-haugenDifferenceStartImpact2012}{}%
+Haugen, Thomas A, Espen Tønnessen, and Stephen K Seiler.
+2012.
+{``The {Difference Is} in the {Start}: {Impact} of {Timing} and {Start Procedure} on {Sprint Running Performance}:''} \emph{Journal of Strength and Conditioning Research} 26 (2): 473--79.
+\url{https://doi.org/10.1519/JSC.0b013e318226030b}.
+
+\leavevmode\hypertarget{ref-jamesIntroductionStatisticalLearning2017}{}%
+James, Gareth, Daniela Witten, Trevor Hastie, and Robert Tibshirani.
+2017.
+\emph{An {Introduction} to {Statistical Learning}: With {Applications} in {R}}.
+1st ed.
+2013, Corr.
+7th printing 2017 edition.
+{New York}: {Springer}.
+
+\leavevmode\hypertarget{ref-jimenez-reyesRelationshipVerticalHorizontal2018}{}%
+Jiménez-Reyes, Pedro, Pierre Samozino, Amador García-Ramos, Víctor Cuadrado-Peñafiel, Matt Brughelli, and Jean-Benoît Morin.
+2018.
+{``Relationship Between Vertical and Horizontal Force-Velocity-Power Profiles in Various Sports and Levels of Practice.''} \emph{PeerJ} 6 (November): e5937.
+\url{https://doi.org/10.7717/peerj.5937}.
+
+\leavevmode\hypertarget{ref-R-shorts}{}%
+Jovanovic, Mladen. 2020. \emph{Shorts: Short Sprints}. \url{https://CRAN.R-project.org/package=shorts}.
+
+\leavevmode\hypertarget{ref-jovanovicBmbstatsBootstrapMagnitudebased2020}{}%
+Jovanović, Mladen. 2020. \emph{Bmbstats: {Bootstrap Magnitude}-Based {Statistics} for {Sports Scientists}}. {Mladen Jovanović}.
+
+\leavevmode\hypertarget{ref-kuhnAppliedPredictiveModeling2018}{}%
+Kuhn, Max, and Kjell Johnson. 2018. \emph{Applied {Predictive Modeling}}. 1st ed. 2013, Corr. 2nd printing 2016 edition. {New York}: {Springer}.
+
+\leavevmode\hypertarget{ref-mangineSpeedForcePower2014}{}%
+Mangine, Gerald T., Jay R.
+Hoffman, Adam M.
+Gonzalez, Adam J.
+Wells, Jeremy R.
+Townsend, Adam R.
+Jajtner, William P.
+McCormack, et al.
+2014.
+{``Speed, {Force}, and {Power Values Produced From Nonmotorized Treadmill Test Are Related} to {Sprinting Performance}:''} \emph{Journal of Strength and Conditioning Research} 28 (7): 1812--19.
+\url{https://doi.org/10.1519/JSC.0000000000000316}.
+
+\leavevmode\hypertarget{ref-marcote-pequenoAssociationForceVelocity2019}{}%
+Marcote-Pequeño, Ramón, Amador García-Ramos, Víctor Cuadrado-Peñafiel, Jorge M.
+González-Hernández, Miguel Ángel Gómez, and Pedro Jiménez-Reyes.
+2019.
+{``Association {Between} the {Force}{{Velocity Profile}} and {Performance Variables Obtained} in {Jumping} and {Sprinting} in {Elite Female Soccer Players}.''} \emph{International Journal of Sports Physiology and Performance} 14 (2): 209--15.
+\url{https://doi.org/10.1123/ijspp.2018-0233}.
+
+\leavevmode\hypertarget{ref-morinSimpleMethodComputing2019}{}%
+Morin, Jean-Benoit, Pierre Samozino, Munenori Murata, Matt R Cross, and Ryu Nagahara.
+2019.
+{``A Simple Method for Computing Sprint Acceleration Kinetics from Running Velocity Data: {Replication} Study with Improved Design.''} \emph{Journal of Biomechanics} 94 (September): 82--87.
+\url{https://doi.org/10.1016/j.jbiomech.2019.07.020}.
+
+\leavevmode\hypertarget{ref-morinInterpretingPowerForceVelocityProfiles2016}{}%
+Morin, Jean-Benoît, and Pierre Samozino.
+2016.
+{``Interpreting {Power}-{Force}-{Velocity Profiles} for {Individualized} and {Specific Training}.''} \emph{International Journal of Sports Physiology and Performance} 11 (2): 267--72.
+\url{https://doi.org/10.1123/ijspp.2015-0638}.
+
+\leavevmode\hypertarget{ref-R-nlme}{}%
+Pinheiro, José, Douglas Bates, and R-core. 2020. \emph{Nlme: Linear and Nonlinear Mixed Effects Models}. \url{https://svn.r-project.org/R-packages/trunk/nlme/}.
+
+\leavevmode\hypertarget{ref-R-base}{}%
+R Core Team. 2020. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\hypertarget{ref-samozinoSimpleMethodMeasuring2016}{}%
+Samozino, P., G.
+Rabita, S.
+Dorel, J.
+Slawinski, N.
+Peyrot, E.
+Saez de Villarreal, and J.-B.
+Morin.
+2016.
+{``A Simple Method for Measuring Power, Force, Velocity Properties, and Mechanical Effectiveness in Sprint Running: {Simple} Method to Compute Sprint Mechanics.''} \emph{Scandinavian Journal of Medicine \& Science in Sports} 26 (6): 648--58.
+\url{https://doi.org/10.1111/sms.12490}.
+
+\leavevmode\hypertarget{ref-stenrothForcevelocityProfilingIce2020}{}%
+Stenroth, Lauri, Paavo Vartiainen, and Pasi A.
+Karjalainen.
+2020.
+{``Force-Velocity Profiling in Ice Hockey Skating: Reliability and Validity of a Simple, Low-Cost Field Method.''} \emph{Sports Biomechanics}, June, 1--16.
+\url{https://doi.org/10.1080/14763141.2020.1770321}.
+
+\leavevmode\hypertarget{ref-vaningenschenauCanCyclePower1991}{}%
+van Ingen Schenau, Gerrit Jan, Ron Jacobs, and Jos J.
+de Koning.
+1991.
+{``Can Cycle Power Predict Sprint Running Performance?''} \emph{European Journal of Applied Physiology and Occupational Physiology} 63 (3-4): 255--60.
+\url{https://doi.org/10.1007/BF00233857}.
+
+\leavevmode\hypertarget{ref-vescoviSprintSpeedCharacteristics2012}{}%
+Vescovi, Jason D.
+2012.
+{``Sprint Speed Characteristics of High-Level {American} Female Soccer Players: {Female Athletes} in {Motion} ({FAiM}) {Study}.''} \emph{Journal of Science and Medicine in Sport} 15 (5): 474--78.
+\url{https://doi.org/10.1016/j.jsams.2012.03.006}.
+
+\leavevmode\hypertarget{ref-vescoviImpactMaximumSpeed2014}{}%
+---------.
+2014.
+{``Impact of {Maximum Speed} on {Sprint Performance During High}-{Level Youth Female Field Hockey Matches}: {Female Athletes} in {Motion} ({FAiM}) {Study}.''} \emph{International Journal of Sports Physiology and Performance} 9 (4): 621--26.
+\url{https://doi.org/10.1123/ijspp.2013-0263}.
+
+\leavevmode\hypertarget{ref-vescoviLocomotorHeartRateMetabolic2016}{}%
+---------.
+2016.
+{``Locomotor, {Heart}-{Rate}, and {Metabolic Power Characteristics} of {Youth Women}'s {Field Hockey}: {Female Athletes} in {Motion} ({FAiM}) {Study}.''} \emph{Research Quarterly for Exercise and Sport} 87 (1): 68--77.
+\url{https://doi.org/10.1080/02701367.2015.1124972}.
+
+\leavevmode\hypertarget{ref-ward-smithEnergyConversionStrategies2001}{}%
+Ward-Smith, A. J. 2001. {``Energy Conversion Strategies During 100 m Sprinting.''} \emph{Journal of Sports Sciences} 19 (9): 701--10. \url{https://doi.org/10.1080/02640410152475838}.
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, Hadley. 2019. \emph{Tidyverse: Easily Install and Load the Tidyverse}. \url{https://CRAN.R-project.org/package=tidyverse}.
+
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+---------. 2020. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
+
+\leavevmode\hypertarget{ref-R-ggplot2}{}%
+Wickham, Hadley, Winston Chang, Lionel Henry, Thomas Lin Pedersen, Kohske Takahashi, Claus Wilke, Kara Woo, Hiroaki Yutani, and Dewey Dunnington.
+2020.
+\emph{Ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics}.
+\url{https://CRAN.R-project.org/package=ggplot2}.
+
+\leavevmode\hypertarget{ref-R-dplyr}{}%
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
+
+\end{CSLReferences}
 

### 2021-04-04 code.R

@@ -1,1681 +1,0 @@
-## ----setup, include=FALSE-------------------------------------------
-require(knitr)
-require(shorts)
-require(tidyverse)
-require(bookdown)
-
-my_random_seed <- 1667
-set.seed(my_random_seed)
-
-knitr::opts_chunk$set(
-  comment = "#>",
-  collapse = TRUE,
-  cache = FALSE,
-  fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
-  dpi = 300,
-  out.width = "100%",
-  fig.align = "center",
-  fig.width = 6,
-  fig.height = 6 * 0.618, # 1 / phi
-  fig.show = "hold",
-  echo = TRUE,
-  warning = FALSE,
-  message = FALSE,
-  width = 70
-)
-
-# automatically create a bib database for R packages
-#knitr::write_bib(c(
-#  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
-#), "packages.bib")
-
-
-# Set rounding
-op <- options()
-
-options(
-  digits = 3,
-  "width" = 70,
-  scipen = 999)
-
-
-## ----four-athletes-table, echo=FALSE--------------------------------
-athletes <- tribble(
-  ~Athlete, ~MSS, ~MAC,
-  "Athlete A", 12, 10,
-  "Athlete B", 12, 6,
-  "Athlete C", 8, 10,
-  "Athlete D", 8, 6
-)
-
-athletes <- athletes %>%
-  mutate(TAU = MSS / MAC)
-
-knitr::kable(
-  athletes,
-  caption = "Four athletes with different MSS and MAC parameters.",
-  digits = 2,
-  booktabs = TRUE
-)
-
-
-## ----four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds."----
-kinematics <- expand_grid(
-  athletes,
-  time = seq(0, 6, length.out = 1000)
-) %>%
-  mutate(
-    distance = predict_distance_at_time(time, MSS, TAU),
-    velocity = predict_velocity_at_time(time, MSS, TAU),
-    acceleration = predict_acceleration_at_time(time, MSS, TAU)
-  )
-
-kinematics_long <- kinematics %>%
-  pivot_longer(
-    cols = c("distance", "velocity", "acceleration"),
-    names_to = "variable"
-  ) %>%
-  mutate(
-    variable = factor(
[…]

### 2021-04-14 paper.Rmd
Reduced version for the PeerJ

@@
   dpi = 300,
-  out.width = "100%",
+  out.width = "90%",
   fig.align = "center",
-  fig.width = 6,
-  fig.height = 6 * 0.618, # 1 / phi
+  fig.width = 5,
+  fig.height = 5 * 0.618, # 1 / phi
   fig.show = "hold",
@@
   message = FALSE,
-  width = 70
+  width = 65
 )
@@
 # automatically create a bib database for R packages
-#knitr::write_bib(c(
+# knitr::write_bib(c(
 #  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
-#), "packages.bib")
+# ), "packages.bib")
 
@@
   "width" = 65,
-  scipen = 999)
+  scipen = 999
+)
 ```
@@
 
-Short sprint performance is one of the most distinguishable and admired physical trait in sports.
+Short sprint performance is one of the most distinguishable and admired physical traits in sports.
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
@@
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
+
 The **shorts** package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+To best of our knowledge, scientist, researchers, and coaches have been performing short sprints modeling using the built-in solver function of Excel (Microsoft Corporation, Redmond, Washington, United States) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothForcevelocityProfilingIce2020; @stenrothSpreadsheetSprintAcceleration2020; @samozinoSimpleMethodMeasuring2016; @clarkNFLCombine40Yard2017; @morinSimpleMethodComputing2019], which makes the **shorts** package major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
 
@@
     legend.position = "top",
-    legend.title = element_blank())
+    legend.title = element_blank()
+  )
 
@@
 If the athlete's body mass (kg) is known, as well as additional air resistance parameters (see [Air resistance and the calculation of force and mechanical power] section of this paper), *Force-Velocity Profile* can be estimated (see [Force-Velocity profile] section of this paper).
-
 ```{r four-athletes-profile, echo=FALSE, fig.cap="Acceleration-Velocity profile of four athletes with different MSS and MAC parameters."}
@@
 Besides providing *residual standard error* (RSE), **shorts** functions provide additional model fit estimators.
-Additional information can be gained by exploring the returned object, particularly object returned from the `nls()` function:
-
-```{r}
-summary(m1)
-```
-
-To extract estimated model parameters, use S3 `coef()` method:
-
-```{r}
-coef(m1)
-```
+Additional information can be gained by exploring the returned object, particularly object returned from the `nls()` function (i.e., by using the S3 `summary()` method).
+To extract estimated model parameters, use S3 `coef()` method.
 
@@
   distance = split_distance,
-  MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU
-)
-
-# Predict acceleration at time
-predict_acceleration_at_time(
-  time = c(0, 1, 2, 3, 4, 5, 6),
   MSS = m1$parameters$MSS,
@@
 )
-
-# To calculate power produced
-predict_power_at_distance(
-  distance = split_distance,
-  MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU,
-  # Additional parameters forwarded to get_air_resistance
-  # Otherwise, defaults are used
-  bodymass = 80,
-  bodyheight = 1.85,
-  barometric_pressure = 780,
-  air_temperature = 20,
-  wind_velocity = 0.5
-)
 ```
@@
 )
-
-head(df)
 ```
@@
 
-Additional information about mixed-effects model performed using the `nlme` package [@R-nlme] can be obtained using `summary()` function:
-
-```{r}
+Additional information about mixed-effects model performed using the `nlme` package [@R-nlme] can be obtained using `summary()` function.
+S3 method `coef()` when applied on mixed-model result will return both the fixed and random effects.
+To create a simple plot of the model, use S3 `plot()` method.
+
+```{r include=FALSE}
 summary(m2)
-```
-
-S3 method `coef()` when applied on mixed-model result will return both the fixed and random effects:
-
-```{r}
 coef(m2)
 ```
-
-To create a simple plot of the model, use S3 `plot()` method:
 
@@
     legend.position = "top",
-    legend.title = element_blank())
+    legend.title = element_blank()
+  )
 ```
@@
 )
-
-m3
-```
-
-```{r}
-plot(m3) + theme_bw(8)
 ```
@@
 )
-
-m3_weighted
 ```
@@
 )
-
-m4
-```
-
-```{r}
-plot(m4) + theme_bw(8)
 ```
@@
 
-```{r}
+```{r results='hide', fig.show='hide'}
 plot(fvp, "time") + theme_bw(8)
@@
     legend.title = element_blank(),
-    legend.position = "top")
-```
-
-The following three models estimate MSS and TAU from the three datasets:
-
-```{r}
+    legend.position = "top"
+  )
+```
+
+The following three models estimate MSS and TAU from the three datasets.
+Tables are created using the `kable()` function from the **knitr** package [@knitr2014; @knitr2015; @R-knitr] and `kable_styling()` function from the **kableExtra** package [@R-kableExtra].
+
+```{r}
+require(knitr)
+require(kableExtra)
+
 # Without synchronization issues
@@
 
-rbind(
+model_df <- rbind(
   data.frame(
@@
 )
+
+kable(model_df, digits = 2, booktabs = TRUE)
 ```
@@
 
-m10
-
 # Mixed model with time correction being random effect
@@
 )
-
-m11
 ```
@@
 In other words, Jack is using a *flying start*, a common scenario when testing field sports athletes.
-Let's see how their sprint outcomes differ.
 
@@
 )
-
-split_times
-```
-
-And here is a graphical representation of the sprint splits:
-
-```{r}
+```
+
+Here is a graphical representation of the sprint splits (please refer to the Supplemental files for the R code):
+
+```{r echo=FALSE}
 plot_df <- split_times %>%
@@
     legend.title = element_blank(),
-    legend.position = "top")
+    legend.position = "top"
+  )
 ```
@@
 
-sprint_parameters
+kable(sprint_parameters, digits = 2, booktabs = TRUE)
 ```
@@
 
-Below is a simulation sprint with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 ($ms^{-1}$ and $ms^{-2}$ respectively), and flying start distance varying from 0 to 1 m.
-
-```{r}
+To explore this further, we have run a simple simulation with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 ($ms^{-1}$ and $ms^{-2}$ respectively), and flying start distance varying from 0 to 1 m (please refer to the Supplemental files for the R code).
+With a simulation dataset, we can check the model estimates and predictions, given the flying start distance.
+
+```{r include=FALSE}
 sim_df <- expand.grid(
@@
     stolen_time = predict_time_at_distance(
-      flying_start_distance, MSS, TAU),
+      flying_start_distance, MSS, TAU
+    ),
     time = true_time - stolen_time
@@
 
-Now when we have a simulation dataset, we can check the model estimates and predictions, given the flying start distance:
-
-```{r}
+```{r include=FALSE}
 # Prediction wrapper
@@
 
-```{r}
+```{r echo=FALSE}
 # Estimates plot
@@
 
-```{r}
+```{r echo=FALSE}
 # MAC
@@
 
-```{r}
+```{r echo=FALSE}
 # PMAX
@@
 
-```{r}
+```{r echo=FALSE}
 # Residuals
@@
     color = flying_start_distance,
-    group = group)
+    group = group
+  )
 ) +
@@
 
-```{r}
+```{r echo=FALSE}
 ggplot(
@@
 
-jack_parameters
+kable(jack_parameters, digits = 2, booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"))
 ```
@@
 
-We will get back to these issues later, but we can examine how these models perform using simulated data with varying flying start distance.
-The following code contains the wrapper that performs all four models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction):
-
-```{r}
+We will get back to these issues later, but we can examine how these models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the Supplemental files for the R code).
+
+```{r echo=FALSE}
 pred_wrapper <- function(data) {
@@
 
-```{r}
+```{r echo=FALSE}
 model_df$model <- factor(
@@
 
-```{r}
+```{r echo=FALSE}
 # time_correction
@@
     y = est_time_correction,
-    color = model)) +
+    color = model
+  )
+) +
   theme_bw(8) +
@@
   geom_line(
-    aes(y = stolen_time), color = "black", linetype = "dashed") +
+    aes(y = stolen_time),
+    color = "black", linetype = "dashed"
+  ) +
   facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
@@
     legend.title = element_blank(),
-    legend.position = "top")
-```
-
-The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents *identity line* since flying start distance is already on the x-axis).
-
-```{r}
+    legend.position = "top"
+  )
+```
+
+The next figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents *identity line* since flying start distance is already on the x-axis).
+
+```{r echo=FALSE}
 # distance_correction
@@
     y = est_distance_correction,
-    color = model)) +
+    color = model
+  )
+) +
   theme_bw(8) +
@@
 Both of these scenarios represent *time lag* that is added to the split times.
-Below we simulate the effect of this time lag on model estimates and predictions.
-
-```{r}
+We have simulated the effect of this time lag on model estimates and predictions (please refer to the Supplemental files for the R code).
+
+```{r include=FALSE}
 sim_df <- expand.grid(
@@
 sim_df$time <- sim_df$time + rand_noise
-```
-
-```{r}
+
 # estimated parameters and predicted time
@@
     legend.title = element_blank(),
-    legend.position = "top")
-```
-
-From the figure below it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
+    legend.position = "top"
+  )
+```
+
+From the next figure it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
 Time correction and time and distance corrections model correctly estimated MAC.
 
-```{r}
+```{r echo=FALSE}
 # MAC
@@
 
-```{r}
+```{r echo=FALSE}
 # distance_correction
@@
 # Print results
-data.frame(
+res_df <- data.frame(
   model = c(
@@
     "With time correction",
-    "With time correction with zeros added"),
+    "With time correction with zeros added"
+  ),
   rbind(
@@
     coef(m_no_zero),
-    coef(m_with_zero))
-)
+    coef(m_with_zero)
+  )
+)
+
+kable(res_df, digits = 2, booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"))
 ```
@@
 )
-
-jack_LOOCV
 ```
@@
 
-Next we plot estimated parameters across LOOCV folds:
-
-```{r}
+Next we plot estimated parameters across LOOCV folds (please refer to the Supplemental files for the R code):
+
+```{r echo=FALSE}
 df <- jack_LOOCV$LOOCV$parameters
@@
 
-```{r}
+```{r echo=FALSE}
 df <- data.frame(
@@
 
-bolt_model
+kable(bolt_model, digits = 2, booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"))
 ```
@@
       MSS = MSS, TAU = TAU,
-      #time_correction = time_correction, 
-      #distance_correction = distance_correction,
+      # time_correction = time_correction,
+      # distance_correction = distance_correction,
       percent = 0.99
@@
       MSS = MSS, TAU = TAU,
-      #time_correction = time_correction, 
+      # time_correction = time_correction,
       percent = 0.99
@@
 
-bolt_model[c(1, 8, 9)]
+kable(bolt_model[c(1, 8, 9)], digits = 2, booktabs = TRUE)
 ```
@@
 
-The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model.
-
-```{r}
+The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model (please refer to the Supplemental files for the R code).
+
+```{r echo=FALSE}
 no_corrections <- mixed_model_using_splits(
@@
 
-```{r}
+```{r echo=FALSE}
 model_fit <- rbind(
@@
 
-```{r}
+```{r echo=FALSE}
 est_params <- rbind(
@@
 If the residuals are within this magnitude band, then the model is good in making practically useful predictions.
-
 Error bars represent residual bias $\pm$ 1 SD.
 
-```{r}
+```{r echo=FALSE}
 model_resid <- rbind(
@@
 
-```{r}
+```{r echo=FALSE}
 df <- model_SESOI %>%
@@
 We are hoping that the **shorts** package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
-
 ```{r include=FALSE}

### 2021-04-14 paper.tex
Reduced version for the PeerJ

@@
 \usepackage{flafter}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
+
 The \textbf{shorts} package (\protect\hyperlink{ref-R-shorts}{Jovanovic 2020}), written in the R language (\protect\hyperlink{ref-R-base}{R Core Team 2020}), represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
@@
 2016}).
+To best of our knowledge, scientist, researchers, and coaches have been performing short sprints modeling using the built-in solver function of Excel (Microsoft Corporation, Redmond, Washington, United States) (\protect\hyperlink{ref-morinSpreadsheetSprintAcceleration2017}{J.
+B.
+Morin 2017}; \protect\hyperlink{ref-morinSpreadsheetSprintAcceleration2019}{Jean-Benoit Morin and Samozino 2019}; \protect\hyperlink{ref-stenrothForcevelocityProfilingIce2020}{Stenroth, Vartiainen, and Karjalainen 2020}; \protect\hyperlink{ref-stenrothSpreadsheetSprintAcceleration2020}{Stenroth and Vartiainen 2020}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
+2016}; \protect\hyperlink{ref-clarkNFLCombine40Yard2017}{Clark et al.
+2017}; \protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+2019}), which makes the \textbf{shorts} package major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/four-athletes-kinematics-1}
+{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/four-athletes-kinematics-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{paper_files/figure-latex/four-athletes-profile-1}
+{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/four-athletes-profile-1}
 
@@
 Besides providing \emph{residual standard error} (RSE), \textbf{shorts} functions provide additional model fit estimators.
-Additional information can be gained by exploring the returned object, particularly object returned from the \texttt{nls()} function:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{summary}\NormalTok{(m1)}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Formula: corrected\_time \textasciitilde{} TAU * I(LambertW::W({-}exp(1)\^{}({-}distance/(MSS * }
-\CommentTok{\#\textgreater{}     TAU) {-} 1))) + distance/MSS + TAU}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Parameters:}
-\CommentTok{\#\textgreater{}     Estimate Std. Error t value Pr(\textgreater{}|t|)    }
-\CommentTok{\#\textgreater{} MSS   9.0121     0.0155     581 0.000003 ***}
-\CommentTok{\#\textgreater{} TAU   1.3083     0.0066     198 0.000025 ***}
-\CommentTok{\#\textgreater{} {-}{-}{-}}
-\CommentTok{\#\textgreater{} Signif.
-codes:  0 \textquotesingle{}***\textquotesingle{} 0.001 \textquotesingle{}**\textquotesingle{} 0.01 \textquotesingle{}*\textquotesingle{} 0.05 \textquotesingle{}.\textquotesingle{} 0.1 \textquotesingle{} \textquotesingle{} 1}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Residual standard error: 0.00249 on 2 degrees of freedom}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Number of iterations to convergence: 4 }
-\CommentTok{\#\textgreater{} Achieved convergence tolerance: 0.00000311}
-\end{Highlighting}
-\end{Shaded}
-
-To extract estimated model parameters, use S3 \texttt{coef()} method:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{coef}\NormalTok{(m1)}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}                9.01                1.31                6.89 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}               15.52                0.00                0.00}
-\end{Highlighting}
-\end{Shaded}
+Additional information can be gained by exploring the returned object, particularly object returned from the \texttt{nls()} function (i.e., by using the S3 \texttt{summary()} method).
+To extract estimated model parameters, use S3 \texttt{coef()} method.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-4-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-2-1} \end{center}
 
@@
 \CommentTok{\#\textgreater{} [1] 2.17 3.43 4.60 5.73}
-
-\CommentTok{\# Predict acceleration at time}
-\FunctionTok{predict\_acceleration\_at\_time}\NormalTok{(}
-  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{, }\DecValTok{3}\NormalTok{, }\DecValTok{4}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{6}\NormalTok{),}
-  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
-  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU}
-\NormalTok{)}
-\CommentTok{\#\textgreater{} [1] 6.8884 3.2075 1.4935 0.6954 0.3238 0.1508 0.0702}
 \end{Highlighting}
@@
 \CommentTok{\#\textgreater{} [1] 119.0  58.6  36.9  28.2}
-
-\CommentTok{\# To calculate power produced}
-\FunctionTok{predict\_power\_at\_distance}\NormalTok{(}
-  \AttributeTok{distance =}\NormalTok{ split\_distance,}
-  \AttributeTok{MSS =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
-  \AttributeTok{TAU =}\NormalTok{ m1}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
-  \CommentTok{\# Additional parameters forwarded to get\_air\_resistance}
-  \CommentTok{\# Otherwise, defaults are used}
-  \AttributeTok{bodymass =} \DecValTok{80}\NormalTok{,}
-  \AttributeTok{bodyheight =} \FloatTok{1.85}\NormalTok{,}
-  \AttributeTok{barometric\_pressure =} \DecValTok{780}\NormalTok{,}
-  \AttributeTok{air\_temperature =} \DecValTok{20}\NormalTok{,}
-  \AttributeTok{wind\_velocity =} \FloatTok{0.5}
-\NormalTok{)}
-\CommentTok{\#\textgreater{} [1] 868 490 323 251}
 \end{Highlighting}
@@
 \NormalTok{)}
-
-\FunctionTok{head}\NormalTok{(df)}
-\CommentTok{\#\textgreater{}   time distance velocity acceleration bodymass}
-\CommentTok{\#\textgreater{} 1 0.00 0.000000   0.0000         6.89       80}
-\CommentTok{\#\textgreater{} 2 0.01 0.000344   0.0686         6.84       80}
-\CommentTok{\#\textgreater{} 3 0.02 0.001371   0.1367         6.78       80}
-\CommentTok{\#\textgreater{} 4 0.03 0.003076   0.2043         6.73       80}
-\CommentTok{\#\textgreater{} 5 0.04 0.005455   0.2714         6.68       80}
-\CommentTok{\#\textgreater{} 6 0.05 0.008502   0.3379         6.63       80}
-\CommentTok{\#\textgreater{}   net\_horizontal\_force air\_resistance horizontal\_force}
-\CommentTok{\#\textgreater{} 1                  551        0.07536              551}
-\CommentTok{\#\textgreater{} 2                  547        0.05609              547}
-\CommentTok{\#\textgreater{} 3                  543        0.03978              543}
-\CommentTok{\#\textgreater{} 4                  539        0.02636              539}
-\CommentTok{\#\textgreater{} 5                  534        0.01576              534}
-\CommentTok{\#\textgreater{} 6                  530        0.00792              530}
-\CommentTok{\#\textgreater{}   horizontal\_force\_relative vertical\_force resultant\_force}
-\CommentTok{\#\textgreater{} 1                      6.89            785             959}
-\CommentTok{\#\textgreater{} 2                      6.84            785             957}
-\CommentTok{\#\textgreater{} 3                      6.78            785             954}
-\CommentTok{\#\textgreater{} 4                      6.73            785             952}
-\CommentTok{\#\textgreater{} 5                      6.68            785             950}
-\CommentTok{\#\textgreater{} 6                      6.63            785             947}
-\CommentTok{\#\textgreater{}   resultant\_force\_relative power relative\_power    RF}
-\CommentTok{\#\textgreater{} 1                     12.0   0.0          0.000 0.575}
-\CommentTok{\#\textgreater{} 2                     12.0  37.5          0.469 0.572}
-\CommentTok{\#\textgreater{} 3                     11.9  74.2          0.928 0.569}
-\CommentTok{\#\textgreater{} 4                     11.9 110.0          1.375 0.566}
-\CommentTok{\#\textgreater{} 5                     11.9 145.0          1.813 0.563}
-\CommentTok{\#\textgreater{} 6                     11.8 179.2          2.241 0.560}
-\CommentTok{\#\textgreater{}   force\_angle}
-\CommentTok{\#\textgreater{} 1        54.9}
-\CommentTok{\#\textgreater{} 2        55.1}
-\CommentTok{\#\textgreater{} 3        55.3}
-\CommentTok{\#\textgreater{} 4        55.5}
-\CommentTok{\#\textgreater{} 5        55.7}
-\CommentTok{\#\textgreater{} 6        55.9}
 \end{Highlighting}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-9-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-7-1} \end{center}
 
@@
 
-Additional information about mixed-effects model performed using the \texttt{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2020}) can be obtained using \texttt{summary()} function:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{summary}\NormalTok{(m2)}
-\CommentTok{\#\textgreater{} Nonlinear mixed{-}effects model fit by maximum likelihood}
-\CommentTok{\#\textgreater{}   Model: corrected\_time \textasciitilde{} TAU * I(LambertW::W({-}exp(1)\^{}({-}distance/(MSS *      TAU) {-} 1))) + distance/MSS + TAU }
-\CommentTok{\#\textgreater{}   Data: train }
-\CommentTok{\#\textgreater{}     AIC   BIC logLik}
-\CommentTok{\#\textgreater{}   {-}75.1 {-}66.7   43.5}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Random effects:}
-\CommentTok{\#\textgreater{}  Formula: list(MSS \textasciitilde{} 1, TAU \textasciitilde{} 1)}
-\CommentTok{\#\textgreater{}  Level: athlete}
-\CommentTok{\#\textgreater{}  Structure: General positive{-}definite, Log{-}Cholesky parametrization}
-\CommentTok{\#\textgreater{}          StdDev Corr }
-\CommentTok{\#\textgreater{} MSS      1.066  MSS  }
-\CommentTok{\#\textgreater{} TAU      0.178  0.877}
-\CommentTok{\#\textgreater{} Residual 0.026       }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Fixed effects:  MSS + TAU \textasciitilde{} 1 }
-\CommentTok{\#\textgreater{}     Value Std.Error DF t{-}value p{-}value}
-\CommentTok{\#\textgreater{} MSS  8.06     0.495 24   16.30       0}
-\CommentTok{\#\textgreater{} TAU  0.66     0.084 24    7.82       0}
-\CommentTok{\#\textgreater{}  Correlation: }
-\CommentTok{\#\textgreater{}     MSS  }
-\CommentTok{\#\textgreater{} TAU 0.874}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Standardized Within{-}Group Residuals:}
-\CommentTok{\#\textgreater{}    Min     Q1    Med     Q3    Max }
-\CommentTok{\#\textgreater{} {-}1.909 {-}0.605  0.154  0.523  1.129 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Number of Observations: 30}
-\CommentTok{\#\textgreater{} Number of Groups: 5}
-\end{Highlighting}
-\end{Shaded}
-
-S3 method \texttt{coef()} when applied on mixed-model result will return both the fixed and random effects:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{coef}\NormalTok{(m2)}
-\CommentTok{\#\textgreater{} $fixed}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}               8.065               0.655              12.309 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}              24.818               0.000               0.000 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} $random}
-\CommentTok{\#\textgreater{}     athlete  MSS   TAU  MAC PMAX time\_correction}
-\CommentTok{\#\textgreater{} 1     James 9.69 0.847 11.4 27.7               0}
-\CommentTok{\#\textgreater{} 2       Jim 7.83 0.505 15.5 30.4               0}
-\CommentTok{\#\textgreater{} 3      John 7.78 0.727 10.7 20.8               0}
-\CommentTok{\#\textgreater{} 4 Kimberley 8.57 0.802 10.7 22.9               0}
-\CommentTok{\#\textgreater{} 5  Samantha 6.45 0.395 16.3 26.4               0}
-\CommentTok{\#\textgreater{}   distance\_correction}
-\CommentTok{\#\textgreater{} 1                   0}
-\CommentTok{\#\textgreater{} 2                   0}
-\CommentTok{\#\textgreater{} 3                   0}
-\CommentTok{\#\textgreater{} 4                   0}
-\CommentTok{\#\textgreater{} 5                   0}
-\end{Highlighting}
-\end{Shaded}
-
-To create a simple plot of the model, use S3 \texttt{plot()} method:
+Additional information about mixed-effects model performed using the \texttt{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2020}) can be obtained using \texttt{summary()} function.
+S3 method \texttt{coef()} when applied on mixed-model result will return both the fixed and random effects.
+To create a simple plot of the model, use S3 \texttt{plot()} method.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-14-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-11-1} \end{center}
 
@@
     \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{,}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{())}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{()}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-12-1} \end{center}
 
@@
 \NormalTok{)}
-
-\NormalTok{m3}
-\CommentTok{\#\textgreater{} Estimated model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}                9.00                1.30                6.92 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}               15.58                0.00                0.00 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit estimators}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}   0.00327   1.00000  {-}0.00406   0.00532   0.00532   0.00276 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}   0.00207       NaN}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{plot}\NormalTok{(m3) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \NormalTok{)}
-
-\NormalTok{m3\_weighted}
-\CommentTok{\#\textgreater{} Estimated model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}                9.00                1.30                6.92 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}               15.58                0.00                0.00 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit estimators}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}   0.00108   1.00000  {-}0.00406   0.00534   0.00534   0.00276 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}   0.00206       NaN}
 \end{Highlighting}
@@
 \NormalTok{)}
-
-\NormalTok{m4}
-\CommentTok{\#\textgreater{} Estimated fixed model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}                8.30                1.01                8.24 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}               17.09                0.00                0.00 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Estimated random model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}     athlete   MSS   TAU  MAC PMAX time\_correction}
-\CommentTok{\#\textgreater{} 1     James 10.00 1.111 9.00 22.5               0}
-\CommentTok{\#\textgreater{} 2       Jim  8.00 0.889 9.00 18.0               0}
-\CommentTok{\#\textgreater{} 3      John  8.00 1.069 7.48 15.0               0}
-\CommentTok{\#\textgreater{} 4 Kimberley  9.01 1.286 7.01 15.8               0}
-\CommentTok{\#\textgreater{} 5  Samantha  6.50 0.685 9.50 15.4               0}
-\CommentTok{\#\textgreater{}   distance\_correction}
-\CommentTok{\#\textgreater{} 1                   0}
-\CommentTok{\#\textgreater{} 2                   0}
-\CommentTok{\#\textgreater{} 3                   0}
-\CommentTok{\#\textgreater{} 4                   0}
-\CommentTok{\#\textgreater{} 5                   0}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit estimators}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}    0.0516    0.9994   {-}0.2191    0.1983    0.2191    0.0516 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}    0.0395       NaN}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{plot}\NormalTok{(m4) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-20-1} \end{center}
+\end{Highlighting}
+\end{Shaded}
 
@@
 Details of the FVP method implemented in the \textbf{shorts} package, as well as the interpretation from a sprint training perspective, are covered elsewhere (\protect\hyperlink{ref-haugenPowerForceVelocityProfilingSprinting2020}{Thomas A.
-Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-morinInterpretingPowerForceVelocityProfiles2016}{Jean-Benoît Morin and Samozino 2016}; \protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-morinInterpretingPowerForceVelocityProfiles2016}{Jean-Benoit Morin and Samozino 2016}; \protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
 2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-22-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
 
@@
 \end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-23-1} \end{center}
 
@@
   \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-24-1} \end{center}
-
-The following three models estimate MSS and TAU from the three datasets:
-
-\begin{Shaded}
-\begin{Highlighting}[]
+    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
+    \AttributeTok{legend.position =} \StringTok{"top"}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-19-1} \end{center}
+
+The following three models estimate MSS and TAU from the three datasets.
+Tables are created using the \texttt{kable()} function from the \textbf{knitr} package (\protect\hyperlink{ref-knitr2014}{Xie 2014}, \protect\hyperlink{ref-knitr2015}{2015}, \protect\hyperlink{ref-R-knitr}{2020}) and \texttt{kable\_styling()} function from the \textbf{kableExtra} package (\protect\hyperlink{ref-R-kableExtra}{Zhu 2021}).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{require}\NormalTok{(knitr)}
+\FunctionTok{require}\NormalTok{(kableExtra)}
+
 \CommentTok{\# Without synchronization issues}
@@
 
-\FunctionTok{rbind}\NormalTok{(}
+\NormalTok{model\_df }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
   \FunctionTok{data.frame}\NormalTok{(}
@@
 \NormalTok{)}
-\CommentTok{\#\textgreater{}                model   MSS  TAU  MAC PMAX time\_correction}
-\CommentTok{\#\textgreater{} 1          True time  9.00 1.30 6.92 15.6               0}
-\CommentTok{\#\textgreater{} 2    Added 0.5s time  9.91 2.34 4.23 10.5               0}
-\CommentTok{\#\textgreater{} 3 Deducted 0.5s time 10.08 1.86 5.43 13.7               0}
-\CommentTok{\#\textgreater{}   distance\_correction}
-\CommentTok{\#\textgreater{} 1                   0}
-\CommentTok{\#\textgreater{} 2                   0}
-\CommentTok{\#\textgreater{} 3                   0}
-\end{Highlighting}
-\end{Shaded}
+
+\FunctionTok{kable}\NormalTok{(model\_df, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{tabular}{lrrrrrr}
+\toprule
+model & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+True time & 9.00 & 1.30 & 6.92 & 15.6 & 0 & 0\\
+Added 0.5s time & 9.91 & 2.34 & 4.23 & 10.5 & 0 & 0\\
+Deducted 0.5s time & 10.08 & 1.86 & 5.43 & 13.7 & 0 & 0\\
+\bottomrule
+\end{tabular}
 
@@
 
-\NormalTok{m10}
-\CommentTok{\#\textgreater{} Estimated fixed model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}                8.30                1.01                8.24 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}               17.10               {-}0.50                0.00 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Estimated random model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}     athlete   MSS   TAU  MAC PMAX time\_correction}
-\CommentTok{\#\textgreater{} 1     James 10.00 1.111 9.00 22.5            {-}0.5}
-\CommentTok{\#\textgreater{} 2       Jim  8.00 0.889 9.00 18.0            {-}0.5}
-\CommentTok{\#\textgreater{} 3      John  8.00 1.069 7.48 15.0            {-}0.5}
-\CommentTok{\#\textgreater{} 4 Kimberley  9.01 1.285 7.01 15.8            {-}0.5}
-\CommentTok{\#\textgreater{} 5  Samantha  6.50 0.685 9.50 15.4            {-}0.5}
-\CommentTok{\#\textgreater{}   distance\_correction}
-\CommentTok{\#\textgreater{} 1                   0}
-\CommentTok{\#\textgreater{} 2                   0}
-\CommentTok{\#\textgreater{} 3                   0}
-\CommentTok{\#\textgreater{} 4                   0}
-\CommentTok{\#\textgreater{} 5                   0}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit estimators}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}    0.0516    0.9994   {-}0.2190    0.1983    0.2190    0.0516 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}    0.0395       Inf}
-
 \CommentTok{\# Mixed model with time correction being random effect}
@@
 \NormalTok{)}
-
-\NormalTok{m11}
-\CommentTok{\#\textgreater{} Estimated fixed model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}                8.30                1.01                8.24 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}               17.10               {-}0.50                0.00 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Estimated random model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}     athlete   MSS   TAU  MAC PMAX time\_correction}
-\CommentTok{\#\textgreater{} 1     James 10.00 1.110 9.00 22.5            {-}0.5}
-\CommentTok{\#\textgreater{} 2       Jim  8.00 0.889 9.00 18.0            {-}0.5}
-\CommentTok{\#\textgreater{} 3      John  8.00 1.069 7.48 15.0            {-}0.5}
-\CommentTok{\#\textgreater{} 4 Kimberley  9.01 1.285 7.01 15.8            {-}0.5}
-\CommentTok{\#\textgreater{} 5  Samantha  6.50 0.685 9.50 15.4            {-}0.5}
-\CommentTok{\#\textgreater{}   distance\_correction}
-\CommentTok{\#\textgreater{} 1                   0}
-\CommentTok{\#\textgreater{} 2                   0}
-\CommentTok{\#\textgreater{} 3                   0}
-\CommentTok{\#\textgreater{} 4                   0}
-\CommentTok{\#\textgreater{} 5                   0}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit estimators}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}    0.0516    0.9994   {-}0.2188    0.1982    0.2188    0.0516 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}    0.0395       Inf}
 \end{Highlighting}
@@
 In other words, Jack is using a \emph{flying start}, a common scenario when testing field sports athletes.
-Let's see how their sprint outcomes differ.
 
@@
 \NormalTok{)}
-
-\NormalTok{split\_times}
-\CommentTok{\#\textgreater{} \# A tibble: 5 x 6}
-\CommentTok{\#\textgreater{}   distance john\_time jack\_distance jack\_true\_time time\_05m}
-\CommentTok{\#\textgreater{}      \textless{}dbl\textgreater{}  \textless{}I\textless{}dbl\textgreater{}\textgreater{}         \textless{}dbl\textgreater{}       \textless{}I\textless{}dbl\textgreater{}\textgreater{} \textless{}I\textless{}dbl\textgreater{}\textgreater{}}
-\CommentTok{\#\textgreater{} 1        5      1.42           5.5           1.50    0.400}
-\CommentTok{\#\textgreater{} 2       10      2.17          10.5           2.23    0.400}
-\CommentTok{\#\textgreater{} 3       20      3.43          20.5           3.49    0.400}
-\CommentTok{\#\textgreater{} 4       30      4.60          30.5           4.65    0.400}
-\CommentTok{\#\textgreater{} 5       40      5.73          40.5           5.78    0.400}
-\CommentTok{\#\textgreater{} \# ... with 1 more variable: jack\_time \textless{}I\textless{}dbl\textgreater{}\textgreater{}}
-\end{Highlighting}
-\end{Shaded}
-
-And here is a graphical representation of the sprint splits:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot\_df }\OtherTok{\textless{}{-}}\NormalTok{ split\_times }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{select}\NormalTok{(distance, john\_time, jack\_time) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{rename}\NormalTok{(}\AttributeTok{John =}\NormalTok{ john\_time, }\AttributeTok{Jack =}\NormalTok{ jack\_time) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"athlete"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"time"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{distance =} \FunctionTok{factor}\NormalTok{(distance))}
-
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  plot\_df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{y =}\NormalTok{ time, }\AttributeTok{color =}\NormalTok{ athlete, }\AttributeTok{group =}\NormalTok{ athlete)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Time (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-30-1} \end{center}
+\end{Highlighting}
+\end{Shaded}
+
+Here is a graphical representation of the sprint splits (please refer to the Supplemental files for the R code):
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-25-1} \end{center}
 
@@
 
-\NormalTok{sprint\_parameters}
-\CommentTok{\#\textgreater{}       MSS   TAU   MAC PMAX time\_correction distance\_correction}
-\CommentTok{\#\textgreater{} John 9.00 1.300  6.92 15.6               0                   0}
-\CommentTok{\#\textgreater{} Jack 8.49 0.704 12.06 25.6               0                   0}
-\end{Highlighting}
-\end{Shaded}
+\FunctionTok{kable}\NormalTok{(sprint\_parameters, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+John & 9.00 & 1.3 & 6.92 & 15.6 & 0 & 0\\
+Jack & 8.49 & 0.7 & 12.06 & 25.6 & 0 & 0\\
+\bottomrule
+\end{tabular}
 
@@
 
-Below is a simulation sprint with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 (\(ms^{-1}\) and \(ms^{-2}\) respectively), and flying start distance varying from 0 to 1 m.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sim\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand.grid}\NormalTok{(}
-  \AttributeTok{MSS =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
-  \AttributeTok{MAC =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
-  \AttributeTok{flying\_start\_distance =} \FunctionTok{c}\NormalTok{(}
-    \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.001}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{),}
-    \FunctionTok{seq}\NormalTok{(}\FloatTok{0.001}\NormalTok{, }\FloatTok{0.01}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{),}
-    \FunctionTok{seq}\NormalTok{(}\FloatTok{0.01}\NormalTok{, }\FloatTok{0.1}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{),}
-    \FunctionTok{seq}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\DecValTok{1}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{20}\NormalTok{)}
-\NormalTok{  ),}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{50}\NormalTok{)}
-\NormalTok{)}
-
-\NormalTok{sim\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{TAU =}\NormalTok{ MSS }\SpecialCharTok{/}\NormalTok{ MAC,}
-    \AttributeTok{PMAX =}\NormalTok{ MSS }\SpecialCharTok{*}\NormalTok{ MAC }\SpecialCharTok{/} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{true\_distance =}\NormalTok{ distance }\SpecialCharTok{+}\NormalTok{ flying\_start\_distance,}
-    \AttributeTok{true\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(true\_distance, MSS, TAU),}
-    \AttributeTok{stolen\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
-\NormalTok{      flying\_start\_distance, MSS, TAU),}
-    \AttributeTok{time =}\NormalTok{ true\_time }\SpecialCharTok{{-}}\NormalTok{ stolen\_time}
-\NormalTok{  )}
-
-\CommentTok{\# Add small noise to allow model fit}
-\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1667}\NormalTok{)}
-\NormalTok{rand\_noise }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(sim\_df), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{5}\NormalTok{)}
-\NormalTok{sim\_df}\SpecialCharTok{$}\NormalTok{time }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df}\SpecialCharTok{$}\NormalTok{time }\SpecialCharTok{+}\NormalTok{ rand\_noise}
-\end{Highlighting}
-\end{Shaded}
-
-Now when we have a simulation dataset, we can check the model estimates and predictions, given the flying start distance:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Prediction wrapper}
-\NormalTok{pred\_wrapper }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data) \{}
-\NormalTok{  model }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time}
-\NormalTok{  )}
-
-\NormalTok{  params }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}\FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(model)))}
-
-\NormalTok{  predicted\_time }\OtherTok{\textless{}{-}} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{MSS =}\NormalTok{ model}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
-    \AttributeTok{TAU =}\NormalTok{ model}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU}
-\NormalTok{  )}
-
-  \FunctionTok{colnames}\NormalTok{(params) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"est\_MSS"}\NormalTok{, }\StringTok{"est\_TAU"}\NormalTok{, }\StringTok{"est\_MAC"}\NormalTok{, }\StringTok{"est\_PMAX"}\NormalTok{,}
-    \StringTok{"est\_time\_correction"}\NormalTok{, }\StringTok{"est\_distance\_correction"}
-\NormalTok{  )}
-
-  \FunctionTok{cbind}\NormalTok{(}
-\NormalTok{    data,}
-\NormalTok{    params,}
-    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{predicted\_time =} \FunctionTok{as.numeric}\NormalTok{(predicted\_time))}
-\NormalTok{  )}
-\NormalTok{\}}
-
-\CommentTok{\# estimated parameters and predicted time}
-\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{do}\NormalTok{(}\FunctionTok{pred\_wrapper}\NormalTok{(.)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{()}
-
-\CommentTok{\# Prediction residuals}
-\NormalTok{model\_df}\SpecialCharTok{$}\NormalTok{residuals }\OtherTok{\textless{}{-}}\NormalTok{ model\_df}\SpecialCharTok{$}\NormalTok{predicted\_time }\SpecialCharTok{{-}}\NormalTok{ model\_df}\SpecialCharTok{$}\NormalTok{time}
-\end{Highlighting}
-\end{Shaded}
+To explore this further, we have run a simple simulation with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 (\(ms^{-1}\) and \(ms^{-2}\) respectively), and flying start distance varying from 0 to 1 m (please refer to the Supplemental files for the R code).
+With a simulation dataset, we can check the model estimates and predictions, given the flying start distance.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Estimates plot}
-\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{slice}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
-    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
-    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{))}
-\NormalTok{  )}
-
-\CommentTok{\# MSS}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MSS, }\AttributeTok{color =}\NormalTok{ MAC\_string)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MSS (m/s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-34-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-29-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# MAC}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MAC, }\AttributeTok{color =}\NormalTok{ MSS\_string)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MAC (m/s/s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-35-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-30-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# PMAX}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_PMAX, }\AttributeTok{color =}\NormalTok{ MSS\_string)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated PMAX (W/kg)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(), }
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-36-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-31-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Residuals}
-\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
-    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
-    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{group =} \FunctionTok{paste}\NormalTok{(MSS, MAC, flying\_start\_distance)}
-\NormalTok{  )}
-
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  model\_df,}
-  \FunctionTok{aes}\NormalTok{(}
-    \AttributeTok{y =}\NormalTok{ residuals,}
-    \AttributeTok{x =}\NormalTok{ distance,}
-    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
-    \AttributeTok{group =}\NormalTok{ group)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MSS\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MAC\_string) }\SpecialCharTok{+}
-  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-32-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  model\_df,}
-  \FunctionTok{aes}\NormalTok{(}
-    \AttributeTok{y =}\NormalTok{ residuals,}
-    \AttributeTok{x =}\NormalTok{ distance,}
-    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
-    \AttributeTok{group =}\NormalTok{ group)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-33-1} \end{center}
 
@@
 
-\NormalTok{jack\_parameters}
-\CommentTok{\#\textgreater{}                                       MSS   TAU   MAC PMAX}
-\CommentTok{\#\textgreater{} John                                 9.00 1.300  6.92 15.6}
-\CommentTok{\#\textgreater{} Jack {-} No corrections                8.49 0.704 12.06 25.6}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1}
-\CommentTok{\#\textgreater{} Jack {-} Estimated time correction     8.96 1.216  7.37 16.5}
-\CommentTok{\#\textgreater{}                                      time\_correction}
-\CommentTok{\#\textgreater{} John                                           0.000}
-\CommentTok{\#\textgreater{} Jack {-} No corrections                          0.000}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)           0.300}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)           0.500}
-\CommentTok{\#\textgreater{} Jack {-} Estimated time correction               0.284}
-\CommentTok{\#\textgreater{}                                      distance\_correction}
-\CommentTok{\#\textgreater{} John                                                   0}
-\CommentTok{\#\textgreater{} Jack {-} No corrections                                  0}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)                   0}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)                   0}
-\CommentTok{\#\textgreater{} Jack {-} Estimated time correction                       0}
-\end{Highlighting}
-\end{Shaded}
+\FunctionTok{kable}\NormalTok{(jack\_parameters, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"scale\_down"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[!h]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+John & 9.00 & 1.30 & 6.92 & 15.6 & 0.00 & 0\\
+Jack - No corrections & 8.49 & 0.70 & 12.06 & 25.6 & 0.00 & 0\\
+Jack - Fixed time correction (+0.3s) & 9.00 & 1.25 & 7.19 & 16.2 & 0.30 & 0\\
+Jack - Fixed time correction (+0.5s) & 9.62 & 1.77 & 5.43 & 13.1 & 0.50 & 0\\
+Jack - Estimated time correction & 8.96 & 1.22 & 7.37 & 16.5 & 0.28 & 0\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 
-\NormalTok{jack\_parameters}
-\CommentTok{\#\textgreater{}                                       MSS   TAU   MAC PMAX}
-\CommentTok{\#\textgreater{} John                                 9.00 1.300  6.92 15.6}
-\CommentTok{\#\textgreater{} Jack {-} No corrections                8.49 0.704 12.06 25.6}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s) 9.00 1.251  7.19 16.2}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s) 9.62 1.770  5.43 13.1}
-\CommentTok{\#\textgreater{} Jack {-} Estimated time correction     8.96 1.216  7.37 16.5}
-\CommentTok{\#\textgreater{} Jack {-} Estimated distance correction 9.00 1.301  6.92 15.6}
-\CommentTok{\#\textgreater{}                                      time\_correction}
-\CommentTok{\#\textgreater{} John                                           0.000}
-\CommentTok{\#\textgreater{} Jack {-} No corrections                          0.000}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)           0.300}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)           0.500}
-\CommentTok{\#\textgreater{} Jack {-} Estimated time correction               0.284}
-\CommentTok{\#\textgreater{} Jack {-} Estimated distance correction           0.400}
-\CommentTok{\#\textgreater{}                                      distance\_correction}
-\CommentTok{\#\textgreater{} John                                               0.000}
-\CommentTok{\#\textgreater{} Jack {-} No corrections                              0.000}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.3s)               0.000}
-\CommentTok{\#\textgreater{} Jack {-} Fixed time correction (+0.5s)               0.000}
-\CommentTok{\#\textgreater{} Jack {-} Estimated time correction                   0.000}
-\CommentTok{\#\textgreater{} Jack {-} Estimated distance correction               0.503}
-\end{Highlighting}
-\end{Shaded}
+\FunctionTok{kable}\NormalTok{(jack\_parameters, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"scale\_down"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[!h]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+John & 9.00 & 1.30 & 6.92 & 15.6 & 0.00 & 0.0\\
+Jack - No corrections & 8.49 & 0.70 & 12.06 & 25.6 & 0.00 & 0.0\\
+Jack - Fixed time correction (+0.3s) & 9.00 & 1.25 & 7.19 & 16.2 & 0.30 & 0.0\\
+Jack - Fixed time correction (+0.5s) & 9.62 & 1.77 & 5.43 & 13.1 & 0.50 & 0.0\\
+Jack - Estimated time correction & 8.96 & 1.22 & 7.37 & 16.5 & 0.28 & 0.0\\
+\addlinespace
+Jack - Estimated distance correction & 9.00 & 1.30 & 6.92 & 15.6 & 0.40 & 0.5\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 
-We will get back to these issues later, but we can examine how these models perform using simulated data with varying flying start distance.
-The following code contains the wrapper that performs all four models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction):
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{pred\_wrapper }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data) \{}
-\NormalTok{  no\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time}
-\NormalTok{  )}
-
-\NormalTok{  fixed\_correction\_short }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
-    \AttributeTok{time\_correction =} \FloatTok{0.3}
-\NormalTok{  )}
-
-\NormalTok{  fixed\_correction\_long }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
-    \AttributeTok{time\_correction =} \FloatTok{0.5}
-\NormalTok{  )}
-
-\NormalTok{  time\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
-    \AttributeTok{control =} \FunctionTok{nls.control}\NormalTok{(}\AttributeTok{tol =} \DecValTok{1}\NormalTok{)}
-\NormalTok{  )}
-
-\NormalTok{  time\_dist\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{time =}\NormalTok{ data}\SpecialCharTok{$}\NormalTok{time,}
-    \AttributeTok{control =} \FunctionTok{nls.control}\NormalTok{(}\AttributeTok{tol =} \DecValTok{1}\NormalTok{)}
-\NormalTok{  )}
-
-\NormalTok{  params }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
-    \FunctionTok{data.frame}\NormalTok{(}
-      \AttributeTok{model =} \StringTok{"No correction"}\NormalTok{,}
-      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(no\_correction))}
-\NormalTok{    ),}
-    \FunctionTok{data.frame}\NormalTok{(}
-      \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(fixed\_correction\_short))}
-\NormalTok{    ),}
-    \FunctionTok{data.frame}\NormalTok{(}
-      \AttributeTok{model =} \StringTok{"Fixed correction +0.5s"}\NormalTok{,}
-      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(fixed\_correction\_long))}
-\NormalTok{    ),}
-    \FunctionTok{data.frame}\NormalTok{(}
-      \AttributeTok{model =} \StringTok{"Time correction"}\NormalTok{,}
-      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(time\_correction))}
-\NormalTok{    ),}
-    \FunctionTok{data.frame}\NormalTok{(}
-      \AttributeTok{model =} \StringTok{"Time and distance correction"}\NormalTok{,}
-      \FunctionTok{t}\NormalTok{(}\FunctionTok{coef}\NormalTok{(time\_dist\_correction))}
-\NormalTok{    )}
-\NormalTok{  )}
-
-  \FunctionTok{colnames}\NormalTok{(params) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"model"}\NormalTok{, }\StringTok{"est\_MSS"}\NormalTok{, }\StringTok{"est\_TAU"}\NormalTok{, }\StringTok{"est\_MAC"}\NormalTok{, }\StringTok{"est\_PMAX"}\NormalTok{,}
-    \StringTok{"est\_time\_correction"}\NormalTok{, }\StringTok{"est\_distance\_correction"}
-\NormalTok{  )}
-
-\NormalTok{  df }\OtherTok{\textless{}{-}} \FunctionTok{expand\_grid}\NormalTok{(}
-\NormalTok{    data,}
-\NormalTok{    params}
-\NormalTok{  )}
-
-\NormalTok{  df}\SpecialCharTok{$}\NormalTok{predicted\_time }\OtherTok{\textless{}{-}} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
-    \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance,}
-    \AttributeTok{MSS =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_MSS,}
-    \AttributeTok{TAU =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_TAU,}
-    \AttributeTok{time\_correction =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_time\_correction,}
-    \AttributeTok{distance\_correction =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{est\_distance\_correction}
-\NormalTok{  )}
-
-\NormalTok{  df}\SpecialCharTok{$}\NormalTok{residuals }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{predicted\_time }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{time}
-  \FunctionTok{return}\NormalTok{(df)}
-\NormalTok{\}}
-
-\CommentTok{\# estimated parameters and predicted time}
-\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{do}\NormalTok{(}\FunctionTok{pred\_wrapper}\NormalTok{(.)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
+We will get back to these issues later, but we can examine how these models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the Supplemental files for the R code).
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_df}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  model\_df}\SpecialCharTok{$}\NormalTok{model,}
-  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"No correction"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.5s"}\NormalTok{,}
-    \StringTok{"Time correction"}\NormalTok{,}
-    \StringTok{"Time and distance correction"}
-\NormalTok{  )}
-\NormalTok{)}
-\CommentTok{\# Estimates plot}
-\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, flying\_start\_distance, model) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{slice}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
-    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
-    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{))}
-\NormalTok{  )}
-
-\CommentTok{\# MSS}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MSS, }\AttributeTok{color =}\NormalTok{ model)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MSS\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MSS (m/s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# MAC}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_MAC, }\AttributeTok{color =}\NormalTok{ model)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MAC (m/s/s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-43-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# PMAX}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ flying\_start\_distance, }\AttributeTok{y =}\NormalTok{ est\_PMAX, }\AttributeTok{color =}\NormalTok{ model)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated PMAX (W/kg)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-39-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# time\_correction}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}
-    \AttributeTok{x =}\NormalTok{ flying\_start\_distance,}
-    \AttributeTok{y =}\NormalTok{ est\_time\_correction,}
-    \AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}
-    \FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ stolen\_time), }\AttributeTok{color =} \StringTok{"black"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated time correction (s)"}\NormalTok{)  }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
-
-The following figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents \emph{identity line} since flying start distance is already on the x-axis).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# distance\_correction}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}
-    \AttributeTok{x =}\NormalTok{ flying\_start\_distance,}
-    \AttributeTok{y =}\NormalTok{ est\_distance\_correction, }
-    \AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_abline}\NormalTok{(}\AttributeTok{slope =} \DecValTok{1}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Flying start distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated distance correction (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-40-1} \end{center}
+
+The next figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents \emph{identity line} since flying start distance is already on the x-axis).
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Residuals}
-\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
-    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
-    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{group =} \FunctionTok{paste}\NormalTok{(MSS, MAC, flying\_start\_distance)}
-\NormalTok{  )}
-
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  model\_df,}
-  \FunctionTok{aes}\NormalTok{(}
-    \AttributeTok{y =}\NormalTok{ residuals,}
-    \AttributeTok{x =}\NormalTok{ distance,}
-    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
-    \AttributeTok{group =}\NormalTok{ group)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{model) }\SpecialCharTok{+}
-  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
 
@@
 Both of these scenarios represent \emph{time lag} that is added to the split times.
-Below we simulate the effect of this time lag on model estimates and predictions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{sim\_df }\OtherTok{\textless{}{-}} \FunctionTok{expand.grid}\NormalTok{(}
-  \AttributeTok{MSS =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
-  \AttributeTok{MAC =} \FunctionTok{c}\NormalTok{(}\DecValTok{6}\NormalTok{, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{),}
-  \AttributeTok{time\_lag =} \FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\AttributeTok{length.out =} \DecValTok{50}\NormalTok{),}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{, }\DecValTok{50}\NormalTok{)}
-\NormalTok{)}
-
-\NormalTok{sim\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{TAU =}\NormalTok{ MSS }\SpecialCharTok{/}\NormalTok{ MAC,}
-    \AttributeTok{PMAX =}\NormalTok{ MSS }\SpecialCharTok{*}\NormalTok{ MAC }\SpecialCharTok{/} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{true\_time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(distance, MSS, TAU),}
-    \AttributeTok{time =}\NormalTok{ true\_time }\SpecialCharTok{+}\NormalTok{ time\_lag}
-\NormalTok{  )}
-
-\CommentTok{\# Add small noise to allow model fit}
-\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1667}\NormalTok{)}
-\NormalTok{rand\_noise }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(sim\_df), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{4}\NormalTok{)}
-\NormalTok{sim\_df}\SpecialCharTok{$}\NormalTok{time }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df}\SpecialCharTok{$}\NormalTok{time }\SpecialCharTok{+}\NormalTok{ rand\_noise}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# estimated parameters and predicted time}
-\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ sim\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, time\_lag) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{do}\NormalTok{(}\FunctionTok{pred\_wrapper}\NormalTok{(.)) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{ungroup}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
+We have simulated the effect of this time lag on model estimates and predictions (please refer to the Supplemental files for the R code).
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_df}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  model\_df}\SpecialCharTok{$}\NormalTok{model,}
-  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"No correction"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.5s"}\NormalTok{,}
-    \StringTok{"Time correction"}\NormalTok{,}
-    \StringTok{"Time and distance correction"}
-\NormalTok{  )}
-\NormalTok{)}
-\CommentTok{\# Estimates plot}
-\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(MSS, TAU, time\_lag, model) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{slice}\NormalTok{(}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
-    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
-    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{))}
-\NormalTok{  )}
-
-\CommentTok{\# MSS}
-\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_MSS, }\AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MSS\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MAC\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MSS (m/s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
-
-From the figure below it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
+
+From the next figure it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
 Time correction and time and distance corrections model correctly estimated MAC.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# MAC}
-\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_MAC, }\AttributeTok{color =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated MAC (m/s/s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# time\_correction}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_time\_correction, }\AttributeTok{color =}\NormalTok{ model)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_abline}\NormalTok{(}\AttributeTok{slope =} \SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated time correction (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# distance\_correction}
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time\_lag, }\AttributeTok{y =}\NormalTok{ est\_distance\_correction, }\AttributeTok{color =}\NormalTok{ model)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.7}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(MAC\_string }\SpecialCharTok{\textasciitilde{}}\NormalTok{ MSS\_string, }\AttributeTok{scales =} \StringTok{"free\_y"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Time lag (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"estimated distance correction (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-53-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Residuals}
-\NormalTok{model\_df }\OtherTok{\textless{}{-}}\NormalTok{ model\_df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \AttributeTok{MSS\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MSS ="}\NormalTok{, MSS),}
-    \AttributeTok{TAU\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"TAU ="}\NormalTok{, TAU),}
-    \AttributeTok{MAC\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"MAC = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(MAC, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{PMAX\_string =} \FunctionTok{paste}\NormalTok{(}\StringTok{"PMAX = "}\NormalTok{, }\FunctionTok{round}\NormalTok{(PMAX, }\DecValTok{2}\NormalTok{)),}
-    \AttributeTok{group =} \FunctionTok{paste}\NormalTok{(MSS, MAC, time\_lag)}
-\NormalTok{  )}
-
-\FunctionTok{ggplot}\NormalTok{(}
-\NormalTok{  model\_df,}
-  \FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ residuals, }\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{color =}\NormalTok{ time\_lag, }\AttributeTok{group =}\NormalTok{ group)}
-\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{model) }\SpecialCharTok{+}
-  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{)  }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+} 
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Time lag"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-54-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-48-1} \end{center}
 
@@
 \NormalTok{m\_no\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance,}
   \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{gate\_time}
@@
 \NormalTok{m\_no\_correction\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance), }
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance),}
   \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{gate\_time)}
@@
 \NormalTok{m\_no\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance,}
   \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{gate\_time}
@@
 \NormalTok{m\_with\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance), }
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance),}
   \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{gate\_time)}
@@
 \CommentTok{\# Print results}
-\FunctionTok{data.frame}\NormalTok{(}
+\NormalTok{res\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
   \AttributeTok{model =} \FunctionTok{c}\NormalTok{(}
     \StringTok{"Without time correction"}\NormalTok{,}
-    \StringTok{"Without time correction with zeros added"}\NormalTok{, }
+    \StringTok{"Without time correction with zeros added"}\NormalTok{,}
     \StringTok{"With time correction"}\NormalTok{,}
-    \StringTok{"With time correction with zeros added"}\NormalTok{),}
+    \StringTok{"With time correction with zeros added"}
+\NormalTok{  ),}
   \FunctionTok{rbind}\NormalTok{(}
@@
     \FunctionTok{coef}\NormalTok{(m\_no\_zero),}
-    \FunctionTok{coef}\NormalTok{(m\_with\_zero))}
-\NormalTok{)}
-\CommentTok{\#\textgreater{}                                      model  MSS  TAU  MAC PMAX}
-\CommentTok{\#\textgreater{} 1                  Without time correction 8.78 1.09 8.06 17.7}
-\CommentTok{\#\textgreater{} 2 Without time correction with zeros added 8.78 1.09 8.06 17.7}
-\CommentTok{\#\textgreater{} 3                     With time correction 9.00 1.30 6.92 15.6}
-\CommentTok{\#\textgreater{} 4    With time correction with zeros added 8.79 1.10 7.97 17.5}
-\CommentTok{\#\textgreater{}   time\_correction distance\_correction}
-\CommentTok{\#\textgreater{} 1         0.00000                   0}
-\CommentTok{\#\textgreater{} 2         0.00000                   0}
-\CommentTok{\#\textgreater{} 3         0.09994                   0}
-\CommentTok{\#\textgreater{} 4         0.00739                   0}
-\end{Highlighting}
-\end{Shaded}
+    \FunctionTok{coef}\NormalTok{(m\_with\_zero)}
+\NormalTok{  )}
+\NormalTok{)}
+
+\FunctionTok{kable}\NormalTok{(res\_df, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"scale\_down"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[!h]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+model & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+Without time correction & 8.78 & 1.09 & 8.06 & 17.7 & 0.00 & 0\\
+Without time correction with zeros added & 8.78 & 1.09 & 8.06 & 17.7 & 0.00 & 0\\
+With time correction & 9.00 & 1.30 & 6.92 & 15.6 & 0.10 & 0\\
+With time correction with zeros added & 8.79 & 1.10 & 7.97 & 17.5 & 0.01 & 0\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 \NormalTok{)}
-
-\NormalTok{jack\_LOOCV}
-\CommentTok{\#\textgreater{} Estimated model parameters}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}                 MSS                 TAU                 MAC }
-\CommentTok{\#\textgreater{}               8.958               1.216               7.367 }
-\CommentTok{\#\textgreater{}                PMAX     time\_correction distance\_correction }
-\CommentTok{\#\textgreater{}              16.499               0.284               0.000 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit estimators}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}   0.00181   1.00000  {-}0.00109   0.00189   0.00189   0.00114 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}   0.00104   0.04981 }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Leave{-}One{-}Out Cross{-}Validation}
-\CommentTok{\#\textgreater{} {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-\CommentTok{\#\textgreater{} Parameters:}
-\CommentTok{\#\textgreater{}    MSS  TAU  MAC PMAX time\_correction distance\_correction}
-\CommentTok{\#\textgreater{} 1 8.98 1.25 7.21 16.2           0.300                   0}
-\CommentTok{\#\textgreater{} 2 8.97 1.22 7.35 16.5           0.284                   0}
-\CommentTok{\#\textgreater{} 3 8.95 1.21 7.40 16.5           0.282                   0}
-\CommentTok{\#\textgreater{} 4 8.96 1.21 7.39 16.5           0.282                   0}
-\CommentTok{\#\textgreater{} 5 8.93 1.20 7.45 16.6           0.278                   0}
-\CommentTok{\#\textgreater{} }
-\CommentTok{\#\textgreater{} Model fit:}
-\CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}        NA   1.00000  {-}0.00639   0.00510   0.00639   0.00401 }
-\CommentTok{\#\textgreater{}       MAE      MAPE }
-\CommentTok{\#\textgreater{}   0.00349   0.18387}
 \end{Highlighting}
@@
 
-Next we plot estimated parameters across LOOCV folds:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{LOOCV}\SpecialCharTok{$}\NormalTok{parameters}
-
-\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{pivot\_longer}\NormalTok{(df, }\AttributeTok{cols =} \DecValTok{1}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"parameter"}\NormalTok{)}
-
-\NormalTok{df}\SpecialCharTok{$}\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  df}\SpecialCharTok{$}\NormalTok{parameter,}
-  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"MSS"}\NormalTok{,}
-    \StringTok{"TAU"}\NormalTok{,}
-    \StringTok{"MAC"}\NormalTok{,}
-    \StringTok{"PMAX"}\NormalTok{,}
-    \StringTok{"time\_correction"}\NormalTok{,}
-    \StringTok{"distance\_correction"}
-\NormalTok{  )}
-\NormalTok{)}
-
-\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ value)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_boxplot}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{parameter, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{axis.ticks.y =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{axis.text.y =} \FunctionTok{element\_blank}\NormalTok{()}
-\NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
+Next we plot estimated parameters across LOOCV folds (please refer to the Supplemental files for the R code):
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
-  \AttributeTok{distance =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{distance,}
-  \AttributeTok{time =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{time,}
-  \AttributeTok{pred\_time =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{pred\_time,}
-  \AttributeTok{LOOCV\_time =}\NormalTok{ jack\_LOOCV}\SpecialCharTok{$}\NormalTok{LOOCV}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{pred\_time}
-\NormalTok{)}
-
-\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ df }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(}\StringTok{"pred\_time"}\NormalTok{, }\StringTok{"LOOCV\_time"}\NormalTok{))}
-
-\NormalTok{df}\SpecialCharTok{$}\NormalTok{resid }\OtherTok{\textless{}{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{value }\SpecialCharTok{{-}}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{time}
-
-\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ distance, }\AttributeTok{y =}\NormalTok{ resid, }\AttributeTok{color =}\NormalTok{ name)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{()) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Predicted {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{theme}\NormalTok{(}
-    \AttributeTok{legend.title =} \FunctionTok{element\_blank}\NormalTok{(),}
-    \AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-58-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
 
@@
 
-\NormalTok{bolt\_model}
-\CommentTok{\#\textgreater{}                      model  MSS   TAU   MAC PMAX time\_correction}
-\CommentTok{\#\textgreater{} 1            No correction 12.1 1.564  7.77 23.6         0.00000}
-\CommentTok{\#\textgreater{} 2       No correction {-} RT 11.7 1.205  9.74 28.6        {-}0.18300}
-\CommentTok{\#\textgreater{} 3          Time correction 11.7 1.202  9.76 28.6        {-}0.18483}
-\CommentTok{\#\textgreater{} 4     Time correction {-} RT 11.7 1.202  9.76 28.6        {-}0.00183}
-\CommentTok{\#\textgreater{} 5      Distance correction 11.6 0.855 13.56 39.3        {-}0.81151}
-\CommentTok{\#\textgreater{} 6 Distance correction {-} RT 11.6 0.855 13.56 39.3        {-}0.62851}
-\CommentTok{\#\textgreater{}   distance\_correction}
-\CommentTok{\#\textgreater{} 1                0.00}
-\CommentTok{\#\textgreater{} 2                0.00}
-\CommentTok{\#\textgreater{} 3                0.00}
-\CommentTok{\#\textgreater{} 4                0.00}
-\CommentTok{\#\textgreater{} 5               {-}3.98}
-\CommentTok{\#\textgreater{} 6               {-}3.98}
-\end{Highlighting}
-\end{Shaded}
+\FunctionTok{kable}\NormalTok{(bolt\_model, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"scale\_down"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[!h]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+model & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+No correction & 12.1 & 1.56 & 7.77 & 23.6 & 0.00 & 0.00\\
+No correction - RT & 11.7 & 1.21 & 9.74 & 28.6 & -0.18 & 0.00\\
+Time correction & 11.7 & 1.20 & 9.76 & 28.6 & -0.18 & 0.00\\
+Time correction - RT & 11.7 & 1.20 & 9.76 & 28.6 & 0.00 & 0.00\\
+Distance correction & 11.6 & 0.85 & 13.56 & 39.3 & -0.81 & -3.98\\
+\addlinespace
+Distance correction - RT & 11.6 & 0.85 & 13.56 & 39.3 & -0.63 & -3.98\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
     \AttributeTok{dist\_99\_MSS =} \FunctionTok{find\_velocity\_critical\_distance}\NormalTok{(}
-      \AttributeTok{MSS =}\NormalTok{ MSS, }\AttributeTok{TAU =}\NormalTok{ TAU, }
-      \CommentTok{\#time\_correction = time\_correction, }
-      \CommentTok{\#distance\_correction = distance\_correction,}
+      \AttributeTok{MSS =}\NormalTok{ MSS, }\AttributeTok{TAU =}\NormalTok{ TAU,}
+      \CommentTok{\# time\_correction = time\_correction,}
+      \CommentTok{\# distance\_correction = distance\_correction,}
       \AttributeTok{percent =} \FloatTok{0.99}
@@
     \AttributeTok{time\_99\_MSS =} \FunctionTok{find\_velocity\_critical\_time}\NormalTok{(}
-      \AttributeTok{MSS =}\NormalTok{ MSS, }\AttributeTok{TAU =}\NormalTok{ TAU, }
-      \CommentTok{\#time\_correction = time\_correction, }
+      \AttributeTok{MSS =}\NormalTok{ MSS, }\AttributeTok{TAU =}\NormalTok{ TAU,}
+      \CommentTok{\# time\_correction = time\_correction,}
       \AttributeTok{percent =} \FloatTok{0.99}
@@
 
-\NormalTok{bolt\_model[}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{)]}
-\CommentTok{\#\textgreater{} \# A tibble: 6 x 3}
-\CommentTok{\#\textgreater{} \# Groups:   model [6]}
-\CommentTok{\#\textgreater{}   model                    dist\_99\_MSS time\_99\_MSS}
-\CommentTok{\#\textgreater{}   \textless{}chr\textgreater{}                          \textless{}dbl\textgreater{}       \textless{}dbl\textgreater{}}
-\CommentTok{\#\textgreater{} 1 No correction                   68.7        7.20}
-\CommentTok{\#\textgreater{} 2 No correction {-} RT              51.1        5.55}
-\CommentTok{\#\textgreater{} 3 Time correction                 51.0        5.54}
-\CommentTok{\#\textgreater{} 4 Time correction {-} RT            51.0        5.54}
-\CommentTok{\#\textgreater{} 5 Distance correction             35.8        3.94}
-\CommentTok{\#\textgreater{} 6 Distance correction {-} RT        35.8        3.94}
-\end{Highlighting}
-\end{Shaded}
+\FunctionTok{kable}\NormalTok{(bolt\_model[}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{)], }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{tabular}{lrr}
+\toprule
+model & dist\_99\_MSS & time\_99\_MSS\\
+\midrule
+No correction & 68.7 & 7.20\\
+No correction - RT & 51.1 & 5.55\\
+Time correction & 51.0 & 5.54\\
+Time correction - RT & 51.0 & 5.54\\
+Distance correction & 35.8 & 3.94\\
+\addlinespace
+Distance correction - RT & 35.8 & 3.94\\
+\bottomrule
+\end{tabular}
 
@@
 
-The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{no\_corrections }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_splits}\NormalTok{(}
-\NormalTok{  df,}
-  \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
-  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
-  \AttributeTok{athlete =} \StringTok{"Athlete"}
-\NormalTok{)}
-
-\NormalTok{fixed\_correction }\OtherTok{\textless{}{-}} \FunctionTok{mixed\_model\_using\_splits}\NormalTok{(}
-\NormalTok{  df,}
-  \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
-  \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
-  \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
-  \AttributeTok{time\_correction =} \FloatTok{0.3}
-\NormalTok{)}
-
-\NormalTok{time\_correction\_fixed }\OtherTok{\textless{}{-}} 
-  \FunctionTok{mixed\_model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-\NormalTok{    df,}
-    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
-    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
-    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
-    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
-\NormalTok{  )}
-
-\NormalTok{time\_correction\_random }\OtherTok{\textless{}{-}} 
-  \FunctionTok{mixed\_model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-\NormalTok{    df,}
-    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
-    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
-    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
-    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
-\NormalTok{  )}
-
-\NormalTok{time\_distance\_correction\_fixed }\OtherTok{\textless{}{-}} 
-  \FunctionTok{mixed\_model\_using\_splits\_with\_corrections}\NormalTok{(}
-\NormalTok{    df,}
-    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
-    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
-    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
-    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
-\NormalTok{  )}
-
-\NormalTok{time\_distance\_correction\_random }\OtherTok{\textless{}{-}} 
-  \FunctionTok{mixed\_model\_using\_splits\_with\_corrections}\NormalTok{(}
-\NormalTok{    df,}
-    \AttributeTok{distance =} \StringTok{"distance"}\NormalTok{,}
-    \AttributeTok{time =} \StringTok{"time"}\NormalTok{,}
-    \AttributeTok{athlete =} \StringTok{"Athlete"}\NormalTok{,}
-    \AttributeTok{random =}\NormalTok{ MSS }\SpecialCharTok{+}\NormalTok{ TAU }\SpecialCharTok{+}\NormalTok{ time\_correction }\SpecialCharTok{+}\NormalTok{ distance\_correction }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}
-\NormalTok{  )}
-\end{Highlighting}
-\end{Shaded}
+The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model (please refer to the Supplemental files for the R code).
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_fit }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"No corrections"}\NormalTok{,}
-    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(no\_corrections}\SpecialCharTok{$}\NormalTok{model\_fit))}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(fixed\_correction}\SpecialCharTok{$}\NormalTok{model\_fit))}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time correction fixed"}\NormalTok{,}
-    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_correction\_fixed}\SpecialCharTok{$}\NormalTok{model\_fit))}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time correction random"}\NormalTok{,}
-    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_correction\_random}\SpecialCharTok{$}\NormalTok{model\_fit))}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_distance\_correction\_fixed}\SpecialCharTok{$}\NormalTok{model\_fit))}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time and distance correction random"}\NormalTok{,}
-    \FunctionTok{t}\NormalTok{(}\FunctionTok{unlist}\NormalTok{(time\_distance\_correction\_random}\SpecialCharTok{$}\NormalTok{model\_fit))}
-\NormalTok{  )}
-\NormalTok{)}
-
-\NormalTok{model\_fit}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  model\_fit}\SpecialCharTok{$}\NormalTok{model,}
-  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
-    \StringTok{"No corrections"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \StringTok{"Time correction fixed"}\NormalTok{,}
-    \StringTok{"Time correction random"}\NormalTok{,}
-    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-    \StringTok{"Time and distance correction random"}
-\NormalTok{  ))}
-\NormalTok{)}
-
-\FunctionTok{ggplot}\NormalTok{(model\_fit, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ RSE, }\AttributeTok{y =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"RSE (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-63-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{est\_params }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"No corrections"}\NormalTok{,}
-\NormalTok{    no\_corrections}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-\NormalTok{    fixed\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time correction fixed"}\NormalTok{,}
-\NormalTok{    time\_correction\_fixed}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time correction random"}\NormalTok{,}
-\NormalTok{    time\_correction\_random}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-\NormalTok{    time\_distance\_correction\_fixed}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time and distance correction random"}\NormalTok{,}
-\NormalTok{    time\_distance\_correction\_random}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{random}
-\NormalTok{  )}
-\NormalTok{)}
-
-\NormalTok{est\_params}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  est\_params}\SpecialCharTok{$}\NormalTok{model,}
-  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
-    \StringTok{"No corrections"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \StringTok{"Time correction fixed"}\NormalTok{,}
-    \StringTok{"Time correction random"}\NormalTok{,}
-    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-    \StringTok{"Time and distance correction random"}
-\NormalTok{  ))}
-\NormalTok{)}
-
-\NormalTok{est\_params }\OtherTok{\textless{}{-}}\NormalTok{ est\_params }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{), }\AttributeTok{names\_to =} \StringTok{"parameter"}\NormalTok{)}
-
-\NormalTok{est\_params}\SpecialCharTok{$}\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  est\_params}\SpecialCharTok{$}\NormalTok{parameter,}
-  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"MSS"}\NormalTok{,}
-    \StringTok{"TAU"}\NormalTok{,}
-    \StringTok{"MAC"}\NormalTok{,}
-    \StringTok{"PMAX"}\NormalTok{,}
-    \StringTok{"time\_correction"}\NormalTok{,}
-    \StringTok{"distance\_correction"}
-\NormalTok{  )}
-\NormalTok{)}
-
-\FunctionTok{ggplot}\NormalTok{(est\_params, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ model, }\AttributeTok{x =}\NormalTok{ value)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_boxplot}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{parameter, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-64-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-58-1} \end{center}
 
@@
 If the residuals are within this magnitude band, then the model is good in making practically useful predictions.
-
 Error bars represent residual bias \(\pm\) 1 SD.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{model\_resid }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"No corrections"}\NormalTok{,}
-\NormalTok{    no\_corrections}\SpecialCharTok{$}\NormalTok{data}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-\NormalTok{    fixed\_correction}\SpecialCharTok{$}\NormalTok{data}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time correction fixed"}\NormalTok{,}
-\NormalTok{    time\_correction\_fixed}\SpecialCharTok{$}\NormalTok{data}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time correction random"}\NormalTok{,}
-\NormalTok{    time\_correction\_random}\SpecialCharTok{$}\NormalTok{data}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-\NormalTok{    time\_distance\_correction\_fixed}\SpecialCharTok{$}\NormalTok{data}
-\NormalTok{  ),}
-  \FunctionTok{data.frame}\NormalTok{(}
-    \AttributeTok{model =} \StringTok{"Time and distance correction random"}\NormalTok{,}
-\NormalTok{    time\_distance\_correction\_random}\SpecialCharTok{$}\NormalTok{data}
-\NormalTok{  )}
-\NormalTok{)}
-
-\NormalTok{model\_resid}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  model\_resid}\SpecialCharTok{$}\NormalTok{model,}
-  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
-    \StringTok{"No corrections"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \StringTok{"Time correction fixed"}\NormalTok{,}
-    \StringTok{"Time correction random"}\NormalTok{,}
-    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-    \StringTok{"Time and distance correction random"}
-\NormalTok{  ))}
-\NormalTok{)}
-
-\NormalTok{model\_resid}\SpecialCharTok{$}\NormalTok{resid }\OtherTok{\textless{}{-}}\NormalTok{ model\_resid}\SpecialCharTok{$}\NormalTok{pred\_time }\SpecialCharTok{{-}}\NormalTok{ model\_resid}\SpecialCharTok{$}\NormalTok{time}
-
-\CommentTok{\# Create SWC / SESOI band}
-\NormalTok{model\_SESOI }\OtherTok{\textless{}{-}}\NormalTok{ model\_resid }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{group\_by}\NormalTok{(model, distance) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{summarise}\NormalTok{(}
-    \AttributeTok{bias =} \FunctionTok{mean}\NormalTok{(resid),}
-    \AttributeTok{variance =} \FunctionTok{sd}\NormalTok{(resid),}
-    \AttributeTok{upper =}\NormalTok{ bias }\SpecialCharTok{+}\NormalTok{ variance,}
-    \AttributeTok{lower =}\NormalTok{ bias }\SpecialCharTok{{-}}\NormalTok{ variance,}
-    \AttributeTok{MAD =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{abs}\NormalTok{(resid)),}
-    \AttributeTok{SESOI\_upper =} \FunctionTok{sd}\NormalTok{(time) }\SpecialCharTok{*} \FloatTok{0.2}\NormalTok{,}
-    \AttributeTok{SESOI\_lower =} \SpecialCharTok{{-}}\FunctionTok{sd}\NormalTok{(time) }\SpecialCharTok{*} \FloatTok{0.2}
-\NormalTok{  )}
-
-\CommentTok{\# Plot}
-\FunctionTok{ggplot}\NormalTok{(model\_resid, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{y =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_vline}\NormalTok{(}
-    \AttributeTok{data =}\NormalTok{ model\_SESOI,}
-    \FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ SESOI\_lower),}
-    \AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}
-\NormalTok{  ) }\SpecialCharTok{+}
-  \FunctionTok{geom\_vline}\NormalTok{(}
-    \AttributeTok{data =}\NormalTok{ model\_SESOI,}
-    \FunctionTok{aes}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ SESOI\_upper),}
-    \AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}
-\NormalTok{  ) }\SpecialCharTok{+}
-  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{color =} \StringTok{"blue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_jitter}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ resid), }\AttributeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{height =} \FloatTok{0.25}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_errorbarh}\NormalTok{(}
-    \AttributeTok{data =}\NormalTok{ model\_SESOI,}
-    \FunctionTok{aes}\NormalTok{(}\AttributeTok{xmin =}\NormalTok{ lower, }\AttributeTok{xmax =}\NormalTok{ upper),}
-    \AttributeTok{height =} \FloatTok{0.1}\NormalTok{, }\AttributeTok{color =} \StringTok{"black"}
-\NormalTok{  ) }\SpecialCharTok{+}
-  \FunctionTok{geom\_point}\NormalTok{(}\AttributeTok{data =}\NormalTok{ model\_SESOI, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ bias), }\AttributeTok{color =} \StringTok{"black"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{distance, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Predicted time {-} observed time (s)"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-65-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-59-1} \end{center}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{df }\OtherTok{\textless{}{-}}\NormalTok{ model\_SESOI }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{), }\AttributeTok{names\_to =} \StringTok{"estimator"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{filter}\NormalTok{(estimator }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"bias"}\NormalTok{, }\StringTok{"variance"}\NormalTok{, }\StringTok{"MAD"}\NormalTok{))}
-
-\NormalTok{df}\SpecialCharTok{$}\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  df}\SpecialCharTok{$}\NormalTok{model,}
-  \AttributeTok{levels =} \FunctionTok{rev}\NormalTok{(}\FunctionTok{c}\NormalTok{(}
-    \StringTok{"No corrections"}\NormalTok{,}
-    \StringTok{"Fixed correction +0.3s"}\NormalTok{,}
-    \StringTok{"Time correction fixed"}\NormalTok{,}
-    \StringTok{"Time correction random"}\NormalTok{,}
-    \StringTok{"Time and distance correction fixed"}\NormalTok{,}
-    \StringTok{"Time and distance correction random"}
-\NormalTok{  ))}
-\NormalTok{)}
-
-\NormalTok{df}\SpecialCharTok{$}\NormalTok{estimator }\OtherTok{\textless{}{-}} \FunctionTok{factor}\NormalTok{(}
-\NormalTok{  df}\SpecialCharTok{$}\NormalTok{estimator,}
-  \AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"bias"}\NormalTok{, }\StringTok{"variance"}\NormalTok{, }\StringTok{"MAD"}\NormalTok{)}
-\NormalTok{)}
-
-\FunctionTok{ggplot}\NormalTok{(df, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ value, }\AttributeTok{y =}\NormalTok{ model)) }\SpecialCharTok{+}
-  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(distance }\SpecialCharTok{\textasciitilde{}}\NormalTok{ estimator, }\AttributeTok{scales =} \StringTok{"free\_x"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{xlab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{ylab}\NormalTok{(}\ConstantTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\includegraphics[width=1\linewidth]{paper_files/figure-latex/unnamed-chunk-66-1} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-60-1} \end{center}
 
@@
 
+\leavevmode\hypertarget{ref-morinSpreadsheetSprintAcceleration2017}{}%
+Morin, J.
+B.
+2017.
+{``A Spreadsheet for {Sprint} Acceleration {Force}-{Velocity}-{Power} Profiling.''} {JB Morin, PhD - Sport Science}.
+December 13, 2017.
+\url{https://jbmorin.net/2017/12/13/a-spreadsheet-for-sprint-acceleration-force-velocity-power-profiling/}.
+
+\leavevmode\hypertarget{ref-morinInterpretingPowerForceVelocityProfiles2016}{}%
+Morin, Jean-Benoit, and Pierre Samozino.
+2016.
+{``Interpreting {Power}-{Force}-{Velocity Profiles} for {Individualized} and {Specific Training}.''} \emph{International Journal of Sports Physiology and Performance} 11 (2): 267--72.
+\url{https://doi.org/10.1123/ijspp.2015-0638}.
+
+\leavevmode\hypertarget{ref-morinSpreadsheetSprintAcceleration2019}{}%
+---------. 2019. {``Spreadsheet for {Sprint} Acceleration Force-Velocity-Power Profiling.''}
+
 \leavevmode\hypertarget{ref-morinSimpleMethodComputing2019}{}%
@@
 \url{https://doi.org/10.1016/j.jbiomech.2019.07.020}.
-
-\leavevmode\hypertarget{ref-morinInterpretingPowerForceVelocityProfiles2016}{}%
-Morin, Jean-Benoît, and Pierre Samozino.
-2016.
-{``Interpreting {Power}-{Force}-{Velocity Profiles} for {Individualized} and {Specific Training}.''} \emph{International Journal of Sports Physiology and Performance} 11 (2): 267--72.
-\url{https://doi.org/10.1123/ijspp.2015-0638}.
 
@@
 
+\leavevmode\hypertarget{ref-stenrothSpreadsheetSprintAcceleration2020}{}%
+Stenroth, Lauri, and Paavo Vartiainen.
+2020.
+{``Spreadsheet for Sprint Acceleration Force-Velocity-Power Profiling with Optimization to Correct Start Time.''} \url{https://doi.org/10.13140/RG.2.2.12841.83045}.
+
 \leavevmode\hypertarget{ref-stenrothForcevelocityProfilingIce2020}{}%
@@
 
+\leavevmode\hypertarget{ref-knitr2014}{}%
+Xie, Yihui.
+2014.
+{``Knitr: A Comprehensive Tool for Reproducible Research in {R}.''} In \emph{Implementing Reproducible Computational Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
+Peng.
+Chapman; Hall/CRC.
+\url{http://www.crcpress.com/product/isbn/9781466561595}.
+
+\leavevmode\hypertarget{ref-knitr2015}{}%
+---------. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.org/knitr/}.
+
+\leavevmode\hypertarget{ref-R-knitr}{}%
+---------. 2020. \emph{Knitr: A General-Purpose Package for Dynamic Report Generation in r}. \url{https://yihui.org/knitr/}.
+
+\leavevmode\hypertarget{ref-R-kableExtra}{}%
+Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+
 \end{CSLReferences}

### 2021-04-14 paper.Rmd
Fixed few typos

@@
 
-The **shorts** package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
-To best of our knowledge, scientist, researchers, and coaches have been performing short sprints modeling using the built-in solver function of Excel (Microsoft Corporation, Redmond, Washington, United States) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothForcevelocityProfilingIce2020; @stenrothSpreadsheetSprintAcceleration2020; @samozinoSimpleMethodMeasuring2016; @clarkNFLCombine40Yard2017; @morinSimpleMethodComputing2019], which makes the **shorts** package major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
+The **shorts** package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sports scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+To best of our knowledge, scientist, researchers, and coaches have been performing short sprints modeling using the built-in solver function of Excel (Microsoft Corporation, Redmond, Washington, United States) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothForcevelocityProfilingIce2020; @stenrothSpreadsheetSprintAcceleration2020; @samozinoSimpleMethodMeasuring2016; @clarkNFLCombine40Yard2017; @morinSimpleMethodComputing2019], which makes the **shorts** package a major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
 

### 2021-04-14 paper.tex
Fixed few typos

@@
 
-The \textbf{shorts} package (\protect\hyperlink{ref-R-shorts}{Jovanovic 2020}), written in the R language (\protect\hyperlink{ref-R-base}{R Core Team 2020}), represents an open-source tool to help sport scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+The \textbf{shorts} package (\protect\hyperlink{ref-R-shorts}{Jovanovic 2020}), written in the R language (\protect\hyperlink{ref-R-base}{R Core Team 2020}), represents an open-source tool to help sports scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
 2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
@@
 2017}; \protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
-2019}), which makes the \textbf{shorts} package major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
+2019}), which makes the \textbf{shorts} package a major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
 

### 2021-07-21 paper.Rmd
Added section to want about predict_XXX issues with distance correction model when distance correction <>0m

@@
 title: "shorts: An R Package for Modeling Short Sprints"
-preprint: false
+preprint: true
 date: "`r Sys.Date()`"
@@
 
+def.chunk.hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def.chunk.hook(x, options)
+  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
+})
+
 knitr::opts_chunk$set(
@@
   message = FALSE,
-  width = 65
+  width = 65,
+  size = "small"
 )
@@
 # automatically create a bib database for R packages
-# knitr::write_bib(c(
-#  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW"
-# ), "packages.bib")
+knitr::write_bib(c(
+  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW", "kableExtra"
+ ), "packages.bib")
 
@@
 
-Although TAU is used in the equations, and later estimated, it is preferred to use MAC instead since it is easier to grasp, particularly for less math inclined coaches.
+Although TAU is used in the equations, and later estimated, we prefer to use MAC instead since it is easier to grasp, particularly for less math inclined coaches.
 
@@
 If the athlete's body mass (kg) is known, as well as additional air resistance parameters (see [Air resistance and the calculation of force and mechanical power] section of this paper), *Force-Velocity Profile* can be estimated (see [Force-Velocity profile] section of this paper).
+
 ```{r four-athletes-profile, echo=FALSE, fig.cap="Acceleration-Velocity profile of four athletes with different MSS and MAC parameters."}
@@
 Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **shorts** package using non-linear least squares regression implemented in the `nls()` function in the **base R** [@R-base] and `nlme()` function in the **nlme** package [@R-nlme] for the mixed-effect models.
-
-## Estimating short sprint parameters using split times
+Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **shorts** package using non-linear least squares regression implemented in the `nls()` function [@batesNonlinearModels1992; @batesNonlinearRegressionAnalysis2007] in the **base R** [@R-base] and `nlme()` function in the **nlme** package [@R-nlme] for the mixed-effect models.
+
+## Estimating short sprint parameters using timing gates split times
 
 Let's consider an example of an athlete with MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$ performing 40m sprint with timing gates positioned at each 10m split.
-For split times, distance is a predictor, and time is the outcome variable, thus the equation \@ref(eq:velocity-time) becomes:
+For split times, distance is a predictor, and time is the outcome variable, thus the equation \@ref(eq:distance-time) becomes:
 
@@
 
-$W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW]. MSS and TAU parameters are estimated using `model_using_splits()` function:
+$W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW].
+Researchers often incorrectly use the equation \@ref(eq:distance-time) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothSpreadsheetSprintAcceleration2020], in which the time is the predictor and distance is the outcome variable, instead of statistically correct equation \@ref(eq:time-distance) [@motulskyIntuitiveBiostatisticsNonmathematical2018, p.
+341].
+
+MSS and TAU parameters are estimated using `model_using_splits()` function:
 
@@
 S3 method `coef()` when applied on mixed-model result will return both the fixed and random effects.
-To create a simple plot of the model, use S3 `plot()` method.
 
@@
 ```
+
+To create a simple plot of the model, use S3 `plot()` method.
 
@@
 
-## Problems at the start when using split times
+## Problems at the start when using timing gates
 
@@
 In other words, Jack is using a *flying start*, a common scenario when testing field sports athletes.
+Flying start distance is often recommended to avoid premature triggering of the timing system by lifted knees or swinging
+arms [@altmannAccuracySingleBeam2018; @altmannDifferentStartingDistances2015; @altmannValiditySingleBeamTiming2017; @haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintRunningPerformance2016].
 
@@
 
-Here is a graphical representation of the sprint splits (please refer to the Supplemental files for the R code):
+Here is a graphical representation of the sprint splits (please refer to the [Supplemental material] for the R code):
 
@@
 
-To explore this further, we have run a simple simulation with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 ($ms^{-1}$ and $ms^{-2}$ respectively), and flying start distance varying from 0 to 1 m (please refer to the Supplemental files for the R code).
-With a simulation dataset, we can check the model estimates and predictions, given the flying start distance.
+To explore this further, we have run a simple simulation by increasing Jack's flying start distance from 0 to 1 m and depicting the estimated MSS, TAU, MAC, and PMAX parameters (please refer to the [Supplemental material] for the R code).
 
@@
 sim_df <- expand.grid(
-  MSS = c(6, 7, 8, 9),
-  MAC = c(6, 7, 8, 9),
-  flying_start_distance = c(
-    seq(0, 0.001, length.out = 20),
-    seq(0.001, 0.01, length.out = 20),
-    seq(0.01, 0.1, length.out = 20),
-    seq(0.1, 1, length.out = 20)
-  ),
-  distance = c(5, 10, 20, 30, 40, 50)
+  MSS = 9,
+  MAC = 9 / 1.3,
+  flying_start_distance = seq(0.0001, 1, length.out = 1000),
+  distance = c(5, 10, 20, 30, 40)
 )
@@
 # Prediction residuals
-model_df$residuals <- model_df$predicted_time - model_df$time
-```
-
-The following figure demonstrates the effect of flying start distance on estimated MSS:
+model_df$residuals <-  model_df$time - model_df$predicted_time
+```
 
@@
 # Estimates plot
-df <- model_df %>%
+plot_df <- model_df %>%
   group_by(MSS, TAU, flying_start_distance) %>%
@@
   mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2))
-  )
-
-# MSS
+    `MSS (m/s)` = est_MSS,
+    `TAU` = est_TAU,
+    `MAC (m/s/s)` = est_MAC,
+    `PMAX (W)` = est_PMAX
+  ) %>%
+  pivot_longer(cols = c("MSS (m/s)", "TAU", "MAC (m/s/s)", "PMAX (W)")) %>%
+  mutate(
+    name = factor(name, levels = c("MSS (m/s)", "TAU", "MAC (m/s/s)", "PMAX (W)"))
+  )
+
+ggplot(plot_df, aes(x = flying_start_distance, y = value)) +
+  theme_bw(8) +
+  geom_line(alpha = 0.7) +
+  facet_wrap(~name, scales = "free_y") + 
+  xlab("Flying start distance (m)") +
+  ylab("Estimated parameter")
+```
+
+As can be seen from the figure, MSS is underestimated as flying start distance increases.
+MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
+Estimated PMAX is also overestimated as flying start distance increases.
+
+Model residuals are also affected by the flying start distance.
+The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40 m versus 5, 15, 30 m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simple simulation:
+
+```{r echo=FALSE}
+# Residuals
+plot_df <- model_df %>%
+  mutate(
+    distance_string = paste0(distance, "m"),
+    distance_string = factor(
+      distance_string,
+      levels = c("5m", "10m", "20m", "30m", "40m")
+    )
+  )
+
 ggplot(
-  df,
-  aes(x = flying_start_distance, y = est_MSS, color = MAC_string)
+  plot_df,
+  aes(
+    y = residuals,
+    x = flying_start_distance,
+  )
 ) +
   theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_wrap(~MSS_string, scales = "free_y") +
+  geom_line() +
+  facet_wrap(~distance_string) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
   xlab("Flying start distance (m)") +
-  ylab("estimated MSS (m/s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-As can be seen from the figure, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
-
-```{r echo=FALSE}
-# MAC
-ggplot(
-  df,
-  aes(x = flying_start_distance, y = est_MAC, color = MSS_string)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_wrap(~MAC_string, scales = "free_y") +
-  xlab("Flying start distance (m)") +
-  ylab("estimated MAC (m/s/s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
-
-And finally, the following image demonstrates the effect of flying start distance on estimated PMAX:
-
-```{r echo=FALSE}
-# PMAX
-ggplot(
-  df,
-  aes(x = flying_start_distance, y = est_PMAX, color = MSS_string)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_wrap(~MAC_string, scales = "free_y") +
-  xlab("Flying start distance (m)") +
-  ylab("estimated PMAX (W/kg)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-Estimated PMAX is also overestimated as flying start distance increases.
-
-Model residuals are also affected by flying start distance.
-The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40 m versus 5, 15, 30 m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simulation:
+  ylab("Observed time - predicted time (s)")
+```
+
+Another way to visualize the effect of the flying start distance on split distance residuals can be found on the next figure:
 
@@
 # Residuals
-model_df <- model_df %>%
-  mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-    group = paste(MSS, MAC, flying_start_distance)
-  )
-
 ggplot(
@@
     color = flying_start_distance,
-    group = group
-  )
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.3) +
-  facet_grid(MSS_string ~ MAC_string) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-  xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)") +
-  theme(legend.position = "top") +
-  labs(color = "Flying start distance")
-```
-
-If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler figure conveying issues with residuals when there is a flying start:
-
-```{r echo=FALSE}
-ggplot(
-  model_df,
-  aes(
-    y = residuals,
-    x = distance,
-    color = flying_start_distance,
-    group = group
+    group = flying_start_distance
   )
@@
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)") +
+  ylab("Observed - predicted time (s)") +
   theme(legend.position = "top") +
@@
 
-In Jack's case, both +0.3 s fixed time correction and time correction estimation yield parameters closer to John's (i.e. true parameters).
-
-Another model definition, which is a novel approach implemented in the **shorts** package, is to utilize *distance correction*, besides time correction. Thus, equation \@ref(eq:time-distance) becomes:
+In Jack's case, both +0.3 s fixed time correction and time correction estimation yield parameters closer to John's (i.e.
+true parameters).
+We have used these two model definitions in retrospective pilot study [@vescoviSprintMechanicalCharacteristics2021], demonstrating statistically significant differences in estimated FVP parameters.
+
+Instead of using time correction as simple intercept in the equation \@ref(eq:time-correction), we can estimate the *distance correction* (i.e., the flying start distance) using the following equation:
 
 \begin{equation}
-  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction (\#eq:distance-correction)
+  \begin{split}
+   t(d) &= (TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU) \\ 
+   &\quad-(TAU \times W(-e^{\frac{distance \; correction}{MSS \times TAU}} - 1) + \frac{distance \; correction}{MSS} + TAU) 
+   \end{split}
+   (\#eq:distance-correction)
 \end{equation}
 
-This model is implemented in `model_using_splits_with_corrections()` function. Below are the model estimates:
-
-```{r}
-jack_profile_distance_correction <- model_using_splits_with_corrections(
+Equation \@ref(eq:distance-correction) was used to generate the data for John and Jack (please refer to the provided R code), as well as for the simple simulation, and can be used as model definition to control for the flying start distance.
+
+To estimate distance correction parameter, we use `model_using_splits_with_distance_correction()` function. Here is how we can estimate Jack parameters using distance correction:
+
+```{r}
+jack_profile_distance_correction <- model_using_splits_with_distance_correction(
   distance = split_times$distance,
@@
 jack_parameters <- rbind(
-  coef(john_profile),
-  coef(jack_profile),
-  coef(jack_profile_fixed_time_short),
-  coef(jack_profile_fixed_time_long),
-  coef(jack_profile_time_estimated),
+  jack_parameters,
   coef(jack_profile_distance_correction)
@@
 
-rownames(jack_parameters) <- c(
-  "John",
-  "Jack - No corrections",
-  "Jack - Fixed time correction (+0.3s)",
-  "Jack - Fixed time correction (+0.5s)",
-  "Jack - Estimated time correction",
-  "Jack - Estimated distance correction"
-)
+rownames(jack_parameters) [6] <- "Jack - Estimated distance correction"
 
@@
 
+Another model definition, which also represents a novel approach implemented in the **shorts** package, is to utilize both time and distance corrections.
+Thus, equation \@ref(eq:time-correction) becomes:
+
+\begin{equation}
+  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction (\#eq:time-distance-correction)
+\end{equation}
+
+This model is implemented in `model_using_splits_with_corrections()` function. Below are the model estimates:
+
+```{r}
+jack_profile_time_distance_correction <- model_using_splits_with_corrections(
+  distance = split_times$distance,
+  time = split_times$jack_time
+)
+
+jack_parameters <- rbind(
+  jack_parameters,
+  coef(jack_profile_time_distance_correction)
+)
+
+rownames(jack_parameters) [7] <- 
+  "Jack - Estimated time & distance corrections"
+
+kable(jack_parameters, digits = 2, booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"))
+```
+
 As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters.
 There are a few issues with this model definition.
-Besides being novel and still not validated with actual data, distance correction model has four parameters to estimate, which implies that at least five sprint splits are needed.
-This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.
+Besides being novel and still not validated with actual data, distance correction model has four parameters to estimate, which implies that at least four sprint splits are needed.
+This imposes practical limitations, since acquiring five timing gates (one for the start and four for splits) might be practically troublesome.
 One strategy that is sometimes implemented is adding zeros to the sample (i.e., $t=0$ and $d=0$), which increase the number of observations.
-Unfortunately, this strategy should not be implemented, as explained later in the [Should we add zero to the sample?] section of this paper.
-
-We will get back to these issues later, but we can examine how these models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the Supplemental files for the R code).
+Unfortunately, this strategy should not be implemented since it will bias the estimated parameters:
+
+```{r}
+jack_profile_zero <- model_using_splits(
+  distance = c(0, split_times$distance),
+  time = c(0, split_times$jack_time)
+)
+
+jack_profile_time_estimated_zero <- model_using_splits_with_time_correction(
+  distance = c(0, split_times$distance),
+  time = c(0, split_times$jack_time)
+)
+
+adding_zero_issue <- rbind(
+  coef(jack_profile),
+  coef(jack_profile_zero),
+  coef(jack_profile_time_estimated),
+  coef(jack_profile_time_estimated_zero)
+)
+
+rownames(adding_zero_issue) <- c(
+  "Jack - No corrections",
+  "Jack - No corrections with zeros added",
+  "Jack - Estimated time correction",
+  "Jack profile - Estimated time correction with added zeros")
+
+kable(adding_zero_issue, digits = 2, booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"))
+```
+
+Adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
+
+It is important to note that using `predict_XXX()` family of functions with distance correction model and using estimated distance correction parameter will yield wrong predictions on the original scale (but not on the *theoretical* scale in which force production starts at d=0m and t=0s).
+This is because `predict_XXX()` family of functions utilize equation \@ref(eq:time-distance-correction) and its derivates and integrals, which works for the uncorrected model (time and distance correction parameters are equal to zero), time correction models (distance correction parameter is equal to zero), and time and distance corrections model.
+You can notice that equations \@ref(eq:time-correction) and \@ref(eq:time-distance-correction) are equivalent to the \@ref(eq:time-distance) equation when time and distance correction parameters are equal to zero.
+Unfortunately, this doesn't work for the distance correction model when estimated distance correction parameter is used since model definition outlined in the equation \@ref(eq:distance-correction) is different that the equation \@ref(eq:time-distance-correction) when estimated distance correction parameter is different than zero.
+The solution is to use generic S3 `predict()` function, but this only predicts split times at split distances.
+
+```{r}
+# Jack's observed times
+split_times$jack_time
+
+# To predict at "theoretical" distance scale (where force application starts at d=0m)
+predict_time_at_distance(
+  distance = split_times$distance, 
+  MSS = jack_profile_time_estimated$parameters$MSS,
+  TAU = jack_profile_time_estimated$parameters$TAU)
+
+predict_time_at_distance(
+  distance = split_times$distance, 
+  MSS = jack_profile_distance_correction$parameters$MSS,
+  TAU = jack_profile_distance_correction$parameters$TAU)
+
+# To predict at original distance scale, use all estimated parameters
+predict_time_at_distance(
+  distance = split_times$distance, 
+  MSS = jack_profile_time_estimated$parameters$MSS,
+  TAU = jack_profile_time_estimated$parameters$TAU,
+  time_correction = jack_profile_time_estimated$parameters$time_correction)
+
+# Unfortunately, this doesn't work for the distance correction model
+predict_time_at_distance(
+  distance = split_times$distance, 
+  MSS = jack_profile_distance_correction$parameters$MSS,
+  TAU = jack_profile_distance_correction$parameters$TAU,
+  distance_correction = jack_profile_distance_correction$parameters$distance_correction)
+
+# Instead use generic S3 predict() function
+predict(jack_profile_distance_correction)
+
+predict(
+  jack_profile_distance_correction$model,
+  newdata = data.frame(distance = split_times$distance))
+```
+
+In the next section we examine how these models (no correction, fixed time correction, estimated time correction, estimated distance correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the [Supplemental material] for the R code).
 
@@
 pred_wrapper <- function(data) {
+
+  merge_model <- function(data, model, model_name) {
+    data$model <- model_name
+    data$est_MSS <- model$parameters$MSS
+    data$est_TAU <- model$parameters$TAU
+    data$est_MAC <- model$parameters$MAC
+    data$est_PMAX <- model$parameters$PMAX
+    data$est_time_correction <- model$parameters$time_correction
+    data$est_distance_correction <- model$parameters$distance_correction
+    data$predicted_time <- model$data$pred_time
+    data$residuals <- data$time - data$predicted_time
+
+    data
+  }
+
   no_correction <- model_using_splits(
@@
     time = data$time,
-    control = nls.control(tol = 1)
+    control = nls.control(warnOnly = TRUE)
+  )
+
+  distance_correction <- model_using_splits_with_distance_correction(
+    distance = data$distance,
+    time = data$time,
+    control = nls.control(warnOnly = TRUE)
   )
@@
     time = data$time,
-    control = nls.control(tol = 1)
-  )
-
-  params <- rbind(
-    data.frame(
-      model = "No correction",
-      t(coef(no_correction))
-    ),
-    data.frame(
-      model = "Fixed correction +0.3s",
-      t(coef(fixed_correction_short))
-    ),
-    data.frame(
-      model = "Fixed correction +0.5s",
-      t(coef(fixed_correction_long))
-    ),
-    data.frame(
-      model = "Time correction",
-      t(coef(time_correction))
-    ),
-    data.frame(
-      model = "Time and distance correction",
-      t(coef(time_dist_correction))
-    )
-  )
-
-  colnames(params) <- c(
-    "model", "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
-    "est_time_correction", "est_distance_correction"
-  )
-
-  df <- expand_grid(
+    control = nls.control(warnOnly = TRUE)
+  )
+
+  rbind(
+    merge_model(
       data, 
-    params
-  )
-
-  df$predicted_time <- predict_time_at_distance(
-    distance = df$distance,
-    MSS = df$est_MSS,
-    TAU = df$est_TAU,
-    time_correction = df$est_time_correction,
-    distance_correction = df$est_distance_correction
-  )
-
-  df$residuals <- df$predicted_time - df$time
-  return(df)
+      no_correction,
+      "No correction"
+    ),
+    merge_model(
+      data, 
+      fixed_correction_short,
+      "Fixed correction +0.3s"
+    ),
+    merge_model(
+      data, 
+      fixed_correction_long,
+      "Fixed correction +0.5s"
+    ),
+    merge_model(
+      data, 
+      time_correction,
+      "Time correction"
+    ),
+    merge_model(
+      data, 
+      distance_correction,
+      "Distance correction"
+    ),
+    merge_model(
+      data, 
+      time_dist_correction,
+      "Time and distance correction"
+    )
+  )
 }
@@
 model_df <- sim_df %>%
+  filter(flying_start_distance > 0) %>%
   group_by(MSS, TAU, flying_start_distance) %>%
@@
 
-As can be seen from the next figure, the estimated time correction model estimates MSS almost perfectly, while the estimated time and distance correction model estimates MSS perfectly.
+As can be seen from the next figure, the estimated time correction model estimates MSS almost perfectly, while the estimated distance and estimate time and distance correction models estimates MSS perfectly.
 
@@
     "Time correction",
+    "Distance correction",
     "Time and distance correction"
@@
 )
+
 # Estimates plot
-df <- model_df %>%
-  group_by(MSS, TAU, flying_start_distance, model) %>%
+plot_df <- model_df %>%
+  group_by(model, MSS, TAU, flying_start_distance) %>%
   slice(1) %>%
   mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2))
-  )
-
-# MSS
-ggplot(
-  df,
-  aes(x = flying_start_distance, y = est_MSS, color = model)
-) +
+    `MSS (m/s)` = est_MSS,
+    `TAU` = est_TAU,
+    `MAC (m/s/s)` = est_MAC,
+    `PMAX (W)` = est_PMAX
+  ) %>%
+  pivot_longer(cols = c("MSS (m/s)", "TAU", "MAC (m/s/s)", "PMAX (W)")) %>%
+  mutate(
+    name = factor(name, levels = c("MSS (m/s)", "TAU", "MAC (m/s/s)", "PMAX (W)"))
+  )
+
+ggplot(plot_df, aes(x = flying_start_distance, y = value, group = model, color = model)) +
   theme_bw(8) +
   geom_line(alpha = 0.7) +
-  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
+  facet_wrap(~name, scales = "free_y") + 
   xlab("Flying start distance (m)") +
-  ylab("estimated MSS (m/s)") +
+  ylab("Estimated parameter") +
   theme(
@@
 
-Similar outcomes are observed for the MAC parameter. The time and distance corrections model performs perfectly, while the time correction model performs almost as good.
-
-```{r echo=FALSE}
-# MAC
-ggplot(
-  df,
-  aes(x = flying_start_distance, y = est_MAC, color = model)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Flying start distance (m)") +
-  ylab("estimated MAC (m/s/s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
+Similar outcomes are observed for the MAC parameter.
+The distance correction model and time and distance corrections models performs perfectly, while the time correction model performs almost as good.
 PMAX demonstrates the same properties as MSS and MAC.
 
-```{r echo=FALSE}
-# PMAX
-ggplot(
-  df,
-  aes(x = flying_start_distance, y = est_PMAX, color = model)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Flying start distance (m)") +
-  ylab("estimated PMAX (W/kg)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the *stolen time*; indicated by the dashed line on the figure).
-
-```{r echo=FALSE}
-# time_correction
-ggplot(
-  df,
-  aes(
-    x = flying_start_distance,
-    y = est_time_correction,
-    color = model
-  )
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  geom_line(
-    aes(y = stolen_time),
-    color = "black", linetype = "dashed"
-  ) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Flying start distance (m)") +
-  ylab("estimated time correction (s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The next figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents *identity line* since flying start distance is already on the x-axis).
-
-```{r echo=FALSE}
-# distance_correction
-ggplot(
-  df,
-  aes(
-    x = flying_start_distance,
-    y = est_distance_correction,
-    color = model
-  )
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  geom_abline(slope = 1, color = "black", linetype = "dashed") +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Flying start distance (m)") +
-  ylab("estimated distance correction (m)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models:
-
-```{r echo=FALSE}
+The next figure depicts residuals across split distances for each simulated flying start distance.
+As can be seen, time correction, distance correction, and time and distance correction models perform much better than the other models.
+
+```{r}
 # Residuals
-model_df <- model_df %>%
-  mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-    group = paste(MSS, MAC, flying_start_distance)
-  )
-
 ggplot(
@@
     color = flying_start_distance,
-    group = group
+    group = flying_start_distance
   )
@@
   geom_line(alpha = 0.3) +
-  facet_wrap(~model) +
   geom_hline(yintercept = 0, linetype = "dashed") +
   scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
+  facet_wrap(~model) +
   xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)") +
+  ylab("Observed - predicted time (s)") +
   theme(legend.position = "top") +
@@
 
-The outcomes from the simulation data clearly demonstrates that the time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model when attempting to overcome the flying start issues.
-Since the time correction model is simpler and requires three parameters to be estimated, it might be practically more useful than the time and distance correction model, which requires four parameters estimation and thus more than five timing gates and sprint splits.
-
-Time correction and time and distance corrections are also implemented in the mixed-models using `mixed_model_using_splits_with_time_correction()` and `mixed_model_using_splits_with_corrections()`.
+The outcomes from this simple simulation data clearly demonstrates that the time correction, distance correction, and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model when attempting to overcome the flying start issues.
+Since the time correction and distance correction models are simpler and requires three parameters to be estimated, they might be practically more useful than the time and distance correction model, which requires four parameters estimation and thus more than four timing gates and sprint splits.
+More elaborate simulation study is currently ongoing and the results will be reported in another paper.
+
+Time correction, distance correction, and time and distance corrections are also implemented in the mixed-models using `mixed_model_using_splits_with_time_correction()`, `mixed_model_using_splits_with_distance_correction()`, and `mixed_model_using_splits_with_corrections()` functions.
 We will showcase their use at the end of this paper.
-
-### Simulation of additional starting issues
-
-Starting behind the initial timing gate represent only one issue (i.e., flying start).
-In this section, we simulate one more issue to check the sensitivity of the presented models to other (less common) perturbations when performing field testing.
-
-One issue that might happen with timing gates is triggering the timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
-This is very similar to the situation when timing starts on a signal (i.e., gun during 100 m sprint race) and there is *reaction time* (RT) involved.
-Both of these scenarios represent *time lag* that is added to the split times.
-We have simulated the effect of this time lag on model estimates and predictions (please refer to the Supplemental files for the R code).
-
-```{r include=FALSE}
-sim_df <- expand.grid(
-  MSS = c(6, 7, 8, 9),
-  MAC = c(6, 7, 8, 9),
-  time_lag = seq(0, 0.5, length.out = 50),
-  distance = c(5, 10, 20, 30, 40, 50)
-)
-
-sim_df <- sim_df %>%
-  mutate(
-    TAU = MSS / MAC,
-    PMAX = MSS * MAC / 4,
-    true_time = predict_time_at_distance(distance, MSS, TAU),
-    time = true_time + time_lag
-  )
-
-# Add small noise to allow model fit
-set.seed(1667)
-rand_noise <- rnorm(nrow(sim_df), 0, 10^-4)
-sim_df$time <- sim_df$time + rand_noise
-
-# estimated parameters and predicted time
-model_df <- sim_df %>%
-  group_by(MSS, TAU, time_lag) %>%
-  do(pred_wrapper(.)) %>%
-  ungroup()
-```
-
-From the figure below it can be seen that time lag affects estimated MSS for the the model without correction and fixed correction model.
-Time correction and time and distance corrections models correctly estimated MSS.
-
-```{r echo=FALSE}
-model_df$model <- factor(
-  model_df$model,
-  levels = c(
-    "No correction",
-    "Fixed correction +0.3s",
-    "Fixed correction +0.5s",
-    "Time correction",
-    "Time and distance correction"
-  )
-)
-# Estimates plot
-df <- model_df %>%
-  group_by(MSS, TAU, time_lag, model) %>%
-  slice(1) %>%
-  mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2))
-  )
-
-# MSS
-ggplot(df, aes(x = time_lag, y = est_MSS, color = model)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MSS_string ~ MAC_string, scales = "free_y") +
-  xlab("Time lag (s)") +
-  ylab("estimated MSS (m/s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-From the next figure it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
-Time correction and time and distance corrections model correctly estimated MAC.
-
-```{r echo=FALSE}
-# MAC
-ggplot(df, aes(x = time_lag, y = est_MAC, color = model)) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Time lag (s)") +
-  ylab("estimated MAC (m/s/s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The figure below depicts correctly identified time lag (i.e. using time correction parameter) using time correction and time and distance corrections models.
-
-```{r echo=FALSE}
-# time_correction
-ggplot(
-  df,
-  aes(x = time_lag, y = est_time_correction, color = model)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  geom_abline(slope = -1, color = "black", linetype = "dashed") +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Time lag (s)") +
-  ylab("estimated time correction (s)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The next figure depicts estimated distance correction for the time and distance correction model.
-The estimated distance correction parameters looks jumpy due to random noise that we have to added to allow model fit, as well as the model estimation error.
-
-```{r echo=FALSE}
-# distance_correction
-ggplot(
-  df,
-  aes(x = time_lag, y = est_distance_correction, color = model)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  facet_grid(MAC_string ~ MSS_string, scales = "free_y") +
-  xlab("Time lag (s)") +
-  ylab("estimated distance correction (m)") +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The following figure depicts residuals (i.e., predicted time minus observed time).
-
-```{r echo=FALSE}
-# Residuals
-model_df <- model_df %>%
-  mutate(
-    MSS_string = paste("MSS =", MSS),
-    TAU_string = paste("TAU =", TAU),
-    MAC_string = paste("MAC = ", round(MAC, 2)),
-    PMAX_string = paste("PMAX = ", round(PMAX, 2)),
-    group = paste(MSS, MAC, time_lag)
-  )
-
-ggplot(
-  model_df,
-  aes(y = residuals, x = distance, color = time_lag, group = group)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.3) +
-  facet_wrap(~model) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  scale_color_gradientn(colours = terrain.colors(5, rev = FALSE)) +
-  xlab("Distance (m)") +
-  ylab("Predicted time - observed time (s)") +
-  theme(legend.position = "top") +
-  labs(color = "Time lag")
-```
-
-There are few other starting issues worth mentioning.
-For example, if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
-In this case, time lag is a negative number since it reduces the split times.
-Another common issue with timing gates in the practical field settings is the bad measurement of the distance and thus bad positions of the timing gates.
-
-The number and distances of the timing gates can also affect the precision of the estimated sprint parameters [@haugenPowerForceVelocityProfilingSprinting2020; @haugenDifferenceStartImpact2012].
-
-In field testing, multiple starting issues can be present.
-For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence.
-More elaborate simulation is beyond the scope of the current paper.
-
-## Should we add zero to the sample?
-
-Fellow sports scientists are often considering adding $t=0$ s at $d=0$ m to the collected split times with the aim of increasing the number of observations.
-The question is whether this strategy is sound and if it should be employed.
-The short answer is no, it shouldn't, particularly if there are flying start issues.
-In the following example, we are demonstrating the issue when zeros are added to the sample when fitting multiple models with the known true parameters:
-
-```{r}
-# Create split times from known MSS and TAU
-df <- tibble(
-  distance = c(5, 10, 20, 30, 40),
-  time = predict_time_at_distance(
-    distance,
-    MSS = 9,
-    TAU = 1.3
-  )
-) %>%
-  mutate(
-    # Add random noise to time
-    time = time + rnorm(n(), 0, 10^-5),
-    gate_time = time - 0.1
-  )
-
-# Model without time correction
-m_no_correction <- model_using_splits(
-  distance = df$distance,
-  time = df$gate_time
-)
-
-# Model without time correction, but with zeros added
-m_no_correction_zero <- model_using_splits(
-  distance = c(0, df$distance),
-  time = c(0, df$gate_time)
-)
-
-# Model without adding zeros for the start
-m_no_zero <- model_using_splits_with_time_correction(
-  distance = df$distance,
-  time = df$gate_time
-)
-
-# Model with added zeros for the start (d=0 and t=0)
-m_with_zero <- model_using_splits_with_time_correction(
-  distance = c(0, df$distance),
-  time = c(0, df$gate_time)
-)
-
-# Print results
-res_df <- data.frame(
-  model = c(
-    "Without time correction",
-    "Without time correction with zeros added",
-    "With time correction",
-    "With time correction with zeros added"
-  ),
-  rbind(
-    coef(m_no_correction),
-    coef(m_no_correction_zero),
-    coef(m_no_zero),
-    coef(m_with_zero)
-  )
-)
-
-kable(res_df, digits = 2, booktabs = TRUE) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"))
-```
-
-As can be seen from the output, only the model with time correction is able to correctly recover true sprint parameters, but adding zeros to the sample in this case, results in MSS and TAU estimates very close to the estimates of the model without time corrections.
-Thus, adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
 
@@
 
-Next we plot estimated parameters across LOOCV folds (please refer to the Supplemental files for the R code):
+Next we plot estimated parameters across LOOCV folds (please refer to the [Supplemental material] for the R code):
 
@@
   xlab("Distance (m)") +
-  ylab("Predicted - observed time (s)") +
+  ylab("Observed - predicted time (s)") +
   theme(
@@
 As expected, the model has more issues predicting unseen split times for both short or long distances.
-Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
+Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least four observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
 LOOCV can also be implemented with the mixed-effects models in the **shorts** package.
@@
 Here is the model estimate of the time and distance it takes for Bolt to reach 99% of MSS.
-Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start, not the measurement scale.
+Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start (i.e., theoretical scale), not the measurement scale.
 
@@
 
-The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players).
+The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players) [@vescoviSprintMechanicalCharacteristics2021].
 Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset.
@@
 
-The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model (please refer to the Supplemental files for the R code).
+The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) time correction model with time correction as fixed effect, (4) time correction model with time correction as random effect, (5) distance correction model with distance correction as fixed effect, (6) distance correction model with distance correction as random effect, (7) time and distance correction model, (8) time and distance correction model with distance correction as fixed effect (and time correction as random effect), and (9) time and distance correction model with distance and time correction as random effects (please refer to the [Supplemental material] for the R code).
 
@@
 
+distance_correction_fixed <-
+  mixed_model_using_splits_with_distance_correction(
+    df,
+    distance = "distance",
+    time = "time",
+    athlete = "Athlete",
+    random = MSS + TAU ~ 1
+  )
+
+distance_correction_random <-
+  mixed_model_using_splits_with_distance_correction(
+    df,
+    distance = "distance",
+    time = "time",
+    athlete = "Athlete",
+    random = MSS + TAU + distance_correction ~ 1, 
+    control = nlme::nlmeControl(tolerance = 0.01)
+  )
+
 time_distance_correction_fixed <-
+  mixed_model_using_splits_with_corrections(
+    df,
+    distance = "distance",
+    time = "time",
+    athlete = "Athlete",
+    random = MSS + TAU ~ 1
+  )
+
+time_distance_correction_time_random <-
   mixed_model_using_splits_with_corrections(
@@
   data.frame(
+    model = "Distance correction fixed",
+    t(unlist(distance_correction_fixed$model_fit))
+  ),
+  data.frame(
+    model = "Distance correction random",
+    t(unlist(distance_correction_random$model_fit))
+  ),
+  data.frame(
     model = "Time and distance correction fixed",
     t(unlist(time_distance_correction_fixed$model_fit))
+  ),
+  data.frame(
+    model = "Time and distance correction (time correction random)",
+    t(unlist(time_distance_correction_time_random$model_fit))
   ),
@@
     "Time correction random",
+    "Distance correction fixed",
+    "Distance correction random",
     "Time and distance correction fixed",
+    "Time and distance correction (time correction random)",
     "Time and distance correction random"
@@
   data.frame(
+    model = "Distance correction fixed",
+    distance_correction_fixed$parameters$random
+  ),
+  data.frame(
+    model = "Distance correction random",
+    distance_correction_random$parameters$random
+  ),
+  data.frame(
     model = "Time and distance correction fixed",
     time_distance_correction_fixed$parameters$random
+  ),
+  data.frame(
+    model = "Time and distance correction (time correction random)",
+    time_distance_correction_time_random$parameters$random
   ),
@@
   data.frame(
+    model = "Time and distance correction (time correction random)",
+    time_distance_correction_time_random$data
+  ),
+  data.frame(
     model = "Time and distance correction random",
@@
     "Time and distance correction fixed",
+    "Time and distance correction (time correction random)",
     "Time and distance correction random"
@@
 
-model_resid$resid <- model_resid$pred_time - model_resid$time
+model_resid$resid <- model_resid$time - model_resid$pred_time
 
@@
   facet_wrap(~distance, scales = "free_x") +
-  xlab("Predicted time - observed time (s)") +
+  xlab("Observed time - predicted time (s)") +
   ylab(NULL)
@@
 
-Which model should should be used?
+Which model should be used?
 Although providing a better fit (using RSE as an estimator of model fit), the time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction).
 Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for timing inconsistencies.
+More thorough theoretical simulation study is currently in development with the aim to explore the behavior of these models under different scenarios.
 
@@
 
+# Supplemental material
+
+The *R Markdown* [@R-bookdown; @R-rmarkdown; @rmarkdown2018; @rmarkdown2020] source code for the paper can be found on the GitHub repository: [https://github.com/mladenjovanovic/shorts-paper](https://github.com/mladenjovanovic/shorts-paper).
+
 # References

### 2021-07-21 paper.tex
Added section to want about predict_XXX issues with distance correction model when distance correction <>0m

@@
 
-\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
+\documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions
 
@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 
-The \textbf{shorts} package (\protect\hyperlink{ref-R-shorts}{Jovanovic 2020}), written in the R language (\protect\hyperlink{ref-R-base}{R Core Team 2020}), represents an open-source tool to help sports scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
+The \textbf{shorts} package (\protect\hyperlink{ref-R-shorts}{Jovanović 2021}), written in the R language (\protect\hyperlink{ref-R-base}{R Core Team 2020}), represents an open-source tool to help sports scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling (\protect\hyperlink{ref-morinSimpleMethodComputing2019}{Jean-Benoit Morin et al.
 2019}; \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.
@@
 Let's consider four athletes with different levels of MSS (high versus low maximal sprinting speed) and MAC (high versus low maximal acceleration; as mentioned previously, using MAC is preferred over using TAU) (Table \ref{tab:four-athletes-table}).
+
+\small
 
@@
 
+\normalsize
+
 Figure \ref{fig:four-athletes-kinematics} depicts distance, velocity, and acceleration over time (from 0 to 6 s).
+
+\small
 
@@
 
+\normalsize
+
 Plotting acceleration against velocity (Figure \ref{fig:four-athletes-profile}), we will get \emph{Acceleration-Velocity Profile}, which is linear, according to the mathematical model.
@@
 
+\small
+
 \begin{figure}
@@
 
+\normalsize
+
 \hypertarget{estimation-using-shorts-package}{%
@@
 Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \textbf{shorts} package using non-linear least squares regression implemented in the \texttt{nls()} function in the \textbf{base R} (\protect\hyperlink{ref-R-base}{R Core Team 2020}) and \texttt{nlme()} function in the \textbf{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2020}) for the mixed-effect models.
-
-\hypertarget{estimating-short-sprint-parameters-using-split-times}{%
-\subsection{Estimating short sprint parameters using split times}\label{estimating-short-sprint-parameters-using-split-times}}
+Estimation of MSS and TAU parameters from equation \eqref{eq:velocity-time} is performed in \textbf{shorts} package using non-linear least squares regression implemented in the \texttt{nls()} function (\protect\hyperlink{ref-batesNonlinearModels1992}{D.
+M.
+Bates and Chambers 1992}; \protect\hyperlink{ref-batesNonlinearRegressionAnalysis2007}{Douglas M.
+Bates and Watts 2007}) in the \textbf{base R} (\protect\hyperlink{ref-R-base}{R Core Team 2020}) and \texttt{nlme()} function in the \textbf{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2021}) for the mixed-effect models.
+
+\hypertarget{estimating-short-sprint-parameters-using-timing-gates-split-times}{%
+\subsection{Estimating short sprint parameters using timing gates split times}\label{estimating-short-sprint-parameters-using-timing-gates-split-times}}
 
 Let's consider an example of an athlete with MSS equal to 9 \(ms^{-1}\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^{-2}\) performing 40m sprint with timing gates positioned at each 10m split.
-For split times, distance is a predictor, and time is the outcome variable, thus the equation \eqref{eq:velocity-time} becomes:
+For split times, distance is a predictor, and time is the outcome variable, thus the equation \eqref{eq:distance-time} becomes:
 
@@
 \(W\) in equation \eqref{eq:time-distance} represents Lambert's W function (\protect\hyperlink{ref-R-LambertW}{Goerg 2020}).
+Researchers often incorrectly use the equation \eqref{eq:distance-time} (\protect\hyperlink{ref-morinSpreadsheetSprintAcceleration2017}{J.
+B.
+Morin 2017}; \protect\hyperlink{ref-morinSpreadsheetSprintAcceleration2019}{Jean-Benoit Morin and Samozino 2019}; \protect\hyperlink{ref-stenrothSpreadsheetSprintAcceleration2020}{Stenroth and Vartiainen 2020}), in which the time is the predictor and distance is the outcome variable, instead of statistically correct equation \eqref{eq:time-distance} (\protect\hyperlink{ref-motulskyIntuitiveBiostatisticsNonmathematical2018}{Motulsky 2018, 341}).
+
 MSS and TAU parameters are estimated using \texttt{model\_using\_splits()} function:
+
+\small
 
@@
 \CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}   0.00249   1.00000  {-}0.00178   0.00265   0.00265   0.00176 }
+\CommentTok{\#\textgreater{}   0.00249   1.00000  {-}0.00265   0.00178   0.00265   0.00176 }
 \CommentTok{\#\textgreater{}       MAE      MAPE }
@@
 \end{Shaded}
+
+\normalsize
 
@@
 
-To create a simple plot of the model, use S3 \texttt{plot()} method, which returns \textbf{ggplot2} (\protect\hyperlink{ref-R-ggplot2}{Wickham, Chang, et al. 2020}) object:
+To create a simple plot of the model, use S3 \texttt{plot()} method, which returns \textbf{ggplot2} (\protect\hyperlink{ref-R-ggplot2}{Wickham, Chang, et al. 2021}) object:
+
+\small
 
@@
 
+\normalsize
+
 Once we have estimated MSS and TAU, we can use \texttt{predict\_XXX()} family of functions to predict various relationships (i.e., time at distance, acceleration at distance, velocity at time, etc.):
+
+\small
 
@@
 
+\normalsize
+
 \hypertarget{air-resistance-and-the-calculation-of-force-and-mechanical-power}{%
@@
 Air resistance (N) is estimated using \texttt{get\_air\_resistance()} function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature (\(C^\circ\)), and wind velocity (\(ms^-1\)) as parameters (please refer to \protect\hyperlink{ref-arsacModelingEnergetics100m2002}{Arsac and Locatelli} (\protect\hyperlink{ref-arsacModelingEnergetics100m2002}{2002}), \protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{Samozino et al.} (\protect\hyperlink{ref-samozinoSimpleMethodMeasuring2016}{2016}), and \protect\hyperlink{ref-vaningenschenauCanCyclePower1991}{van Ingen Schenau, Jacobs, and de Koning} (\protect\hyperlink{ref-vaningenschenauCanCyclePower1991}{1991}) for more information):
+
+\small
 
@@
 
+\normalsize
+
 When estimating force and power, the air resistance parameters can be set using \texttt{"..."}, which are forwarded to the \texttt{get\_air\_resistance()}:
+
+\small
 
@@
 
+\normalsize
+
 The easiest way to get all kinematics and kinetics for short sprints is to use \texttt{predict\_kinematics()} function:
+
+\small
 
@@
 
+\normalsize
+
 Plotting the model predictions can be done once we convert data from wide to long with the help of \textbf{ggplot2} (\protect\hyperlink{ref-R-ggplot2}{Wickham, Chang, et al.
-2020}), \textbf{dplyr} (\protect\hyperlink{ref-R-dplyr}{Wickham, François, et al.
-2020}), \textbf{tidyr} (\protect\hyperlink{ref-R-tidyr}{Wickham 2020}), and \textbf{tidyverse} (\protect\hyperlink{ref-R-tidyverse}{Wickham 2019}) packages:
+2021}), \textbf{dplyr} (\protect\hyperlink{ref-R-dplyr}{Wickham, François, et al.
+2021}), \textbf{tidyr} (\protect\hyperlink{ref-R-tidyr}{Wickham 2021a}), and \textbf{tidyverse} (\protect\hyperlink{ref-R-tidyverse}{Wickham 2021b}) packages:
+
+\small
 
@@
 
+\normalsize
+
 These kinematic and kinetic variables are utilized in \protect\hyperlink{force-velocity-profile}{Force-Velocity profile} estimation, which is covered later in this paper.
@@
 To identify these values, \textbf{shorts} package comes with \texttt{find\_XXX()} family of functions:
+
+\small
 
@@
 
+\normalsize
+
 \hypertarget{mixed-effects-model}{%
@@
 To demonstrate this functionality, we load the \texttt{split\_times} dataset provided in the \textbf{shorts} package:
+
+\small
 
@@
 \CommentTok{\#\textgreater{}       RSE R\_squared    minErr    maxErr maxAbsErr      RMSE }
-\CommentTok{\#\textgreater{}    0.0260    0.9998   {-}0.0293    0.0496    0.0496    0.0214 }
+\CommentTok{\#\textgreater{}    0.0260    0.9998   {-}0.0496    0.0293    0.0496    0.0214 }
 \CommentTok{\#\textgreater{}       MAE      MAPE }
@@
 
-Additional information about mixed-effects model performed using the \texttt{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2020}) can be obtained using \texttt{summary()} function.
+\normalsize
+
+Additional information about mixed-effects model performed using the \texttt{nlme} package (\protect\hyperlink{ref-R-nlme}{Pinheiro, Bates, and R-core 2021}) can be obtained using \texttt{summary()} function.
 S3 method \texttt{coef()} when applied on mixed-model result will return both the fixed and random effects.
+
 To create a simple plot of the model, use S3 \texttt{plot()} method.
 
+\small
+
 \begin{Shaded}
@@
 
+\normalsize
+
 The following figure contains kinematics for all athletes in \texttt{split\_times} dataset. Please note that power calculation takes default parameters for each individual:
+
+\small
 
@@
 
+\normalsize
+
 \hypertarget{estimating-short-sprint-parameters-using-radar-gun}{%
@@
 Let's consider the same example of an athlete with MSS equal to 9 \(ms^{-1}\), TAU equal to 1.3, and MAC equal to 6.92 \(ms^{-2}\) performing 40m sprint with velocity estimated using radar run (in this case with 1 Hz sampling rate).
+
+\small
 
@@
 
+\normalsize
+
 Both split and radar gun models allow the use of \emph{weighted} non-linear regression.
@@
 Weighted non-linear regression is performed by setting \texttt{weights} parameter:
+
+\small
 
@@
 
+\normalsize
+
 \hypertarget{mixed-effects-model-1}{%
@@
 Mixed-effects model using radar data is done using \texttt{mixed\_model\_using\_radar()} function. To perform mixed model, let's load data that comes with \textbf{shorts} package.
+
+\small
 
@@
 2016}).
+
+\small
 
@@
 
+\normalsize
+
 To plot FVP kinematics and kinetics (which are exactly the same as generated by the \texttt{predict\_kinematics()} function), use S3 \texttt{plot()} function.
@@
 
+\small
+
 \begin{Shaded}
@@
 
+\normalsize
+
 To plot FVP estimated kinetics against time, use \texttt{type\ =\ "time"} parameter:
 
+\small
+
 \begin{Shaded}
@@
 Let's use our athlete and add and deduct 0.5 s to simulate an error in synchronization and its effect on estimated MSS and TAU.
+
+\small
 
@@
 
+\normalsize
+
 The following three models estimate MSS and TAU from the three datasets.
-Tables are created using the \texttt{kable()} function from the \textbf{knitr} package (\protect\hyperlink{ref-knitr2014}{Xie 2014}, \protect\hyperlink{ref-knitr2015}{2015}, \protect\hyperlink{ref-R-knitr}{2020}) and \texttt{kable\_styling()} function from the \textbf{kableExtra} package (\protect\hyperlink{ref-R-kableExtra}{Zhu 2021}).
+Tables are created using the \texttt{kable()} function from the \textbf{knitr} package (\protect\hyperlink{ref-knitr2014}{Xie 2014}, \protect\hyperlink{ref-knitr2015}{2015}, \protect\hyperlink{ref-R-knitr}{2021b}) and \texttt{kable\_styling()} function from the \textbf{kableExtra} package (\protect\hyperlink{ref-R-kableExtra}{Zhu 2021}).
+
+\small
 
@@
 
+\normalsize
+
 As can be seen from the example, all estimated parameters are affected by an error in synchronization of time with velocity (with MSS being the least affected in this example).
@@
 This model is incorporated in the \texttt{model\_using\_radar\_with\_time\_correction()} function:
+
+\small
 
@@
 
+\normalsize
+
 When using \texttt{predict\_XXX()} family of functions, one can provide estimated time correction to get predictions at original time scale.
+
+\small
 
@@
 
+\normalsize
+
 \hypertarget{mixed-model-approach}{%
@@
 When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the \texttt{mixed\_model\_using\_radar\_with\_time\_correction()} function.
+
+\small
 
@@
 
-\hypertarget{problems-at-the-start-when-using-split-times}{%
-\subsection{Problems at the start when using split times}\label{problems-at-the-start-when-using-split-times}}
+\normalsize
+
+\hypertarget{problems-at-the-start-when-using-timing-gates}{%
+\subsection{Problems at the start when using timing gates}\label{problems-at-the-start-when-using-timing-gates}}
 
@@
 In other words, Jack is using a \emph{flying start}, a common scenario when testing field sports athletes.
+Flying start distance is often recommended to avoid premature triggering of the timing system by lifted knees or swinging
+arms (\protect\hyperlink{ref-altmannAccuracySingleBeam2018}{Altmann et al.
+2018}, \protect\hyperlink{ref-altmannDifferentStartingDistances2015}{2015}, \protect\hyperlink{ref-altmannValiditySingleBeamTiming2017}{2017}; \protect\hyperlink{ref-haugenPowerForceVelocityProfilingSprinting2020}{Thomas A.
+Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-haugenSprintRunningPerformance2016}{T.
+Haugen and Buchheit 2016}).
+
+\small
 
@@
 
-Here is a graphical representation of the sprint splits (please refer to the Supplemental files for the R code):
+\normalsize
+
+Here is a graphical representation of the sprint splits (please refer to the \protect\hyperlink{supplemental-material}{Supplemental material} for the R code):
+
+\small
 
@@
 
+\normalsize
+
 Using the following code, we can see the differences in estimated MSS and TAU parameters:
+
+\small
 
@@
 
+\normalsize
+
 As can be seen from the results, a flying start yields biased estimates, particularly for the TAU, MAC and PMAX.
 
-To explore this further, we have run a simple simulation with 5, 10, 20, 30, 40, and 50 m splits, with MSS and MAC varying from 6 to 9 (\(ms^{-1}\) and \(ms^{-2}\) respectively), and flying start distance varying from 0 to 1 m (please refer to the Supplemental files for the R code).
-With a simulation dataset, we can check the model estimates and predictions, given the flying start distance.
-
-The following figure demonstrates the effect of flying start distance on estimated MSS:
+To explore this further, we have run a simple simulation by increasing Jack's flying start distance from 0 to 1 m and depicting the estimated MSS, TAU, MAC, and PMAX parameters (please refer to the \protect\hyperlink{supplemental-material}{Supplemental material} for the R code).
+
+\small
 
@@
 
-As can be seen from the figure, MSS is underestimated as flying start distance increases. The following image demonstrates the effect of flying start distance on estimated MAC:
+\normalsize
+
+As can be seen from the figure, MSS is underestimated as flying start distance increases.
+MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
+Estimated PMAX is also overestimated as flying start distance increases.
+
+Model residuals are also affected by the flying start distance.
+The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40 m versus 5, 15, 30 m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simple simulation:
+
+\small
 
@@
 
-MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
-
-And finally, the following image demonstrates the effect of flying start distance on estimated PMAX:
+\normalsize
+
+Another way to visualize the effect of the flying start distance on split distance residuals can be found on the next figure:
+
+\small
 
@@
 
-Estimated PMAX is also overestimated as flying start distance increases.
-
-Model residuals are also affected by flying start distance.
-The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40 m versus 5, 15, 30 m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simulation:
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-32-1} \end{center}
-
-If we merge individual facets (i.e., combinations of MSS and MAC), we can get simpler figure conveying issues with residuals when there is a flying start:
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-33-1} \end{center}
+\normalsize
 
@@
 Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5 s) or estimated time correction:
+
+\small
 
@@
 
+\normalsize
+
 In Jack's case, both +0.3 s fixed time correction and time correction estimation yield parameters closer to John's (i.e.~true parameters).
-
-Another model definition, which is a novel approach implemented in the \textbf{shorts} package, is to utilize \emph{distance correction}, besides time correction.
-Thus, equation \eqref{eq:time-distance} becomes:
+We have used these two model definitions in retrospective pilot study (\protect\hyperlink{ref-vescoviSprintMechanicalCharacteristics2021}{Vescovi and Jovanovi'c 2021}), demonstrating statistically significant differences in estimated FVP parameters.
+
+Instead of using time correction as simple intercept in the equation \eqref{eq:time-correction}, we can estimate the \emph{distance correction} (i.e., the flying start distance) using the following equation:
 
 \begin{equation}
-  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction \label{eq:distance-correction}
+  \begin{split}
+   t(d) &= (TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU) \\ 
+   &\quad-(TAU \times W(-e^{\frac{distance \; correction}{MSS \times TAU}} - 1) + \frac{distance \; correction}{MSS} + TAU) 
+   \end{split}
+   \label{eq:distance-correction}
 \end{equation}
 
-This model is implemented in \texttt{model\_using\_splits\_with\_corrections()} function. Below are the model estimates:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{jack\_profile\_distance\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
+Equation \eqref{eq:distance-correction} was used to generate the data for John and Jack (please refer to the provided R code), as well as for the simple simulation, and can be used as model definition to control for the flying start distance.
+
+To estimate distance correction parameter, we use \texttt{model\_using\_splits\_with\_distance\_correction()} function. Here is how we can estimate Jack parameters using distance correction:
+
+\small
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jack\_profile\_distance\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_distance\_correction}\NormalTok{(}
   \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
@@
 \NormalTok{jack\_parameters }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
-  \FunctionTok{coef}\NormalTok{(john\_profile),}
-  \FunctionTok{coef}\NormalTok{(jack\_profile),}
-  \FunctionTok{coef}\NormalTok{(jack\_profile\_fixed\_time\_short),}
-  \FunctionTok{coef}\NormalTok{(jack\_profile\_fixed\_time\_long),}
-  \FunctionTok{coef}\NormalTok{(jack\_profile\_time\_estimated),}
+\NormalTok{  jack\_parameters,}
   \FunctionTok{coef}\NormalTok{(jack\_profile\_distance\_correction)}
@@
 
-\FunctionTok{rownames}\NormalTok{(jack\_parameters) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-  \StringTok{"John"}\NormalTok{,}
-  \StringTok{"Jack {-} No corrections"}\NormalTok{,}
-  \StringTok{"Jack {-} Fixed time correction (+0.3s)"}\NormalTok{,}
-  \StringTok{"Jack {-} Fixed time correction (+0.5s)"}\NormalTok{,}
-  \StringTok{"Jack {-} Estimated time correction"}\NormalTok{,}
-  \StringTok{"Jack {-} Estimated distance correction"}
-\NormalTok{)}
+\FunctionTok{rownames}\NormalTok{(jack\_parameters) [}\DecValTok{6}\NormalTok{] }\OtherTok{\textless{}{-}} \StringTok{"Jack {-} Estimated distance correction"}
 
@@
 \addlinespace
-Jack - Estimated distance correction & 9.00 & 1.30 & 6.92 & 15.6 & 0.40 & 0.5\\
+Jack - Estimated distance correction & 9.00 & 1.30 & 6.92 & 15.6 & 0.00 & 0.5\\
 \bottomrule
@@
 
-As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters.
-There are a few issues with this model definition.
-Besides being novel and still not validated with actual data, distance correction model has four parameters to estimate, which implies that at least five sprint splits are needed.
-This imposes practical limitations, since acquiring six timing gate (one for the start and five for splits) might be practically troublesome.
-One strategy that is sometimes implemented is adding zeros to the sample (i.e., \(t=0\) and \(d=0\)), which increase the number of observations.
-Unfortunately, this strategy should not be implemented, as explained later in the \protect\hyperlink{should-we-add-zero-to-the-sample}{Should we add zero to the sample?} section of this paper.
-
-We will get back to these issues later, but we can examine how these models (no correction, fixed time correction, estimated time correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the Supplemental files for the R code).
-
-As can be seen from the next figure, the estimated time correction model estimates MSS almost perfectly, while the estimated time and distance correction model estimates MSS perfectly.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-37-1} \end{center}
-
-Similar outcomes are observed for the MAC parameter. The time and distance corrections model performs perfectly, while the time correction model performs almost as good.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
-
-PMAX demonstrates the same properties as MSS and MAC.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-39-1} \end{center}
-
-The following figure depicts estimated time correction, and as can be seen, only the time and distance correction model estimated the time correction correctly (i.e., the \emph{stolen time}; indicated by the dashed line on the figure).
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-40-1} \end{center}
-
-The next figure depicts estimated distance correction, and same as with the time correction, only the time and distance correction model estimated the distance correction correctly (i.e., flying start distance; indicated by the dashed line on the figure, which represents \emph{identity line} since flying start distance is already on the x-axis).
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
-
-The following figure depicts model residuals against the distance, and as can be seen, time correction and time and distance correction models performs much better than no correction and fixed correction models:
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
-
-The outcomes from the simulation data clearly demonstrates that the time correction and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model when attempting to overcome the flying start issues.
-Since the time correction model is simpler and requires three parameters to be estimated, it might be practically more useful than the time and distance correction model, which requires four parameters estimation and thus more than five timing gates and sprint splits.
-
-Time correction and time and distance corrections are also implemented in the mixed-models using \texttt{mixed\_model\_using\_splits\_with\_time\_correction()} and \texttt{mixed\_model\_using\_splits\_with\_corrections()}.
-We will showcase their use at the end of this paper.
-
-\hypertarget{simulation-of-additional-starting-issues}{%
-\subsubsection{Simulation of additional starting issues}\label{simulation-of-additional-starting-issues}}
-
-Starting behind the initial timing gate represent only one issue (i.e., flying start).
-In this section, we simulate one more issue to check the sensitivity of the presented models to other (less common) perturbations when performing field testing.
-
-One issue that might happen with timing gates is triggering the timing system before the sprint is initiated (e.g., by cutting the beam with an arm swing prematurely).
-This is very similar to the situation when timing starts on a signal (i.e., gun during 100 m sprint race) and there is \emph{reaction time} (RT) involved.
-Both of these scenarios represent \emph{time lag} that is added to the split times.
-We have simulated the effect of this time lag on model estimates and predictions (please refer to the Supplemental files for the R code).
-
-From the figure below it can be seen that time lag affects estimated MSS for the the model without correction and fixed correction model.
-Time correction and time and distance corrections models correctly estimated MSS.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-44-1} \end{center}
-
-From the next figure it can be seen that time lag affects estimated MAC for the the model without correction and fixed correction models.
-Time correction and time and distance corrections model correctly estimated MAC.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-45-1} \end{center}
-
-The figure below depicts correctly identified time lag (i.e.~using time correction parameter) using time correction and time and distance corrections models.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-46-1} \end{center}
-
-The next figure depicts estimated distance correction for the time and distance correction model.
-The estimated distance correction parameters looks jumpy due to random noise that we have to added to allow model fit, as well as the model estimation error.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
-
-The following figure depicts residuals (i.e., predicted time minus observed time).
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-48-1} \end{center}
-
-There are few other starting issues worth mentioning.
-For example, if the initial timing gate has a time delay (i.e., once triggered, there is a time delay before the timing starts).
-In this case, time lag is a negative number since it reduces the split times.
-Another common issue with timing gates in the practical field settings is the bad measurement of the distance and thus bad positions of the timing gates.
-
-The number and distances of the timing gates can also affect the precision of the estimated sprint parameters (\protect\hyperlink{ref-haugenPowerForceVelocityProfilingSprinting2020}{Thomas A.
-Haugen, Breitschädel, and Samozino 2020}; \protect\hyperlink{ref-haugenDifferenceStartImpact2012}{Thomas A.
-Haugen, Tønnessen, and Seiler 2012}).
-
-In field testing, multiple starting issues can be present.
-For example, one might have a bad position of the initial gate, athlete might be moved back but also manage to trigger the gate before the start commence.
-More elaborate simulation is beyond the scope of the current paper.
-
-\hypertarget{should-we-add-zero-to-the-sample}{%
-\subsection{Should we add zero to the sample?}\label{should-we-add-zero-to-the-sample}}
-
-Fellow sports scientists are often considering adding \(t=0\) s at \(d=0\) m to the collected split times with the aim of increasing the number of observations.
-The question is whether this strategy is sound and if it should be employed.
-The short answer is no, it shouldn't, particularly if there are flying start issues.
-In the following example, we are demonstrating the issue when zeros are added to the sample when fitting multiple models with the known true parameters:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Create split times from known MSS and TAU}
-\NormalTok{df }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{20}\NormalTok{, }\DecValTok{30}\NormalTok{, }\DecValTok{40}\NormalTok{),}
-  \AttributeTok{time =} \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
-\NormalTok{    distance,}
-    \AttributeTok{MSS =} \DecValTok{9}\NormalTok{,}
-    \AttributeTok{TAU =} \FloatTok{1.3}
-\NormalTok{  )}
-\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
-  \FunctionTok{mutate}\NormalTok{(}
-    \CommentTok{\# Add random noise to time}
-    \AttributeTok{time =}\NormalTok{ time }\SpecialCharTok{+} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{n}\NormalTok{(), }\DecValTok{0}\NormalTok{, }\DecValTok{10}\SpecialCharTok{\^{}{-}}\DecValTok{5}\NormalTok{),}
-    \AttributeTok{gate\_time =}\NormalTok{ time }\SpecialCharTok{{-}} \FloatTok{0.1}
-\NormalTok{  )}
-
-\CommentTok{\# Model without time correction}
-\NormalTok{m\_no\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance,}
-  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{gate\_time}
-\NormalTok{)}
-
-\CommentTok{\# Model without time correction, but with zeros added}
-\NormalTok{m\_no\_correction\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance),}
-  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{gate\_time)}
-\NormalTok{)}
-
-\CommentTok{\# Model without adding zeros for the start}
-\NormalTok{m\_no\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-  \AttributeTok{distance =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{distance,}
-  \AttributeTok{time =}\NormalTok{ df}\SpecialCharTok{$}\NormalTok{gate\_time}
-\NormalTok{)}
-
-\CommentTok{\# Model with added zeros for the start (d=0 and t=0)}
-\NormalTok{m\_with\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
-  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{distance),}
-  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, df}\SpecialCharTok{$}\NormalTok{gate\_time)}
-\NormalTok{)}
-
-\CommentTok{\# Print results}
-\NormalTok{res\_df }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
-  \AttributeTok{model =} \FunctionTok{c}\NormalTok{(}
-    \StringTok{"Without time correction"}\NormalTok{,}
-    \StringTok{"Without time correction with zeros added"}\NormalTok{,}
-    \StringTok{"With time correction"}\NormalTok{,}
-    \StringTok{"With time correction with zeros added"}
-\NormalTok{  ),}
-  \FunctionTok{rbind}\NormalTok{(}
-    \FunctionTok{coef}\NormalTok{(m\_no\_correction),}
-    \FunctionTok{coef}\NormalTok{(m\_no\_correction\_zero),}
-    \FunctionTok{coef}\NormalTok{(m\_no\_zero),}
-    \FunctionTok{coef}\NormalTok{(m\_with\_zero)}
-\NormalTok{  )}
-\NormalTok{)}
-
-\FunctionTok{kable}\NormalTok{(res\_df, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+\normalsize
+
+Another model definition, which also represents a novel approach implemented in the \textbf{shorts} package, is to utilize both time and distance corrections.
+Thus, equation \eqref{eq:time-correction} becomes:
+
+\begin{equation}
+  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction \label{eq:time-distance-correction}
+\end{equation}
+
+This model is implemented in \texttt{model\_using\_splits\_with\_corrections()} function. Below are the model estimates:
+
+\small
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jack\_profile\_time\_distance\_correction }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_corrections}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance,}
+  \AttributeTok{time =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{jack\_time}
+\NormalTok{)}
+
+\NormalTok{jack\_parameters }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+\NormalTok{  jack\_parameters,}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_time\_distance\_correction)}
+\NormalTok{)}
+
+\FunctionTok{rownames}\NormalTok{(jack\_parameters) [}\DecValTok{7}\NormalTok{] }\OtherTok{\textless{}{-}} 
+  \StringTok{"Jack {-} Estimated time \& distance corrections"}
+
+\FunctionTok{kable}\NormalTok{(jack\_parameters, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"scale\_down"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{))}
@@
 \toprule
-model & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+  & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
 \midrule
-Without time correction & 8.78 & 1.09 & 8.06 & 17.7 & 0.00 & 0\\
-Without time correction with zeros added & 8.78 & 1.09 & 8.06 & 17.7 & 0.00 & 0\\
-With time correction & 9.00 & 1.30 & 6.92 & 15.6 & 0.10 & 0\\
-With time correction with zeros added & 8.79 & 1.10 & 7.97 & 17.5 & 0.01 & 0\\
+John & 9.00 & 1.30 & 6.92 & 15.6 & 0.00 & 0.0\\
+Jack - No corrections & 8.49 & 0.70 & 12.06 & 25.6 & 0.00 & 0.0\\
+Jack - Fixed time correction (+0.3s) & 9.00 & 1.25 & 7.19 & 16.2 & 0.30 & 0.0\\
+Jack - Fixed time correction (+0.5s) & 9.62 & 1.77 & 5.43 & 13.1 & 0.50 & 0.0\\
+Jack - Estimated time correction & 8.96 & 1.22 & 7.37 & 16.5 & 0.28 & 0.0\\
+\addlinespace
+Jack - Estimated distance correction & 9.00 & 1.30 & 6.92 & 15.6 & 0.00 & 0.5\\
+Jack - Estimated time \& distance corrections & 9.00 & 1.30 & 6.92 & 15.6 & 0.40 & 0.5\\
 \bottomrule
@@
 
-As can be seen from the output, only the model with time correction is able to correctly recover true sprint parameters, but adding zeros to the sample in this case, results in MSS and TAU estimates very close to the estimates of the model without time corrections.
-Thus, adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
+\normalsize
+
+As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters.
+There are a few issues with this model definition.
+Besides being novel and still not validated with actual data, distance correction model has four parameters to estimate, which implies that at least four sprint splits are needed.
+This imposes practical limitations, since acquiring five timing gates (one for the start and four for splits) might be practically troublesome.
+One strategy that is sometimes implemented is adding zeros to the sample (i.e., \(t=0\) and \(d=0\)), which increase the number of observations.
+Unfortunately, this strategy should not be implemented since it will bias the estimated parameters:
+
+\small
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{jack\_profile\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits}\NormalTok{(}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, split\_times}\SpecialCharTok{$}\NormalTok{distance),}
+  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, split\_times}\SpecialCharTok{$}\NormalTok{jack\_time)}
+\NormalTok{)}
+
+\NormalTok{jack\_profile\_time\_estimated\_zero }\OtherTok{\textless{}{-}} \FunctionTok{model\_using\_splits\_with\_time\_correction}\NormalTok{(}
+  \AttributeTok{distance =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, split\_times}\SpecialCharTok{$}\NormalTok{distance),}
+  \AttributeTok{time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, split\_times}\SpecialCharTok{$}\NormalTok{jack\_time)}
+\NormalTok{)}
+
+\NormalTok{adding\_zero\_issue }\OtherTok{\textless{}{-}} \FunctionTok{rbind}\NormalTok{(}
+  \FunctionTok{coef}\NormalTok{(jack\_profile),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_zero),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_time\_estimated),}
+  \FunctionTok{coef}\NormalTok{(jack\_profile\_time\_estimated\_zero)}
+\NormalTok{)}
+
+\FunctionTok{rownames}\NormalTok{(adding\_zero\_issue) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+  \StringTok{"Jack {-} No corrections"}\NormalTok{,}
+  \StringTok{"Jack {-} No corrections with zeros added"}\NormalTok{,}
+  \StringTok{"Jack {-} Estimated time correction"}\NormalTok{,}
+  \StringTok{"Jack profile {-} Estimated time correction with added zeros"}\NormalTok{)}
+
+\FunctionTok{kable}\NormalTok{(adding\_zero\_issue, }\AttributeTok{digits =} \DecValTok{2}\NormalTok{, }\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"scale\_down"}\NormalTok{, }\StringTok{"hold\_position"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{table}[!h]
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MSS & TAU & MAC & PMAX & time\_correction & distance\_correction\\
+\midrule
+Jack - No corrections & 8.49 & 0.70 & 12.06 & 25.6 & 0.00 & 0\\
+Jack - No corrections with zeros added & 8.49 & 0.70 & 12.06 & 25.6 & 0.00 & 0\\
+Jack - Estimated time correction & 8.96 & 1.22 & 7.37 & 16.5 & 0.28 & 0\\
+Jack profile - Estimated time correction with added zeros & 8.51 & 0.73 & 11.65 & 24.8 & 0.02 & 0\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\normalsize
+
+Adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
+
+It is important to note that using \texttt{predict\_XXX()} family of functions with distance correction model and using estimated distance correction parameter will yield wrong predictions on the original scale (but not on the \emph{theoretical} scale in which force production starts at d=0m and t=0s).
+This is because \texttt{predict\_XXX()} family of functions utilize equation \eqref{eq:time-distance-correction} and its derivates and integrals, which works for the uncorrected model (time and distance correction parameters are equal to zero), time correction models (distance correction parameter is equal to zero), and time and distance corrections model.
+You can notice that equations \eqref{eq:time-correction} and \eqref{eq:time-distance-correction} are equivalent to the \eqref{eq:time-distance} equation when time and distance correction parameters are equal to zero.
+Unfortunately, this doesn't work for the distance correction model when estimated distance correction parameter is used since model definition outlined in the equation \eqref{eq:distance-correction} is different that the equation \eqref{eq:time-distance-correction} when estimated distance correction parameter is different than zero.
+The solution is to use generic S3 \texttt{predict()} function, but this only predicts split times at split distances.
+
+\small
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Jack\textquotesingle{}s observed times}
+\NormalTok{split\_times}\SpecialCharTok{$}\NormalTok{jack\_time}
+\CommentTok{\#\textgreater{} [1] 1.10 1.83 3.09 4.25 5.39}
+
+\CommentTok{\# To predict at "theoretical" distance scale (where force application starts at d=0m)}
+\FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{MSS =}\NormalTok{ jack\_profile\_time\_estimated}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ jack\_profile\_time\_estimated}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU)}
+\CommentTok{\#\textgreater{} [1] 1.38 2.12 3.37 4.54 5.67}
+
+\FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{MSS =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU)}
+\CommentTok{\#\textgreater{} [1] 1.42 2.17 3.43 4.60 5.73}
+
+\CommentTok{\# To predict at original distance scale, use all estimated parameters}
+\FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{MSS =}\NormalTok{ jack\_profile\_time\_estimated}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ jack\_profile\_time\_estimated}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \AttributeTok{time\_correction =}\NormalTok{ jack\_profile\_time\_estimated}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{time\_correction)}
+\CommentTok{\#\textgreater{} [1] 1.10 1.84 3.09 4.25 5.39}
+
+\CommentTok{\# Unfortunately, this doesn\textquotesingle{}t work for the distance correction model}
+\FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
+  \AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance, }
+  \AttributeTok{MSS =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{MSS,}
+  \AttributeTok{TAU =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
+  \AttributeTok{distance\_correction =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{distance\_correction)}
+\CommentTok{\#\textgreater{} [1] 1.50 2.23 3.49 4.65 5.78}
+
+\CommentTok{\# Instead use generic S3 predict() function}
+\FunctionTok{predict}\NormalTok{(jack\_profile\_distance\_correction)}
+\CommentTok{\#\textgreater{} [1] 1.10 1.83 3.09 4.25 5.39}
+
+\FunctionTok{predict}\NormalTok{(}
+\NormalTok{  jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{model,}
+  \AttributeTok{newdata =} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{distance =}\NormalTok{ split\_times}\SpecialCharTok{$}\NormalTok{distance))}
+\CommentTok{\#\textgreater{} [1] 1.10 1.83 3.09 4.25 5.39}
+\end{Highlighting}
+\end{Shaded}
+
+\normalsize
+
+In the next section we examine how these models (no correction, fixed time correction, estimated time correction, estimated distance correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the \protect\hyperlink{supplemental-material}{Supplemental material} for the R code).
+
+\small
+
+\normalsize
+
+As can be seen from the next figure, the estimated time correction model estimates MSS almost perfectly, while the estimated distance and estimate time and distance correction models estimates MSS perfectly.
+
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-38-1} \end{center}
+
+\normalsize
+
+Similar outcomes are observed for the MAC parameter.
+The distance correction model and time and distance corrections models performs perfectly, while the time correction model performs almost as good.
+PMAX demonstrates the same properties as MSS and MAC.
+
+The next figure depicts residuals across split distances for each simulated flying start distance.
+As can be seen, time correction, distance correction, and time and distance correction models perform much better than the other models.
+
+\small
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Residuals}
+\FunctionTok{ggplot}\NormalTok{(}
+\NormalTok{  model\_df,}
+  \FunctionTok{aes}\NormalTok{(}
+    \AttributeTok{y =}\NormalTok{ residuals,}
+    \AttributeTok{x =}\NormalTok{ distance,}
+    \AttributeTok{color =}\NormalTok{ flying\_start\_distance,}
+    \AttributeTok{group =}\NormalTok{ flying\_start\_distance}
+\NormalTok{  )}
+\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme\_bw}\NormalTok{(}\DecValTok{8}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.3}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{scale\_color\_gradientn}\NormalTok{(}\AttributeTok{colours =} \FunctionTok{terrain.colors}\NormalTok{(}\DecValTok{5}\NormalTok{, }\AttributeTok{rev =} \ConstantTok{FALSE}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{facet\_wrap}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{model) }\SpecialCharTok{+}
+  \FunctionTok{xlab}\NormalTok{(}\StringTok{"Distance (m)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{ylab}\NormalTok{(}\StringTok{"Observed {-} predicted time (s)"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"top"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{color =} \StringTok{"Flying start distance"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-39-1} \end{center}
+
+\normalsize
+
+The outcomes from this simple simulation data clearly demonstrates that the time correction, distance correction, and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model when attempting to overcome the flying start issues.
+Since the time correction and distance correction models are simpler and requires three parameters to be estimated, they might be practically more useful than the time and distance correction model, which requires four parameters estimation and thus more than four timing gates and sprint splits.
+More elaborate simulation study is currently ongoing and the results will be reported in another paper.
+
+Time correction, distance correction, and time and distance corrections are also implemented in the mixed-models using \texttt{mixed\_model\_using\_splits\_with\_time\_correction()}, \texttt{mixed\_model\_using\_splits\_with\_distance\_correction()}, and \texttt{mixed\_model\_using\_splits\_with\_corrections()} functions.
+We will showcase their use at the end of this paper.
 
@@
 
+\normalsize
+
 The model print output provides training dataset estimates and model performance, as well as LOOCV estimates and model performance.
 
-Next we plot estimated parameters across LOOCV folds (please refer to the Supplemental files for the R code):
-
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-51-1} \end{center}
+Next we plot estimated parameters across LOOCV folds (please refer to the \protect\hyperlink{supplemental-material}{Supplemental material} for the R code):
+
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-41-1} \end{center}
+
+\normalsize
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-52-1} \end{center}
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-42-1} \end{center}
+
+\normalsize
 
 As expected, the model has more issues predicting unseen split times for both short or long distances.
-Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
+Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least four observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
 LOOCV can also be implemented with the mixed-effects models in the \textbf{shorts} package.
@@
 For the sake of this analysis, only 10 m splits over 60 m race distance are used.
+
+\small
 
@@
 
+\normalsize
+
 Here is the model estimate of the time and distance it takes for Bolt to reach 99\% of MSS.
-Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start, not the measurement scale.
+Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start (i.e., theoretical scale), not the measurement scale.
+
+\small
 
@@
 
+\normalsize
+
 \hypertarget{vescovi-data}{%
@@
 
-The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players).
+The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players) (\protect\hyperlink{ref-vescoviSprintMechanicalCharacteristics2021}{Vescovi and Jovanovi'c 2021}).
 Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset.
@@
 Below is the mixed-effects models analysis of the dataset.
+
+\small
 
@@
 
-The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) estimated time correction as a fixed effect model, (4) estimated time correction as a random effect model, (5) estimated distance correction as fixed effect model (and time correction as random effect), and (6) estimated distance correction as random effect model (please refer to the Supplemental files for the R code).
+\normalsize
+
+The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) time correction model with time correction as fixed effect, (4) time correction model with time correction as random effect, (5) distance correction model with distance correction as fixed effect, (6) distance correction model with distance correction as random effect, (7) time and distance correction model, (8) time and distance correction model with distance correction as fixed effect (and time correction as random effect), and (9) time and distance correction model with distance and time correction as random effects (please refer to the \protect\hyperlink{supplemental-material}{Supplemental material} for the R code).
+
+\small
+
+\normalsize
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-57-1} \end{center}
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-47-1} \end{center}
+
+\normalsize
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-58-1} \end{center}
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-48-1} \end{center}
+
+\normalsize
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-59-1} \end{center}
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-49-1} \end{center}
+
+\normalsize
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-60-1} \end{center}
-
-Which model should should be used?
+\small
+
+\begin{center}\includegraphics[width=0.9\linewidth]{paper_files/figure-latex/unnamed-chunk-50-1} \end{center}
+
+\normalsize
+
+Which model should be used?
 Although providing a better fit (using RSE as an estimator of model fit), the time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction).
 Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for timing inconsistencies.
+More thorough theoretical simulation study is currently in development with the aim to explore the behavior of these models under different scenarios.
 
 We are hoping that the \textbf{shorts} package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
+
+\hypertarget{supplemental-material}{%
+\section{Supplemental material}\label{supplemental-material}}
+
+The \emph{R Markdown} (\protect\hyperlink{ref-R-bookdown}{Xie 2021a}; \protect\hyperlink{ref-R-rmarkdown}{Allaire et al.
+2021}; \protect\hyperlink{ref-rmarkdown2018}{Xie, Allaire, and Grolemund 2018}; \protect\hyperlink{ref-rmarkdown2020}{Xie, Dervieux, and Riederer 2020}) source code for the paper can be found on the GitHub repository: \url{https://github.com/mladenjovanovic/shorts-paper}.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-R-rmarkdown}{}%
+Allaire, JJ, Yihui Xie, Jonathan McPherson, Javier Luraschi, Kevin Ushey, Aron Atkins, Hadley Wickham, Joe Cheng, Winston Chang, and Richard Iannone.
+2021.
+\emph{Rmarkdown: Dynamic Documents for r}.
+\url{https://CRAN.R-project.org/package=rmarkdown}.
+
+\leavevmode\hypertarget{ref-altmannDifferentStartingDistances2015}{}%
+Altmann, Stefan, Marian Hoffmann, Gunther Kurz, Rainer Neumann, Alexander Woll, and Sascha Haertel.
+2015.
+{``Different {Starting Distances Affect} 5-m {Sprint Times}.''} \emph{Journal of Strength and Conditioning Research} 29 (8): 2361--66.
+\url{https://doi.org/10.1519/JSC.0000000000000865}.
+
+\leavevmode\hypertarget{ref-altmannValiditySingleBeamTiming2017}{}%
+Altmann, Stefan, Max Spielmann, Florian Azad Engel, Rainer Neumann, Steffen Ringhof, Doris Oriwol, and Sascha Haertel.
+2017.
+{``Validity of {Single}-{Beam Timing Lights} at {Different Heights}.''} \emph{Journal of Strength and Conditioning Research} 31 (7): 1994--99.
+\url{https://doi.org/10.1519/JSC.0000000000001889}.
+
+\leavevmode\hypertarget{ref-altmannAccuracySingleBeam2018}{}%
+Altmann, Stefan, Max Spielmann, Florian Azad Engel, Steffen Ringhof, Doris Oriwol, Sascha Härtel, and Rainer Neumann.
+2018.
+{``Accuracy of Single Beam Timing Lights for Determining Velocities in a Flying 20-m Sprint: {Does} Timing Light Height Matter?''} \emph{Journal of Human Sport and Exercise} 13 (3).
+\url{https://doi.org/10.14198/jhse.2018.133.10}.
+
 \leavevmode\hypertarget{ref-arsacModelingEnergetics100m2002}{}%
@@
 \url{https://doi.org/10.1152/japplphysiol.00754.2001}.
+
+\leavevmode\hypertarget{ref-batesNonlinearModels1992}{}%
+Bates, D. M., and J. M. Chambers. 1992. {``Nonlinear Models.''} In \emph{Statistical {Models} in {S}}. {Wadsworth \& Brooks/Cole}.
+
+\leavevmode\hypertarget{ref-batesNonlinearRegressionAnalysis2007}{}%
+Bates, Douglas M., and Donald G. Watts. 2007. \emph{Nonlinear Regression Analysis and Its Applications}. Wiley Series in Probability and Mathematical Statistics. {New York, NY}: {Wiley}.
 
@@
 
+\leavevmode\hypertarget{ref-haugenSprintRunningPerformance2016}{}%
+Haugen, Thomas, and Martin Buchheit.
+2016.
+{``Sprint {Running Performance Monitoring}: {Methodological} and {Practical Considerations}.''} \emph{Sports Medicine} 46 (5): 641--56.
+\url{https://doi.org/10.1007/s40279-015-0446-0}.
+
 \leavevmode\hypertarget{ref-jamesIntroductionStatisticalLearning2017}{}%
@@
 
-\leavevmode\hypertarget{ref-R-shorts}{}%
-Jovanovic, Mladen. 2020. \emph{Shorts: Short Sprints}. \url{https://CRAN.R-project.org/package=shorts}.
-
 \leavevmode\hypertarget{ref-jovanovicBmbstatsBootstrapMagnitudebased2020}{}%
 Jovanović, Mladen. 2020. \emph{Bmbstats: {Bootstrap Magnitude}-Based {Statistics} for {Sports Scientists}}. {Mladen Jovanović}.
+
+\leavevmode\hypertarget{ref-R-shorts}{}%
+---------. 2021. \emph{Shorts: Short Sprints}. \url{https://mladenjovanovic.github.io/shorts/}.
 
@@
 
+\leavevmode\hypertarget{ref-motulskyIntuitiveBiostatisticsNonmathematical2018}{}%
+Motulsky, Harvey. 2018. \emph{Intuitive Biostatistics: A Nonmathematical Guide to Statistical Thinking}. Fourth edition. {New York}: {Oxford University Press}.
+
 \leavevmode\hypertarget{ref-R-nlme}{}%
-Pinheiro, José, Douglas Bates, and R-core. 2020. \emph{Nlme: Linear and Nonlinear Mixed Effects Models}. \url{https://svn.r-project.org/R-packages/trunk/nlme/}.
+Pinheiro, José, Douglas Bates, and R-core. 2021. \emph{Nlme: Linear and Nonlinear Mixed Effects Models}. \url{https://svn.r-project.org/R-packages/trunk/nlme/}.
 
@@
 
+\leavevmode\hypertarget{ref-vescoviSprintMechanicalCharacteristics2021}{}%
+Vescovi, Jason D., and Mladen Jovanovi'c.
+2021.
+{``Sprint {Mechanical Characteristics} of {Female Soccer Players}: {A Retrospective Pilot Study} to {Examine} a {Novel Approach} for {Correction} of {Timing Gate Starts}.''} \emph{Frontiers in Sports and Active Living} 3 (May): 629694.
+\url{https://doi.org/10.3389/fspor.2021.629694}.
+
 \leavevmode\hypertarget{ref-ward-smithEnergyConversionStrategies2001}{}%
@@
 
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+Wickham, Hadley. 2021a. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
+
 \leavevmode\hypertarget{ref-R-tidyverse}{}%
-Wickham, Hadley. 2019. \emph{Tidyverse: Easily Install and Load the Tidyverse}. \url{https://CRAN.R-project.org/package=tidyverse}.
-
-\leavevmode\hypertarget{ref-R-tidyr}{}%
----------. 2020. \emph{Tidyr: Tidy Messy Data}. \url{https://CRAN.R-project.org/package=tidyr}.
+---------. 2021b. \emph{Tidyverse: Easily Install and Load the Tidyverse}. \url{https://CRAN.R-project.org/package=tidyverse}.
 
@@
 Wickham, Hadley, Winston Chang, Lionel Henry, Thomas Lin Pedersen, Kohske Takahashi, Claus Wilke, Kara Woo, Hiroaki Yutani, and Dewey Dunnington.
-2020.
+2021.
 \emph{Ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics}.
@@
 \leavevmode\hypertarget{ref-R-dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2020. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2021. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
 
@@
 
+\leavevmode\hypertarget{ref-R-bookdown}{}%
+---------. 2021a. \emph{Bookdown: Authoring Books and Technical Documents with r Markdown}. \url{https://CRAN.R-project.org/package=bookdown}.
+
 \leavevmode\hypertarget{ref-R-knitr}{}%
----------. 2020. \emph{Knitr: A General-Purpose Package for Dynamic Report Generation in r}. \url{https://yihui.org/knitr/}.
+---------. 2021b. \emph{Knitr: A General-Purpose Package for Dynamic Report Generation in r}. \url{https://yihui.org/knitr/}.
+
+\leavevmode\hypertarget{ref-rmarkdown2018}{}%
+Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
+
+\leavevmode\hypertarget{ref-rmarkdown2020}{}%
+Xie, Yihui, Christophe Dervieux, and Emily Riederer. 2020. \emph{R Markdown Cookbook}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown-cookbook}.
 
 \leavevmode\hypertarget{ref-R-kableExtra}{}%
-Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with Kable and Pipe Syntax}.
 

### 2021-07-24 paper.Rmd
Recompiled

@@
 
-# To predict at "theoretical" distance scale (where force application starts at d=0m)
+# To predict at "theoretical" distance scale
+#  (where force application starts at d=0m)
 predict_time_at_distance(
@@
   TAU = jack_profile_distance_correction$parameters$TAU,
-  distance_correction = jack_profile_distance_correction$parameters$distance_correction)
+  distance_correction = 
+    jack_profile_distance_correction$parameters$distance_correction)
 

### 2021-07-24 paper.tex
Recompiled

@@
 
-\CommentTok{\# To predict at "theoretical" distance scale (where force application starts at d=0m)}
+\CommentTok{\# To predict at "theoretical" distance scale}
+\CommentTok{\#  (where force application starts at d=0m)}
 \FunctionTok{predict\_time\_at\_distance}\NormalTok{(}
@@
   \AttributeTok{TAU =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{TAU,}
-  \AttributeTok{distance\_correction =}\NormalTok{ jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{distance\_correction)}
+  \AttributeTok{distance\_correction =} 
+\NormalTok{    jack\_profile\_distance\_correction}\SpecialCharTok{$}\NormalTok{parameters}\SpecialCharTok{$}\NormalTok{distance\_correction)}
 \CommentTok{\#\textgreater{} [1] 1.50 2.23 3.49 4.65 5.78}

### 2021-11-20 shorts-paper.Rmd
Paper prepared for submission to International Journal of Strength and Conditioning after review comments

--- 2021-07-21 paper.Rmd
@@
 title: "shorts: An R Package for Modeling Short Sprints"
-preprint: true
 date: "`r Sys.Date()`"
 author: 
-  - name: Mladen Jovanović
-    affiliation: 1
-    corresponding: true
-    email: coach.mladen.jovanovic@gmail.com
-  - name: Jason D. Vescovi
-    affiliation: 2
-    corresponding: false
-    email: vescovij@gmail.com
-affiliation:
-  - code: 1
-    address: Faculty of Sport and Physical Education, University of Belgrade, Serbia
-  - code: 2
-    address: Faculty of Kinesiology and Physical Education, Graduate School of Exercise Science, Toronto, ON Canada
+  - Mladen Jovanović^[Faculty of Sport and Physical Education, University of Belgrade, Serbia, coach.mladen.jovanovic@gmail.com]
+  - Jason D. Vescovi^[Faculty of Kinesiology and Physical Education, Graduate School of Exercise Science, Toronto, ON Canada, vescovij@gmail.com]
 abstract: >
@@
 In this paper, we: 1) provide the **shorts** package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the **shorts** package to help alleviate errors within the subsequent parameter outcomes.
-header-includes: >
-  \usepackage{lipsum}
-  \usepackage[htt]{hyphenat}
-  \usepackage{booktabs}
-  \usepackage{microtype}
-preamble: >
-  \usepackage{amsmath}
 citation_package: natbib
@@
 output:
-  bookdown::pdf_book:
-    base_format: rticles::peerj_article # for using bookdown features like \@ref()
-    extra_dependencies: ["flafter"]
-  rticles::peerj_article: default
+  bookdown::word_document2:
+    number_sections: FALSE
 link-citations: yes
@@
   fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
-  dpi = 300,
+  dpi = 600,
   out.width = "90%",
@@
 Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (<5 s) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the type of sport.
+Peak anaerobic power is achieved within the first few seconds (<5 s) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the athlete and their sport.
 For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)[@brownAssessmentLinearSprinting2004].
@@
 Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time) originally proposed by @doi:10.1098/rspb.1927.0035, and more recently popularized by @clarkNFLCombine40Yard2017, and @samozinoSimpleMethodMeasuring2016.
-Equation \@ref(eq:velocity-time) represents function for instantaneous horizontal velocity $v$ given the time $t$ and two model parameters:
+Equation \@ref(eq:velocity-time) represents the function for instantaneous horizontal velocity $v$ given the time $t$ and two model parameters:
 
@@
 
-By derivating equation \@ref(eq:velocity-time), we can get equation for horizontal acceleration \@ref(eq:acceleration-time).
+By derivating equation \@ref(eq:velocity-time), we can get the equation for horizontal acceleration \@ref(eq:acceleration-time).
 
@@
 
-By integrating equation \@ref(eq:velocity-time), we can get equation for distance covered \@ref(eq:distance-time).
+By integrating equation \@ref(eq:velocity-time), we can get the equation for distance covered \@ref(eq:distance-time).
 
@@
   xlab(expression("Velocity (ms"^-1 * ")")) +
+  xlim(c(0, 12)) +
   theme(
@@
 
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), (2) getting a velocity trace using a radar gun.
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), or (2) getting a velocity trace using a radar gun.
 Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **shorts** package using non-linear least squares regression implemented in the `nls()` function [@batesNonlinearModels1992; @batesNonlinearRegressionAnalysis2007] in the **base R** [@R-base] and `nlme()` function in the **nlme** package [@R-nlme] for the mixed-effect models.
@@
 $W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW].
-Researchers often incorrectly use the equation \@ref(eq:distance-time) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothSpreadsheetSprintAcceleration2020], in which the time is the predictor and distance is the outcome variable, instead of statistically correct equation \@ref(eq:time-distance) [@motulskyIntuitiveBiostatisticsNonmathematical2018, p.
+Researchers often incorrectly use the equation \@ref(eq:distance-time) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothSpreadsheetSprintAcceleration2020], in which the time is the predictor and distance is the outcome variable, instead of statistically correct equation \@ref(eq:time-distance).
+This practice should be avoided because switching of the predictor and outcome variables in the regression model might produce biased estimated parameters [@motulskyIntuitiveBiostatisticsNonmathematical2018, p.
 341].
+This might not necessarily be the case with the short sprints profiling, but it is nevertheless a bad statistical practice.
 
@@
 Both split and radar gun models allow the use of *weighted* non-linear regression.
-For example, we can give more weight to shorter distance or faster velocities.
-Weighted non-linear regression is performed by setting `weights` parameter:
+Weighting in regression is utilized when the observations have unequal error variance [@gelmanRegressionOtherStories2020], which can happen due to instrumental error (e.g., multiplicative error instead of additive error) or due to biological variability (e.g., higher split time variance on shorter distances versus longer distances).
+In this case, observations with higher error variance get lower weight when fitting the model [@gelmanRegressionOtherStories2020].
+According to @gelmanRegressionOtherStories2020, unequal variances are not typically a major issue for the goal of estimating regression parameters, but they can become more important when making predictions about individual observations.
+Further exploring the topic of unequal error variances is beyond the scope of this paper, but the **shorts** package provides weighting feature for the potential research of this topic in the future.
+
+Weighted non-linear regression is performed by setting `weights` parameter. For example, we can give more weight to shorter distance or faster velocities:
 
@@
 require(knitr)
-require(kableExtra)
 
@@
 
-kable(jack_parameters, digits = 2, booktabs = TRUE) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"))
+kable(jack_parameters, digits = 2, booktabs = TRUE)
 ```
@@
 
-kable(jack_parameters, digits = 2, booktabs = TRUE) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"))
+kable(jack_parameters, digits = 2, booktabs = TRUE) 
 ```
@@
 
-kable(adding_zero_issue, digits = 2, booktabs = TRUE) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"))
+kable(adding_zero_issue, digits = 2, booktabs = TRUE)
 ```
@@
 You can notice that equations \@ref(eq:time-correction) and \@ref(eq:time-distance-correction) are equivalent to the \@ref(eq:time-distance) equation when time and distance correction parameters are equal to zero.
-Unfortunately, this doesn't work for the distance correction model when estimated distance correction parameter is used since model definition outlined in the equation \@ref(eq:distance-correction) is different that the equation \@ref(eq:time-distance-correction) when estimated distance correction parameter is different than zero.
+Unfortunately, this doesn't work for the distance correction model when the estimated distance correction parameter is used since model definition outlined in the equation \@ref(eq:distance-correction) is different that the equation \@ref(eq:time-distance-correction) when the estimated distance correction parameter is different than zero.
 The solution is to use generic S3 `predict()` function, but this only predicts split times at split distances.
@@
 
-To estimate parameter stability, model over-fitting, and performance on the unseen data, **shorts** model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
+To estimate parameter stability, model over-fitting, and performance on unseen data, **shorts** model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
 LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation.
 This process is repeated for each observation in the model dataset.
-LOOCV allows one to check estimated parameters stability, and model performance on the unseen data.
+LOOCV allows one to check estimated parameters stability, and model performance on unseen data.
 
@@
 Let's utilize demonstrated functionalities of the **shorts** package using real-world data.
-The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the **shorts** package (see `?vescovi`).
+The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the **shorts** package (see the dataset documentation by executing `?vescovi` command in the R console).
 
@@
 
-kable(bolt_model, digits = 2, booktabs = TRUE) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"))
+kable(bolt_model, digits = 2, booktabs = TRUE)
 ```
@@
 
-The following models were used: (1) no corrections model, (2) fixed time correction model (using 0.3s heuristic rule of thumb), (3) time correction model with time correction as fixed effect, (4) time correction model with time correction as random effect, (5) distance correction model with distance correction as fixed effect, (6) distance correction model with distance correction as random effect, (7) time and distance correction model, (8) time and distance correction model with distance correction as fixed effect (and time correction as random effect), and (9) time and distance correction model with distance and time correction as random effects (please refer to the [Supplemental material] for the R code).
+The following models were used: (1) no corrections model, (2) fixed time correction model (using +0.3s heuristic rule of thumb), (3) time correction model with time correction as fixed effect, (4) time correction model with time correction as random effect, (5) distance correction model with distance correction as fixed effect, (6) distance correction model with distance correction as random effect, (7) time and distance correction model, (8) time and distance correction model with distance correction as fixed effect (and time correction as random effect), and (9) time and distance correction model with distance and time correction as random effects (please refer to the [Supplemental material] for the R code).
 

### 2021-11-20 shorts-paper.Rmd
Few extra changes

@@
 Both split and radar gun models allow the use of *weighted* non-linear regression.
-Weighting in regression is utilized when the observations have unequal error variance [@gelmanRegressionOtherStories2020], which can happen due to instrumental error (e.g., multiplicative error instead of additive error) or due to biological variability (e.g., higher split time variance on shorter distances versus longer distances).
+Weighting in regression is utilized when the observations have unequal error variance [@gelmanRegressionOtherStories2020], which can happen due to instrumental error (e.g., a multiplicative error instead of additive error) or due to biological variability (e.g., higher split time variance on shorter distances versus longer distances).
 In this case, observations with higher error variance get lower weight when fitting the model [@gelmanRegressionOtherStories2020].
 According to @gelmanRegressionOtherStories2020, unequal variances are not typically a major issue for the goal of estimating regression parameters, but they can become more important when making predictions about individual observations.
-Further exploring the topic of unequal error variances is beyond the scope of this paper, but the **shorts** package provides weighting feature for the potential research of this topic in the future.
-
-Weighted non-linear regression is performed by setting `weights` parameter. For example, we can give more weight to shorter distance or faster velocities:
+Further exploring the topic of unequal error variances is beyond the scope of this paper, but the **shorts** package provides a weighting feature for the potential research of this topic in the future.
+
+Weighted non-linear regression is performed by setting `weights` parameter. For example, we can give more weight to shorter distances or faster velocities:
 

### 2022-06-20 shorts-paper.Rmd
Updated to use {short} 1.2

@@
 ```{r}
-# Since this is a perfect simulation and stats::nls will complain
-# we need to add very small noise, or measurement error to the times
-set.seed(1667)
-rand_noise <- rnorm(nrow(split_times), 0, 10^-5)
-split_times$john_time <- split_times$john_time + rand_noise
-split_times$jack_time <- split_times$jack_time + rand_noise
-
 john_profile <- model_using_splits(
@@
   )
-
-# Add small noise to allow model fit
-set.seed(1667)
-rand_noise <- rnorm(nrow(sim_df), 0, 10^-5)
-sim_df$time <- sim_df$time + rand_noise
 ```
@@
     distance = data$distance,
-    time = data$time,
-    control = nls.control(warnOnly = TRUE)
+    time = data$time
   )

### 2022-07-07 shorts-paper.Rmd
Rewrote to be used with {shorts} 2.0.0

@@
 ---
-title: "shorts: An R Package for Modeling Short Sprints"
+title: "**{shorts}**: An R Package for Modeling Short Sprints"
 date: "`r Sys.Date()`"
@@
 abstract: >
-  Short sprint performance is one of the most distinguishable and admired physical traits in sports.
+  Short sprint performance is one of the sport's most distinguishable and admired physical traits.
 Short sprints have been modeled using the mono-exponential equation that involves two parameters: (1) maximum sprinting speed (MSS) and (2) relative acceleration (TAU).
-The most common methods to assess short sprint performance are with a radar gun or timing gates.
-In this paper, we: 1) provide the **shorts** package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the **shorts** package to help alleviate errors within the subsequent parameter outcomes.
+The most common methods to assess short sprint performance are radar guns or timing gates.
+In this paper, we: 1) provide the **{shorts}** package that can model sprint timing data from these two sources; 2) discuss potential issues with assessing sprint time (synchronization and flying start, respectively); and 3) provide model definitions within the **{shorts}** package to help alleviate errors within the subsequent parameter outcomes.
 citation_package: natbib
@@
   x <- def.chunk.hook(x, options)
-  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
+  ifelse(options$size != "normalsize", paste0("\n \\", options$size, "\n\n", x, "\n\n \\normalsize"), x)
 })
@@
 # automatically create a bib database for R packages
-knitr::write_bib(c(
-  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "nlme", "LambertW", "kableExtra"
- ), "packages.bib")
+# knitr::write_bib(c(
+#  .packages(), "bookdown", "knitr", "rmarkdown", "shorts", "LambertW", "kableExtra", #"minpack.lm"
+# ), "packages.bib")
 
@@
   "width" = 65,
-  scipen = 999
+  scipen = 999,
+  knitr.kable.NA = ""
 )
@@
 
-Short sprint performance is one of the most distinguishable and admired physical traits in sports.
-Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand still position over a distance that doesn't result in deceleration at the end.
-Peak anaerobic power is achieved within the first few seconds (<5 s) of maximal efforts [@mangineSpeedForcePower2014], whereas the ability to achieve maximal sprint speed varies based on the athlete and their sport.
-For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 m) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 m)[@brownAssessmentLinearSprinting2004].
+Short sprint performance is one of the sport's most distinguishable and admired physical traits.
+Short sprints, commonly performed in most team sports (e.g., soccer, field hockey, handball, football, etc.), are defined as maximal running from a stand-still position over a distance that does not result in deceleration at the end.
+Peak anaerobic power is achieved within the first few seconds (<5 $s$) of maximal efforts [@mangineSpeedForcePower2014].
+In contrast, achieving maximal sprint speed varies based on the athlete and sport.
+For example, track and field sprinters are trained to achieve maximal speed later in a race (i.e., 50-60 $m$) [@ward-smithEnergyConversionStrategies2001], but team sport athletes have sport-specific attributes and reach it much sooner (i.e., 30-40 $m$)[@brownAssessmentLinearSprinting2004].
 Regardless of the differences in kinematics between athletes, evaluating short sprint performance is routinely included within a battery of fitness tests for a wide range of sports
 
-The use of force plates is considered the gold standard for assessing mechanical properties of sprinting; however, there are logistical and financial challenges to capturing the profile of an entire sprint [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
-Radar and laser technology are frequently used laboratory-grade methods [@buchheitMechanicalDeterminantsAcceleration2014; @edwardsSprintAccelerationCharacteristics2020; @jimenez-reyesRelationshipVerticalHorizontal2018; @marcote-pequenoAssociationForceVelocity2019] but not normally accessible to practitioners working in sports.
-Undoubtedly, the most common method available and used to evaluate sprint performance are timing gates.
-Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 m), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
-This approach presents an advantage to practitioners who can use the outcomes to describe individual differences, quantify the effects of training interventions, and better understanding the limiting factors of performance.
-
-The **shorts** package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sports scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
-To best of our knowledge, scientist, researchers, and coaches have been performing short sprints modeling using the built-in solver function of Excel (Microsoft Corporation, Redmond, Washington, United States) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothForcevelocityProfilingIce2020; @stenrothSpreadsheetSprintAcceleration2020; @samozinoSimpleMethodMeasuring2016; @clarkNFLCombine40Yard2017; @morinSimpleMethodComputing2019], which makes the **shorts** package a major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
-
-In the current paper, we will provide an explanation of one commonly used mathematical equation to model short sprints, modeling applications using the **shorts** package, issues that can arise during measurement and estimation, and potential solutions to those problems.
+The use of force plates is considered the gold standard for assessing the mechanical properties of sprinting; however, there are logistical and financial challenges to capturing the profile of an entire sprint [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+Radar and laser technology are frequently used laboratory-grade methods [@buchheitMechanicalDeterminantsAcceleration2014; @edwardsSprintAccelerationCharacteristics2020; @jimenez-reyesRelationshipVerticalHorizontal2018; @marcote-pequenoAssociationForceVelocity2019] but are not ordinarily accessible to practitioners working in sports.
+Undoubtedly, the most common method available and used to evaluate sprint performance is the use of timing gates.
+Often multiple gates are positioned at varying distances to capture split times (e.g., 5, 10, 20 $m$), which can now be incorporated into the method for determining sprint mechanical properties [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+This approach presents an advantage to practitioners using the outcomes to describe individual differences, quantify the effects of training interventions, and better understand performance limiting factors.
+
+The **{shorts}** package [@R-shorts], written in the R language [@R-base], represents an open-source tool to help sports scientists translate raw timing data into detailed mechanical outcomes through mathematical modeling [@morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+To the best of our knowledge, scientist, researchers, and coaches have been performing short sprints modeling using the built-in solver function of Microsoft Excel (Microsoft Corporation, Redmond, Washington, United States) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothForcevelocityProfilingIce2020; @stenrothSpreadsheetSprintAcceleration2020; @samozinoSimpleMethodMeasuring2016; @clarkNFLCombine40Yard2017; @morinSimpleMethodComputing2019], which makes the **{shorts}** package a major improvement in ease-of-use, speed, transparency, reproducibility, and more feature-rich model fitting.
+
+The current paper will explain one commonly used mathematical equation to model short sprints, modeling applications using the **{shorts}** package, issues arising during measurement and estimation, and potential solutions to those problems.
 
@@
 
-Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time) originally proposed by @doi:10.1098/rspb.1927.0035, and more recently popularized by @clarkNFLCombine40Yard2017, and @samozinoSimpleMethodMeasuring2016.
+Short sprints have been modeled using the mono-exponential equation \@ref(eq:velocity-time), originally proposed by @doi:10.1098/rspb.1927.0035 and more recently popularized by @clarkNFLCombine40Yard2017 and @samozinoSimpleMethodMeasuring2016.
 Equation \@ref(eq:velocity-time) represents the function for instantaneous horizontal velocity $v$ given the time $t$ and two model parameters:
@@
 
-The parameters of the equation \@ref(eq:velocity-time) are *maximum sprinting speed* (MSS; expressed in $ms^{-1}$) and *relative acceleration* (TAU).
-Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; *maximal acceleration*, expressed in $ms^{-2}$) \@ref(eq:maximal-acceleration).
+The parameters of equation \@ref(eq:velocity-time) are *maximum sprinting speed* (MSS; expressed in $ms^{-1}$) and *relative acceleration* (TAU; expressed in $s$).
+Mathematically, TAU represents the ratio of MSS to initial acceleration (MAC; *maximal acceleration*, expressed in $ms^{-2}$) (equation \@ref(eq:maximal-acceleration)).
+Given the equation \@ref(eq:velocity-time), TAU can be interpreted as the time required to reach a sprinting velocity equal to `r round(100 * (1 - exp(-1)), 1)`% of MSS.
 
@@
 
-Although TAU is used in the equations, and later estimated, we prefer to use MAC instead since it is easier to grasp, particularly for less math inclined coaches.
+Although TAU is used in the equations and later estimated, we prefer MAC since it is easier to grasp, particularly for less math-inclined coaches.
 
@@
 
-Let's consider four athletes with different levels of MSS (high versus low maximal sprinting speed) and MAC (high versus low maximal acceleration; as mentioned previously, using MAC is preferred over using TAU) (Table \@ref(tab:four-athletes-table)).
+Let us consider four athletes with different levels of MSS (high versus low maximal sprinting speed) and MAC (high versus low maximal acceleration; as mentioned previously, using MAC is preferred over using TAU):
 
@@
   athletes,
-  caption = "Four athletes with different MSS and MAC parameters.",
+  #caption = "Four athletes with different MSS and MAC parameters.",
   digits = 2,
@@
 
-Figure \@ref(fig:four-athletes-kinematics) depicts distance, velocity, and acceleration over time (from 0 to 6 s).
-
-```{r four-athletes-kinematics, echo=FALSE, fig.cap="Kinematic characteristic of four athletes with different MSS and MAC parameters over a period of 0 to 6 seconds."}
+Next figure depicts distance, velocity, and acceleration over time (from 0 to 6 $s$):
+
+```{r four-athletes-kinematics, echo=FALSE}
 kinematics <- expand_grid(
@@
   mutate(
-    distance = predict_distance_at_time(time, MSS, TAU),
-    velocity = predict_velocity_at_time(time, MSS, TAU),
-    acceleration = predict_acceleration_at_time(time, MSS, TAU)
+    distance = predict_distance_at_time(time, MSS, MAC),
+    velocity = predict_velocity_at_time(time, MSS, MAC),
+    acceleration = predict_acceleration_at_time(time, MSS, MAC)
   )
@@
 
-Plotting acceleration against velocity (Figure \@ref(fig:four-athletes-profile)), we will get *Acceleration-Velocity Profile*, which is linear, according to the mathematical model.
+Plotting acceleration against velocity in the next figure, we will get the *Acceleration-Velocity Profile*, which is linear, according to the mathematical model.
 If the athlete's body mass (kg) is known, as well as additional air resistance parameters (see [Air resistance and the calculation of force and mechanical power] section of this paper), *Force-Velocity Profile* can be estimated (see [Force-Velocity profile] section of this paper).
 
-```{r four-athletes-profile, echo=FALSE, fig.cap="Acceleration-Velocity profile of four athletes with different MSS and MAC parameters."}
+```{r four-athletes-profile, echo=FALSE}
 gg <- ggplot(
@@
 
-# Estimation using shorts package
-
-Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 m), or (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **shorts** package using non-linear least squares regression implemented in the `nls()` function [@batesNonlinearModels1992; @batesNonlinearRegressionAnalysis2007] in the **base R** [@R-base] and `nlme()` function in the **nlme** package [@R-nlme] for the mixed-effect models.
+# Estimation using **{shorts}** package
+
+Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 $m$), or (2) getting a velocity trace using a radar gun.
+Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **{shorts}** package for the **R language** [@R-base] using *non-linear least squares regression* implemented in the `nlsLM()` function from the **{minpack.lm}** package [@R-minpack.lm].
 
@@
 
-Let's consider an example of an athlete with MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$ performing 40m sprint with timing gates positioned at each 10m split.
-For split times, distance is a predictor, and time is the outcome variable, thus the equation \@ref(eq:distance-time) becomes:
+For timing gates split times, distance is a predictor, and time is the outcome variable. Thus, equation \@ref(eq:distance-time) becomes:
 
@@
 
-$W$ in equation \@ref(eq:time-distance) represents Lambert's W function [@R-LambertW].
-Researchers often incorrectly use the equation \@ref(eq:distance-time) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothSpreadsheetSprintAcceleration2020], in which the time is the predictor and distance is the outcome variable, instead of statistically correct equation \@ref(eq:time-distance).
-This practice should be avoided because switching of the predictor and outcome variables in the regression model might produce biased estimated parameters [@motulskyIntuitiveBiostatisticsNonmathematical2018, p.
+In equation \@ref(eq:time-distance), $W$ represents Lambert's W function [@R-LambertW].
+Researchers often incorrectly use the equation \@ref(eq:distance-time) [@morinSpreadsheetSprintAcceleration2017; @morinSpreadsheetSprintAcceleration2019; @stenrothSpreadsheetSprintAcceleration2020], in which the time is the predictor and distance is the outcome variable, instead of the statistically correct equation \@ref(eq:time-distance).
+This practice should be avoided because switching the predictor and outcome variables in the regression model might produce biased estimated parameters [@motulskyIntuitiveBiostatisticsNonmathematical2018, p.
 341].
-This might not necessarily be the case with the short sprints profiling, but it is nevertheless a bad statistical practice.
-
-MSS and TAU parameters are estimated using `model_using_splits()` function:
+These biases might not necessarily be significant with the short sprints profiling, but it is nevertheless a bad statistical practice.
+
+Here is an example using a built-in dataset.
+We are going to filter out one athlete (e.g., John) with the help of the **tidyverse** [@R-tidyverse] package and estimate MSS, TAU, and MAC parameters using the `model_timing_gates()` function:
 
@@
 require(shorts)
-
-split_distance <- c(10, 20, 30, 40)
-
-split_time <- c(2.17, 3.43, 4.60, 5.73)
-
-m1 <- model_using_splits(
-  distance = split_distance,
-  time = split_time
+require(tidyverse)
+
+# Load built-in dataset
+data(split_times)
+
+# Filter timing gates splits for John
+john_TG_data <- split_times %>%
+  filter(athlete == "John")
+
+john_TG_data
+
+# Estimate John's MSS, TAU, MAC, and PMAX
+m1 <- model_timing_gates(
+  distance = john_TG_data$distance,
+  time = john_TG_data$time
 )
@@
 
-Maximal relative power (PMAX) from the output is estimated using $\frac{MSS \times MAC}{4}$, which disregards the air resistance.
-`time_correction` and `distance_corection` parameters will be covered later in the paper.
-
-Besides providing *residual standard error* (RSE), **shorts** functions provide additional model fit estimators.
-Additional information can be gained by exploring the returned object, particularly object returned from the `nls()` function (i.e., by using the S3 `summary()` method).
+Maximal relative power (PMAX) from the output is estimated using $\frac{MSS \times MAC}{4}$, disregarding the air resistance.
+
+Besides providing *residual standard error* (RSE), **{shorts}** functions provide additional model fit estimators.
+Additional information can be gained by exploring the returned object, particularly the object returned from the `nlsLM()` function (i.e., by using the S3 `summary()` method).
 To extract estimated model parameters, use S3 `coef()` method.
 
-To create a simple plot of the model, use S3 `plot()` method, which returns **ggplot2** [@R-ggplot2] object:
+To create a simple plot of the model, use the S3 `plot()` method, which returns **ggplot2** [@R-ggplot2] object:
 
@@
 
-Once we have estimated MSS and TAU, we can use `predict_XXX()` family of functions to predict various relationships (i.e., time at distance, acceleration at distance, velocity at time, etc.):
+Once we have estimated MSS and MAC, we can use `predict_XXX()` family of functions to predict various relationships (i.e., time at distance, acceleration at distance, velocity at time, etc.):
 
@@
 predict_time_at_distance(
-  distance = split_distance,
+  distance = john_TG_data$distance,
   MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU
-)
-```
-
-### Air resistance and the calculation of force and mechanical power
+  MAC = m1$parameters$MAC
+)
+```
+
+## Estimating short sprint parameters using the radar gun
+
+Estimating the short sprint profile using radar gun data takes time as a predictor and velocity as the outcome variable.
+Estimation using radar gun data is implemented in the **{shorts}** package using `model_radar_gun()` function.
+
+Here is an example using built-in dataset:
+
+```{r}
+# Load built-in dataset
+data(radar_gun_data)
+
+# Filter radar gun data for John
+john_RG_data <- radar_gun_data %>%
+  filter(athlete == "John")
+
+head(john_RG_data)
+
+# Estimate John's MSS, TAU, MAC, and PMAX
+m2 <- model_radar_gun(
+  velocity = john_RG_data$velocity,
+  time = john_RG_data$time
+)
+
+m2
+```
+
+There is the additional parameter in the output, TC, that is estimated using the `model_radar_gun()` function.
+The function of this parameter is explained in the [Problems with time sync with the radar gun] section of this paper.
+
+Both timing gates and radar gun models allow *weighted* non-linear regression.
+Weighting in regression is utilized when the observations have unequal error variance [@gelmanRegressionOtherStories2020], which can happen due to instrumental error (e.g., a multiplicative error instead of additive error) or due to biological variability (e.g., higher split time variance on shorter distances versus longer distances).
+In this case, observations with higher error variance get lower weight when fitting the model [@gelmanRegressionOtherStories2020].
+According to @gelmanRegressionOtherStories2020, unequal variances are not typically a significant issue for the goal of estimating regression parameters, but they can become more important when making predictions about individual observations.
+Further exploring the topic of unequal error variances is beyond the scope of this paper.
+However, the **{shorts}** package provides a weighting feature for potential research of this topic in the future.
+
+Weighted non-linear regression is performed by setting the `weights` parameter. For example, we can give more weight to shorter distances or faster velocities:
+
+```{r}
+m2_weighted <- model_radar_gun(
+  velocity = john_RG_data$velocity,
+  time = john_RG_data$time,
+  weights = 1 / (john_RG_data$velocity + 1)
+)
+
+m2_weighted
+```
+
+## Air resistance and the calculation of force and mechanical power
 
 To estimate force production at distance or time (using `predict_force_at_distance()` and `predict_force_at_time()` functions), as well as power production (using `predict_power_at_distance()` and `predict_power_at_time()` functions), one needs to take into account the air resistance.
-Air resistance (N) is estimated using `get_air_resistance()` function, which takes velocity, body mass (kg), body height (m), barometric pressure (Torr), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
+Air resistance (measured in Newtons, $N$) is estimated using `get_air_resistance()` function, which takes velocity, body mass ($kg$), body height ($m$), barometric pressure ($Torr$), air temperature ($C^\circ$), and wind velocity ($ms^-1$) as parameters (please refer to @arsacModelingEnergetics100m2002, @samozinoSimpleMethodMeasuring2016, and @vaningenschenauCanCyclePower1991 for more information):
 
@@
 predict_force_at_distance(
-  distance = split_distance,
+  distance = john_TG_data$distance,
   MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU,
+  MAC = m1$parameters$MAC,
+  # Additional parameters forwarded to get_air_resistance
+  # Otherwise, defaults are used
+  bodymass = john_TG_data$bodyweight,
+  bodyheight = 1.85,
+  barometric_pressure = 780,
+  air_temperature = 20,
+  wind_velocity = 0.5
+)
+```
+
+The easiest way to get all kinematics and kinetics for short sprints is to use `predict_kinematics()` function:
+
+```{r}
+df <- predict_kinematics(
+  m1,
+  max_time = 6,
+  frequency = 100,
+  # Additional parameters forwarded to get_air_resistance
+  # Otherwise, defaults are used
+  bodymass = john_TG_data$bodyweight[1],
+  bodyheight = 1.85,
+  barometric_pressure = 780,
+  air_temperature = 20,
+  wind_velocity = 0.5
+)
+```
+
+Plotting the model predictions can be done once we convert data from wide to long with the help of **ggplot2** [@R-ggplot2], **dplyr** [@R-dplyr], and **tidyr** [@R-tidyr] packages (loaded already using the **tidyverse** [@R-tidyverse] package):
+
+```{r}
+variable_names <- colnames(df)
+
+df <- pivot_longer(data = df, cols = -2) %>%
+  mutate(name = factor(name, levels = variable_names))
+
+ggplot(df, aes(x = distance, y = value)) +
+  theme_bw(8) +
+  facet_wrap(~name, scales = "free_y") +
+  geom_line(alpha = 0.7) +
+  ylab(NULL) +
+  xlab("Distance (m)")
+```
+
+These kinematic and kinetic variables are utilized in [Force-Velocity profile] estimation, which is covered later in this paper.
+
+## Utility functions
+
+Another valuable addition for sports scientists and coaches is the ability to determine the distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
+To identify these values, the **{shorts}** package comes with the `find_XXX()` family of functions:
+
+```{r}
+# Finds distance where 90% of maximum sprinting speed is reached
+find_velocity_critical_distance(
+  MSS = m1$parameters$MSS,
+  MAC = m1$parameters$MAC,
+  percent = 0.9
+)
+
+# Finds maximal power and distance (this time using air resistance)
+find_max_power_distance(
+  MSS = m1$parameters$MSS,
+  MAC = m1$parameters$MAC,
   # Additional parameters forwarded to get_air_resistance
@@
 )
-```
-
-The easiest way to get all kinematics and kinetics for short sprints is to use `predict_kinematics()` function:
-
-```{r}
-df <- predict_kinematics(
-  m1,
-  max_time = 6,
-  frequency = 100,
+
+# Finds distance over 90% power range
+find_power_critical_distance(
+  MSS = m1$parameters$MSS,
+  MAC = m1$parameters$MAC,
   # Additional parameters forwarded to get_air_resistance
@@
 
-Plotting the model predictions can be done once we convert data from wide to long with the help of **ggplot2** [@R-ggplot2], **dplyr** [@R-dplyr], **tidyr** [@R-tidyr], and **tidyverse** [@R-tidyverse] packages:
-
-```{r}
-require(tidyverse)
-
-variable_names <- colnames(df)
-
-df <- pivot_longer(data = df, cols = -2) %>%
-  mutate(name = factor(name, levels = variable_names))
-
-ggplot(df, aes(x = distance, y = value)) +
-  theme_bw(8) +
-  facet_wrap(~name, scales = "free_y") +
-  geom_line(alpha = 0.7) +
-  ylab(NULL) +
-  xlab("Distance (m)")
-```
-
-These kinematic and kinetic variables are utilized in [Force-Velocity profile] estimation, which is covered later in this paper.
-
-### Utility functions
-
-Another valuable addition for sport scientists and coaches is the ability to determine the distances and times where 90% of maximum sprinting speed is reached, or where peak power is within 90% range.
-To identify these values, **shorts** package comes with `find_XXX()` family of functions:
-
-```{r}
-# Finds distance where 90% of maximum sprinting speed is reached
-find_velocity_critical_distance(
+## Force-Velocity profile
+
+To create *Force-Velocity Profile* (FVP) using single athlete estimated sprint model parameters (i.e., MSS and MAC), you can use the `make_FV_profile()` function.
+When estimating FVP, athlete body mass (kg) can be set using `bodymass` parameter, while the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance()` function.
+Details of the FVP method implemented in the **{shorts}** package and the interpretation from a sprint training perspective are covered elsewhere [@haugenPowerForceVelocityProfilingSprinting2020; @morinInterpretingPowerForceVelocityProfiles2016; @morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
+
+```{r}
+# To create Force-Velocity Profile
+fvp <- make_FV_profile(
   MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU,
-  percent = 0.9
-)
-
-# Finds maximal power and distance (this time using air resistance)
-find_max_power_distance(
-  MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU,
-  # Additional parameters forwarded to get_air_resistance
-  # Otherwise, defaults are used
-  bodymass = 80,
-  bodyheight = 1.85,
-  barometric_pressure = 780,
-  air_temperature = 20,
-  wind_velocity = 0.5
-)
-
-# Finds distance over 90% power range
-find_power_critical_distance(
-  MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU,
-  # Additional parameters forwarded to get_air_resistance
-  # Otherwise, defaults are used
-  bodymass = 80,
-  bodyheight = 1.85,
-  barometric_pressure = 780,
-  air_temperature = 20,
-  wind_velocity = 0.5
-)
-```
-
-### Mixed-effects model
-
-Sprint performance is often evaluated with a group of athletes (e.g., soccer club) representing a single strata of interest.
-Sports scientists can estimate individual profiles, or utilize mixed-effects models.
-To perform mixed-effects models in **shorts** for split times, one can use `mixed_model_using_splits()` function.
-To demonstrate this functionality, we load the `split_times` dataset provided in the **shorts** package:
-
-```{r}
-data(split_times)
-
-# Mixed model
-m2 <- mixed_model_using_splits(
-  data = split_times,
-  distance = "distance",
-  time = "time",
-  athlete = "athlete",
-
-  # Select random effects
-  # Default is MSS and TAU
-  random = MSS + TAU ~ 1
-)
-
-m2
-```
-
-Additional information about mixed-effects model performed using the `nlme` package [@R-nlme] can be obtained using `summary()` function.
-S3 method `coef()` when applied on mixed-model result will return both the fixed and random effects.
-
-```{r include=FALSE}
-summary(m2)
-coef(m2)
-```
-
-To create a simple plot of the model, use S3 `plot()` method.
-
-```{r}
-plot(m2) + theme_bw(8)
-```
-
-The following figure contains kinematics for all athletes in `split_times` dataset. Please note that power calculation takes default parameters for each individual:
-
-```{r}
-df <- predict_kinematics(m2, max_time = 10)
-
-variable_names <- colnames(df)
-
-df <- pivot_longer(df, cols = c(-1, -3)) %>%
-  mutate(name = factor(name, levels = variable_names))
-
-ggplot(
-  filter(df, distance < 40),
-  aes(x = distance, y = value, group = athlete, color = athlete)
-) +
-  theme_bw(8) +
-  facet_wrap(~name, scales = "free_y") +
-  geom_line(alpha = 0.7) +
-  ylab(NULL) +
-  xlab("Distance (m)") +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank()
-  )
-```
-
-## Estimating short sprint parameters using radar gun
-
-Estimation of the short sprint profile using radar gun data takes time as predictor and velocity as the outcome variable. Thus equation \@ref(eq:velocity-time) is used to estimate MSS and TAU.
-
-Let's consider the same example of an athlete with MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$ performing 40m sprint with velocity estimated using radar run (in this case with 1 Hz sampling rate).
-
-```{r}
-sprint_time <- seq(0, 6, 1)
-
-sprint_velocity <- c(0.00, 4.83, 7.07, 8.10, 8.59, 8.81, 8.91)
-
-m3 <- model_using_radar(
-  velocity = sprint_velocity,
-  time = sprint_time
-)
-```
-
-Both split and radar gun models allow the use of *weighted* non-linear regression.
-Weighting in regression is utilized when the observations have unequal error variance [@gelmanRegressionOtherStories2020], which can happen due to instrumental error (e.g., a multiplicative error instead of additive error) or due to biological variability (e.g., higher split time variance on shorter distances versus longer distances).
-In this case, observations with higher error variance get lower weight when fitting the model [@gelmanRegressionOtherStories2020].
-According to @gelmanRegressionOtherStories2020, unequal variances are not typically a major issue for the goal of estimating regression parameters, but they can become more important when making predictions about individual observations.
-Further exploring the topic of unequal error variances is beyond the scope of this paper, but the **shorts** package provides a weighting feature for the potential research of this topic in the future.
-
-Weighted non-linear regression is performed by setting `weights` parameter. For example, we can give more weight to shorter distances or faster velocities:
-
-```{r}
-m3_weighted <- model_using_radar(
-  velocity = sprint_velocity,
-  time = sprint_time,
-  weights = 1 / (sprint_velocity + 1)
-)
-```
-
-### Mixed-effects model
-
-Mixed-effects model using radar data is done using `mixed_model_using_radar()` function. To perform mixed model, let's load data that comes with **shorts** package.
-
-```{r}
-data("radar_gun_data")
-
-m4 <- mixed_model_using_radar(
-  radar_gun_data,
-  time = "time",
-  velocity = "velocity",
-  athlete = "athlete"
-)
-```
-
-## Force-Velocity profile
-
-To create *Force-Velocity Profile* (FVP) using single athlete estimated sprint model parameters (i.e., TAU and MSS), you can use `get_FV_profile()` function.
-When estimating FVP, athlete body mass (kg) can be set using `bodymass` parameter, while the air resistance parameters can be set using `"..."`, which are forwarded to the `get_air_resistance()` function.
-Details of the FVP method implemented in the **shorts** package, as well as the interpretation from a sprint training perspective, are covered elsewhere [@haugenPowerForceVelocityProfilingSprinting2020; @morinInterpretingPowerForceVelocityProfiles2016; @morinSimpleMethodComputing2019; @samozinoSimpleMethodMeasuring2016].
-
-```{r}
-# To create Force-Velocity Profile
-fvp <- get_FV_profile(
-  MSS = m1$parameters$MSS,
-  TAU = m1$parameters$TAU,
+  MAC = m1$parameters$MAC,
   bodymass = 80,
@@
 
-To plot FVP kinematics and kinetics (which are exactly the same as generated by the `predict_kinematics()` function), use S3 `plot()` function.
-By default, FVP estimated kinetics are plotted against velocity (on x-axis).
-
-```{r}
+To plot FVP kinematics and kinetics (which are the same as generated by the `predict_kinematics()` function), use the S3 `plot()` function.
+FVP estimated kinetics are default plotted against velocity (on the x-axis).
+
+```{r results='hide', fig.show='hide'}
 plot(fvp) + theme_bw(8)
@@
 
-There is a challenge when collecting sprint data that could have a substantial impact on modeled outcomes.
-To ensure accurate parameter outcomes, the initial force production must be synced with start time [@haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintMechanicalVariables2019].
-Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the **shorts** package.
-
-## Problems with time sync with radar gun
-
-One source of error in the modeled estimation using a radar gun is the time synchronization.
-In theory, synchronization is ideal when a sprint is initiated at $t=0$ (i.e., $v(t=0) = 0$).
+There is a challenge when collecting sprint data that could substantially impact modeled outcomes.
+To ensure accurate parameter outcomes, the initial force production must be synced with the start time [@haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintMechanicalVariables2019].
+Below we describe this challenge when using radar guns or timing gates and suggest potential solutions within the **{shorts}** package.
+
+## Problems with time sync with the radar gun
+
+Time synchronization is one error in the modeled estimation using a radar gun.
+In theory, synchronization is ideal when a sprint is initiated at $t=0$ $s$ (i.e., $v(t=0) = 0$ $ms^-1$).
 In practice, this is often not the case.
-Let's use our athlete and add and deduct 0.5 s to simulate an error in synchronization and its effect on estimated MSS and TAU.
-
-```{r}
-df <- tibble(
-  `true time` = sprint_time,
-  velocity = sprint_velocity,
-  `0.5s added` = `true time` + 0.5,
-  `0.5s deducted` = `true time` - 0.5
-)
-
-plot_df <- pivot_longer(df, cols = -2, names_to = "Sync issue")
-
-ggplot(
-  plot_df,
-  aes(x = value, y = velocity, color = `Sync issue`)
-) +
-  theme_bw(8) +
-  geom_line(alpha = 0.7) +
-  xlab("Time (s)") +
-  ylab(expression("Velocity (" * ms^-1 * ")")) +
-  theme(
-    legend.title = element_blank(),
-    legend.position = "top"
-  )
-```
-
-The following three models estimate MSS and TAU from the three datasets.
-Tables are created using the `kable()` function from the **knitr** package [@knitr2014; @knitr2015; @R-knitr] and `kable_styling()` function from the **kableExtra** package [@R-kableExtra].
-
-```{r}
-require(knitr)
-
-# Without synchronization issues
-m5 <- model_using_radar(
-  velocity = df$velocity,
-  time = df$`true time`
-)
-
-# With time added
-m6 <- model_using_radar(
-  velocity = df$velocity,
-  time = df$`0.5s added`
-)
-
-# With time deducted
-m7 <- model_using_radar(
-  velocity = df$velocity,
-  time = df$`0.5s deducted`
-)
-
-model_df <- rbind(
-  data.frame(
-    model = "True time",
-    t(coef(m5))
-  ),
-  data.frame(
-    model = "Added 0.5s time",
-    t(coef(m6))
-  ),
-  data.frame(
-    model = "Deducted 0.5s time",
-    t(coef(m7))
-  )
-)
-
-kable(model_df, digits = 2, booktabs = TRUE)
-```
-
-As can be seen from the example, all estimated parameters are affected by an error in synchronization of time with velocity (with MSS being the least affected in this example).
-The potential solution incorporated into the **shorts** package involves estimation of the *time correction* parameter using the following equation:
+This might happen when the measurement starts before the sprint starts.
+Although these data should be *trimmed*, the time variable might not be synchronized perfectly (i.e., sprint starts exactly at $t=0$ $s$) with the sprint initiation afterward.
+
+The potential solution incorporated into the **{shorts}** package involves estimating the *time correction* (TC) parameter using equation \@ref(eq:velocity-time-correction).
+TC parameter serves as *intercept*, similar to linear regression, and allows the model to be fitted correctly in the scenario explained.
 
 \begin{equation}
-  v(t) = MSS \times (1 - e^{-\frac{t + time \; correction}{TAU}}) (\#eq:velocity-time-correction)
+  v(t) = MSS \times (1 - e^{-\frac{t + TC}{TAU}}) (\#eq:velocity-time-correction)
 \end{equation}
 
-This model is incorporated in the `model_using_radar_with_time_correction()` function:
-
-```{r}
-# With time added
-m8 <- model_using_radar_with_time_correction(
-  velocity = df$velocity,
-  time = df$`0.5s added`
-)
-coef(m8)
-
-# With time deducted
-m9 <- model_using_radar_with_time_correction(
-  velocity = df$velocity,
-  time = df$`0.5s deducted`
-)
-coef(m9)
-```
-
-When using `predict_XXX()` family of functions, one can provide estimated time correction to get predictions at original time scale.
-
-```{r}
-# Using the true time
-predict_velocity_at_time(
-  time = df$`true time`,
-  MSS = m5$parameters$MSS,
-  TAU = m5$parameters$TAU
-)
-
-# Using time with sync issues
-predict_velocity_at_time(
-  time = df$`0.5s added`,
-  MSS = m8$parameters$MSS,
-  TAU = m8$parameters$TAU,
-  time_correction = m8$parameters$time_correction
-)
-```
-
-### Mixed-model approach
-
-When it comes to mixed-model approach, time correction can be modeled as a fixed effect or random effect using the `mixed_model_using_radar_with_time_correction()` function.
-
-```{r}
-# Adding 0.5s to radar_gun_data
-radar_gun_data$time <- radar_gun_data$time + 0.5
-
-# Mixed model with time correction being fixed effect
-m10 <- mixed_model_using_radar_with_time_correction(
-  radar_gun_data,
-  time = "time",
-  velocity = "velocity",
-  athlete = "athlete",
-  random = MSS + TAU ~ 1
-)
-
-# Mixed model with time correction being random effect
-m11 <- mixed_model_using_radar_with_time_correction(
-  radar_gun_data,
-  time = "time",
-  velocity = "velocity",
-  athlete = "athlete",
-  random = MSS + TAU + time_correction ~ 1
-)
-```
+This model is incorporated in the `model_radar_gun()` function (see [Estimating short sprint parameters using the radar gun]), and TC is estimated in addition to MSS, MAC, and TAU parameters.
 
@@
 
-Let's imagine we have two twin brothers with same short sprint characteristics: MSS equal to 9 $ms^{-1}$, TAU equal to 1.3, and MAC equal to 6.92 $ms^{-2}$.
-Let's call them John and Jack.
-They both perform 40m sprint using timing gates set at 5, 10, 20, 30, and 40 m.
-The initial timing gate at the start (i.e., $d=0$ m) serves to activate the timing system (i.e., when they cross the beam).
+Let us imagine we have two twin brothers with the same short sprint characteristics: MSS equal to 9 $ms^{-1}$ and MAC equal to 8 $ms^{-2}$.
+Let us call them John and Jack.
+They perform 40 $m$ sprints using timing gates set at 5, 10, 20, 30, and 40 $m$.
+The initial timing gate at the start (i.e., $d=0$ $m$) activates the timing system (i.e., when they cross the beam).
 
 John represents the *theoretical model*, in which we assume that the initial force production and the timing initiation are perfectly synchronized.
-Jack, on the other hand, represents a *practical model*, and decides to move slightly behind the initial timing gate (i.e.
-for 0.5 m) and use body rocking to initiate the sprint start.
-In other words, Jack is using a *flying start*, a common scenario when testing field sports athletes.
-Flying start distance is often recommended to avoid premature triggering of the timing system by lifted knees or swinging
-arms [@altmannAccuracySingleBeam2018; @altmannDifferentStartingDistances2015; @altmannValiditySingleBeamTiming2017; @haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintRunningPerformance2016].
-
-```{r}
-MSS <- 9
-TAU <- 1.3
-MAC <- MSS / TAU
-
+On the other hand, Jack represents a *practical model* and decides to move slightly behind the initial timing gate (i.e., for 0.5 $m$) and use body rocking to initiate the sprint start.
+In other words, Jack uses a *flying start*, a common scenario when testing field sports athletes.
+Flying start distance is often recommended to avoid premature timing system triggering by lifting knees or swinging arms [@altmannAccuracySingleBeam2018; @altmannDifferentStartingDistances2015; @altmannValiditySingleBeamTiming2017; @haugenPowerForceVelocityProfilingSprinting2020; @haugenSprintRunningPerformance2016].
+
+Data for this scenario is generated using `create_timing_gates_splits()` function:
+
+```{r}
 split_times <- tibble(
   distance = c(5, 10, 20, 30, 40),
-  john_time = predict_time_at_distance(distance, MSS, TAU),
-
-  # Jack's performance
-  jack_distance = distance + 0.5,
-  jack_true_time = predict_time_at_distance(jack_distance, MSS, TAU),
-  time_05m = predict_time_at_distance(0.5, MSS, TAU),
-  jack_time = jack_true_time - time_05m
-)
-```
-
-Here is a graphical representation of the sprint splits (please refer to the [Supplemental material] for the R code):
+  john_time = create_timing_gates_splits(
+    MSS = 9,
+    MAC = 8,
+    gates = distance
+  ),
+  jack_time = create_timing_gates_splits(
+    MSS = 9,
+    MAC = 8,
+    gates = distance,
+    FD = 0.5
+  )
+)
+```
+
+Here is a graphical representation of the sprint splits (please refer to the [Supplemental Material] for the R code):
 
@@
 
-Using the following code, we can see the differences in estimated MSS and TAU parameters:
-
-```{r}
-john_profile <- model_using_splits(
+We can see the differences in estimated MSS and MAC parameters using the following code. For better table output, we are going to use the `kable()`function from the **{knitr}** package [@R-knitr].
+
+```{r}
+john_profile <- model_timing_gates(
   distance = split_times$distance,
@@
 
-jack_profile <- model_using_splits(
+jack_profile <- model_timing_gates(
   distance = split_times$distance,
@@
 
-As can be seen from the results, a flying start yields biased estimates, particularly for the TAU, MAC and PMAX.
-
-To explore this further, we have run a simple simulation by increasing Jack's flying start distance from 0 to 1 m and depicting the estimated MSS, TAU, MAC, and PMAX parameters (please refer to the [Supplemental material] for the R code).
+As can be seen from the results, a flying start yields biased estimates, particularly for the TAU, MAC, and PMAX.
+
+### Simple simulation
+
+To explore this further, we have run a simple simulation by increasing Jack's flying start distance from 0 to 1 $m$ and depicting the estimated MSS, TAU, MAC, and PMAX parameters (please refer to the [Supplemental Material] for the R code).
 
@@
   MSS = 9,
-  MAC = 9 / 1.3,
-  flying_start_distance = seq(0.0001, 1, length.out = 1000),
+  MAC = 8,
+  flying_start_distance = seq(0, 1, length.out = 1000),
   distance = c(5, 10, 20, 30, 40)
@@
     PMAX = MSS * MAC / 4,
-    true_distance = distance + flying_start_distance,
-    true_time = predict_time_at_distance(true_distance, MSS, TAU),
-    stolen_time = predict_time_at_distance(
-      flying_start_distance, MSS, TAU
-    ),
-    time = true_time - stolen_time
+    time = create_timing_gates_splits(
+      MSS = MSS,
+      MAC = MAC,
+      gates = distance,
+      FD = flying_start_distance
+    )
   )
@@
 pred_wrapper <- function(data) {
-  model <- model_using_splits(
+  model <- model_timing_gates(
     distance = data$distance,
@@
 
-  predicted_time <- predict_time_at_distance(
-    distance = data$distance,
-    MSS = model$parameters$MSS,
-    TAU = model$parameters$TAU
-  )
+  predicted_time <- model$data$pred_time
 
   colnames(params) <- c(
-    "est_MSS", "est_TAU", "est_MAC", "est_PMAX",
-    "est_time_correction", "est_distance_correction"
+    "est_MSS", "est_TAU", "est_MAC", "est_PMAX"
   )
@@
 model_df <- sim_df %>%
-  group_by(MSS, TAU, flying_start_distance) %>%
+  group_by(MSS, MAC, flying_start_distance) %>%
   do(pred_wrapper(.)) %>%
@@
 plot_df <- model_df %>%
-  group_by(MSS, TAU, flying_start_distance) %>%
+  group_by(MSS, MAC, flying_start_distance) %>%
   slice(1) %>%
@@
     `MSS (m/s)` = est_MSS,
-    `TAU` = est_TAU,
+    `TAU (s)` = est_TAU,
     `MAC (m/s/s)` = est_MAC,
@@
   ) %>%
-  pivot_longer(cols = c("MSS (m/s)", "TAU", "MAC (m/s/s)", "PMAX (W)")) %>%
+  pivot_longer(cols = c("MSS (m/s)", "TAU (s)", "MAC (m/s/s)", "PMAX (W)")) %>%
   mutate(
-    name = factor(name, levels = c("MSS (m/s)", "TAU", "MAC (m/s/s)", "PMAX (W)"))
+    name = factor(name, levels = c("MSS (m/s)", "TAU (s)", "MAC (m/s/s)", "PMAX (W)"))
   )
@@
 
-As can be seen from the figure, MSS is underestimated as flying start distance increases.
-MAC (and also TAU) are highly affected by the flying start distance, and from the figure we can notice that MAC is overestimated as flying start distance increases.
-Estimated PMAX is also overestimated as flying start distance increases.
+As can be seen from the figure, MSS and TAU are underestimated, while MAC and PMAX are overestimated as the flying start distance increases.
 
 Model residuals are also affected by the flying start distance.
-The shape of residuals distribution depends on number and splits utilized (e.g., 10, 20, 30, 40 m versus 5, 15, 30 m), but here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simple simulation:
+The shape of the distribution of the residuals depends on the number and splits utilized (e.g., 10, 20, 30, 40 $m$ versus 5, 15, 30 $m$).
+However, here we can see the effect of the flying start distance on the model residuals per split distance utilized in our simple simulation:
 
@@
 
-Another way to visualize the effect of the flying start distance on split distance residuals can be found on the next figure:
+Another way to visualize the effect of the flying start distance on split distance residuals can be found in the following figure:
 
@@
 
-Clearly, any type of flying start where there is a difference between initial force production and start time can result in biased parameters and predictions.
-Since maximal sprint speed is difficult to improve, the effects of start inconsistencies can mask effects of the training intervention.
-It is thus crucial to standardize the start when testing and implementing the following techniques when using the **shorts** package.
+Any flying start with a difference between initial force production and start time can result in biased parameters and predictions.
+Since maximal sprint speed is difficult to improve, the effects of start inconsistencies can mask the effects of the training intervention.
+It is thus crucial to standardize the start when testing and implementing the following techniques when using the **{shorts}** package.
 
@@
 
-A potential solution is to use a correction factor - the recommendation in the literature is +0.5 s [@haugenSprintMechanicalProperties2020; @haugenSprintMechanicalVariables2019].
-Interestingly, the average difference between using timing gates and a block start for 40 m sprint time was 0.27 s [@haugenDifferenceStartImpact2012].
+A potential solution is to use a correction factor - the recommendation in the literature is +0.5 $s$ [@haugenSprintMechanicalProperties2020; @haugenSprintMechanicalVariables2019].
+Interestingly, the average difference between timing gates and a block start for 40 $m$ sprint time was 0.27 $s$ [@haugenDifferenceStartImpact2012].
 So, while a timing correction factor is warranted to avoid subsequent errors in estimates of kinetic variables (e.g., overestimate power), a correction factor that is too large will have the opposite effect (e.g., underestimate power).
 
-Rather than providing *apriori* time correction from the literature, **shorts** package provides an estimation of this parameter from the data provided, together with MSS and TAU.
-Exactly the same method is suggested by @stenrothForcevelocityProfilingIce2020, named *time shift method*, and the estimated parameter named *time shift parameter*.
-We have named this parameter *time correction* to be in agreement with the parameter introduced in [Problems with time sync with radar gun] section of this paper, as well as the available literature.
+Rather than providing *apriori* time correction from the literature, the **{shorts}** package provides an estimation of this parameter from the data provided, together with MSS and MAC.
+The same method is suggested by @stenrothForcevelocityProfilingIce2020, named the *time shift method*, and the estimated parameter is named the *time shift parameter*.
+We have named this parameter *time correction* (TC) to agree with the parameter introduced in the [Problems with time sync with the radar gun] section of this paper and the available literature.
+The model that implements the TC parameter is termed the *estimated TC model*.
 
@@
 \begin{equation}
-  t(d) = TAU \times W(-e^{\frac{-d}{MSS \times TAU}} - 1) + \frac{d}{MSS} + TAU - time \; correction (\#eq:time-correction)
+  t(d) = TAU \times W(-e^{\frac{-d}{MSS \times TAU}} - 1) + \frac{d}{MSS} + TAU - TC (\#eq:time-correction)
 \end{equation}
 
-To estimate time correction parameter, we use `model_using_splits_with_time_correction()` function.
-Here is how we can estimate Jack parameters using either provided time correction (e.g., +0.3 and +0.5 s) or estimated time correction:
-
-```{r}
-jack_profile_fixed_time_short <- model_using_splits(
+In **{shorts}** package, TC model is implemented in the `model_timing_gates_TC()` function.
+Here is how we can estimate Jack parameters using fixed time corrections (e.g., +0.3 and +0.5 $s$) or using estimated TC model:
+
+```{r}
+jack_profile_fixed_time_short <- model_timing_gates(
   distance = split_times$distance,
-  time = split_times$jack_time,
-  time_correction = 0.3
-)
-
-jack_profile_fixed_time_long <- model_using_splits(
+  time = split_times$jack_time + 0.3
+)
+
+jack_profile_fixed_time_long <- model_timing_gates(
   distance = split_times$distance,
-  time = split_times$jack_time,
-  time_correction = 0.5
-)
-
-jack_profile_time_estimated <- model_using_splits_with_time_correction(
+  time = split_times$jack_time + 0.5
+)
+
+jack_profile_time_estimated <- model_timing_gates_TC(
   distance = split_times$distance,
@@
 
-jack_parameters <- rbind(
-  coef(john_profile),
-  coef(jack_profile),
-  coef(jack_profile_fixed_time_short),
-  coef(jack_profile_fixed_time_long),
-  coef(jack_profile_time_estimated)
-)
-
-rownames(jack_parameters) <- c(
-  "John",
-  "Jack - No corrections",
-  "Jack - Fixed time correction (+0.3s)",
-  "Jack - Fixed time correction (+0.5s)",
-  "Jack - Estimated time correction"
+jack_parameters <- full_join(
+  rbind(
+    data.frame(athlete = "John", t(coef(john_profile))),
+    data.frame(athlete = "Jack - No corrections", t(coef(jack_profile))),
+    data.frame(athlete = "Jack - Fixed TC (+0.3s)", t(coef(jack_profile_fixed_time_short))),
+    data.frame(athlete = "Jack - Fixed TC (+0.5s)", t(coef(jack_profile_fixed_time_long)))
+  ),
+  data.frame(athlete = "Jack - Estimated TC", t(coef(jack_profile_time_estimated)))
 )
@@
 
-In Jack's case, both +0.3 s fixed time correction and time correction estimation yield parameters closer to John's (i.e.
-true parameters).
-We have used these two model definitions in retrospective pilot study [@vescoviSprintMechanicalCharacteristics2021], demonstrating statistically significant differences in estimated FVP parameters.
-
-Instead of using time correction as simple intercept in the equation \@ref(eq:time-correction), we can estimate the *distance correction* (i.e., the flying start distance) using the following equation:
+In Jack's case, both +0.3 $s$ fixed time correction and estimated TC model yield parameters closer to John's (i.e., *true* parameters).
+We have used these models in a retrospective pilot study [@vescoviSprintMechanicalCharacteristics2021], demonstrating statistically significant differences in estimated FVP parameters.
+
+Instead of using time correction as a simple intercept in equation \@ref(eq:time-correction), we can estimate the *flying distance* (FD) using the following equation:
 
@@
   \begin{split}
-   t(d) &= (TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU) \\ 
-   &\quad-(TAU \times W(-e^{\frac{distance \; correction}{MSS \times TAU}} - 1) + \frac{distance \; correction}{MSS} + TAU) 
+   t(d) &= (TAU \times W(-e^{\frac{-d + FD}{MSS \times TAU}} - 1) + \frac{d + FD}{MSS} + TAU) \\ 
+   &\quad-(TAU \times W(-e^{\frac{FD}{MSS \times TAU}} - 1) + \frac{FD}{MSS} + TAU) 
    \end{split}
@@
 
-Equation \@ref(eq:distance-correction) was used to generate the data for John and Jack (please refer to the provided R code), as well as for the simple simulation, and can be used as model definition to control for the flying start distance.
-
-To estimate distance correction parameter, we use `model_using_splits_with_distance_correction()` function. Here is how we can estimate Jack parameters using distance correction:
-
-```{r}
-jack_profile_distance_correction <- model_using_splits_with_distance_correction(
+We have named this model *estimated FD model*, and in the **{shorts}** package it is implemented in the `model_timing_gates_FD()` function.
+Here is how we can estimate Jack parameters using estimated FD model:
+
+```{r}
+jack_profile_FD <- model_timing_gates_FD(
   distance = split_times$distance,
@@
 
-jack_parameters <- rbind(
+jack_parameters <- full_join(
   jack_parameters,
-  coef(jack_profile_distance_correction)
-)
-
-rownames(jack_parameters) [6] <- "Jack - Estimated distance correction"
+  data.frame(athlete = "Jack - Estimated FD", t(coef(jack_profile_FD)))
+)
 
@@
 
-Another model definition, which also represents a novel approach implemented in the **shorts** package, is to utilize both time and distance corrections.
-Thus, equation \@ref(eq:time-correction) becomes:
-
-\begin{equation}
-  t(d) = TAU \times W(-e^{\frac{-d + distance \; correction}{MSS \times TAU}} - 1) + \frac{d + distance \; correction}{MSS} + TAU - time \; correction (\#eq:time-distance-correction)
-\end{equation}
-
-This model is implemented in `model_using_splits_with_corrections()` function. Below are the model estimates:
-
-```{r}
-jack_profile_time_distance_correction <- model_using_splits_with_corrections(
-  distance = split_times$distance,
-  time = split_times$jack_time
-)
-
-jack_parameters <- rbind(
-  jack_parameters,
-  coef(jack_profile_time_distance_correction)
-)
-
-rownames(jack_parameters) [7] <- 
-  "Jack - Estimated time & distance corrections"
-
-kable(jack_parameters, digits = 2, booktabs = TRUE)
-```
-
-As can be seen from the results, adding distance correction results in correctly estimating Jack's sprint parameters.
+As can be seen from the results, the estimated FD model correctly estimated Jack's sprint parameters.
 There are a few issues with this model definition.
-Besides being novel and still not validated with actual data, distance correction model has four parameters to estimate, which implies that at least four sprint splits are needed.
-This imposes practical limitations, since acquiring five timing gates (one for the start and four for splits) might be practically troublesome.
-One strategy that is sometimes implemented is adding zeros to the sample (i.e., $t=0$ and $d=0$), which increase the number of observations.
-Unfortunately, this strategy should not be implemented since it will bias the estimated parameters:
-
-```{r}
-jack_profile_zero <- model_using_splits(
-  distance = c(0, split_times$distance),
-  time = c(0, split_times$jack_time)
-)
-
-jack_profile_time_estimated_zero <- model_using_splits_with_time_correction(
-  distance = c(0, split_times$distance),
-  time = c(0, split_times$jack_time)
-)
-
-adding_zero_issue <- rbind(
-  coef(jack_profile),
-  coef(jack_profile_zero),
-  coef(jack_profile_time_estimated),
-  coef(jack_profile_time_estimated_zero)
-)
-
-rownames(adding_zero_issue) <- c(
-  "Jack - No corrections",
-  "Jack - No corrections with zeros added",
-  "Jack - Estimated time correction",
-  "Jack profile - Estimated time correction with added zeros")
-
-kable(adding_zero_issue, digits = 2, booktabs = TRUE)
-```
-
-Adding zeros to the sample nullifies the potential benefits of using time correction model and should be avoided in practice.
-
-It is important to note that using `predict_XXX()` family of functions with distance correction model and using estimated distance correction parameter will yield wrong predictions on the original scale (but not on the *theoretical* scale in which force production starts at d=0m and t=0s).
-This is because `predict_XXX()` family of functions utilize equation \@ref(eq:time-distance-correction) and its derivates and integrals, which works for the uncorrected model (time and distance correction parameters are equal to zero), time correction models (distance correction parameter is equal to zero), and time and distance corrections model.
-You can notice that equations \@ref(eq:time-correction) and \@ref(eq:time-distance-correction) are equivalent to the \@ref(eq:time-distance) equation when time and distance correction parameters are equal to zero.
-Unfortunately, this doesn't work for the distance correction model when the estimated distance correction parameter is used since model definition outlined in the equation \@ref(eq:distance-correction) is different that the equation \@ref(eq:time-distance-correction) when the estimated distance correction parameter is different than zero.
-The solution is to use generic S3 `predict()` function, but this only predicts split times at split distances.
-
-```{r}
-# Jack's observed times
-split_times$jack_time
-
-# To predict at "theoretical" distance scale (where force application starts at d=0m)
-predict_time_at_distance(
-  distance = split_times$distance, 
-  MSS = jack_profile_time_estimated$parameters$MSS,
-  TAU = jack_profile_time_estimated$parameters$TAU)
-
-predict_time_at_distance(
-  distance = split_times$distance, 
-  MSS = jack_profile_distance_correction$parameters$MSS,
-  TAU = jack_profile_distance_correction$parameters$TAU)
-
-# To predict at original distance scale, use all estimated parameters
-predict_time_at_distance(
-  distance = split_times$distance, 
-  MSS = jack_profile_time_estimated$parameters$MSS,
-  TAU = jack_profile_time_estimated$parameters$TAU,
-  time_correction = jack_profile_time_estimated$parameters$time_correction)
-
-# Unfortunately, this doesn't work for the distance correction model
-predict_time_at_distance(
-  distance = split_times$distance, 
-  MSS = jack_profile_distance_correction$parameters$MSS,
-  TAU = jack_profile_distance_correction$parameters$TAU,
-  distance_correction = jack_profile_distance_correction$parameters$distance_correction)
-
-# Instead use generic S3 predict() function
-predict(jack_profile_distance_correction)
-
-predict(
-  jack_profile_distance_correction$model,
-  newdata = data.frame(distance = split_times$distance))
-```
-
-In the next section we examine how these models (no correction, fixed time correction, estimated time correction, estimated distance correction, and estimated time and distance correction) perform using simulated data with varying flying start distance (please refer to the [Supplemental material] for the R code).
+Besides being novel and still not validated with actual data, the estimated FD model has three parameters to estimate, which implies that at least four sprint splits are needed.
+This imposes practical limitations since acquiring five timing gates (one for a start and four for splits) might be practically troublesome (which is also the case with the estimated TC model).
+In addition, the estimated FD model can be ill-defined in scenarios involving reaction time and no actual flying sprint involved.
+This is often the case when a gunshot is used to initiate the timing system.
+
+### Simple simulation using proposed models
+
+In this section, we examine how these models (no correction, fixed time correction, estimated TC, and estimated FD) perform using simulated data with varying flying start distances (please refer to the [Supplemental Material] for the R code), assuming *true* MSS equal to 9 $ms^{-1}$, and MAC equal to 8 $ms^{-2}$ (i.e., sprint characteristics of Jack and John).
+Timing gates are set at 5, 10, 20, 30, and 40 $m$.
 
@@
 pred_wrapper <- function(data) {
-
   merge_model <- function(data, model, model_name) {
@@
     data$est_PMAX <- model$parameters$PMAX
-    data$est_time_correction <- model$parameters$time_correction
-    data$est_distance_correction <- model$parameters$distance_correction
     data$predicted_time <- model$data$pred_time
-    data$residuals <- data$time - data$predicted_time
+    data$residuals <- model$data$time - data$predicted_time
 
@@
 
-  no_correction <- model_using_splits(
+  no_correction <- model_timing_gates(
     distance = data$distance,
@@
 
-  fixed_correction_short <- model_using_splits(
+  fixed_correction_short <- model_timing_gates(
     distance = data$distance,
-    time = data$time,
-    time_correction = 0.3
-  )
-
-  fixed_correction_long <- model_using_splits(
+    time = data$time + 0.3
+  )
+
+  fixed_correction_long <- model_timing_gates(
     distance = data$distance,
-    time = data$time,
-    time_correction = 0.5
-  )
-
-  time_correction <- model_using_splits_with_time_correction(
+    time = data$time + 0.5
+  )
+
+  time_correction <- model_timing_gates_TC(
     distance = data$distance,
@@
 
-  distance_correction <- model_using_splits_with_distance_correction(
-    distance = data$distance,
-    time = data$time
-  )
-
-  time_dist_correction <- model_using_splits_with_corrections(
+  distance_correction <- model_timing_gates_FD(
     distance = data$distance,
@@
       fixed_correction_short,
-      "Fixed correction +0.3s"
+      "Fixed TC +0.3s"
     ),
@@
       fixed_correction_long,
-      "Fixed correction +0.5s"
+      "Fixed TC +0.5s"
     ),
@@
       time_correction,
-      "Time correction"
+      "Estimated TC"
     ),
@@
       distance_correction,
-      "Distance correction"
-    ),
-    merge_model(
-      data, 
-      time_dist_correction,
-      "Time and distance correction"
+      "Estimated FD"
     )
@@
   filter(flying_start_distance > 0) %>%
-  group_by(MSS, TAU, flying_start_distance) %>%
+  group_by(MSS, MAC, flying_start_distance) %>%
   do(pred_wrapper(.)) %>%
@@
 
-As can be seen from the next figure, the estimated time correction model estimates MSS almost perfectly, while the estimated distance and estimate time and distance correction models estimates MSS perfectly.
+As can be seen from the following figure, the estimated TC model estimates sprint parameters with increasing bias as the flying distance increases.
+In contrast, the estimated FD model estimates sprint parameters perfectly.
+If correct TC is not applied, fixed time correction models have a much larger bias.
+This value depends on sprint characteristics, the flying distance, and timing gate splits.
+It is thus impossible to provide a single fixed time correction value that will serve its corrective purpose across different scenarios and athletes measured.
 
@@
     "No correction",
-    "Fixed correction +0.3s",
-    "Fixed correction +0.5s",
-    "Time correction",
-    "Distance correction",
-    "Time and distance correction"
+    "Fixed TC +0.3s",
+    "Fixed TC +0.5s",
+    "Estimated TC",
+    "Estimated FD"
   )
@@
 plot_df <- model_df %>%
-  group_by(model, MSS, TAU, flying_start_distance) %>%
+  group_by(model, MSS, MAC, flying_start_distance) %>%
   slice(1) %>%
@@
 
-Similar outcomes are observed for the MAC parameter.
-The distance correction model and time and distance corrections models performs perfectly, while the time correction model performs almost as good.
-PMAX demonstrates the same properties as MSS and MAC.
-
-The next figure depicts residuals across split distances for each simulated flying start distance.
-As can be seen, time correction, distance correction, and time and distance correction models perform much better than the other models.
+The following figure depicts residuals across split distances for each simulated flying start distance.
+As can be seen, estimated TC and FD models perform much better than the no correction and fixed time correction models.
 
@@
 
-The outcomes from this simple simulation data clearly demonstrates that the time correction, distance correction, and time and distance correction models represent sound improvements in parameter estimation and model fit compared to no corrections model and fixed correction model when attempting to overcome the flying start issues.
-Since the time correction and distance correction models are simpler and requires three parameters to be estimated, they might be practically more useful than the time and distance correction model, which requires four parameters estimation and thus more than four timing gates and sprint splits.
-More elaborate simulation study is currently ongoing and the results will be reported in another paper.
-
-Time correction, distance correction, and time and distance corrections are also implemented in the mixed-models using `mixed_model_using_splits_with_time_correction()`, `mixed_model_using_splits_with_distance_correction()`, and `mixed_model_using_splits_with_corrections()` functions.
-We will showcase their use at the end of this paper.
-
-# Leave-one-out Cross-Validation
-
-To estimate parameter stability, model over-fitting, and performance on unseen data, **shorts** model function comes with implemented *leave-one-out cross validation* (LOOCV) [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
-LOOCV involves a simple, yet powerful procedure, of removing each observation, rebuilding the model, and making predictions for that removed observation.
+The results of this simple simulation demonstrate that the estimated TC and FD models represent sound improvements in parameter estimation and model fit compared to the no corrections model and fixed TC models when attempting to overcome the flying start issues.
+A more detailed simulation study is ongoing, and the results will be reported in another paper.
+
+# Cross-Validation
+
+To estimate parameter stability, model over-fitting, and performance on unseen data, **{shorts}** model functions come with implemented *leave-one-out cross-validation* (LOOCV) for the timing gates models and *n-fold cross-validation* (CV) for the radar gun models [@jamesIntroductionStatisticalLearning2017; @jovanovicBmbstatsBootstrapMagnitudebased2020; @kuhnAppliedPredictiveModeling2018].
+LOOCV involves a simple yet powerful procedure of removing each observation, rebuilding the model, and making predictions for that removed observation.
 This process is repeated for each observation in the model dataset.
-LOOCV allows one to check estimated parameters stability, and model performance on unseen data.
-
-Let's perform LOOCV using Jack's data and the time correction model:
-
-```{r}
-jack_LOOCV <- model_using_splits_with_time_correction(
+LOOCV allows one to check estimated parameters' stability and model performance on unseen data.
+
+Let's perform LOOCV using Jack's data and the estimated TC model:
+
+```{r}
+jack_LOOCV <- model_timing_gates_TC(
   distance = split_times$distance,
@@
 
-Next we plot estimated parameters across LOOCV folds (please refer to the [Supplemental material] for the R code):
+Next, we plot estimated parameters across LOOCV folds (please refer to the [Supplemental Material] for the R code):
 
 ```{r echo=FALSE}
-df <- jack_LOOCV$LOOCV$parameters
-
-df <- pivot_longer(df, cols = 1:6, names_to = "parameter")
+df <- jack_LOOCV$CV$parameters
+
+df <- pivot_longer(df, cols = 1:5, names_to = "parameter")
 
@@
     "PMAX",
-    "time_correction",
-    "distance_correction"
+    "FD"
   )
@@
   pred_time = jack_LOOCV$data$pred_time,
-  LOOCV_time = jack_LOOCV$LOOCV$data$pred_time
+  LOOCV_time = jack_LOOCV$CV$data$pred_time
 )
@@
 
-As expected, the model has more issues predicting unseen split times for both short or long distances.
-Please note, that since LOOCV removes one observation, if the model estimates three parameters, then at least four observations are needed, since we need to make sure the model can be estimated once a single observation is removed.
-LOOCV can also be implemented with the mixed-effects models in the **shorts** package.
+As expected, the model has more issues predicting unseen split times for short or long distances.
+Please note that since LOOCV removes one observation, if the model estimates three parameters, then at least five observations are needed since we need to ensure the model can be estimated once a single observation is removed.
+
+Since there are much more observations in the radar gun data, n-fold CV is implemented instead of LOOCV.
+CV procedure randomly splits the dataset into n-folds, leave one fold-out for prediction, and makes a model using the remaining folds.
+This is then repeated for all the folds.
+
+```{r}
+# Estimate John's MSS, TAU, MAC, and PMAX
+m2_CV <- model_radar_gun(
+  velocity = john_RG_data$velocity,
+  time = john_RG_data$time,
+  # Eneter number of folds
+  CV = 10
+)
+
+# Estimated parameters for each fold
+m2_CV$CV$parameters
+```
 
@@
 
-Let's utilize demonstrated functionalities of the **shorts** package using real-world data.
-The first dataset comes from Usain Bolt's performance from IAAF World Championship held in London, 2017, and the second dataset involve Jason Vescovi's sample data for 52 female soccer and field hockey athletes which comes with the **shorts** package (see the dataset documentation by executing `?vescovi` command in the R console).
-
-## Usain Bolt's run from London 2017
-
-The following dataset represents Usain Bolt's race in the finals at the IAAF World Championship held in London, 2017.
-Since reaction time enters the splits, we want to see how that will affect the model estimates, and particularly, if the estimated time correction model will pick-up reaction time.
-
-For the sake of this analysis, only 10 m splits over 60 m race distance are used.
+Let us use real-world data to demonstrate the functionalities of the **{shorts}** package. We will use the dataset from Usain Bolt's performance at the IAAF World Championship finals in London, 2017.
+
+Since reaction time enters the splits, we want to see how that will affect the model estimates, mainly if the estimated time correction model will pick up reaction time.
+
+For the sake of this analysis, only 10 $m$ splits over 60 $m$ race distance are used.
 
@@
 # No corrections model
-bolt_m1 <- model_using_splits(
+bolt_m1 <- model_timing_gates(
   distance = bolt_distance,
@@
 # Model with reaction time as fixed time correction
-bolt_m2 <- model_using_splits(
+bolt_m2 <- model_timing_gates(
   distance = bolt_distance,
-  time = bolt_time,
-  time_correction = -bolt_reaction_time
+  time = bolt_time - bolt_reaction_time
 )
@@
 # Model with estimated time correction
-bolt_m3 <- model_using_splits_with_time_correction(
+bolt_m3 <- model_timing_gates_TC(
   distance = bolt_distance,
@@
 
-# Model with estimated time correction, but deducted reaction time
-bolt_m4 <- model_using_splits_with_time_correction(
-  distance = bolt_distance,
-  time = bolt_time - bolt_reaction_time
-)
-
-# Model with estimated time and distance corrections
-bolt_m5 <- model_using_splits_with_corrections(
-  distance = bolt_distance,
-  time = bolt_time
-)
-
-# Model with estimated time and distance corrections and
-#  deducted reaction time
-bolt_m6 <- model_using_splits_with_corrections(
-  distance = bolt_distance,
-  time = bolt_time - bolt_reaction_time
-)
-
-bolt_model <- rbind(
-  data.frame(
-    model = "No correction",
-    t(coef(bolt_m1))
+# Model with flying distance correction
+# THIS CANNOT BE ESTIMATED SINCE IT IS ILL-DEFINED MODEL
+# bolt_m4 <- model_timing_gates_FD(
+#   distance = bolt_distance,
+#   time = bolt_time
+# )
+
+bolt_model <- full_join(
+  rbind(
+    data.frame(model = "No correction", t(coef(bolt_m1))),
+    data.frame(model = "No correction - RT", t(coef(bolt_m2)))
   ),
-  data.frame(
-    model = "No correction - RT",
-    t(coef(bolt_m2))
-  ),
-  data.frame(
-    model = "Time correction",
-    t(coef(bolt_m3))
-  ),
-  data.frame(
-    model = "Time correction - RT",
-    t(coef(bolt_m4))
-  ),
-  data.frame(
-    model = "Distance correction",
-    t(coef(bolt_m5))
-  ),
-  data.frame(
-    model = "Distance correction - RT",
-    t(coef(bolt_m6))
-  )
+  data.frame(model = "Estimated TC", t(coef(bolt_m3)))
 )
@@
 Here is the model estimate of the time and distance it takes for Bolt to reach 99% of MSS.
-Please note that we are not using distance and time correction parameters, since we want these estimates to be on the time/distance scale aligned with the actual sprint start (i.e., theoretical scale), not the measurement scale.
+These estimated times and distances represent *true* time and distance when the sprint is initiated at $t=0$ $s$ and $d=0$ $m$.
 
@@
   mutate(
-    dist_99_MSS = find_velocity_critical_distance(
-      MSS = MSS, TAU = TAU,
-      # time_correction = time_correction,
-      # distance_correction = distance_correction,
+    `99% MSS (m)` = find_velocity_critical_distance(
+      MSS = MSS, MAC = MAC,
       percent = 0.99
     ),
-    time_99_MSS = find_velocity_critical_time(
-      MSS = MSS, TAU = TAU,
-      # time_correction = time_correction,
+    `99% MSS (s)` = find_velocity_critical_time(
+      MSS = MSS, MAC = MAC,
       percent = 0.99
@@
 
-kable(bolt_model[c(1, 8, 9)], digits = 2, booktabs = TRUE)
-```
-
-## Vescovi data
-
-The data from Vescovi represents a sub-set of data from a total of 220 high-level female athletes (151 soccer players and 69 field hockey players) [@vescoviSprintMechanicalCharacteristics2021].
-Using a random number generator, a total of 52 players (35 soccer and 17 field hockey) were selected for the sample dataset.
-
-The protocol for assessing linear sprint speed has been described previously [@vescoviImpactMaximumSpeed2014; @vescoviLocomotorHeartRateMetabolic2016; @vescoviSprintSpeedCharacteristics2012] and was identical for each cohort.
-Briefly, all athletes performed a standardized warm-up that included general exercises such as jogging, shuffling, multi-directional movements, and dynamic stretching exercises.
-Infrared timing gates (Brower Timing, Utah) were positioned at the start line and at 5, 10, 20, 30, and 35 m at a height of approximately 1.0 m.
-Participants stood with their lead foot positioned approximately 5 cm behind the initial infrared beam (i.e., start line).
-Only forward movement was permitted (no leaning or rocking backwards) and timing started when the laser of the starting gate was triggered.
-The best 35 m time, and all associated split times were kept for analysis.
-
-Below is the mixed-effects models analysis of the dataset.
-
-```{r}
-data("vescovi")
-
-# Convert data to long
-df <- vescovi %>%
-  select(1:13) %>%
-  # slice(1:10) %>%
-  pivot_longer(
-    cols = 9:13,
-    names_to = "distance",
-    values_to = "time"
-  ) %>%
-  mutate(
-    distance = as.numeric(str_extract(distance, "^[0-9]+"))
-  )
-```
-
-The following models were used: (1) no corrections model, (2) fixed time correction model (using +0.3s heuristic rule of thumb), (3) time correction model with time correction as fixed effect, (4) time correction model with time correction as random effect, (5) distance correction model with distance correction as fixed effect, (6) distance correction model with distance correction as random effect, (7) time and distance correction model, (8) time and distance correction model with distance correction as fixed effect (and time correction as random effect), and (9) time and distance correction model with distance and time correction as random effects (please refer to the [Supplemental material] for the R code).
-
-```{r echo=FALSE}
-no_corrections <- mixed_model_using_splits(
-  df,
-  distance = "distance",
-  time = "time",
-  athlete = "Athlete"
-)
-
-fixed_correction <- mixed_model_using_splits(
-  df,
-  distance = "distance",
-  time = "time",
-  athlete = "Athlete",
-  time_correction = 0.3
-)
-
-time_correction_fixed <-
-  mixed_model_using_splits_with_time_correction(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU ~ 1
-  )
-
-time_correction_random <-
-  mixed_model_using_splits_with_time_correction(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU + time_correction ~ 1
-  )
-
-distance_correction_fixed <-
-  mixed_model_using_splits_with_distance_correction(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU ~ 1
-  )
-
-distance_correction_random <-
-  mixed_model_using_splits_with_distance_correction(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU + distance_correction ~ 1, 
-    control = nlme::nlmeControl(tolerance = 0.01)
-  )
-
-time_distance_correction_fixed <-
-  mixed_model_using_splits_with_corrections(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU ~ 1
-  )
-
-time_distance_correction_time_random <-
-  mixed_model_using_splits_with_corrections(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU + time_correction ~ 1
-  )
-
-time_distance_correction_random <-
-  mixed_model_using_splits_with_corrections(
-    df,
-    distance = "distance",
-    time = "time",
-    athlete = "Athlete",
-    random = MSS + TAU + time_correction + distance_correction ~ 1
-  )
-```
-
-The following image represents model fit estimator RSE for each model. As can be seen, RSE is reduced the more flexible the model.
-
-```{r echo=FALSE}
-model_fit <- rbind(
-  data.frame(
-    model = "No corrections",
-    t(unlist(no_corrections$model_fit))
-  ),
-  data.frame(
-    model = "Fixed correction +0.3s",
-    t(unlist(fixed_correction$model_fit))
-  ),
-  data.frame(
-    model = "Time correction fixed",
-    t(unlist(time_correction_fixed$model_fit))
-  ),
-  data.frame(
-    model = "Time correction random",
-    t(unlist(time_correction_random$model_fit))
-  ),
-  data.frame(
-    model = "Distance correction fixed",
-    t(unlist(distance_correction_fixed$model_fit))
-  ),
-  data.frame(
-    model = "Distance correction random",
-    t(unlist(distance_correction_random$model_fit))
-  ),
-  data.frame(
-    model = "Time and distance correction fixed",
-    t(unlist(time_distance_correction_fixed$model_fit))
-  ),
-  data.frame(
-    model = "Time and distance correction (time correction random)",
-    t(unlist(time_distance_correction_time_random$model_fit))
-  ),
-  data.frame(
-    model = "Time and distance correction random",
-    t(unlist(time_distance_correction_random$model_fit))
-  )
-)
-
-model_fit$model <- factor(
-  model_fit$model,
-  levels = rev(c(
-    "No corrections",
-    "Fixed correction +0.3s",
-    "Time correction fixed",
-    "Time correction random",
-    "Distance correction fixed",
-    "Distance correction random",
-    "Time and distance correction fixed",
-    "Time and distance correction (time correction random)",
-    "Time and distance correction random"
-  ))
-)
-
-ggplot(model_fit, aes(x = RSE, y = model)) +
-  theme_bw(8) +
-  geom_point() +
-  xlab("RSE (s)") +
-  ylab(NULL)
-```
-
-The following image depicts estimated parameters for each model:
-
-```{r echo=FALSE}
-est_params <- rbind(
-  data.frame(
-    model = "No corrections",
-    no_corrections$parameters$random
-  ),
-  data.frame(
-    model = "Fixed correction +0.3s",
-    fixed_correction$parameters$random
-  ),
-  data.frame(
-    model = "Time correction fixed",
-    time_correction_fixed$parameters$random
-  ),
-  data.frame(
-    model = "Time correction random",
-    time_correction_random$parameters$random
-  ),
-  data.frame(
-    model = "Distance correction fixed",
-    distance_correction_fixed$parameters$random
-  ),
-  data.frame(
-    model = "Distance correction random",
-    distance_correction_random$parameters$random
-  ),
-  data.frame(
-    model = "Time and distance correction fixed",
-    time_distance_correction_fixed$parameters$random
-  ),
-  data.frame(
-    model = "Time and distance correction (time correction random)",
-    time_distance_correction_time_random$parameters$random
-  ),
-  data.frame(
-    model = "Time and distance correction random",
-    time_distance_correction_random$parameters$random
-  )
-)
-
-est_params$model <- factor(
-  est_params$model,
-  levels = rev(c(
-    "No corrections",
-    "Fixed correction +0.3s",
-    "Time correction fixed",
-    "Time correction random",
-    "Distance correction fixed",
-    "Distance correction random",
-    "Time and distance correction fixed",
-    "Time and distance correction (time correction random)",
-    "Time and distance correction random"
-  ))
-)
-
-est_params <- est_params %>%
-  pivot_longer(cols = -(1:2), names_to = "parameter")
-
-est_params$parameter <- factor(
-  est_params$parameter,
-  levels = c(
-    "MSS",
-    "TAU",
-    "MAC",
-    "PMAX",
-    "time_correction",
-    "distance_correction"
-  )
-)
-
-ggplot(est_params, aes(y = model, x = value)) +
-  theme_bw(8) +
-  geom_boxplot() +
-  facet_wrap(~parameter, scales = "free_x") +
-  xlab(NULL) +
-  ylab(NULL)
-```
-
-The following image depicts model residuals across distance splits.
-To provide practical magnitude of the residuals, we have used between subject observed time SD multiplied with 0.2 and -0.2.
-This provides practical anchor for the residual magnitude, often referred to as *smallest worthwhile change* (SWC) or *smallest effect size of interest* (SESOI) [@jovanovicBmbstatsBootstrapMagnitudebased2020].
-If the residuals are within this magnitude band, then the model is good in making practically useful predictions.
-Error bars represent residual bias $\pm$ 1 SD.
-
-```{r echo=FALSE}
-model_resid <- rbind(
-  data.frame(
-    model = "No corrections",
-    no_corrections$data
-  ),
-  data.frame(
-    model = "Fixed correction +0.3s",
-    fixed_correction$data
-  ),
-  data.frame(
-    model = "Time correction fixed",
-    time_correction_fixed$data
-  ),
-  data.frame(
-    model = "Time correction random",
-    time_correction_random$data
-  ),
-  data.frame(
-    model = "Time and distance correction fixed",
-    time_distance_correction_fixed$data
-  ),
-  data.frame(
-    model = "Time and distance correction (time correction random)",
-    time_distance_correction_time_random$data
-  ),
-  data.frame(
-    model = "Time and distance correction random",
-    time_distance_correction_random$data
-  )
-)
-
-model_resid$model <- factor(
-  model_resid$model,
-  levels = rev(c(
-    "No corrections",
-    "Fixed correction +0.3s",
-    "Time correction fixed",
-    "Time correction random",
-    "Time and distance correction fixed",
-    "Time and distance correction (time correction random)",
-    "Time and distance correction random"
-  ))
-)
-
-model_resid$resid <- model_resid$time - model_resid$pred_time
-
-# Create SWC / SESOI band
-model_SESOI <- model_resid %>%
-  group_by(model, distance) %>%
-  summarise(
-    bias = mean(resid),
-    variance = sd(resid),
-    upper = bias + variance,
-    lower = bias - variance,
-    MAD = mean(abs(resid)),
-    SESOI_upper = sd(time) * 0.2,
-    SESOI_lower = -sd(time) * 0.2
-  )
-
-# Plot
-ggplot(model_resid, aes(y = model)) +
-  theme_bw(8) +
-  geom_vline(
-    data = model_SESOI,
-    aes(xintercept = SESOI_lower),
-    color = "blue", alpha = 0.5, linetype = "dashed"
-  ) +
-  geom_vline(
-    data = model_SESOI,
-    aes(xintercept = SESOI_upper),
-    color = "blue", alpha = 0.5, linetype = "dashed"
-  ) +
-  geom_vline(xintercept = 0, color = "blue", alpha = 0.5) +
-  geom_jitter(aes(x = resid), alpha = 0.1, height = 0.25) +
-  geom_errorbarh(
-    data = model_SESOI,
-    aes(xmin = lower, xmax = upper),
-    height = 0.1, color = "black"
-  ) +
-  geom_point(data = model_SESOI, aes(x = bias), color = "black") +
-  facet_wrap(~distance, scales = "free_x") +
-  xlab("Observed time - predicted time (s)") +
-  ylab(NULL)
-```
-
-The following figure depicts model residuals estimators (bias, or mean residual; variance, or SD of the residuals, and MAD, or mean absolute difference).
-
-```{r echo=FALSE}
-df <- model_SESOI %>%
-  pivot_longer(cols = -(1:2), names_to = "estimator") %>%
-  filter(estimator %in% c("bias", "variance", "MAD"))
-
-df$model <- factor(
-  df$model,
-  levels = rev(c(
-    "No corrections",
-    "Fixed correction +0.3s",
-    "Time correction fixed",
-    "Time correction random",
-    "Time and distance correction fixed",
-    "Time and distance correction (time correction random)",
-    "Time and distance correction random"
-  ))
-)
-
-df$estimator <- factor(
-  df$estimator,
-  levels = c("bias", "variance", "MAD")
-)
-
-ggplot(df, aes(x = value, y = model)) +
-  theme_bw(8) +
-  geom_point() +
-  facet_grid(distance ~ estimator, scales = "free_x") +
-  xlab(NULL) +
-  ylab(NULL)
-```
+kable(bolt_model[c(1, 7, 8)], digits = 2, booktabs = TRUE)
+```
+
+# Conclusion
 
 Which model should be used?
-Although providing a better fit (using RSE as an estimator of model fit), the time and distance correction models often estimate these parameters that are harder to interpret (e.g., negative distance correction).
-Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice, against gold-standard methods, assessing their agreement, as well as their power in detecting and adjusting for timing inconsistencies.
-More thorough theoretical simulation study is currently in development with the aim to explore the behavior of these models under different scenarios.
-
-We are hoping that the **shorts** package will help fellow sports scientists and coaches in exploring short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes, but also robust enough to take into account potential sprint initiation and timing inconsistencies.
+Although providing novel theoretical models in this paper, we acknowledge the need for validating them in practice against gold-standard methods, assessing their agreement, and their power in detecting and adjusting for timing inconsistencies.
+A more thorough theoretical simulation study is currently in development, intending to explore the behavior of these models under different scenarios.
+
+We hope that the **{shorts}** package will help fellow sports scientists and coaches explore short sprint profiles and help in driving research, particularly in devising measuring protocols that are sensitive enough to capture training intervention changes but also robust enough to take into account potential sprint initiation and timing inconsistencies.
 ```{r include=FALSE}
@@
 
-# Supplemental material
+# Supplemental Material
 

### 2022-07-21 shorts-paper.Rmd
Removed visible chunk of code that shouldn't be there

@@
 Here is an example using a built-in dataset.
-We are going to filter out one athlete (e.g., John) with the help of the **tidyverse** [@R-tidyverse] package and estimate MSS, TAU, and MAC parameters using the `model_timing_gates()` function:
+We are going to filter out one athlete (e.g., John) with the help of the **{tidyverse}** [@R-tidyverse] package and estimate MSS, TAU, and MAC parameters using the `model_timing_gates()` function:
 
@@
 
-To create a simple plot of the model, use the S3 `plot()` method, which returns **ggplot2** [@R-ggplot2] object:
+To create a simple plot of the model, use the S3 `plot()` method, which returns **{ggplot2}** [@R-ggplot2] object:
 
@@
 
-Plotting the model predictions can be done once we convert data from wide to long with the help of **ggplot2** [@R-ggplot2], **dplyr** [@R-dplyr], and **tidyr** [@R-tidyr] packages (loaded already using the **tidyverse** [@R-tidyverse] package):
+Plotting the model predictions can be done once we convert data from wide to long with the help of **{ggplot2}** [@R-ggplot2], **{dplyr}** [@R-dplyr], and **{tidyr}** [@R-tidyr] packages (loaded already using the **{tidyverse}** [@R-tidyverse] package):
 
@@
 
-### How to overcome missing the initial force production when using timing gates?
+## How to overcome missing the initial force production when using timing gates?
 

### 2022-07-21 shorts-paper.Rmd
small edits

@@
 Short sprints profiling is usually performed by: (1) measuring split times using timing gates (i.e., positioned at various distances, e.g., 5, 10, 20, 30, 40 $m$), or (2) getting a velocity trace using a radar gun.
-Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **{shorts}** package for the **R language** [@R-base] using *non-linear least squares regression* implemented in the `nlsLM()` function from the **{minpack.lm}** package [@R-minpack.lm].
+Estimation of MSS and TAU parameters from equation \@ref(eq:velocity-time) is performed in **{shorts}** package using *non-linear least squares regression* implemented in the `nlsLM()` function from the **{minpack.lm}** package [@R-minpack.lm].
 
@@
 There is the additional parameter in the output, TC, that is estimated using the `model_radar_gun()` function.
-The function of this parameter is explained in the [Problems with time sync with the radar gun] section of this paper.
+The utility of this parameter is explained in the [Problems with time sync with the radar gun] section of this paper.
 
